### 2018-04-30 paper/paper.tex
Committing a first paper structure.

\documentclass{llncs}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,rotating}
\usepackage[font=small]{subfig,caption}
\usepackage{algorithm2e}
\usepackage{todonotes,lineno,paralist}
\usepackage{placeins,rotating,hyperref}
\graphicspath{{figures/}}

\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
\title{A Heuristic Approach towards Drawings of Graph with High Crossing Resolution}

\institute{
Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
\texttt{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de}\\
\texttt{\{christian-marius.geckeler,amadaeus.spallek,jan.splett\}@student.uni-tuebingen.de}
}

% ============================================================
\begin{document}
\maketitle
% ==================================================================

\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:introduction}

In Graph Drawing, there exist a really reach literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg}. However, the task of drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is difficult and challenging, and as a result the proposed techniques are significantly fewer.

\section{Related Work}
\label{sec:relatedwork}

\section{Description of our Heuristic Approach}
\label{sec:algorithm}

\section{Experimental Evaluation}
\label{sec:experiments}

\section{Conclusions}
\label{sec:conclusions}

\bibliographystyle{abbrvurl}
\bibliography{references}

\end{document}

### 2018-04-30 paper/relationships.tex
Committing a first paper structure.

In this section, we consider relationships between the classes of graphs that admit smooth orthogonal $k$-drawings and octilinear $k$-drawings, $k \geq 1$, denoted as $SC_k$ and $8C_k$, respectively. Our findings are also summarized in Fig.~\ref{fig:graphClasses}.

By definition, $SC_1 \subseteq SC_2$ and $8C_1 \subseteq 8C_2 \subseteq 8C_3$ hold. Since each planar graph of max-degree~$8$ admits an octilinear $3$-drawing~\cite{slopes}, class $8C_3$ coincides with the class of planar graphs of max-degree~$8$. Similarly, class $SC_2$ coincides with the class of planar graph of max-degree~$4$, as these graphs admit smooth~orthogonal $2$-drawings~\cite{smog2}. This also implies that $SC_2 \subseteq 8C_2$, since each planar graph of max-degree~$4$ admits an octilinear $2$-drawing~\cite{octi}. The relationship $8C_2 \neq 8C_3$ follows from~\cite{octi}, where it was proven that there exist planar graphs of max-degree~$6$ that do not admit octilinear $2$-drawings. The relationship $SC_2 \neq 8C_2$ follows from~\cite{octi-2}, where it was shown that there exist planar graphs of max-degree~$5$ that admit octilinear $2$-drawings and no octilinear $1$-drawings, and the fact that planar graphs of max-degree~$5$ cannot be drawn in the smooth orthogonal model. The octahedron graph admits neither a bendless smooth orthogonal drawing~\cite{smog1} nor a bendless octilinear drawing~\cite{octi-2}. However, since it is of max-degree~$4$, it admits $2$-drawings in both models~\cite{smog2,octi}. Hence, it belongs to $8C_2 \cap SC_2 \setminus (8C_1 \cup SC_1)$. To prove that $8C_1 \setminus SC_2 \neq \emptyset$, observe that a caterpillar whose spine vertices are of degree 8 clearly admits an octilinear $1$-drawing, however, due to its degree it cannot be drawn in the smooth orthogonal model.

To complete the discussion of the relationships of Fig.~\ref{fig:graphClasses}, we have to show~that $SC_1$ and $8C_1$ are incomparable. This is the most interesting part of our proof, as usually one can ``easily'' convert a bendless octilinear drawing of a planar graph of max-degree~$4$ to a corresponding bendless smooth orthogonal one (e.g., by replacing diagonal segments with quarter circular arcs), and vice versa; see, e.g., Figs.~\ref{fig:octilinear}-\ref{fig:smooth}. Since the endpoints of each edge of a bendless smooth orthogonal or octilinear drawing are along a line with slope $0$, $1$, $-1$ or~$\infty$, such conversions are in principle possible. Two difficulties that might arise are to preserve planarity and to guarantee that no two edges enter a vertex using the same port. Clearly, however, there exist infinitely many (even $4$-regular) planar graphs that admit both drawings in both models; Fig.~\ref{fig:trains} shows how one can construct them; see also Appendix~\ref{app:relationships}. We summarize this observation in the~following theorem.

\begin{figure}[t!]
	\centering
	\begin{minipage}[b]{0.45\textwidth}
		\centering
		\subfloat[\label{fig:trainsSC1} {A smooth orthogonal $1$-drawing}]{
		\includegraphics[scale=0.65,page=3]{relationships}}
	\end{minipage}
	\hfil
	\begin{minipage}[b]{0.45\textwidth}
		\subfloat[\label{fig:trains8C1} {An octilinear $1$-drawing}]{
		\includegraphics[scale=0.65,page=2]{relationships}}
	\end{minipage}
	\caption{Illustrations for the proof of Theorem~\ref{theo:union}.}
\label{fig:trains}
\end{figure}

\newcommand{\union}{There is an infinitely large family of $4$-regular planar graphs that admit both bendless smooth orthogonal and bendless octilinear drawings.}
\begin{theorem}
\union
\label{theo:union}
\end{theorem}

\noindent In the next two theorems we show that $SC_1$ and $8C_1$ are incomparable.

\begin{theorem}
There is an infinitely large family of $4$-regular planar graphs that admit bendless smooth orthogonal drawings but no bendless octilinear drawings.
\label{theo:smoothAndNotOcti}
\end{theorem}
\begin{proof}
Consider the planar graph $C$ of Fig.~\ref{fig:4gonOctahedronComponentSC1}, which is drawn bendless smooth orthogonal. We claim that $C$ admits no bendless octilinear drawing. If one substitutes its degree-$2$ vertex (denoted by~$c$ in Fig.~\ref{fig:4gonOctahedronComponentSC1}) by an edge connecting its two neighbors, then the resulting graph is triconnected, which admits an unique embedding (up to the choice of its outerface). Hence, in the presence of the~degree-$2$ vertex, graph $C$ has exactly two embeddings; see Figs.~\ref{fig:4gonOctahedronComponentSC1}-\ref{fig:4gonOctahedronComponentEmbedding2}. Now, observe that the outerface of any octilinear drawing of graph $C$ (if any) has length at most $5$ (Constraint~$1$). In addition, each vertex of this outerface (except for $c$, which is of degree~$2$) must have two ports pointing in the interior of this drawing, because every vertex of $C$ is of degree~$4$ except for $c$. This implies that the angle formed by any two consecutive edges of this outerface is at most $225^\circ$, except for the pair of edges incident to $c$ (Constraint~$2$). But if we want to satisfy both constraints, then at least an edge of this outerface must be drawn with a bend; see Fig.~\ref{fig:4gonOctahedronComponentNotOcC1}. Hence, graph $C$ does not admit a bendless octilinear drawing.

\begin{figure}[t!]
	\centering
	\begin{minipage}[b]{0.18\textwidth}
		\centering
		\subfloat[\label{fig:4gonOctahedronComponentSC1} {}]{
		\includegraphics[scale=0.5,page=5]{relationships}}
	\end{minipage}
	\hfil
	\begin{minipage}[b]{0.18\textwidth}
		\centering
		\subfloat[\label{fig:4gonOctahedronComponentEmbedding2} {}]{
		\includegraphics[scale=0.5,page=6]{relationships}}
	\end{minipage}
	\hfil
	\begin{minipage}[b]{0.18\textwidth}
		\centering
		\subfloat[\label{fig:4gonOctahedronComponentNotOcC1} {}]{
		\includegraphics[scale=0.5,page=7]{relationships}}
	\end{minipage}
	\hfil
	\begin{minipage}[b]{0.36\textwidth}
		\centering
		\subfloat[\label{fig:4gonOctahedronChain} {}]{
		\includegraphics[scale=0.5,page=4]{relationships}}
	\end{minipage}
	\caption{%
	Illustrations for the proof of Theorem~\ref{theo:smoothAndNotOcti}.}
\label{fig:4gonOctahedron}
\end{figure}

Based on graph $C$, for each $k \in \mathbb{N}_0$ we construct a $4$-regular planar graph $G_k$ consisting of $k + 2$ biconnected components $C_1,\ldots,C_{k+2}$ arranged in a chain; see Fig.~\ref{fig:4gonOctahedronChain} for the case $k=1$. Clearly, $G_k$ admits a bendless smooth orthogonal drawing for any $k$. Since the end-components of the chain (i.e., $C_1$ and $C_{k+2}$)~are isomorphic to $C$, $G_k$ does not admit a bendless octilinear drawing for any $k$.
\end{proof}

\begin{theorem}
There is an infinitely large family of $4$-regular planar graphs that admit bendless octilinear drawings but no bendless smooth orthogonal drawings.
\label{theo:octiAndNotSmooth}
\end{theorem}
\begin{proof}[sketch]
Consider the planar graph $B$ of Fig.~\ref{fig:necklaceGem}, which is drawn bendless octilinear. Graph $B$ has two separation pairs (i.e., $\{t_1,t_2\}$ and $\{p_1,p_2\}$ in Fig.~\ref{fig:necklaceGem}). If we require the outerface of $B$ to be the one of Fig.~\ref{fig:necklaceGem}, then all possible planar embeddings of $B$ are isomorphic to the one of Fig.~\ref{fig:necklaceGem}. We exploit this property in Lemma~\ref{lem:necklaceGemNotOcti} of Appendix~\ref{app:relationships} to show that $B$ does not admit a bendless smooth~orthogonal drawing with this outerface. The detailed proof is based on an exhaustive consideration of all bendless smooth orthogonal drawings of subgraphs of $B$, which we incrementally augment by adding more vertices~to~them.

\begin{figure}[h]
	\centering
	\subfloat[\label{fig:necklaceGem} {}]{
	\includegraphics[scale=0.38,page=8]{relationships}}
	\hfil
	\subfloat[\label{fig:necklaceConstruction} {}]{
	\includegraphics[scale=0.38,page=9]{relationships}}
	\caption{Illustrations for the proof of Theorem~\ref{theo:octiAndNotSmooth}.}
	\label{fig:necklace}
\end{figure}

Based on graph $B$, for each $k \in \mathbb{N}_0$ we construct a $4$-regular planar graph $G_k$ consisting of $2k + 2$ copies of $B$ arranged in a cycle; see Fig.~\ref{fig:necklaceConstruction} where each copy of $B$ is drawn as a gray-shaded parallelogram. By construction, $G_k$ admits a bendless octilinear drawing for any $k$. Since by planarity at least one copy of graph $B$ must be embedded with the outerface of Fig.~\ref{fig:necklaceGem}, for any~$k$, graph $G_k$ does not admit a bendless smooth orthogonal drawing, as well.
\end{proof}

% This completes our comparison of graph classes which shows that density does not
% divide $SC_1$ and $OC_1$ and that while there is an infinitely large
% intersection, the two classes are incomparable.

### 2018-05-02 paper/paper.tex
Working a bit more on the paper and add a new K_5 example

@@
 \author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
-\title{A Heuristic Approach towards Drawings of Graph with High Crossing Resolution}
+\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
 
@@
 \begin{abstract}
+The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed at the crossing points of the edges.
+Recent experiments have shown that the larger the crossing resolution, the easier it is to read and interpret a drawing.
+However, maximizing the crossing resolution of a graph turns out to be a difficult problem and only heuristic algorithms are known.
 
+In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem.
+The core of our algorithm is based on a randomized procedure, which starting from an arbitrary input drawing, yields one with better crossing resolution by moving ...
 \end{abstract}
@@
 
-In Graph Drawing, there exist a really reach literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg}.
-However, the task of drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is difficult and challenging, and as a result the proposed techniques are significantly fewer.
+In Graph Drawing, there exist a really reach literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
+As a result, the proposed techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
+
+In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important statistical evidence (through a series of eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and therefore the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
+In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality and readability of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
+As a result, a new and prominent research direction was initiated, which is unofficially recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/jgaa/BekosKM18,DBLP:journals/corr/abs-1804-07257} for recent surveys.
 
@@
 \label{sec:relatedwork}
+
+%Spring embedders, energy based methods, which changed our viewpoint towards edge crossings
 

### 2018-05-02 paper/paper.tex
Working a bit more on the paper.

@@
 The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed at the crossing points of the edges.
-Recent experiments have shown that the larger the crossing resolution, the easier it is to read and interpret a drawing.
-However, maximizing the crossing resolution of a graph turns out to be a difficult problem and only heuristic algorithms are known.
+Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
+However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting spring embedding algorithms.
 
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem.
-The core of our algorithm is based on a randomized procedure, which starting from an arbitrary input drawing, yields one with better crossing resolution by moving ...
+In this paper, we propose a new --randomization-based-- heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it again the known ones from the literature.
+Our experimental evaluation indicates that the new heuristic significantly outperforms the known ones in terms of the quality of the produced layouts (in terms of the actual crossing resolution), but this comes at the cost of slightly higher running time, especially for large and dense graphs.
 \end{abstract}
@@
 
+Formally, the minimum angle formed by any two crossed edges in a drawing is refer to as its \emph{crossing resolution}.
+Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is refer to as \emph{right-angle-crossing} graph or \emph{RAC} graph, for short.
+For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
+Notably, RAC graph are sparse (they contain at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, where $n$ denotes the number of  vertices), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+
+The latter result is already an indication that the crossing resolution maximization problem might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
+Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
+
+From a practical point of view, we are only aware of two methods that aim in drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms.
+The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
+Common in both algorithms are appropriate forces, which are applied on the endvertices of every pair of crossing edges.
+Of course, each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
+
+In this paper, we attack the crossing resolution maximization problem by following a completely different approach.
+We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which in practise turned out to be quite efficient\footnote{Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to computed the expected quality of the produced drawings.}.
+However, since we could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
+To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
+Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of spring embedding algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly higher running time, especially for large and dense graphs.
+
+The remainder of this paper is structured as follows.
+Section~\ref{sec:relatedwork} overviews related results.
+In Section~\ref{sec:preliminaries}, we introduce preliminary notions and notation used throughout the paper.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
+We conclude in Section~\ref{sec:conclusions} with open problems and future direction.
+
 \section{Related Work}
@@
 
-%Spring embedders, energy based methods, which changed our viewpoint towards edge crossings
+\section{Preliminaries}
+\label{sec:preliminaries}
 

### 2018-05-05 src/Experiment/Experiment.java
Working experimente with contest files.
Working experimente with contest files.

@@ -1,0 +1,353 @@
+package Experiment;
+
+import algorithms.graphs.GridGraph;
+import com.yworks.yfiles.geometry.RectD;
+import com.yworks.yfiles.graph.IEdge;
+import com.yworks.yfiles.graph.IGraph;
+import com.yworks.yfiles.graph.INode;
+import com.yworks.yfiles.graph.LayoutUtilities;
+import com.yworks.yfiles.graphml.GraphMLIOHandler;
+import com.yworks.yfiles.layout.organic.OrganicLayout;
+import io.ContestIOHandler;
+import io.GraphIOHandler;
+import layout.algo.randommovement.RandomMovementLayout;
+import layout.algo.utils.PositionMap;
+import main.MainFrame;
+import sidepanel.SidePanelTab;
+import util.BoundingBox;
+
+import java.io.*;
+import java.nio.file.*;
+import java.util.ArrayList;
+import java.util.Optional;
+
+/**
+ * Created by Ama on 27.04.2018.
+ */
+public class Experiment {
+
+    private static String outputDirectory, inputDirectory, fileName, filePrefix;
+
+    private MainFrame mainFrame;
+    private IGraph graph;
+    private long sartTime, endTime, calcTime;
+    private boolean reached90Degree = false;
+    private PrintWriter pw;
+    private StringBuilder sb;
+    private int maxIterations = 1000;
+    private int iterations = 0;
+    private double bestAngle = 0;
+
+    private int totalVertecies, totalEdges;
+    private double currentGraphSizeX, currentGraphSizeY, minimumAngle, testAngle;
+
+    public  Experiment (MainFrame mainFrame, String fileName, String inputDir, String outputDir) {
+        this.mainFrame = mainFrame;
+        this.graph = mainFrame.graph;
+        this.fileName = fileName;
+        this.filePrefix = getPrefixString(fileName);
+        this.inputDirectory = inputDir;
+        this.outputDirectory = outputDir;
+        this.calcTime = 0;
+
+        //sb = new StringBuilder();
+
+        try {
+            pw = new PrintWriter(new File(outputDirectory + "\\" + filePrefix + ".csv"));
+
+            sb = new StringBuilder();
+            sb.append("Graph Name");
+            sb.append(';');
+            sb.append("Time");
+            sb.append(';');
+            sb.append("Iterations");
+            sb.append(';');
+            sb.append("Angle");
+            sb.append(';');
+            sb.append('\n');
+
[…]

### 2018-05-05 src/Experiment/Experiments.java

@@ -1,0 +1,204 @@
+package Experiment;
+
+import algorithms.graphs.GridGraph;
+import com.yworks.yfiles.algorithms.GraphChecker;
+import com.yworks.yfiles.geometry.RectD;
+import com.yworks.yfiles.graph.*;
+
+/**
+ * Created by Ama on 24.04.2018.
+ */
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.IOException;
+import java.io.PrintWriter;
+import java.nio.file.Paths;
+import java.util.ArrayList;
+import java.util.Optional;
+import java.util.stream.Collectors;
+import java.util.stream.IntStream;
+
+import com.yworks.yfiles.layout.YGraphAdapter;
+import com.yworks.yfiles.layout.organic.OrganicLayout;
+import com.yworks.yfiles.view.GraphComponent;
+import io.ContestIOHandler;
+import layout.algo.randommovement.RandomMovementLayout;
+import layout.algo.utils.PositionMap;
+import main.MainFrame;
+import sidepanel.SidePanelTab;
+import util.BoundingBox;
+
+import javax.swing.*;
+
+public class Experiments {
+
+    //private y.view.Graph2DView view;
+     private GraphComponent comp;
+    //private static String inputDirectory  = "E:\\graph\\origin\\";
+
+//    private static String inputDirectory  = "E:\\graph\\north\\north_graphml\\north\\";
+  //  private static String outputDirectory = "E:\\graph\\afterRandom\\north\\";
+
+   private static String inputDirectory  = "E:\\graph\\graphml\\";
+    private static String outputDirectory = "E:\\graph\\afterRandom\\graphml\\";
+
+
+
+
+    /* Config. for the type of experiment */
+    private final boolean time = false;
+    private final boolean iterations = false;
+
+    private PrintWriter pw;
+    private StringBuilder sb;
+    private int numThread = 6;
+    private int numOfIteration = 100;
+    private int numOfIterationFactor = 1000;
+    private long maxCalcTime = 6000; //in mili. sec
+    private boolean withPlanarGraphs = true;
+    private boolean withConnectedGraphs = true;
+    private String[] childernP;
+
+
+    public Experiments()
+    {
+        this.comp = new GraphComponent();
+    }
+
+    public synchronized void run()
+    {
+            java.io.File dir = new java.io.File(inputDirectory);
+         /*   System.out.println(dir.getName());
+
+
+            sb = new StringBuilder();
+            sb.append("Graph Name");
[…]

### 2018-05-05 src/Experiment/Experiments.java
Added simple GraphMl IO-handler (read)

@@ -100,7 +100,8 @@ public class Experiments {
             else {
 
                     childernP = children;
-                    for(int i = 0; i<children.length; i++){
+                    for(int i = 0; i<1; i++){
+                    //for(int i = 0; i<children.length; i++){
                     //    openFrames(i, children.length, inputDirectory, "pattern");
                         openFrames(i, i, inputDirectory, children[i]);
                         //System.out.println("Iteration:   "+ i);
@@ -134,7 +135,7 @@ public class Experiments {
     private void openFrames(int startIndex, int endIndex, String folderPath, String pattern) {
         IntStream.range(startIndex, endIndex + 1)
                 .parallel()
-                .forEach(index -> MainFrame.start(WindowConstants.DISPOSE_ON_CLOSE, false, mainFrame -> {
+                .forEach(index -> MainFrame.start(WindowConstants.DISPOSE_ON_CLOSE, true, mainFrame -> {
 
                     //String fileName = index +"";
                     String fileName = pattern;
@@ -142,7 +143,7 @@ public class Experiments {
 
                     Experiment experiment = new Experiment(mainFrame, fileName, inputDirectory, outputDirectory);
                     experiment.loadGraph();
-                    System.out.println("Was da los    1");
+                   System.out.println("Was da los    1");
                     boolean isNotFinished = true;
                     int iterationFactor = 1;
                     boolean reached90Deg = false;
@@ -177,8 +178,9 @@ public class Experiments {
                     experiment.writeGraph();
                     mainFrame.dispose();
 
+                }
 
-                }}));
+    }));
     }
 
     private String getPrefixString(String str){

### 2018-05-05 src/io/SimpleGraphmlIOHandler.java

@@ -1,0 +1,164 @@
+package io;
+
+import com.yworks.yfiles.algorithms.INodeMap;
+import com.yworks.yfiles.geometry.PointD;
+import com.yworks.yfiles.graph.IEdge;
+import com.yworks.yfiles.graph.IGraph;
+import com.yworks.yfiles.graph.IMapper;
+import com.yworks.yfiles.graph.INode;
+import com.yworks.yfiles.graphml.GraphMLIOHandler;
+
+import java.io.IOException;
+import java.nio.file.Files;
+import java.nio.file.Path;
+import java.nio.file.Paths;
+import java.util.List;
+
+/**
+ * Created by Ama on 05.05.2018.
+ */
+public class SimpleGraphmlIOHandler extends GraphIOHandler{
+
+    static List<String> lines;
+
+    public static void read(IGraph g, String inputFileName) throws  IOException {
+        Path path = Paths.get(inputFileName);
+        lines = Files.readAllLines(path);
+
+        g.clear();
+
+        int phase = 0;
+        int nodesCount = -1;
+        int maxNodes = 0;
+        int edgesCount = -1;
+        int nodesAdded = 0;
+
+
+        for(String line: lines) {
+            line = line.trim();
+
+            if (line.isEmpty() || line.equals("") || line.equals("\n")) {
+                continue;
+            }else if(line.charAt(1) == 'n') {
+                maxNodes++;
+            }
+        }
+
+        INode[] nodes = new INode[maxNodes];
+        String[] nodesId = new String[maxNodes];
+
+
+        for(String line: lines){
+            line = line.trim();
+
+            if(line.isEmpty() || line.equals("") || line.equals("\n")) { continue;}
+            else if(line.charAt(1) == 'n') {
+                int index = line.indexOf("\"");
+                line = line.substring(index+1);
+                index = line.indexOf("\"");
+                String objString  = line.substring(0,index);
+                nodesCount++;
+                INode node = g.createNode(new PointD(nodesCount, nodesCount));
+                node.setTag(objString);
+              //  System.out.println("nodeCount  " +nodesCount);
+                nodes[nodesCount] = node;
+                nodesId[nodesCount] = objString;
+
+            }
+            else if(line.charAt(1) == 'e') {
+
+                int index = line.indexOf("\"");
+                line = line.substring(index+1);
+                index = line.indexOf("\"");
+                String edgeID  = line.substring(0,index);
+
+
+                line = line.substring(index+1);
[…]

### 2018-05-05 src/main/MainFrame.java

@@ -14,6 +14,7 @@ import com.yworks.yfiles.view.*;
 import com.yworks.yfiles.view.input.*;
 import graphoperations.Scaling;
 import io.ContestIOHandler;
+import io.SimpleGraphmlIOHandler;
 import layout.algo.execution.ILayout;
 import layout.algo.utils.BestSolutionMonitor;
 import sidepanel.InitSidePanel;
@@ -373,7 +374,8 @@ public class MainFrame extends JFrame {
     public void openFile(String fileNamePath) {
         initSidePanel.removeDefaultListeners();
         try {
-            view.importFromGraphML(fileNamePath);
+          //  view.importFromGraphML(fileNamePath);
+            SimpleGraphmlIOHandler.read(graph, fileNamePath);
             view.fitGraphBounds();
             view.updateUI();
             bestSolution.reset();

### 2018-05-05 src/Experiment/Experiment.java
SimpleGraphIOHandler and Experiment
The IOHandler handls now edges without an id.
Experimnts got new constructors and some abort condition.

@@ -72,9 +72,8 @@ public class Experiment {
         catch(FileNotFoundException ex){
             System.out.println(ex);
         }
-
-
     }
+
     public void run(){
 
         loadGraph();
@@ -100,6 +99,7 @@ public class Experiment {
         }
 
     }
+
     public void writeGraphInformations(){
         sb = new StringBuilder();
         sb.append(this.filePrefix);
@@ -155,6 +155,8 @@ public class Experiment {
     }
 
     public void loadGraph() {
+        this.reached90Degree = false;
+
         if (this.fileName.endsWith(".txt")){
             this.mainFrame.openContestFile(this.inputDirectory + this.fileName);
             System.out.println("Try to open .txt");
@@ -165,7 +167,7 @@ public class Experiment {
 
         mainFrame.minimumAngleMonitor.updateMinimumAngleInfoBar();
         this.minimumAngle = Math.round(1000.0 * mainFrame.minimumAngleMonitor.getMinAngle()) / 1000.0;
-        System.out.println("laod Graph " + this.minimumAngle);
+        System.out.println("load Graph " + this.minimumAngle);
     }
 
     public void writeGraph() {
@@ -177,28 +179,39 @@ public class Experiment {
     }
 
     public  void runAlgorithms() {
+        //System.out.println("++ 1");
         this.mainFrame.initSidePanel.removeDefaultListeners();
+        // System.out.println("++ 2");
         LayoutUtilities.applyLayout(this.mainFrame.graph, new OrganicLayout());
+        //System.out.println("++ 3");
         this.mainFrame.initSidePanel.addDefaultListeners();
-
+        //System.out.println("++ 4");
         Optional<SidePanelTab> tab = this.mainFrame.getTabForAlgorithm(RandomMovementLayout.class);
+        // System.out.println("++ 5");
         if (!tab.isPresent()) {
             return;
         }
+        //System.out.println("++ 6");
         tab.get().getExecutor().setMaxIterations(maxIterations);
+        //System.out.println("++ 7");
         long tmpTime = System.currentTimeMillis();
+        // System.out.println("++ 8");
         tab.get().startPauseExecution();
         boolean value = true;
-        while(!tab.get().getExecutor().isFinished())
-        {
 
+       // System.out.println("++ 8.5");
+        //while(!tab.get().getExecutor().isFinished()){
[…]

### 2018-05-05 src/Experiment/Experiments.java

@@ -1,33 +1,22 @@
 package Experiment;
 
-import algorithms.graphs.GridGraph;
 import com.yworks.yfiles.algorithms.GraphChecker;
-import com.yworks.yfiles.geometry.RectD;
 import com.yworks.yfiles.graph.*;
 
 /**
  * Created by Ama on 24.04.2018.
  */
 
-import java.io.File;
-import java.io.FileNotFoundException;
-import java.io.IOException;
 import java.io.PrintWriter;
-import java.nio.file.Paths;
-import java.util.ArrayList;
 import java.util.Optional;
-import java.util.stream.Collectors;
 import java.util.stream.IntStream;
 
 import com.yworks.yfiles.layout.YGraphAdapter;
 import com.yworks.yfiles.layout.organic.OrganicLayout;
 import com.yworks.yfiles.view.GraphComponent;
-import io.ContestIOHandler;
 import layout.algo.randommovement.RandomMovementLayout;
-import layout.algo.utils.PositionMap;
 import main.MainFrame;
 import sidepanel.SidePanelTab;
-import util.BoundingBox;
 
 import javax.swing.*;
 
@@ -40,24 +29,17 @@ public class Experiments {
 //    private static String inputDirectory  = "E:\\graph\\north\\north_graphml\\north\\";
   //  private static String outputDirectory = "E:\\graph\\afterRandom\\north\\";
 
-   private static String inputDirectory  = "E:\\graph\\graphml\\";
-    private static String outputDirectory = "E:\\graph\\afterRandom\\graphml\\";
-
-
+    private String inputDirectory  = "E:\\graph\\graphml\\";
+    private String outputDirectory = "E:\\graph\\afterRandom\\graphml2\\";
 
 
     /* Config. for the type of experiment */
-    private final boolean time = false;
-    private final boolean iterations = false;
-
-    private PrintWriter pw;
-    private StringBuilder sb;
     private int numThread = 6;
     private int numOfIteration = 100;
     private int numOfIterationFactor = 1000;
     private long maxCalcTime = 6000; //in mili. sec
-    private boolean withPlanarGraphs = true;
-    private boolean withConnectedGraphs = true;
+    private boolean planarGraphsAllowed = false;
+    private boolean unconnectedGraphsAllowed = true;
     private String[] childernP;
 
 
@@ -66,23 +48,36 @@ public class Experiments {
         this.comp = new GraphComponent();
     }
 
[…]

### 2018-05-05 src/Experiment/mainExperiment.java

@@ -6,7 +6,13 @@ package Experiment;
 public class mainExperiment {
     public static void main(String[] args)
     {
-        Experiments e = new Experiments();
+        //Experiments e = new Experiments();
+
+       // Experiments e = new Experiments(("E:\\graph\\graphml\\"), "E:\\graph\\afterRandom\\graphml2\\");
+        Experiments e = new Experiments(
+                "E:\\graph\\rome\\rome\\", "E:\\graph\\afterRandom\\rome\\",
+                10000, 100, 1000,
+                false, false);
         e.run();
     }
 }

### 2018-05-05 src/io/SimpleGraphmlIOHandler.java

@@ -56,28 +56,36 @@ public class SimpleGraphmlIOHandler extends GraphIOHandler{
                 int index = line.indexOf("\"");
                 line = line.substring(index+1);
                 index = line.indexOf("\"");
-                String objString  = line.substring(0,index);
+                String objString  = line.substring(1,index);
                 nodesCount++;
                 INode node = g.createNode(new PointD(nodesCount, nodesCount));
                 node.setTag(objString);
               //  System.out.println("nodeCount  " +nodesCount);
+                System.out.println(objString + "  node");
                 nodes[nodesCount] = node;
                 nodesId[nodesCount] = objString;
 
             }
             else if(line.charAt(1) == 'e') {
-
+                edgesCount++;
                 int index = line.indexOf("\"");
-                line = line.substring(index+1);
-                index = line.indexOf("\"");
-                String edgeID  = line.substring(0,index);
+
+                /* Test if en edge ID exists */
+                String idTest = line.substring(index-3, index);
+                String edgeID = edgesCount +"";
+               if(idTest.equals("id=")){
+                    line = line.substring(index+1);
+                    index = line.indexOf("\"");
+                    edgeID  = line.substring(0,index);
+                    line = line.substring(index+1);
+                    index = line.indexOf("\"");
+               }
 
 
                 line = line.substring(index+1);
                 index = line.indexOf("\"");
-                line = line.substring(index+1);
-                index = line.indexOf("\"");
-                String objString  = line.substring(0,index);
+                String objString  = line.substring(1,index);
+                System.out.println(objString + "  source");
                 int indexN1 = searchIndex(nodesId, objString);
                 if(indexN1 == -1){System.out.println("Maping Node1 failed");}
 
@@ -86,12 +94,13 @@ public class SimpleGraphmlIOHandler extends GraphIOHandler{
                 index = line.indexOf("\"");
                 line = line.substring(index+1);
                 index = line.indexOf("\"");
-                objString = line.substring(0,index);
+                objString = line.substring(1,index);
+                System.out.println(objString + "  target");
                 int indexN2 = searchIndex(nodesId, objString);
                 if(indexN1 == -1){System.out.println("Maping Node2 failed");}
 
 
-                edgesCount++;
+
                 IEdge edge = g.createEdge(nodes[indexN1], nodes[indexN2]);
[…]

### 2018-05-05 paper/paper.tex
Finished with the introduction.

@@
 In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality and readability of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, which is unofficially recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/jgaa/BekosKM18,DBLP:journals/corr/abs-1804-07257} for recent surveys.
+As a result, a new and prominent research direction was initiated, which is unofficially recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
 
@@
 
-From a practical point of view, we are only aware of two methods that aim in drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms.
+From a practical point of view, we are only aware of two methods that aim in drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms~\cite{DBLP:journals/congnum/Eades84}.
 The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
 Common in both algorithms are appropriate forces, which are applied on the endvertices of every pair of crossing edges.
-Of course, each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
+Each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
 
 In this paper, we attack the crossing resolution maximization problem by following a completely different approach.
-We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which in practise turned out to be quite efficient\footnote{Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to computed the expected quality of the produced drawings.}.
+We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which in practise turned out to be quite efficient\footnote{Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to computed the expected quality of the produced drawings.
+A notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.}.
 However, since we could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
@@
 The remainder of this paper is structured as follows.
-Section~\ref{sec:relatedwork} overviews related results.
+Section~\ref{sec:relatedwork} overviews related works.
 In Section~\ref{sec:preliminaries}, we introduce preliminary notions and notation used throughout the paper.
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
-We conclude in Section~\ref{sec:conclusions} with open problems and future direction.
+We conclude in Section~\ref{sec:conclusions} with interesting open problems and future~directions.
 
@@
 
+As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (propertied of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
+The study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
+Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
+Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
+Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} completely characterized the complete bipartite RAC graphs.
+Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
+Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
+Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
+Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
+Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
+
+To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
+More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ has at most $(3n-6)\pi/\alpha$ edges.
+Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
+
+An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem (as naturally expected, of course).
+The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
+Notable exceptions are of courses the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}.
+According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium state of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
+
 \section{Preliminaries}
 \label{sec:preliminaries}
+
+Unless otherwise specified, in this paper we consider simple undirected graphs.
+Let $G=(V,E)$ be such a graph.
+The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
+The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
+
+Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p_u=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
+For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p_u$ to $p_v$ is denoted as $\overrightarrow{p_up_v}$, while the line segment delimited by $p_u$ and $p_v$ as $\overline{p_up_v}$.
 

### 2018-05-08 paper/paper.tex
Working a bit more on the algorithm's description (not finished yet).

@@
 
-In this paper, we propose a new --randomization-based-- heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it again the known ones from the literature.
-Our experimental evaluation indicates that the new heuristic significantly outperforms the known ones in terms of the quality of the produced layouts (in terms of the actual crossing resolution), but this comes at the cost of slightly higher running time, especially for large and dense graphs.
+In this paper, we propose a new --randomization-based-- heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known ones from the literature.
+Our experimental evaluation indicates that the new heuristic produces drawings with significantly better crossing resolution, but this comes at the cost of slightly higher running time, especially when the input graph is large and dense.
 \end{abstract}
@@
 The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-Notable exceptions are of courses the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}.
+Exceptions are the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}.
 According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium state of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
@@
 
+In this section, we describe our heuristic algorithm for obtaining drawings with high crossing resolution.
+Without loss of generality, we assume that the input of our algorithm consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ that is of some crossing resolution $c(\Gamma_0)$.
+We may assume that no two edges of $G$ overlap in $\Gamma_0$, that is, $c(\Gamma_0)>0$.
+Note that a drawing $\Gamma_0$ of $G$ meeting these preconditions can be, e.g., a circular drawing of $G$ or a drawing obtained by applying a standard spring embedding algorithm on $G$.
+
+Our algorithm is iterative and at each iteration step performs some operations that are mainly based on randomization.
+More precisely, at the $i$-th iteration step, we assume that we have computed a drawing $\Gamma_{i-1}$ of some crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$, where $\Gamma_0$ is our initial drawing.
+In other words, we assume, as an invariant property of our algorithm, that the crossing resolution cannot be decreased at some iteration step.
+Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrary at random from a so-called \emph{vertex-pool}, which may contain:
+
+\begin{itemize}
+\item either all the vertices of drawing $\Gamma_{i-1}$, or
+\item a prespecified subset of the vertices of drawing $\Gamma_{i-1}$, which we call \emph{critical}.
+\end{itemize}
+
+To formally define the critical vertices, we first need to introduced the notion of critical edge-pairs.
+A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is either equal (or very close, in some cases) to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
+The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
+
+Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
+As a result, their role in our algorithm is crucial.
+The reason is that by slightly changing the location of a critical vertex appropriately, one naturally expects to improve the crossing resolution of the current drawing.
+In fact, what quickly we realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local minima after a number of iterations.
+Hence, special care is needed in order to avoid these bottlenecks, especially when the graph to be drawn is large.
+More precisely, in such a case an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of the critical vertices to move.
+We discuss heuristics to avoid such bottlenecks in  Section~\ref{ssec:minima}.
+
+So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex from the vertex-pool of the current drawing to move, so to improve the crossing resolution.
+We have not described, however, how the chosen vertex is moved, i.e., how to compute its new position.
+To this end, let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm.
+
+\begin{itemize}
+\item Description of the randomized algorithm.
+\item Big graphs $\rightarrow$ focus on critical edges.
+\item Mimics the human. 
+\item Avoiding local minima.
+\item Crossing calculation. 
+\end{itemize}
+
+\subsection{Avoiding local minima.}
+\label{ssec:minima}
+
+\subsection{Grid drawings: an interesting variant.}
+\label{ssec:grid}
+
 \section{Experimental Evaluation}
 \label{sec:experiments}
+
+\begin{itemize}
+\item Experimental setup.
+\item Benchmarks + degree-3.
+\item Aesthetics:
+
+\begin{itemize}
+\item Number of crossings (\# of crossings).
+\item Average size of crossing angles (angle size).
+\item Average of edge lengths (edge length).
+\item Angular resolution (angular res.).
+\end{itemize}
+
+\item Results.
+\item Discussion.
+\end{itemize}
 

### 2018-05-08 paper/paper.tex
Working a bit more one the algorithm; finished the description of the main part.

@@
 \usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist}
+\usepackage{todonotes,lineno,paralist,soul}
 \usepackage{placeins,rotating,hyperref}
@@
 
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
+\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
 \title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
@@
 \texttt{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{christian-marius.geckeler,amadaeus.spallek,jan.splett\}@student.uni-tuebingen.de}
+\texttt{\{christian-marius.geckeler,jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
 }
@@
 
-Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p_u=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p_u$ to $p_v$ is denoted as $\overrightarrow{p_up_v}$, while the line segment delimited by $p_u$ and $p_v$ as $\overline{p_up_v}$.
+Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
+For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
 
@@
 So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex from the vertex-pool of the current drawing to move, so to improve the crossing resolution.
-We have not described, however, how the chosen vertex is moved, i.e., how to compute its new position.
-To this end, let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm.
+We have not described, however, how the chosen vertex is moved, i.e., how we compute its new position in the next drawing.
+
+Let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm.
+Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
+To compute the position of vertex $v_i$ in drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from the point $p(v_i)$ of vertex $v_i$ in drawing $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
+These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$.
+The position of vertex $v_i$ in drawing $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
+More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[0,\delta]$, where $\delta$ is also a parameter of the algorithm.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$\todo{Add a figure illustrating the approach.}.
+
+If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the new position of vertex $v_i$ in drawing $\Gamma_i$ is $p(v_i)$, that is, the same as in drawing $\Gamma_{i-1}$, since the crossing resolution of drawing $\Gamma_i$ must be at least as large as the one of $\Gamma_{i-1}$.
+On the other hand, if there is one or more feasible points (among $\pi_0,\pi_1,\ldots,\pi_{\rho-1}$), then one may consider two different approaches to determine the new position of vertex $v_i$ in drawing $\Gamma_i$.
+The most natural is to chose this feasible point that maximizes the crossing resolution of the obtained drawing.
+As an alternative, one may rely again on randomization and chose uniformly at random one the feasible points as the new position of vertex $v_i$ in drawing $\Gamma_i$.
+In our implementation, we did not observe any significant difference in these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
+In Section~\ref{ssec:termination}, we describe conditions  The iteration steps
 
 \begin{itemize}
-\item Description of the randomized algorithm.
-\item Big graphs $\rightarrow$ focus on critical edges.
-\item Mimics the human. 
+\item \st{Description of the randomized algorithm}.
+\item \st{Big graphs $\rightarrow$ focus on critical edges}.
 \item Avoiding local minima.
@@
 \label{ssec:minima}
+
+From the above description, it is not difficult to see that our algorithm in a sense mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify...
+
+\subsection{Conditions for termination.}
+\label{ssec:termination}
 

### 2018-05-08 src/sidepanel/InitSidePanel.java
Beginning the box
Actually simply the class, as the size of the box is missing.

@@ -19,6 +19,7 @@ import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.randommovement.RandomMovementConfigurator;
 import layout.algo.randommovement.RandomMovementLayout;
 import main.MainFrame;
+import view.visual.DrawBoundingBox;
 import view.visual.DrawScale;
 
 import javax.swing.*;
@@ -74,6 +75,7 @@ public class InitSidePanel {
     private void initDefault() {
         //show scale
         mainFrame.view.getBackgroundGroup().addChild(new DrawScale(mainFrame.view), ICanvasObjectDescriptor.VISUAL);
+        mainFrame.view.getBackgroundGroup().addChild(new DrawBoundingBox(mainFrame.view), ICanvasObjectDescriptor.VISUAL);
 
         masterEnableMinimumAngle = new JCheckBox("Show minimum angle");
         masterEnableMinimumAngle.addItemListener(this::masterMinAngleDisplayEnabled);

### 2018-05-11 paper/paper.tex
Working a bit more on the presentation; add also some figures

@@
 
-In Graph Drawing, there exist a really reach literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+In Graph Drawing, there exist a really rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
 However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
@@
 Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is refer to as \emph{right-angle-crossing} graph or \emph{RAC} graph, for short.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is refer to as \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; refer to Figure~\ref{fig:examples} for an illustration.
 For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
 Notably, RAC graph are sparse (they contain at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, where $n$ denotes the number of  vertices), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+
+\begin{figure}[t!]
+	\centering
+	\subfloat[\label{fig:k5} {}]{
+	\includegraphics[page=1]{figures/examples}}
+	\hfil
+	\subfloat[\label{fig:k6} {}]{
+	\includegraphics[page=2]{figures/examples}}
+	\caption{%
+	(a)~A RAC drawing of the complete graph $K_5$, and
+	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
+	\label{fig:examples}
+\end{figure}
 
@@
 The position of vertex $v_i$ in drawing $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[0,\delta]$, where $\delta$ is also a parameter of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$\todo{Add a figure illustrating the approach.}.
+More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$.
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[page=1]{figures/algorithm}
+	\caption{%
+	Illustration of an iteration step of our algorithm: 
+	The chosen vertex is the white one; 
+	the computed rays $r_1,\ldots,r_8$ have been rotated by $8^\circ$; 
+	the black-colored points along these rays are points $\pi_1,\ldots,\pi_8$; 
+	among them, $\pi_4$ yields the best solution.}
+	\label{fig:algo}
+\end{figure}
 
@@
 In Section~\ref{ssec:termination}, we describe conditions  The iteration steps
+ s
 
@@
 
-From the above description, it is not difficult to see that our algorithm in a sense mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify...
+From the above description, it is not difficult to see that our algorithm in a sense mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify the pair of edges that define the crossing resolution of the drawing (we call them critical edges); then, one would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
+Of course, we cannot investigate all possible positions.
+Therefore, we discretized them by considering a number of rays around the chosen vertex and by randomly choosing a single point at each of them.
+
+While the aforementioned procedure
 

### 2018-05-14 src/Experiment/Experiment.java
Only one frame, new output Layout, force algorithm,

@@ -1,15 +1,17 @@
 package Experiment;
 
 import algorithms.graphs.GridGraph;
+import algorithms.graphs.MinimumAngle;
+import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.geometry.RectD;
-import com.yworks.yfiles.graph.IEdge;
-import com.yworks.yfiles.graph.IGraph;
-import com.yworks.yfiles.graph.INode;
-import com.yworks.yfiles.graph.LayoutUtilities;
+import com.yworks.yfiles.graph.*;
 import com.yworks.yfiles.graphml.GraphMLIOHandler;
 import com.yworks.yfiles.layout.organic.OrganicLayout;
 import io.ContestIOHandler;
 import io.GraphIOHandler;
+import layout.algo.execution.ILayout;
+import layout.algo.forcealgorithm.ForceAlgorithm;
+import layout.algo.genetic.GeneticForceAlgorithmLayout;
 import layout.algo.randommovement.RandomMovementLayout;
 import layout.algo.utils.PositionMap;
 import main.MainFrame;
@@ -30,34 +32,76 @@ public class Experiment {
 
     private MainFrame mainFrame;
     private IGraph graph;
-    private long sartTime, endTime, calcTime;
+    private long sartTime, endTime, calcTime, minimumAngleTime, minimumAngleIterations;
+    private long maxTimeForAlgo = 100000;
     private boolean reached90Degree = false;
+    private boolean isInInvLoop = false;
     private PrintWriter pw;
     private StringBuilder sb;
     private int maxIterations = 1000;
     private int iterations = 0;
+    private int numOfUnchangedAngle = 0;
+    private int numOfCrossings = -1;
     private double bestAngle = 0;
+    private int algorithmNum = 1; // 1: Random Algo. 2: Force Algo
+    private Mapper<INode, PointD> positions;
+    Optional<SidePanelTab> tab;
 
     private int totalVertecies, totalEdges;
     private double currentGraphSizeX, currentGraphSizeY, minimumAngle, testAngle;
 
-    public  Experiment (MainFrame mainFrame, String fileName, String inputDir, String outputDir) {
+    public  Experiment (MainFrame mainFrame, String fileName, String inputDir, String outputDir, int boxSize) {
         this.mainFrame = mainFrame;
         this.graph = mainFrame.graph;
         this.fileName = fileName;
-        this.filePrefix = getPrefixString(fileName);
[…]

### 2018-05-14 src/Experiment/Experiments.java

@@ -7,13 +7,14 @@ import com.yworks.yfiles.graph.*;
  * Created by Ama on 24.04.2018.
  */
 
-import java.io.PrintWriter;
 import java.util.Optional;
-import java.util.stream.IntStream;
 
 import com.yworks.yfiles.layout.YGraphAdapter;
 import com.yworks.yfiles.layout.organic.OrganicLayout;
 import com.yworks.yfiles.view.GraphComponent;
+import layout.algo.execution.ILayout;
+import layout.algo.forcealgorithm.ForceAlgorithm;
+import layout.algo.genetic.GeneticForceAlgorithmLayout;
 import layout.algo.randommovement.RandomMovementLayout;
 import main.MainFrame;
 import sidepanel.SidePanelTab;
@@ -38,8 +39,10 @@ public class Experiments {
     private int numOfIteration = 100;
     private int numOfIterationFactor = 1000;
     private long maxCalcTime = 6000; //in mili. sec
+    private int boxSize = 10000;
     private boolean planarGraphsAllowed = false;
     private boolean unconnectedGraphsAllowed = true;
+    private static boolean isFrameFinished = false;
     private String[] childernP;
 
 
@@ -55,22 +58,24 @@ public class Experiments {
         this.outputDirectory = outputDirectory;
     }
 
-    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps){
+    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize){
         this.comp = new GraphComponent();
         this.inputDirectory = inputDirectory;
         this.outputDirectory = outputDirectory;
         this.maxCalcTime = maxCalcTime;
         this.numOfIteration = numOfIterationPerStep;
         this.numOfIterationFactor = numOfSteps;
+        this.boxSize = boxSize;
     }
 
-    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, boolean planarGraphsAllowed, boolean unconnectedGraphsAllowed){
+    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize, boolean planarGraphsAllowed, boolean unconnectedGraphsAllowed){
         this.comp = new GraphComponent();
[…]

### 2018-05-14 src/Experiment/MainExperiment.java

@@ -1,0 +1,37 @@
+package Experiment;
+
+/**
+ * Created by Ama on 27.04.2018.
+ */
+public class MainExperiment {
+    public static void main(String[] args) {
+        //Experiments e = new Experiments();
+
+        // Experiments e = new Experiments(("E:\\graph\\graphml\\"), "E:\\graph\\afterRandom\\graphml2\\");
+
+        Experiments e = new Experiments(
+                "E:\\graph\\rome\\smallSet\\", "E:\\graph\\experiment\\rome\\smallSet\\random\\",
+                10000, 100, 1000, 10000,
+                false, false);
+        e.runOnlyRandom();
+
+        e = new Experiments(
+                "E:\\graph\\rome\\smallSet\\", "E:\\graph\\experiment\\rome\\smallSet\\force\\",
+                10000, 100, 1000, 10000,
+                false, false);
+        e.runOnlyForce();
+
+    }
+
+/*        for(int i = 0; i < 2; i++){
+            Experiments e = new Experiments(
+                    "E:\\graph\\rome\\rome\\", "E:\\graph\\afterRandom\\rome\\",
+                    10000, 100, 1000, (10000 * (10 * i)) ,
+                    false, false);
+
+            e.run();
+        }
+*/
+
+
+}

### 2018-05-14 src/Experiment/mainExperiment.java

@@ -1,18 +1,0 @@
-package Experiment;
-
-/**
- * Created by Ama on 27.04.2018.
- */
-public class mainExperiment {
-    public static void main(String[] args)
-    {
-        //Experiments e = new Experiments();
-
-       // Experiments e = new Experiments(("E:\\graph\\graphml\\"), "E:\\graph\\afterRandom\\graphml2\\");
-        Experiments e = new Experiments(
-                "E:\\graph\\rome\\rome\\", "E:\\graph\\afterRandom\\rome\\",
-                10000, 100, 1000,
-                false, false);
-        e.run();
-    }
-}

### 2018-05-14 src/io/SimpleGraphmlIOHandler.java

@@ -61,7 +61,7 @@ public class SimpleGraphmlIOHandler extends GraphIOHandler{
                 INode node = g.createNode(new PointD(nodesCount, nodesCount));
                 node.setTag(objString);
               //  System.out.println("nodeCount  " +nodesCount);
-                System.out.println(objString + "  node");
+               // System.out.println(objString + "  node");
                 nodes[nodesCount] = node;
                 nodesId[nodesCount] = objString;
 
@@ -85,7 +85,7 @@ public class SimpleGraphmlIOHandler extends GraphIOHandler{
                 line = line.substring(index+1);
                 index = line.indexOf("\"");
                 String objString  = line.substring(1,index);
-                System.out.println(objString + "  source");
+             //   System.out.println(objString + "  source");
                 int indexN1 = searchIndex(nodesId, objString);
                 if(indexN1 == -1){System.out.println("Maping Node1 failed");}
 
@@ -95,7 +95,7 @@ public class SimpleGraphmlIOHandler extends GraphIOHandler{
                 line = line.substring(index+1);
                 index = line.indexOf("\"");
                 objString = line.substring(1,index);
-                System.out.println(objString + "  target");
+              //  System.out.println(objString + "  target");
                 int indexN2 = searchIndex(nodesId, objString);
                 if(indexN1 == -1){System.out.println("Maping Node2 failed");}
 

### 2018-05-14 src/main/MainFrame.java

@@ -34,7 +34,7 @@ import java.util.function.Consumer;
 public class MainFrame extends JFrame {
 
     /* Box related issue*/
-    public static final double BOX_SIZE = 10000;
+    public static double BOX_SIZE = 10000;
 
     /* Graph Drawing related objects */
     public GraphComponent view;
@@ -66,9 +66,9 @@ public class MainFrame extends JFrame {
     /**
      * Creates new form MainFrame
      */
-    private MainFrame() {}
+    public MainFrame() {}
 
-    private void init() {
+    public void init() {
         this.initComponents();
         this.initMenuBar();
 
@@ -386,4 +386,5 @@ public class MainFrame extends JFrame {
             initSidePanel.addDefaultListeners();
         }
     }
+
 }

### 2018-05-14 paper/paper.tex
Finished with the main part of the description.

@@
-\documentclass{llncs}
+\documentclass[runningheads]{llncs}
 \usepackage{amsmath,amssymb}
@@
 \author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
+\authorrunning{M.~A.~Bekos et al.}
 \title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
+\titlerunning{Drawings of Graphs with High Crossing Resolution}
 
@@
 Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-As a result, their role in our algorithm is crucial.
+As a result, their role in our algorithm is crucial\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
 The reason is that by slightly changing the location of a critical vertex appropriately, one naturally expects to improve the crossing resolution of the current drawing.
-In fact, what quickly we realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+In fact, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
 However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local minima after a number of iterations.
 Hence, special care is needed in order to avoid these bottlenecks, especially when the graph to be drawn is large.
-More precisely, in such a case an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of the critical vertices to move.
 We discuss heuristics to avoid such bottlenecks in  Section~\ref{ssec:minima}.
@@
 To compute the position of vertex $v_i$ in drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from the point $p(v_i)$ of vertex $v_i$ in drawing $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$.
+These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; refer to Fig.~\ref{fig:algo} for an illustration.
 The position of vertex $v_i$ in drawing $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
 More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$, and there is no vertex of drawing $\Gamma_{i-1}$ at point $\pi_j$.
 
@@
 	\centering
-	\includegraphics[page=1]{figures/algorithm}
+	\subfloat[\label{fig:algo-rays} {}]{	
+	\includegraphics[page=1]{figures/algorithm}}
+	\hfil
+	\subfloat[\label{fig:algo-move} {}]{
+	\includegraphics[page=2]{figures/algorithm}}
 	\caption{%
 	Illustration of an iteration step of our algorithm: 
-	The chosen vertex is the white one; 
+	(a)~The chosen vertex is the white one; 
 	the computed rays $r_1,\ldots,r_8$ have been rotated by $8^\circ$; 
 	the black-colored points along these rays are points $\pi_1,\ldots,\pi_8$; 
-	among them, $\pi_4$ yields the best solution.}
+	among them, $\pi_4$ yields the best solution.
+	(b)~The resulting drawing after moving the vertex at position $\pi_3$.}
 	\label{fig:algo}
@@
 In our implementation, we did not observe any significant difference in these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-In Section~\ref{ssec:termination}, we describe conditions  The iteration steps
- s
-
-\begin{itemize}
-\item \st{Description of the randomized algorithm}.
-\item \st{Big graphs $\rightarrow$ focus on critical edges}.
-\item Avoiding local minima.
-\item Crossing calculation. 
-\end{itemize}
+In Section~\ref{ssec:termination}, we describe conditions under which we determine that our algorithm has converged.
 
@@
 
-From the above description, it is not difficult to see that our algorithm in a sense mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify the pair of edges that define the crossing resolution of the drawing (we call them critical edges); then, one would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot investigate all possible positions.
+From the above description, it is not difficult to see that our algorithm, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify the pair of edges that define the crossing resolution of the drawing (we call them critical edges); then, one would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
+Of course, we cannot consider all possible positions for the vertex to be moved.
 Therefore, we discretized them by considering a number of rays around the chosen vertex and by randomly choosing a single point at each of them.
 
-While the aforementioned procedure
+To avoid getting trapped to locally optimal solutions with the aforementioned procedure, we mainly investigated two different approaches.
+Both of them are parametrizable by two input parameters $\zeta$ and $\zeta'$.
+The first one mimics again the human behaviour.
+What would one do to get away from a locally optimal solution?
+First, she would stop trying to move the endvertices of the edges defining the crossing resolution; she would start moving ``irrelevant'' vertices hoping that by changing the embedding of the graph or its crossing structure, a better solution would be easier to be computed again.
+Our algorithm is mimicking this idea by allowing the vertex-pool to also contain all the vertices of the graph, as follows:
+
+\begin{itemize}
+%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices. 
+\item If during the last $\zeta$ iterations the crossing resolution has not been improved, then the vertex-pool becomes \emph{wider} containing all the vertices of the graph, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations.
+\item After these $\zeta'$ iterations, the vertex-pool switches back to the critical vertices.
+\end{itemize}
+%
+From our experimental evaluation, we observed that, while this approach turned to be quite effective for medium-size graphs, for larger and denser graph, unfortunately, it was not so efficient.
+What we observed is that in larger graphs it was unlike to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not at the part of the drawing where the critical vertices resided.
+
+The second approach that we investigated to avoid locally optimal solutions was based on the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
+Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastical'' or ``sharp'' move may help to get untrapped.
+We turned this idea into an algorithmic step as follows:
+
+\begin{itemize}
+\item If during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these parameters for $\zeta'$ iterations.
+\item After these $\zeta'$ iterations, the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to their initial (input) values.
+\end{itemize}
+%
+Of course, this approach may lead to drawings with larger area, but this is somehow expected, as it turns out that drawings with high crossing resolution may require large drawing area; see, e.g.,~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
 
@@
 \label{ssec:termination}
+
+The termination condition of our algorithm is simple and depends also on a input parameter $\tau$, such that $\tau \geq \zeta + \zeta'$.
+More precisely, if the crossing resolution during the last $\tau$ iterations has not been improved, then we assume that the algorithm has converged and we do not proceed any more.
+
+\subsection{Complexity Related Issues.}
+\label{ssec:complexity}
+
+A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points among the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing).
+Given  a drawing, a naive approach to compute its crossing points requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively.
+
+However, if the algorithm had to compute all crossing points and the corresponding angles at these points at each iteration step, then our algorithm would be not be useful in practice.
+Instead, the algorithm is computing all the crossing points only once, at the beginning of the algorithm, and stores them.
+Then, at each iteration step, it is enough to compute the crossing angles only for the edges that are affected by a move of the chosen vertex.
+By adopting this simple approach, the computation of the crossing points can be improved drastically.
+\todo{More details should be given.}
 

### 2018-05-15 src/main/InitMenuBar.java
File loading

@@ -375,6 +375,13 @@ public class InitMenuBar {
         openContestItem.setText("Open Contest File");
         openContestItem.addActionListener(this::openContestItemActionPerformed);
         fileMenu.add(openContestItem);
+        
+        JMenuItem openAMItem = new JMenuItem();
+//      openContestItem.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_U, InputEvent.CTRL_MASK));
+        openAMItem.setIcon(new ImageIcon(getClass().getResource("/resources/open-16.png")));
+        openAMItem.setText("Open AM");
+        openAMItem.addActionListener(this::openAMItemActionPerformed);
+        fileMenu.add(openAMItem);
 
         JMenuItem openContest2018Item = new JMenuItem();
         openContest2018Item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_U, InputEvent.CTRL_MASK));
@@ -745,6 +752,28 @@ public class InitMenuBar {
         if (chooser.showOpenDialog(null) == JFileChooser.APPROVE_OPTION) {
             this.fileNamePath = chooser.getSelectedFile().toString();
             mainFrame.openContestFile(fileNamePath);
+            mainFrame.removedNodes.clear();
+            this.mainFrame.removedChains.clear();
+            this.fileNamePathFolder = chooser.getSelectedFile().getParent();
+        }
+    }
+    
+    private void openAMItemActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {
+        JFileChooser chooser = new JFileChooser(this.fileNamePathFolder);
+        chooser.setFileFilter(new FileFilter() {
+            public boolean accept(File file) {
+                return (file.isDirectory() || file.toString().toLowerCase().endsWith("txt"));
+            }
+
+            public String getDescription() {
+                return "ASCII Files [.txt]";
+            }
+
+        });
+
+        if (chooser.showOpenDialog(null) == JFileChooser.APPROVE_OPTION) {
+            this.fileNamePath = chooser.getSelectedFile().toString();
+            mainFrame.openAMFile(fileNamePath);
             mainFrame.removedNodes.clear();
             this.mainFrame.removedChains.clear();
             this.fileNamePathFolder = chooser.getSelectedFile().getParent();

### 2018-05-15 src/main/MainFrame.java

@@ -15,6 +15,7 @@ import com.yworks.yfiles.view.input.*;
 import graphoperations.RemovedChains;
 import graphoperations.RemovedNodes;
 import graphoperations.Scaling;
+import io.AdjacencyMatrixHandler;
 import io.Contest2018IOHandler;
 import io.ContestIOHandler;
 import layout.algo.execution.ILayout;
@@ -372,4 +373,20 @@ public class MainFrame extends JFrame {
             initSidePanel.addDefaultListeners();
         }
     }
+
+	public void openAMFile(String fileNamePath) {
+        initSidePanel.removeDefaultListeners();
+        try {
+            AdjacencyMatrixHandler.read(graph, fileNamePath);
+            view.fitGraphBounds();
+            view.updateUI();
+            bestSolution.reset();
+            removedNodes = new RemovedNodes(graph);
+            setTitle(Paths.get(fileNamePath).getFileName().toString());
+        } catch (IOException e) {
+            infoLabel.setText("An error occured while reading the input file.");
+        } finally {
+            initSidePanel.addDefaultListeners();
+        }
+	}
 }

### 2018-05-16 paper/paper.tex
Description of the crossing computation (to be finished with Lukas) and of the gridding.

@@
 
-\subsection{Grid drawings: an interesting variant.}
+\subsection{Grid drawings: An interesting variant.}
 \label{ssec:grid}
+
+The algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
+However, it can be easily adjusted to produced such drawings.
+Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ feasible points $\pi_1,\ldots,\pi_\rho$.
+If we round these points to their closest grid points, say $\overline{\pi}_1,\ldots,\overline{\pi}_\rho$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, so to be at grid points).
+
+An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
+Critical here is not to affect the crossing resolution too much during the conversion.
+To this end, we can use a technique that is also based on randomization.
+For each vertex in the computed drawing, we can chose uniformly at random one of the grid points that are neighboring the vertex, and move the vertex to this point, assuming of course that there is no other vertex at this point.
 

### 2018-05-17 src/Experiment/MainExperiment.java
Fix center from aspect-ratio

@@ -24,7 +24,7 @@ public class MainExperiment {
 
 */
             Experiments e = new Experiments(
-                    "E:\\graph\\rome\\smallSet\\", "E:\\graph\\afterRandom\\romeRF\\",
+                    "E:\\graph\\rome\\smallSet\\", "E:\\graph\\afterRandom\\romeRF2\\",
                     10000, 100, 1000, 10000,
                     false, false);
 

### 2018-05-17 src/graphoperations/GraphOperations.java

@@ -19,11 +19,11 @@ public class GraphOperations {
     IEdge sEdge = getSmallstEdge(g);
     INode sSource = sEdge.getSourceNode();
     INode sTarget = sEdge.getTargetNode();
-    double sLength  = euclidDist(sSource.getLayout().getX(), sSource.getLayout().getY(), sTarget.getLayout().getX(), sTarget.getLayout().getY());
+    double sLength  = euclidDist(sSource.getLayout().getCenter().getX(), sSource.getLayout().getCenter().getY(), sTarget.getLayout().getCenter().getX(), sTarget.getLayout().getCenter().getY());
     IEdge lEdge = getLongestEdge(g);
     INode lSource = lEdge.getSourceNode();
     INode lTarget = lEdge.getTargetNode();
-    double lLength  = euclidDist(lSource.getLayout().getX(), lSource.getLayout().getY(), lTarget.getLayout().getX(), lTarget.getLayout().getY());
+    double lLength  = euclidDist(lSource.getLayout().getCenter().getX(), lSource.getLayout().getCenter().getY(), lTarget.getLayout().getCenter().getX(), lTarget.getLayout().getCenter().getY());
     return lLength / sLength;
   }
 
@@ -34,7 +34,7 @@ public class GraphOperations {
       IEdge edge = edgeList.getItem(0);
       INode source = edge.getSourceNode();
       INode target = edge.getTargetNode();
-      double length  = euclidDist(source.getLayout().getX(), source.getLayout().getY(), target.getLayout().getX(), target.getLayout().getY());
+      double length  = euclidDist(source.getLayout().getCenter().getX(), source.getLayout().getCenter().getY(), target.getLayout().getCenter().getX(), target.getLayout().getCenter().getY());
       double smallLength = length;
       IEdge smallEdge = edge;
 
@@ -42,7 +42,7 @@ public class GraphOperations {
         edge = edgeList.getItem(i);
         source = edge.getSourceNode();
         target = edge.getTargetNode();
-        length = euclidDist(source.getLayout().getX(), source.getLayout().getY(), target.getLayout().getX(), target.getLayout().getY());
+        length = euclidDist(source.getLayout().getCenter().getX(), source.getLayout().getCenter().getY(), target.getLayout().getCenter().getX(), target.getLayout().getCenter().getY());
         if(length < smallLength){
           smallLength = length;
           smallEdge = edge;
@@ -64,7 +64,7 @@ public class GraphOperations {
[…]

### 2018-05-17 paper/paper.tex
A bit more on the description

--- 2018-05-14 paper/paper.tex
@@
 
-\subsection{Grid drawings: an interesting variant.}
+\subsection{Grid drawings: An interesting variant.}
 \label{ssec:grid}
+
+The algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
+However, it can be easily adjusted to produced such drawings.
+Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ candidate points $\pi_1,\ldots,\pi_\rho$.
+If we round these points to their closest grid points, say $\overline{\pi}_1,\ldots,\overline{\pi}_\rho$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
+
+An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
+Critical here is not to affect the crossing resolution too much during the conversion.
+Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
+In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
+For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
+If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points.
 

### 2018-05-17 paper/paper.tex
Finished the description of the algorithm with the help of Lukas.

@@
 
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
+\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
 \authorrunning{M.~A.~Bekos et al.}
@@
 \begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed at the crossing points of the edges.
+The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed by any pair of crossing edges.
 Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
@@
 
-In this paper, we propose a new --randomization-based-- heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known ones from the literature.
+In this paper, we propose a new randomization-based heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known ones from the literature.
 Our experimental evaluation indicates that the new heuristic produces drawings with significantly better crossing resolution, but this comes at the cost of slightly higher running time, especially when the input graph is large and dense.
@@
 
-In Graph Drawing, there exist a really rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+In Graph Drawing, there exists a really rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
 However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the proposed techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
+As a result, the enstablished techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
 
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important statistical evidence (through a series of eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and therefore the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
+In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence (through a series of eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and therefore the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
 In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality and readability of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, which is unofficially recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
+As a result, a new and prominent research direction was initiated, which is recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
 
-Formally, the minimum angle formed by any two crossed edges in a drawing is refer to as its \emph{crossing resolution}.
+Formally, the minimum angle formed by any pair of crossing edges in a drawing is refered to as its \emph{crossing resolution}.
 Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is refer to as \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; refer to Figure~\ref{fig:examples} for an illustration.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; refer to Figure~\ref{fig:examples} for an illustration.
 For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
@@
 
-The latter result is already an indication that the crossing resolution maximization problem might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
+The latter result is already an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
 Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
 
-From a practical point of view, we are only aware of two methods that aim in drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms are appropriate forces, which are applied on the endvertices of every pair of crossing edges.
+From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms~\cite{DBLP:journals/congnum/Eades84}.
+The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one is due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
+Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
 Each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
 
-In this paper, we attack the crossing resolution maximization problem by following a completely different approach.
-We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which in practise turned out to be quite efficient\footnote{Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to computed the expected quality of the produced drawings.
-A notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.}.
-However, since we could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
+In this paper, we approach the crossing resolution maximization problem from a different perspective.
+We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
+How would one try to increase the crossing resolution of a given drawing?
+First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
+Of course, we cannot consider all possible positions for the vertex to be moved.
+Therefore, we discretize them by considering a number of rays around the chosen vertex and by choosing (at random again) a single point at each of them.
+
+In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to compute the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
+Since we could not provide any theoretical guarantee on the expected quality of the produced drawings as well (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
 To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of spring embedding algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly higher running time, especially for large and dense graphs.
+Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of spring embedding algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
 
-The remainder of this paper is structured as follows.
-Section~\ref{sec:relatedwork} overviews related works.
-In Section~\ref{sec:preliminaries}, we introduce preliminary notions and notation used throughout the paper.
+\paragraph{Preliminaries:}
+Unless otherwise specified, in this paper we consider simple undirected graphs.
+Let $G=(V,E)$ be such a graph.
+The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
+The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
+%
+Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
+%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
+
+\paragraph{Structure of the paper:}
+The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works. 
+%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper. 
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
@@
 
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (propertied of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
+As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
 The study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
@@
 To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ has at most $(3n-6)\pi/\alpha$ edges.
+More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
 Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
 
-An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem (as naturally expected, of course).
-The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
+An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
+The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
 Exceptions are the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}.
-According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium state of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-
-\section{Preliminaries}
-\label{sec:preliminaries}
-
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-
-Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
+According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
 
@@
 In this section, we describe our heuristic algorithm for obtaining drawings with high crossing resolution.
-Without loss of generality, we assume that the input of our algorithm consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ that is of some crossing resolution $c(\Gamma_0)$.
-We may assume that no two edges of $G$ overlap in $\Gamma_0$, that is, $c(\Gamma_0)>0$.
+We assume that the input of our algorithm consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ that is of some crossing resolution $c(\Gamma_0)$.
+We also assume that no two edges of $G$ overlap in $\Gamma_0$, that is, $c(\Gamma_0)>0$.
 Note that a drawing $\Gamma_0$ of $G$ meeting these preconditions can be, e.g., a circular drawing of $G$ or a drawing obtained by applying a standard spring embedding algorithm on $G$.
@@
 
-To formally define the critical vertices, we first need to introduced the notion of critical edge-pairs.
-A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is either equal (or very close, in some cases) to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
+To formally define the critical vertices, we first need to introduce the notion of critical edge-pairs.
+A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is either equal to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
 The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
@@
 Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-As a result, their role in our algorithm is crucial\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
+As a result, their role in our algorithm is central\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
 The reason is that by slightly changing the location of a critical vertex appropriately, one naturally expects to improve the crossing resolution of the current drawing.
 In fact, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local minima after a number of iterations.
+However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
 Hence, special care is needed in order to avoid these bottlenecks, especially when the graph to be drawn is large.
-We discuss heuristics to avoid such bottlenecks in  Section~\ref{ssec:minima}.
+We discuss heuristics to avoid such bottlenecks in  Section~\ref{ssec:maxima}.
 
@@
 More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$, and there is no vertex of drawing $\Gamma_{i-1}$ at point $\pi_j$.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$, and there is no vertex of drawing $\Gamma_{i-1}$ at point $\pi_j$, i.e., we do not introduce any vertex-vertex overlap.
 
@@
 	(a)~The chosen vertex is the white one; 
-	the computed rays $r_1,\ldots,r_8$ have been rotated by $8^\circ$; 
-	the black-colored points along these rays are points $\pi_1,\ldots,\pi_8$; 
+	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$; 
+	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$; 
 	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_3$.}
+	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
 	\label{fig:algo}
@@
 On the other hand, if there is one or more feasible points (among $\pi_0,\pi_1,\ldots,\pi_{\rho-1}$), then one may consider two different approaches to determine the new position of vertex $v_i$ in drawing $\Gamma_i$.
-The most natural is to chose this feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one the feasible points as the new position of vertex $v_i$ in drawing $\Gamma_i$.
+The most natural is to chose the feasible point that maximizes the crossing resolution of the obtained drawing.
+As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the new position of vertex $v_i$ in drawing $\Gamma_i$.
 In our implementation, we did not observe any significant difference in these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
@@
 
-\subsection{Avoiding local minima.}
-\label{ssec:minima}
-
-From the above description, it is not difficult to see that our algorithm, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify the pair of edges that define the crossing resolution of the drawing (we call them critical edges); then, one would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Therefore, we discretized them by considering a number of rays around the chosen vertex and by randomly choosing a single point at each of them.
+\subsection{Avoiding local maxima.}
+\label{ssec:maxima}
 
@@
 Both of them are parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first one mimics again the human behaviour.
+The first one mimics the human behaviour.
 What would one do to get away from a locally optimal solution?
@@
 A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points among the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing).
-Given  a drawing, a naive approach to compute its crossing points requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively.
+Given  a drawing, a naive approach to compute its crossing points requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; refer, e.g., to~\cite{DBLP:books/lib/BergCKO08}.
 
-However, if the algorithm had to compute all crossing points and the corresponding angles at these points at each iteration step, then our algorithm would be not be useful in practice.
-Instead, the algorithm is computing all the crossing points only once, at the beginning of the algorithm, and stores them.
-Then, at each iteration step, it is enough to compute the crossing angles only for the edges that are affected by a move of the chosen vertex.
-By adopting this simple approach, the computation of the crossing points can be improved drastically.
-\todo{More details should be given.}
+However, if the algorithm had to compute all crossing points and the corresponding angles at these points for each candidate position at each iteration step, then our algorithm would not be useful in practice.
+Instead, we adopted a different approach, which turned out to be quite efficient in practice.
+Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate points to move $v_i$.
+Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
+Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in drawing $\Gamma_{i-1}$.
+Let $\phi_i$ be the minimum crossing angle computed with this procedure; this is our reference angle.
+Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that vertex $v_i$ is at position $\pi_j$.
+Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
+Clearly, in our context, position $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
+Note that the complexity of this approach is $O(deg(v_i) \cdot m)$, which in worst-case is $O(nm)$.
 
@@
 However, it can be easily adjusted to produced such drawings.
-Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ candidate points $\pi_1,\ldots,\pi_\rho$.
-If we round these points to their closest grid points, say $\overline{\pi}_1,\ldots,\overline{\pi}_\rho$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
+Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ candidate points $\pi_0,\ldots,\pi_{\rho-1}$.
+If we round these points to their closest grid points, say $\overline{\pi}_0,\ldots,\overline{\pi}_{\rho-1}$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
 

### 2018-05-22 src/layout/algo/forcealgorithm/forces/TotalResolutionForce.java
Separate force total resolution, tune parameters

@@ -18,6 +18,15 @@ public class TotalResolutionForce implements IForce {
     private final IGraph graph;
     private AbstractLayoutInterfaceItem<Double> threshold;
     private AbstractLayoutInterfaceItem<Boolean> activated;
+    private AbstractLayoutInterfaceItem<Boolean> activated1;
+    private AbstractLayoutInterfaceItem<Boolean> activated2;
+    private AbstractLayoutInterfaceItem<Boolean> activated3;
+    private AbstractLayoutInterfaceItem<Boolean> activated4;
+
+    private AbstractLayoutInterfaceItem<Double> force1;
+    private AbstractLayoutInterfaceItem<Double> force2;
+    private AbstractLayoutInterfaceItem<Double> force3;
+    private AbstractLayoutInterfaceItem<Double> force4;
 
     public TotalResolutionForce(IGraph graph) {
         this.graph = graph;
@@ -31,6 +40,34 @@ public class TotalResolutionForce implements IForce {
         activated = itemFactory.toggleableParameter(threshold);
         activated.setValue(true);
         toggleableParameters.add(activated);
+
+        force1 = itemFactory.doubleParameter("Crossing1", 0.0, 0.1);
+        force1.setValue(0.05);
+
+        activated1 = itemFactory.toggleableParameter(force1);
+        activated1.setValue(true);
+        toggleableParameters.add(activated1);
+
+        force2 = itemFactory.doubleParameter("Crossing2", 0.0, 0.1);
+        force2.setValue(0.05);
+
+        activated2 = itemFactory.toggleableParameter(force2);
+        activated2.setValue(true);
+        toggleableParameters.add(activated2);
+
+        force3 = itemFactory.doubleParameter("Angular1", 0.0, 0.1);
+        force3.setValue(0.05);
+
+        activated3 = itemFactory.toggleableParameter(force3);
+        activated3.setValue(true);
+        toggleableParameters.add(activated3);
+
+        force4 = itemFactory.doubleParameter("Angular2", 0.0, 0.1);
+        force4.setValue(0.05);
+
+        activated4 = itemFactory.toggleableParameter(force4);
+        activated4.setValue(true);
+        toggleableParameters.add(activated4);
     }
 
     /**
@@ -43,12 +80,16 @@ public class TotalResolutionForce implements IForce {
         }
 
[…]

### 2018-05-22 src/layout/algo/utils/LayoutUtils.java
nodes not on top of another (needs testing)

@@ -25,7 +25,7 @@ public static Boolean overlap(PointD position, Mapper<INode, PointD> positions) 
 	PointD tmp;
 		while(It.hasNext()){
 			tmp=It.next().getValue();
-			if (position.hits(tmp, 0.5)) {
+			if (position.hits(tmp, 0.5) || position.equals(tmp)) {
 				value= false;
 			}
 		}

### 2018-05-22 src/view/visual/DrawBoundingBox.java
Bounding box can be drawn until any distance

@@ -35,13 +35,11 @@ public class DrawBoundingBox implements IVisual{
         double c_x = MainFrame.BOX_SIZE[0];
         double c_y = MainFrame.BOX_SIZE[1];
 
-        if ((int) (100 / z) > 20) { //dont draw for < 20
-            graphics2D.setStroke(new BasicStroke((float) (1 / z)));
-            graphics2D.setColor(this.color);
-            graphics2D.drawLine((int) (0), (int) (0), (int) (c_x), (int) (0));
-            graphics2D.drawLine((int) (0), (int) (0), (int) (0), (int) (c_y));
-            graphics2D.drawLine((int) (0), (int) (c_y), (int) (c_x), (int) (c_y));
-            graphics2D.drawLine((int) (c_x), (int) (0), (int) (c_x), (int) (c_y));
-        }
+        graphics2D.setStroke(new BasicStroke((float) (1 / z)));
+        graphics2D.setColor(this.color);
+        graphics2D.drawLine((int) (0), (int) (0), (int) (c_x), (int) (0));
+        graphics2D.drawLine((int) (0), (int) (0), (int) (0), (int) (c_y));
+        graphics2D.drawLine((int) (0), (int) (c_y), (int) (c_x), (int) (c_y));
+        graphics2D.drawLine((int) (c_x), (int) (0), (int) (c_x), (int) (c_y));
     }
 }

### 2018-05-22 src/layout/algo/randommovement/RandomMovementConfigurator.java
Add option random algo to disable increasing step size

@@ -13,6 +13,7 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<Integer> numSamplesForJumping;
   AbstractLayoutInterfaceItem<Boolean> toggleNodeDistributions;
   AbstractLayoutInterfaceItem<Boolean> onlyGridPositions;
+  AbstractLayoutInterfaceItem<Boolean> allowIncreaseStepSize;
 
   @Override
   public void init(ILayoutInterfaceItemFactory itemFactory) {
@@ -32,13 +33,16 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
     jumpOnLocalMaximum.setValue(false);
 
     onlyGridPositions = itemFactory.booleanParameter("Only use grid coordinates");
-    onlyGridPositions.setValue(true);
+    onlyGridPositions.setValue(false);
 
     useGaussianDistribution = itemFactory.booleanParameter("Focus on critical nodes");
     useGaussianDistribution.setValue(true);
 
     toggleNodeDistributions = itemFactory.booleanParameter("Automatically toggle focusing on critical nodes");
     toggleNodeDistributions.setValue(true);
+
+    allowIncreaseStepSize = itemFactory.booleanParameter("Automatically increase step size");
+    allowIncreaseStepSize.setValue(false);
   }
 
 }

### 2018-05-22 src/layout/algo/randommovement/RandomMovementLayout.java

@@ -148,10 +148,12 @@ public class RandomMovementLayout implements ILayout {
   }
 
   private boolean resolveLocalMaximum() {
-    if (!configurator.jumpOnLocalMaximum.getValue()) {
+    if (configurator.allowIncreaseStepSize.getValue()) {
       double boxDiagonal =
-          Math.sqrt(boundingBox.getWidth() * boundingBox.getWidth() + boundingBox.getHeight() * boundingBox.getHeight());
+              Math.sqrt(boundingBox.getWidth() * boundingBox.getWidth() + boundingBox.getHeight() * boundingBox.getHeight());
       configurator.maxStepSize.setValue(Math.min(configurator.maxStepSize.getValue() * 2, boxDiagonal));
+    }
+    if (!configurator.jumpOnLocalMaximum.getValue()) {
       return false;
     }
 

### 2018-05-23 src/algorithms/graphs/MinimumAngle.java
Updated Chains with random movement every x chains, needs debugging

@@ -99,7 +99,7 @@ public class MinimumAngle {
   }
 
   public static double getMinimumAngleForNode(Mapper<INode, PointD> positions, INode node, IGraph graph) {
-    return getMinimumAngleForNode(graph, node, positions).orElse(Double.POSITIVE_INFINITY);
+    return m.getMinimumAngleForNode(graph, node, positions).orElse(Double.POSITIVE_INFINITY);
   }
 
   public static Optional<Double> getMinimumAngleForNode(IGraph graph, INode node, @Nullable Mapper<INode, PointD> nodePositions) {
@@ -107,7 +107,7 @@ public class MinimumAngle {
   }
 
   public static Optional<Intersection> getMinimumAngleCrossing(IGraph graph) {
-    return getMinimumAngleCrossing(graph, null);
+    return m.getMinimumAngleCrossing(graph, null);
   }
 
   public static Optional<Intersection> getMinimumAngleCrossing(IGraph graph, @Nullable IMapper<INode, PointD> np){

### 2018-05-23 src/graphoperations/RemovedChains.java

@@ -35,11 +35,16 @@ public final class RemovedChains {
     }
   }
 
-  public ArrayList<Set<INode>> reinsert(int chainNum) {
+  //lineMode: reinsert vertices on a line between the endpoint vertices
+  public ArrayList<Set<INode>> reinsert(int chainNum, boolean lineMode) {
     ArrayList<Set<INode>> chains = new ArrayList<>();
     for (int i = 0; i < chainNum; i++) {
 //        removedChains.pop().reinsertNodes();
-      chains.add(removedChains.pop().reinsertNodesLine(startEnd.get(i)));
+      if (lineMode) {
+        chains.add(removedChains.pop().reinsertNodesLine(startEnd.get(i)));
+      } else {
+        chains.add(removedChains.pop().reinsertNodes());
+      }
     }
     return chains;
   }

### 2018-05-23 src/layout/algo/clinchlayout/ClinchLayoutConfigurator.java

@@ -5,6 +5,9 @@ import layout.algo.layoutinterface.AbstractLayoutInterfaceItem;
 import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.layoutinterface.ILayoutInterfaceItemFactory;
 
+import java.util.Map;
+import java.util.Optional;
+
 public class ClinchLayoutConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<IGraphSelection> selection;
   AbstractLayoutInterfaceItem<Double> initialStepSize;
@@ -18,5 +21,11 @@ public class ClinchLayoutConfigurator implements ILayoutConfigurator {
 
     initialStepSize.setValue(2.5);
     stepSizeMultiplier.setValue(1.5);
+  }
+
+
+  @Override
+  public Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters() {
+    return Optional.empty();
   }
 }

### 2018-05-23 src/layout/algo/forcealgorithm/ForceAlgorithmConfigurator.java

@@ -8,10 +8,7 @@ import layout.algo.layoutinterface.AbstractLayoutInterfaceItem;
 import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.layoutinterface.ILayoutInterfaceItemFactory;
 
-import java.util.Collection;
-import java.util.HashSet;
-import java.util.LinkedList;
-import java.util.List;
+import java.util.*;
 
 public class ForceAlgorithmConfigurator implements ILayoutConfigurator {
   public List<IForce> forces = new LinkedList<>();
@@ -39,5 +36,10 @@ public class ForceAlgorithmConfigurator implements ILayoutConfigurator {
 
   AbstractLayoutInterfaceItem<Mapper<INode, PointD>> getDebugVectors() {
     return debugVectors;
+  }
+
+  @Override
+  public Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters() {
+    return Optional.empty();
   }
 }

### 2018-05-23 src/layout/algo/genetic/GeneticForceAlgorithmConfigurator.java

@@ -5,6 +5,8 @@ import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.layoutinterface.ILayoutInterfaceItemFactory;
 
 import java.util.ArrayList;
+import java.util.Map;
+import java.util.Optional;
 
 public class GeneticForceAlgorithmConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<Integer> iterationsPerGeneration;
@@ -22,6 +24,11 @@ public class GeneticForceAlgorithmConfigurator implements ILayoutConfigurator {
 
     iterationsPerGeneration.setValue(100);
     stepSize.setValue(0.1);
+  }
+
+  @Override
+  public Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters() {
+    return Optional.empty();
   }
 
 }

### 2018-05-23 src/layout/algo/gridding/GridderConfigurator.java

@@ -4,6 +4,9 @@ import layout.algo.layoutinterface.AbstractLayoutInterfaceItem;
 import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.layoutinterface.ILayoutInterfaceItemFactory;
 
+import java.util.Map;
+import java.util.Optional;
+
 public class GridderConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<Boolean> respectMinimumAngle;
   AbstractLayoutInterfaceItem<Double> allowDecreasingBy;
@@ -39,5 +42,10 @@ public class GridderConfigurator implements ILayoutConfigurator {
 
     forceGridAfterStop = itemFactory.booleanParameter("Force grid after stop was pressed");
     forceGridAfterStop.setValue(true);
+  }
+
+  @Override
+  public Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters() {
+    return Optional.empty();
   }
 }

### 2018-05-23 src/layout/algo/layoutinterface/ILayoutConfigurator.java

@@ -1,5 +1,8 @@
 package layout.algo.layoutinterface;
 
+import java.util.Map;
+import java.util.Optional;
+
 /**
  * The interface for algorithm configurators. Classes implementing this interface should not contain any complex logic
  * but should instead serve as simple container for all parameters ({@link AbstractLayoutInterfaceItem}s) of the
@@ -15,4 +18,6 @@ public interface ILayoutConfigurator {
    *                    become invalid after the invocation
    */
   void init(ILayoutInterfaceItemFactory itemFactory);
+
+  Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters();
 }

### 2018-05-23 src/layout/algo/randommovement/RandomMovementConfigurator.java

@@ -4,6 +4,10 @@ import layout.algo.layoutinterface.AbstractLayoutInterfaceItem;
 import layout.algo.layoutinterface.ILayoutConfigurator;
 import layout.algo.layoutinterface.ILayoutInterfaceItemFactory;
 
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Optional;
+
 public class RandomMovementConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<Double> minStepSize;
   AbstractLayoutInterfaceItem<Double> maxStepSize;
@@ -43,6 +47,14 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
 
     toggleNodeDistributions = itemFactory.booleanParameter("Automatically toggle focusing on critical nodes");
     toggleNodeDistributions.setValue(false);
+  }
+
+  @Override
+  public Optional<Map<String, AbstractLayoutInterfaceItem<Boolean>>> getBooleanParameters() {
+    Map<String, AbstractLayoutInterfaceItem<Boolean>> paramaters = new HashMap<>();
+    paramaters.put("useReinsertChainNodes", useReinsertChainNodes);
+    paramaters.put("toggleNodeDistributions", toggleNodeDistributions);
+    return Optional.of(paramaters);
   }
 
 }

### 2018-05-23 src/layout/algo/randommovement/RandomMovementLayout.java

@@ -59,6 +59,7 @@ public class RandomMovementLayout implements ILayout {
     }
 
     fixNodes = new HashSet<>();
+//    reinsertedChain = new HashSet<>();
   }
 
   @Override
@@ -68,7 +69,16 @@ public class RandomMovementLayout implements ILayout {
 
   @Override
   public void setVarNodes(Set<INode> variableNodes) {
+      if (reinsertedChain == null) {            //cause modifygraph calls init (ie pause unpause)
+          reinsertedChain = new HashSet<>();
+      }
     this.reinsertedChain = variableNodes;
+//      this.reinsertedChain.clear();
+//      this.reinsertedChain.addAll(variableNodes);
+      System.out.println("Setting it "+variableNodes.size());
+      for (INode x : reinsertedChain) {
+          System.out.println(x);
+      }
   }
 
   private ArrayList<Sample> initSampleDirections() {
@@ -252,8 +262,17 @@ public class RandomMovementLayout implements ILayout {
   }
 
   private Optional<INode> chainNodeSelection() {
-      Object[] arr = reinsertedChain.toArray();
-      return Optional.of((INode)arr[ThreadLocalRandom.current().nextInt(0,arr.length)]);
+//      Object[] arr = reinsertedChain.toArray();
+//      return Optional.of((INode)arr[ThreadLocalRandom.current().nextInt(0,arr.length)]);
+//      System.out.println("selecting: "+reinsertedChain.size());
+//      for (INode x : reinsertedChain) {
+//          System.out.println(reinsertedChain.stream().skip(random.nextInt(reinsertedChain.size())).findFirst());
+//      }
+      return reinsertedChain.stream().skip(random.nextInt(reinsertedChain.size())).findFirst();
+//      return StreamSupport.stream(reinsertedChain.spliterator(), false)
+//              .filter(node -> !fixNodes.contains(node))
+//              .skip(random.nextInt(reinsertedChain.size()))
+//              .findFirst();
   }
 
   private Optional<INode> selectRandomNode(Iterable<INode> nodes, int size) {

### 2018-05-23 src/main/InitMenuBar.java

@@ -517,6 +517,7 @@ public class InitMenuBar {
 
                     if (useChains.isSelected()) {
                         mainFrame.removedChains = chains.remove(numVertices, mainFrame.removedChains);
+                        mainFrame.removedChains.setStartEnd(chains.getStartEnd());
                     } else if (useHighest.isSelected()) {
                         mainFrame.removedNodes.removeHighestDegree(numVertices);
                     } else {
@@ -624,7 +625,7 @@ public class InitMenuBar {
                         throw new NumberFormatException("Too many chains");
                     }
 
-                    mainFrame.removedChains.reinsert(numChains);
+                    mainFrame.removedChains.reinsert(numChains, false);
 
                     Scaling.scaleNodeSizes(view);
                 } catch (NumberFormatException exc) {

### 2018-05-23 src/main/MinimumAngleMonitor.java

@@ -46,7 +46,8 @@ public class MinimumAngleMonitor {
     Optional<Intersection> minAngleCr = MinimumAngle.getMinimumAngleCrossing(graph);
 
     if (minAngleCr.isPresent()){
-      if (oldAngle <= minAngleCr.get().angle){
+      if (oldAngle < minAngleCr.get().angle){
+        System.out.println("set best with nodes: "+ graph.getNodes().size()+ " angle: "+ oldAngle + " new : "+ minAngleCr.get().angle);
         oldAngle = minAngleCr.get().angle;
 
         bestSolution.setBestMinimumAngle(oldAngle, graph.getNodes().size());

### 2018-05-23 src/sidepanel/SidePanelTab.java

@@ -22,7 +22,9 @@ import layout.algo.FraysseixPachPollack;
 import layout.algo.NodeSwapper;
 import layout.algo.execution.BasicIGraphLayoutExecutor;
 import layout.algo.execution.ILayout;
+import layout.algo.layoutinterface.AbstractLayoutInterfaceItem;
 import layout.algo.layoutinterface.ILayoutConfigurator;
+import layout.algo.randommovement.RandomMovementLayout;
 import layout.algo.utils.PositionMap;
 import main.MainFrame;
 import util.BoundingBox;
@@ -475,6 +477,7 @@ public class SidePanelTab {
 //--------------------------------------------------------------------
             initSidePanel.addDefaultListeners();
         });
+        System.out.println("rmeoved "+removedChains.size());
     }
 
     private void reinsertChainItemActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {
@@ -486,8 +489,8 @@ public class SidePanelTab {
             return;
         }
 
-        modifyGraph(() -> {
-            initSidePanel.removeDefaultListeners();
+//        modifyGraph(() -> {
+//            initSidePanel.removeDefaultListeners();
 
 //--------------------------------------------------------------------
 //            //TODO reinsert chains (currently regular reinsert 1 chain)
@@ -508,10 +511,10 @@ public class SidePanelTab {
             reinsertVerticesItem();
 //--------------------------------------------------------------------
 
-            Scaling.scaleNodeSizes(initSidePanel.mainFrame.view);
-
-            initSidePanel.addDefaultListeners();
-        });
+//            Scaling.scaleNodeSizes(initSidePanel.mainFrame.view);
+//
+//            initSidePanel.addDefaultListeners();
+//        });
     }
 
 
@@ -541,45 +544,80 @@ public class SidePanelTab {
             return;
         }
         Thread automaticInsertion = new Thread(() -> {
-                double startingAngle = initSidePanel.mainFrame.minimumAngleMonitor.getMinimumAngle();
-                double epsilon = startingAngle/100;
-                int last_iteration = executor.getCurrentIteration();
-                if (!algorithmName.equals("Random Movement")) {
-                    setOutputTextArea("Recommended to use Random Movement!");
-                    System.out.println("Recommended to use Random Movement!");
+            double startingAngle = initSidePanel.mainFrame.minimumAngleMonitor.getMinimumAngle();
+            double epsilon = startingAngle/10;
+            boolean runRandomAll = false;
[…]

### 2018-05-23 paper/paper.tex

@@
 Of course, we cannot consider all possible positions for the vertex to be moved.
-Therefore, we discretize them by considering a number of rays around the chosen vertex and by choosing (at random again) a single point at each of them.
+Instead, we consider a small set of randomly generated ones.
+Then, if there exists a position among them, that does not lead to a reduction of the crossing resolution, we simply move the vertex to this~position.
 
@@
 %
-From our experimental evaluation, we observed that, while this approach turned to be quite effective for medium-size graphs, for larger and denser graph, unfortunately, it was not so efficient.
-What we observed is that in larger graphs it was unlike to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not at the part of the drawing where the critical vertices resided.
+From our extensive practical analysis, we quickly realized that, while this approach turned out to be quite effective for medium-size graphs, for larger and denser graph, unfortunately, it was not so efficient.
+What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
 
@@
 
-The termination condition of our algorithm is simple and depends also on a input parameter $\tau$, such that $\tau \geq \zeta + \zeta'$.
+The termination condition of our algorithm is simple and depends also on an input parameter $\tau$, such that $\tau \geq \zeta + \zeta'$.
 More precisely, if the crossing resolution during the last $\tau$ iterations has not been improved, then we assume that the algorithm has converged and we do not proceed any more.
@@
 The algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produced such drawings.
+However, it can be easily adjusted to produce such drawings.
 Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ candidate points $\pi_0,\ldots,\pi_{\rho-1}$.
@@
 For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points.
+If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighborhood.
 

### 2018-05-23 src/io/Contest2018IOHandler.java
JSON-2018 add node tags for saving manual graphs

@@ -132,6 +132,14 @@ public class Contest2018IOHandler {
      * Writing graph in JSON file with contest 2018 format
      */
     public void write(IGraph graph, String outputFileName, JTextArea outputTextArea) throws IOException {
+        //allow saving of non-contest files in contest format
+        int tagNum = 0;
+        for (INode u : graph.getNodes()) {
+            if (u.getTag() == null) {
+                u.setTag(tagNum);
+            }
+            tagNum++;
+        }
         //Get Nodes
         ArrayList<Contest2018IOHandler.Node> nodes = new ArrayList<>(graph.getNodes().size());
         StringBuilder errorText = new StringBuilder();

### 2018-05-27 src/algorithms/graphs/MinimumAngle.java
Closes #162 Working Angular (Total) Resolution in Random Movement -Default is minimum angle -display angle and random movement angle not synced

@@ -3,15 +3,17 @@ package algorithms.graphs;
 import com.sun.istack.internal.Nullable;
 import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.graph.*;
-import com.yworks.yfiles.graph.styles.*;
-import com.yworks.yfiles.view.*;
+import com.yworks.yfiles.graph.styles.IEdgeStyle;
+import com.yworks.yfiles.graph.styles.PolylineEdgeStyle;
+import com.yworks.yfiles.view.Pen;
+import layout.algo.utils.PositionMap;
+import util.Util;
+import util.graph2d.Intersection;
+import util.graph2d.LineSegment;
 
 import java.util.*;
-import java.util.stream.*;
-
-import layout.algo.utils.PositionMap;
-import util.*;
-import util.graph2d.*;
+import java.util.stream.Collectors;
+import java.util.stream.Stream;
 
 // to allow overriding for caching, we have an explicit singleton object, which implements all methods. Static calls just use the singleton.
 // This is because you can't derive static methods to member-methods.
@@ -99,7 +101,7 @@ public class MinimumAngle {
   }
 
   public static double getMinimumAngleForNode(Mapper<INode, PointD> positions, INode node, IGraph graph) {
-    return getMinimumAngleForNode(graph, node, positions).orElse(Double.POSITIVE_INFINITY);
+    return m.getMinimumAngleForNode(graph, node, positions).orElse(Double.POSITIVE_INFINITY);
   }
 
   public static Optional<Double> getMinimumAngleForNode(IGraph graph, INode node, @Nullable Mapper<INode, PointD> nodePositions) {

### 2018-05-27 src/experiment/Experiment.java

@@ -382,7 +382,7 @@ public class Experiment {
         mainFrame.minimumAngleMonitor.updateMinimumAngleInfoBar();
 
 
-        this.actAngle = Math.round(1000.0 * mainFrame.minimumAngleMonitor.getActAngle()) / 1000.0;
+        this.actAngle = Math.round(1000.0 * mainFrame.minimumAngleMonitor.getCurrentAngle()) / 1000.0;
         if(this.minimumAngle >= this.actAngle){
             this.numOfUnchangedAngle++;
             System.out.println("!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!numebrs   " + this.numOfUnchangedAngle);

### 2018-05-27 src/layout/algo/randommovement/RandomMovementConfigurator.java

@@ -18,6 +18,8 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
   AbstractLayoutInterfaceItem<Boolean> onlyGridPositions;
   ArrayList<AbstractLayoutInterfaceItem> itemList;
   AbstractLayoutInterfaceItem<Boolean> allowIncreaseStepSize;
+  AbstractLayoutInterfaceItem<Boolean> useCrossingResolution;
+  AbstractLayoutInterfaceItem<Boolean> useAngularResolution;
 
 
   @Override
@@ -60,6 +62,12 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
 
     allowIncreaseStepSize = itemFactory.booleanParameter("Automatically increase step size");
     allowIncreaseStepSize.setValue(false);
+
+    useCrossingResolution = itemFactory.booleanParameter("Use Crossing Resolution (Default)");
+    useCrossingResolution.setValue(true);
+
+    useAngularResolution = itemFactory.booleanParameter("Use Angular Resolution");
+    useAngularResolution.setValue(false);
   }
 
   @Override

### 2018-05-27 src/layout/algo/randommovement/RandomMovementLayout.java

@@ -1,5 +1,6 @@
 package layout.algo.randommovement;
 
+import algorithms.graphs.AngularResolution;
 import algorithms.graphs.MinimumAngle;
 import com.yworks.yfiles.algorithms.Bfs;
 import com.yworks.yfiles.algorithms.Node;
@@ -94,7 +95,7 @@ public class RandomMovementLayout implements ILayout {
 
     Collection<Sample> samples = selectRandomSamples();
 
-    double originalAngle = MinimumAngle.getMinimumAngleForNode(positions, node, graph);
+    double originalAngle = getAngleForNode(positions, node, graph);
     PointD originalPosition = positions.getValue(node);
 
     Sample[] goodSamples = samples.stream()
@@ -112,7 +113,7 @@ public class RandomMovementLayout implements ILayout {
         .filter(sample -> LayoutUtils.overlap(sample.position, positions))
         .peek(sample -> {
           positions.setValue(node, sample.position);
-          sample.minimumAngle = MinimumAngle.getMinimumAngleForNode(positions, node, graph);
+          sample.minimumAngle = getAngleForNode(positions, node, graph);
         })
         .filter(sample -> sample.minimumAngle > originalAngle)
         .toArray(Sample[]::new);
@@ -200,7 +201,7 @@ public class RandomMovementLayout implements ILayout {
         .limit(configurator.numSamplesForJumping.getValue())
         .max(Comparator.comparingDouble(position -> {
           positions.setValue(node, position);
-          return MinimumAngle.getMinimumAngleForNode(positions, node, graph);
+          return getAngleForNode(positions, node, graph);
         }))
         .ifPresent(pointD -> positions.setValue(node, pointD));
   }
@@ -254,4 +255,15 @@ public class RandomMovementLayout implements ILayout {
   public Mapper<INode, PointD> getNodePositions() {
     return positions;
   }
+
+  private double getAngleForNode(Mapper<INode, PointD> positions, INode node, IGraph graph) {
+    if (configurator.useCrossingResolution.getValue() && configurator.useAngularResolution.getValue()) {
+      return Math.min(MinimumAngle.getMinimumAngleForNode(positions, node, graph), AngularResolution.getAngularResolutionForNode(positions, node, graph));
+    } else if (configurator.useAngularResolution.getValue()) {
+      return AngularResolution.getAngularResolutionForNode(positions, node, graph);
+    } else {
+     return MinimumAngle.getMinimumAngleForNode(positions, node, graph);
+    }
+  }
+
 }

### 2018-05-27 src/main/InitMenuBar.java

@@ -819,7 +819,8 @@ public class InitMenuBar {
                 this.mainFrame.removedNodes.clear();
                 this.mainFrame.removedChains.clear();
                 this.mainFrame.bestSolution.reset();
-                this.mainFrame.minimumAngleMonitor.updateMinimumAngleInfoBar();
+//                this.mainFrame.minimumAngleMonitor.updateMinimumAngleInfoBar();
+                this.mainFrame.minimumAngleMonitor.updateAngleInfoBar();
             } catch (IOException ioe) {
                 this.infoLabel.setText("An error occured while reading the input file.");
             } finally {
@@ -1003,7 +1004,9 @@ public class InitMenuBar {
     }
 
     private void minimumCrossingAngleMenuActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {
-        minimumAngleMonitor.showMinimumAngle(graph, view, infoLabel, true);
+//        minimumAngleMonitor.showMinimumAngle(graph, view, infoLabel, true);
+//        minimumAngleMonitor.updateMinimumAngleInfoBar();
+        minimumAngleMonitor.updateAngleInfoBar();
     }
 
     private void overlappingNodesMenuActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {

### 2018-05-27 src/main/MinimumAngleMonitor.java

@@ -1,7 +1,7 @@
 package main;
 
+import algorithms.graphs.AngularResolution;
 import algorithms.graphs.MinimumAngle;
-
 import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.graph.*;
 import com.yworks.yfiles.utils.IEventListener;
@@ -19,46 +19,60 @@ public class MinimumAngleMonitor {
   private IGraph graph;
   private JLabel infoLabel;
   private GraphComponent view;
-  private double oldAngle;
-  private double actAngle;
+  private double oldMinAngle;
+  private double oldAngularRes;
+  private double currentAngle;
+  private boolean useCrossingResolution;
+  private boolean useAngularResolution;
 
   private BestSolutionMonitor bestSolution;
 
-  private IEventListener<ItemEventArgs<IEdge>> minimumAngleEdgeCreatedListener = (o, ItemEventArgs) -> showMinimumAngle(graph, view, infoLabel, false);
-  private IEventListener<EdgeEventArgs> minimumAngleEdgeRemovedListener = (o, EdgeEventArgs) -> showMinimumAngle(graph, view, infoLabel, false);
-  private IEventListener<ItemEventArgs<INode>> minimumAngleNodeCreatedListener = (o, ItemEventArgs) -> showMinimumAngle(graph, view, infoLabel, false);
-  private IEventListener<NodeEventArgs> minimumAngleNodeRemovedListener = (o, NodeEventArgs) -> showMinimumAngle(graph, view, infoLabel, false);
-  private INodeLayoutChangedHandler minimumAngleLayoutChangedHandler = (o, iNode, rectD) -> showMinimumAngle(graph, view, infoLabel, false);
+  private IEventListener<ItemEventArgs<IEdge>> minimumAngleEdgeCreatedListener = (o, ItemEventArgs) -> updateAngleInfoBar();
+  private IEventListener<EdgeEventArgs> minimumAngleEdgeRemovedListener = (o, EdgeEventArgs) -> updateAngleInfoBar();
+  private IEventListener<ItemEventArgs<INode>> minimumAngleNodeCreatedListener = (o, ItemEventArgs) -> updateAngleInfoBar();
+  private IEventListener<NodeEventArgs> minimumAngleNodeRemovedListener = (o, NodeEventArgs) -> updateAngleInfoBar();
[…]

### 2018-05-27 src/sidepanel/InitSidePanel.java

@@ -38,6 +38,8 @@ public class InitSidePanel {
     public JCheckBox masterEnableMinimumAngle;
     private boolean stateEnableMinimumAngle;
     public JCheckBox masterAllowClickCreateNodeEdge;
+    public JCheckBox masterEnableCrossingResolution;
+    public JCheckBox masterEnableAngularResolution;
 
 
     public InitSidePanel(MainFrame mainFrame) {
@@ -85,12 +87,22 @@ public class InitSidePanel {
         masterAllowClickCreateNodeEdge.addItemListener(this::masterAllowClickCreateNodeEdgeActionPerformed);
         masterAllowClickCreateNodeEdge.setSelected(false);
 
+        masterEnableCrossingResolution = new JCheckBox("Crossing Resolution");
+        masterEnableCrossingResolution.addItemListener(this::masterEnableCrossingResolutionActionPerformed);
+        masterEnableCrossingResolution.setSelected(true);
+
+        masterEnableAngularResolution = new JCheckBox("Angular Resolution");
+        masterEnableAngularResolution.addItemListener(this::masterEnableAngularResolutionActionPerformed);
+        masterEnableAngularResolution.setSelected(false);
+
         sidePanelTabs = new ArrayList<>();
 
         tabbedSidePane.addChangeListener(changeEvent -> {
             int selectedTab = tabbedSidePane.getSelectedIndex();
             sidePanelTabs.get(selectedTab).setEnableMinimumAngleDisplay(masterEnableMinimumAngle.isSelected());
             sidePanelTabs.get(selectedTab).setAllowClickCreateNodeEdge(masterAllowClickCreateNodeEdge.isSelected());
+            sidePanelTabs.get(selectedTab).setEnableCrossingResolution(masterEnableCrossingResolution.isSelected());
+            sidePanelTabs.get(selectedTab).setEnableAngularResolution(masterEnableAngularResolution.isSelected());
             for (int i = 0; i < sidePanelTabs.size() - 1; i++) {    //exclude misc
                 if (sidePanelTabs.get(i).getExecutor().isRunning()) {
                     sidePanelTabs.get(selectedTab).setOutputTextArea(sidePanelTabs.get(i).algorithmName + " is Still Running!");
@@ -178,7 +190,7 @@ public class InitSidePanel {
         if (stateEnableMinimumAngle && removedListeners) {
[…]

### 2018-05-27 src/util/DisplayMessagesGui.java

@@ -36,4 +36,10 @@ public class DisplayMessagesGui {
                 " after " + (iterations +1) + " iterations.";
         return text;
     }
+
+    public static String createAngularResMsg(double angularRes, int nodes, BestSolutionMonitor bestSolution) {
+        String text = "Angular Resolution: " + angularRes;
+        text += " | Stored Best: " + bestSolution.getBestAngularResolutionForNodes(nodes).orElse(Double.POSITIVE_INFINITY);
+        return text;
+    }
 }

### 2018-05-27 paper/paper.tex
A bit more on the related work section.

@@
 Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting spring embedding algorithms.
+However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
 
@@
 
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms~\cite{DBLP:journals/congnum/Eades84}.
+From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
 The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one is due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
@@
 To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of spring embedding algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
+Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of force-directed algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
 
@@
 An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-Exceptions are the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}.
-According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
+Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
+%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
+
+More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
+%
+In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
+%
+However, the obtained drawings are not straight-line.
+For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
+%
+Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the force-directed technique.
 

### 2018-05-27 paper/paper.tex
line numbers

@@
 \maketitle
+\linenumbers
 % ==================================================================
@@
 %In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper. 
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
 We conclude in Section~\ref{sec:conclusions} with interesting open problems and future~directions.
@@
 %
-Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the force-directed technique.
+Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are also based on the force-directed technique.
 

### 2018-05-27 paper/paper.tex
Addressing some comments by Michael.

@@
 \begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed by any pair of crossing edges.
+The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
 Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
@@
 
-In this paper, we propose a new randomization-based heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known ones from the literature.
+In this paper, we propose a new randomization-based heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
 Our experimental evaluation indicates that the new heuristic produces drawings with significantly better crossing resolution, but this comes at the cost of slightly higher running time, especially when the input graph is large and dense.
@@
 In Graph Drawing, there exists a really rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restricted), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the enstablished techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
+However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
+As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
 
@@
 
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to compute the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
+In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
 Since we could not provide any theoretical guarantee on the expected quality of the produced drawings as well (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
@@
 Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} completely characterized the complete bipartite RAC graphs.
+Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
+%Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} completely characterized the complete bipartite RAC graphs.
 Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
@@
 We also assume that no two edges of $G$ overlap in $\Gamma_0$, that is, $c(\Gamma_0)>0$.
-Note that a drawing $\Gamma_0$ of $G$ meeting these preconditions can be, e.g., a circular drawing of $G$ or a drawing obtained by applying a standard spring embedding algorithm on $G$.
+A circular drawing of $G$ or a drawing obtained by applying a standard force-directed algorithm on $G$ clearly meets these preconditions.
 
@@
 In other words, we assume, as an invariant property of our algorithm, that the crossing resolution cannot be decreased at some iteration step.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrary at random from a so-called \emph{vertex-pool}, which may contain:
+Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random from a so-called \emph{vertex-pool}, which may contain:
 
@@
 
-To formally define the critical vertices, we first need to introduce the notion of critical edge-pairs.
-A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is either equal to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
+Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
+To formally define them, we first need to introduce the notion of critical edge-pairs.
+A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
 The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
 
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-As a result, their role in our algorithm is central\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
+The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
 The reason is that by slightly changing the location of a critical vertex appropriately, one naturally expects to improve the crossing resolution of the current drawing.
@@
 %
-From our extensive practical analysis, we quickly realized that, while this approach turned out to be quite effective for medium-size graphs, for larger and denser graph, unfortunately, it was not so efficient.
+From our extensive practical analysis, we quickly realized that, while this approach turned out to be quite effective for medium-size graphs, for larger and denser graphs, unfortunately, it was not so efficient.
 What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.

### 2018-05-29 src/layout/algo/execution/BasicIGraphLayoutExecutor.java
Fixes Slow Forces with many nodes -also means that there is no nice node-by-node movement for medium graphs, just one big movement per iteration

@@ -167,11 +167,13 @@ public class BasicIGraphLayoutExecutor {
   }
 
   private void updateGraph(Mapper<INode, PointD> nodePositions) {
+    propertyChange.firePropertyChange("removeListeners", false, true);
     synchronized (graph) {
       for (Map.Entry<INode, PointD> entry : nodePositions.getEntries()) {
         graph.setNodeCenter(entry.getKey(), entry.getValue());
       }
     }
+    propertyChange.firePropertyChange("removeListeners", true, false);
   }
 
   protected void updateProgress(int iteration) {

### 2018-05-29 src/experiment/Experiment.java
Reset the output crossing angle to the best one, termination after 2 iterations with an epsilon and other things

@@ -2,6 +2,7 @@ package experiment;
 
 import algorithms.graphs.GridGraph;
 import algorithms.graphs.MinimumAngle;
+import com.sun.corba.se.impl.orbutil.graph.Graph;
 import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.geometry.RectD;
 import com.yworks.yfiles.graph.*;
@@ -33,7 +34,6 @@ public class Experiment {
     private long maxTimeForAlgo = 2000; //default 100 sec
     private boolean reached90Degree = false;
     private boolean isInInvLoop = false;
-    private boolean firstOrganicLayout = true;
     private PrintWriter pw;
     private StringBuilder sb;
     private int maxIterations = 1000;
@@ -43,8 +43,8 @@ public class Experiment {
     private int  minimumAngleNumOfCrossings;
     private double aspect_ratio, minimumAngleAspectratio;
     private double angular_resolution, minimumAngleAngularResolution;
+    private final double approx_epsilon = 0.001; // If the angle plus epsilon doesn't change  after number of iterations, then it is converged
     private double actAngle = 0;
-    private int algorithmNum = 1; // 1: Random Algo. 2: Force Algo
     private Mapper<INode, PointD> positions;
     Optional<SidePanelTab> tab;
 
@@ -262,10 +262,6 @@ public class Experiment {
      */
     public  void runAlgorithms() {
         this.mainFrame.initSidePanel.removeDefaultListeners();
-        if(this.firstOrganicLayout){
-            LayoutUtilities.applyLayout(this.mainFrame.graph, new OrganicLayout());
-            this.firstOrganicLayout = false;
-        }
 
         this.mainFrame.initSidePanel.addDefaultListeners();
 
@@ -290,13 +286,14 @@ public class Experiment {
         this.positions = this.tab.get().getExecutor().getLayout().getNodePositions();
         System.out.println(fileName + "OUT 13");
         this.endTime = endTime + (System.currentTimeMillis() - tmpTime);
+        this.iterations += maxIterations;
+        calcGraphInformations();
     }
 
     /**
      * Loads a new graph and reset the parameters
      */
     public void loadGraph() {
-        this.firstOrganicLayout = true;
         this.reached90Degree = false;
         this.isInInvLoop = false;
         this.numOfUnchangedAngle = 0;
[…]

### 2018-05-29 src/experiment/Experiments.java

@@ -45,11 +45,14 @@ public class Experiments {
     private int numOfIterationFactor = 1000;
     private long maxCalcTime = Long.MAX_VALUE; //in mili. sec
     private int boxSize = 10000;
-    private int maxNumberOfUnchangedAngle = 50;
+    private int maxNumberOfUnchangedAngle = 2;
     private boolean planarGraphsAllowed = false;
     private boolean unconnectedGraphsAllowed = true;
     boolean reached90Deg = false;
     private String[] childernP;
+    private boolean withCrossingRes = true;
+    private boolean withAngularRes = false;
+
 
 
     public Experiments()
@@ -75,7 +78,7 @@ public class Experiments {
 
     }
 
-    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize, boolean planarGraphsAllowed, boolean unconnectedGraphsAllowed){
+    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize, boolean planarGraphsAllowed, boolean unconnectedGraphsAllowed, boolean withCrossingRes, boolean withAngularRes){
         this.comp = new GraphComponent();
         this.inputDirectory = inputDirectory;
         this.outputDirectory = outputDirectory;
@@ -85,6 +88,8 @@ public class Experiments {
         this.boxSize = boxSize;
         this.planarGraphsAllowed = planarGraphsAllowed;
         this.unconnectedGraphsAllowed = unconnectedGraphsAllowed;
+        this.withCrossingRes = withCrossingRes;
+        this.withAngularRes = withAngularRes;
 
     }
 
@@ -275,24 +280,33 @@ try {
                 frame.dispose();
                 return;}//TODO: Nur zum testen
 
+
+
             if(algorithmName.equals(this.randomMovementString)){
                 experiment.setAlgorithm(RandomMovementLayout.class);
                                     /* Set config. parameters */
                 experiment.getTab().get().configurator.getItems().get(0).setValue(10.0); // Minimum step size
                 experiment.getTab().get().configurator.getItems().get(1).setValue(150.0); // Maximum step size
-                experiment.getTab().get().configurator.getItems().get(2).setValue(50); // Faild iteration ... detect loacal maximum
+                experiment.getTab().get().configurator.getItems().get(2).setValue(50); // Failed iteration ... detect loacal maximum
[…]

### 2018-05-29 src/layout/algo/randommovement/RandomMovementConfigurator.java

@@ -67,9 +67,11 @@ public class RandomMovementConfigurator implements ILayoutConfigurator {
 
     useCrossingResolution = itemFactory.booleanParameter("Use Crossing Resolution (Default)");
     useCrossingResolution.setValue(true);
+    itemList.add(useCrossingResolution);
 
     useAngularResolution = itemFactory.booleanParameter("Use Angular Resolution");
     useAngularResolution.setValue(false);
+    itemList.add(useAngularResolution);
 
   }
 

### 2018-05-29 paper/paper.tex

@@
 
-Formally, the minimum angle formed by any pair of crossing edges in a drawing is refered to as its \emph{crossing resolution}.
+Formally, the value of the minimum angle formed by any pair of crossing edges in a drawing is refered to as its \emph{crossing resolution}.
 Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
@@
 Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-%Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} completely characterized the complete bipartite RAC graphs.
+Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
 Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
@@
 In other words, we assume, as an invariant property of our algorithm, that the crossing resolution cannot be decreased at some iteration step.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random from a so-called \emph{vertex-pool}, which may contain:
+Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
 
@@
 The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
-The reason is that by slightly changing the location of a critical vertex appropriately, one naturally expects to improve the crossing resolution of the current drawing.
-In fact, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+The reason is that by slightly changing the location of a critical vertex or the location of one of its neighbours appropriately, one naturally expects to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a weighted random selection procedure, in which the vertices which are at distance $i$ from the vertices of the vertex-pool in the graph have higher weights than the corresponding vertices at distance $j$, when $ \leq <i<j$.
+Hence, if the vertex-pool contains critical vertices, then the closer a vertex is to a critical vertex, the more likely it is to be chosen.
+Otherwise, each vertex of the graph can be chosen with the same probability.
+
+With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
 However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
@@
 
-So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex from the vertex-pool of the current drawing to move, so to improve the crossing resolution.
+So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
 We have not described, however, how the chosen vertex is moved, i.e., how we compute its new position in the next drawing.
 
-Let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm.
+Let $v_i$ be the vertex of the drawing $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration of our algorithm.
 Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
@@
 However, it can be easily adjusted to produce such drawings.
-Recall that, at the $i$-th iteration step, our algorithm chooses uniformly at random a vertex $v_i$ from the vertex-pool and determines a set of $\rho$ candidate points $\pi_0,\ldots,\pi_{\rho-1}$.
+Recall that, at the $i$-th iteration step, our algorithm chooses at random a vertex $v_i$ (based on the content of the vertex-pool), and determines a set of $\rho$ candidate points $\pi_0,\ldots,\pi_{\rho-1}$.
 If we round these points to their closest grid points, say $\overline{\pi}_0,\ldots,\overline{\pi}_{\rho-1}$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
@@
 For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighborhood.
+If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
 

### 2018-05-30 src/layout/algo/randommovement/RandomMovementLayout.java
Only check overlap if gridding enabled, node-edge overlap minor

@@ -110,7 +110,7 @@ public class RandomMovementLayout implements ILayout {
           }
         })
         .filter(sample -> boundingBox.contains(sample.position))
-        .filter(sample -> LayoutUtils.overlap(sample.position, positions, node, graph))
+        .filter(sample -> configurator.onlyGridPositions.getValue() || LayoutUtils.overlap(sample.position, positions, node, graph))
         .peek(sample -> {
           positions.setValue(node, sample.position);
           sample.minimumAngle = getAngleForNode(positions, node, graph);

### 2018-05-30 src/main/MinimumAngleMonitor.java
update current angle if getting but listeners not enabled

@@ -69,7 +69,6 @@ public class MinimumAngleMonitor {
     if (!bestSolution.getBestMinimumAngleForNodes(graph.getNodes().size()).isPresent()) {
       oldMinAngle = 0;
     }
-    computeAngularResolution();
 
     Optional<Intersection> minAngleCr = computeMinimumAngle();
 
@@ -185,10 +184,12 @@ public class MinimumAngleMonitor {
   }
 
   public double getCurrentCrossingResolution() {
+    computeMinimumAngle();
     return currentCrossingResolution;
   }
 
   public double getCurrentAngularResolution() {
+    computeAngularResolution();
     return currentAngularResolution;
   }
 }

### 2018-05-30 src/sidepanel/SidePanelTab.java
minor chain fix

@@ -580,10 +580,10 @@ public class SidePanelTab {
                     try{Thread.sleep(100);} catch(InterruptedException c){}
                 }
 //                if (executor.getCurrentIteration() >= last_iteration + 1000 && !runRandomAll) {
-                if (!runRandomAll && false) {
+//                if (!runRandomAll && false) {
+                if (!runRandomAll) {
 //                    if (minAngle < startingAngle - epsilon) {
                     if (count >= 20 ) {
-                        System.out.println("its habbening");
                         Map<String, AbstractLayoutInterfaceItem<Boolean>> map = configurator.getBooleanParameters().orElse(null);
                         try {
                             if (map != null) {

### 2018-05-30 src/layout/algo/utils/MutualityCheck.java
Mutual crossings get resolved via button

@@ -1,46 +1,111 @@
 package layout.algo.utils;
 
+import java.util.Arrays;
 import java.util.Iterator;
 import java.util.List;
+import java.util.Random;
+
 import util.graph2d.Intersection;
 
 
 import com.yworks.yfiles.geometry.PointD;
+import com.yworks.yfiles.graph.IEdge;
 import com.yworks.yfiles.graph.IGraph;
 import com.yworks.yfiles.graph.IMapper;
 import com.yworks.yfiles.graph.INode;
+import com.yworks.yfiles.graph.Mapper;
+import com.yworks.yfiles.utils.IListEnumerable;
 
 
 public class MutualityCheck {
+	
+	
 	public static class MutualityCheckHelper{
 		
-		public void detector (IGraph graph, IMapper<INode, PointD> nodePositions) {
+		public static void detector (IGraph graph, Mapper<INode, PointD> nodePositions) {
 			List<Intersection> getCrossings = algorithms.graphs.yFilesSweepLine.getCrossings(graph, true, nodePositions);
-			int amountOfNodes = graph.getNodes().size();
-			int neighborsWithCrossings[][]= new int [amountOfNodes][amountOfNodes];
+			int amountOfEdges = graph.getEdges().size();
+			PointD crossingEdges[][]= new PointD [amountOfEdges][amountOfEdges];
+			Boolean[][] crossingEdgesvalid = new Boolean [amountOfEdges][amountOfEdges];
+			//Arrays.fill(crossingEdgesvalid, Boolean.FALSE);
+			IListEnumerable<IEdge> edges = graph.getEdges();
 			Iterator<Intersection> crossings= getCrossings.iterator();
 			while (crossings.hasNext()) {
 				Intersection crossing = crossings.next();
-				neighborsWithCrossings[Integer.parseInt(crossing.segment1.n1.getTag().toString())][Integer.parseInt(crossing.segment1.n2.getTag().toString())]=1; // 1-2
-				neighborsWithCrossings[Integer.parseInt(crossing.segment1.n1.getTag().toString())][Integer.parseInt(crossing.segment2.n1.getTag().toString())]=1; // 1-3
-				neighborsWithCrossings[Integer.parseInt(crossing.segment1.n1.getTag().toString())][Integer.parseInt(crossing.segment2.n2.getTag().toString())]=1; // 1-4
[…]

### 2018-05-30 src/main/MainFrame.java

@@ -43,7 +43,7 @@ public class MainFrame extends JFrame {
 
     /* Graph Drawing related objects */
     public GraphComponent view;
-    public IGraph graph;
+    public static IGraph graph;
     private OrganicLayout defaultLayouter;
     public GraphEditorInputMode graphEditorInputMode;
     private GridVisualCreator gridVisualCreator;

### 2018-05-30 src/sidepanel/SidePanelTab.java

@@ -18,6 +18,8 @@ import com.yworks.yfiles.view.ISelectionModel;
 import graphoperations.Chains;
 import graphoperations.RemovedChains;
 import graphoperations.Scaling;
+import layout.algo.utils.MutualityCheck;
+import layout.algo.utils.MutualityCheck.MutualityCheckHelper;
 import layout.algo.FraysseixPachPollack;
 import layout.algo.NodeSwapper;
 import layout.algo.execution.BasicIGraphLayoutExecutor;
@@ -330,6 +332,13 @@ public class SidePanelTab {
         cCustomPanel.gridy = ++cCustomPanelY;
         swapperItem.addActionListener(this::swapperItemActionPerformed);
         custom.add(swapperItem, cCustomPanel);
+        
+        JButton mutuallyItem = new JButton("Mutually");
+        cCustomPanel.fill = GridBagConstraints.HORIZONTAL;
+        cCustomPanel.gridx = 1;
+        cCustomPanel.gridy = cCustomPanelY;
+        mutuallyItem.addActionListener(this::mutuallyItemActionPerformed);
+        custom.add(mutuallyItem, cCustomPanel);
 
         cSidePanel.fill = GridBagConstraints.HORIZONTAL;
         cSidePanel.gridx = 0;
@@ -796,6 +805,28 @@ public class SidePanelTab {
         initSidePanel.mainFrame.view.updateUI();
         initSidePanel.addDefaultListeners();
     }
+    
+    private void mutuallyItemActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {
+      initSidePanel.removeDefaultListeners();
+      IGraph graph = initSidePanel.mainFrame.graph;
+      Mapper<INode, PointD> nodePositions = PositionMap.FromIGraph(graph);
+      MutualityCheckHelper.detector(graph, nodePositions); 
+      PositionMap.applyToGraph(graph, nodePositions);
+      initSidePanel.mainFrame.view.fitGraphBounds();
+      
+      initSidePanel.mainFrame.view.updateUI();
+      initSidePanel.addDefaultListeners();
+    }
+    
+//    private void mutuallyItemActionPerformed{
+//        initSidePanel.removeDefaultListeners();
+//        IGraph graph=MainFrame.graph;
+//        MutualityCheckHelper.detector(graph, MainFrame.nodePositions); 
+//        
+//        
+//        initSidePanel.mainFrame.view.updateUI();
[…]

### 2018-05-30 src/experiment/Experiment.java
Added extra Folder for the graph results

@@ -337,7 +337,7 @@ public class Experiment {
      */
     public void writeGraph() {
         try {
-            ContestIOHandler.write(graph, outputDirectory + "out_" + filePrefix + ".txt", mainFrame.initSidePanel.getOutputTextArea());
+            ContestIOHandler.write(graph, outputDirectory + "resultGraph\\" + "out_" + filePrefix + ".txt", mainFrame.initSidePanel.getOutputTextArea());
         } catch (IOException ioe) {
             System.out.println("An error occured while exporting the graph.");
         }

### 2018-05-30 src/experiment/MainExperiment.java

@@ -5,37 +5,28 @@ package experiment;
  */
 
 public class MainExperiment {
+    static String inputDir = "E:\\graph\\rome\\rome\\";
+    static String outputDir = "E:\\graph\\results\\TESTrome290518\\";
 
     public static void main(String[] args) {
         //Experiments e = new Experiments();
 
         // Experiments e = new Experiments(("E:\\graph\\graphml\\"), "E:\\graph\\afterRandom\\graphml2\\");
 
-      Experiments e = new Experiments(
-                "E:\\graph\\rome\\rome\\", "E:\\graph\\results\\rome290518\\",
+      Experiments e = new Experiments(inputDir, outputDir,
               Long.MAX_VALUE, 100, 1000, 10000,
-                false, false, true, false);
+              false, false, true, false);
        e.runOnlyRandom();
 
-     e = new Experiments(
-                "E:\\graph\\rome\\rome\\", "E:\\graph\\results\\rome290518\\",
-                Long.MAX_VALUE, 100, 1000, 10000,
-                false, false, true, false);
+       e = new Experiments(inputDir, outputDir,
+               Long.MAX_VALUE, 100, 1000, 10000,
+               false, false, true, false);
         e.runOnlyForce();
-        e = new Experiments(
-                "E:\\graph\\rome\\rome\\", "E:\\graph\\results\\rome290518\\",
+
+        e = new Experiments(inputDir, outputDir,
                 Long.MAX_VALUE, 100, 1000, 10000,
                 false, false, true, false);
         e.runOnlyTotalResolutionForce();
 
-
-
-     /*       Experiments e = new Experiments(
-                    "E:\\graph\\rome\\smallSet\\", "E:\\graph\\results\\rome\\",
-                    10000, 100, 1000, 10000,
-                    false, false);
-
-            e.run();
-
-    */}
+    }
 }

### 2018-05-30 src/layout/algo/randommovement/RandomMovementLayout.java
Add Gaussian Node selection for angular resolution

@@ -217,31 +217,60 @@ public class RandomMovementLayout implements ILayout {
   }
 
   private Optional<INode> gaussianNodeSelection() {
-    return MinimumAngle.getMinimumAngleCrossing(graph, positions)
-        .flatMap(crossing -> {
-          YGraphAdapter graphAdapter = new YGraphAdapter(graph);
 
-          NodeList nodesOfCrossing = new NodeList();
-          nodesOfCrossing.add(graphAdapter.getCopiedNode(crossing.segment1.n1));
-          nodesOfCrossing.add(graphAdapter.getCopiedNode(crossing.segment1.n2));
-          nodesOfCrossing.add(graphAdapter.getCopiedNode(crossing.segment2.n1));
-          nodesOfCrossing.add(graphAdapter.getCopiedNode(crossing.segment2.n2));
+    if (configurator.useAngularResolution.getValue() ||
+            (configurator.useAngularResolution.getValue() && configurator.useCrossingResolution.getValue()
+                    && MinimumAngle.getMinimumAngle(graph, positions).orElse(Double.MAX_VALUE) < AngularResolution.getAngularResolution(graph))) {
+      YGraphAdapter graphAdapter = new YGraphAdapter(graph);
 
-          NodeList[] layers = Bfs.getLayers(graphAdapter.getYGraph(), nodesOfCrossing);
+      NodeList nodesOfCrossing = new NodeList();
+      INode[] tmpList = AngularResolution.getCriticalNodes(graph);
+      for (INode u : tmpList) {
+        nodesOfCrossing.add(graphAdapter.getCopiedNode(u));
+      }
 
-          // select a layer according to a normal distribution with a variance that is equal to half of the number
-          // of layers. The reason for this number is that more than 95% of all values are within the interval
-          // [-2 * sigma; 2 * sigma]:
-          //     2 * sigma = #layers
-          // <=>     sigma = #layers / 2
-          int layerIndex = Math.min((int) Math.floor(Math.abs(random.nextGaussian() * layers.length / 2d)), layers.length - 1);
+      NodeList[] layers = Bfs.getLayers(graphAdapter.getYGraph(), nodesOfCrossing);
 
-          Set<INode> nodesOfLayer = layers[layerIndex].stream()
+      // select a layer according to a normal distribution with a variance that is equal to half of the number
+      // of layers. The reason for this number is that more than 95% of all values are within the interval
+      // [-2 * sigma; 2 * sigma]:
+      //     2 * sigma = #layers
[…]

### 2018-05-30 paper/paper.tex
Shortening the intro; mention the side results for the angular and the total resolution.

@@
 
-In this paper, we propose a new randomization-based heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
+In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
 Our experimental evaluation indicates that the new heuristic produces drawings with significantly better crossing resolution, but this comes at the cost of slightly higher running time, especially when the input graph is large and dense.
@@
 
-In Graph Drawing, there exists a really rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the edge crossings which negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
+In Graph Drawing, there exists a really rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
+As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
 
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence (through a series of eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and therefore the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality and readability of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, which is recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017} and mainly focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
+In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
+% (through eye-tracking experiments), 
+that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
+In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
+As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
 
-Formally, the value of the minimum angle formed by any pair of crossing edges in a drawing is refered to as its \emph{crossing resolution}.
-Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; refer to Figure~\ref{fig:examples} for an illustration.
+Formally, the value of the minimum angle formed by any two crossing edges in a drawing is refered to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; see Figure~\ref{fig:examples}.
 For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
@@
 	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1]{figures/examples}}
+	\includegraphics[page=1,scale=0.85]{figures/examples}}
 	\hfil
 	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2]{figures/examples}}
+	\includegraphics[page=2,scale=0.85]{figures/examples}}
 	\caption{%
@@
 From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one is due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}.
+The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
 Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
+Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
 
-In this paper, we approach the crossing resolution maximization problem from a different perspective.
-We suggest a simple and quite intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one try to increase the crossing resolution of a given drawing?
+In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
+% for computing drawings with high crossing resolution, 
+which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
+How would one increase the crossing resolution of a given drawing?
 First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
@@
 Instead, we consider a small set of randomly generated ones.
-Then, if there exists a position among them, that does not lead to a reduction of the crossing resolution, we simply move the vertex to this~position.
+If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
 
 In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we could not provide any theoretical guarantee on the expected quality of the produced drawings as well (mainly due to the nature of the problem itself), we decided to follow a more practical approach.
-To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of force-directed algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
+Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
+%(mainly due to the nature of the problem itself), 
+we followed a more practical approach.
+We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
+% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms. 
+Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
+%, especially 
+for large and dense graphs.
+Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
 
-\paragraph{Preliminaries:}
+\medskip\noindent{\it Preliminaries:}
 Unless otherwise specified, in this paper we consider simple undirected graphs.
@@
 
-\paragraph{Structure of the paper:}
+\medskip\noindent{\it Structure of the paper:}
 The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works. 
@@
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
-We conclude in Section~\ref{sec:conclusions} with interesting open problems and future~directions.
+We conclude in Section~\ref{sec:conclusions} with open problems.
 

### 2018-05-30 paper/paper.tex
Shortening a bit the related work section.

@@
 
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-The study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
+As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
+An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
+Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
+Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
+Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
+The results are fewer, when the right-angle constraint is relaxed.
+Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
+Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
+
+An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
+Most of the approaches that could be useful in practise are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
+COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
+%
+Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
+%
+For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings. 
+%
+Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+
+%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
+Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
 Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
@@
 
-To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
+%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
 More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
@@
 
-An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
+%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
 Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
 %Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-
-More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
 %
-In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
+%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
 %
-However, the obtained drawings are not straight-line.
+%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
+%
+%However, the obtained drawings are not straight-line.
 For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
 %
-Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are also based on the force-directed technique.
+%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are also based on the force-directed technique.
 

### 2018-05-31 src/experiment/Experiment.java
added total resolution to Monitor

@@ -42,14 +42,16 @@ public class Experiment {
     private int numOfCrossings = -1;
     private int  minimumAngleNumOfCrossings;
     private double aspect_ratio, minimumAngleAspectratio;
-    private double angular_resolution, minimumAngleAngularResolution;
+    private double angular_resolution, minimumAngleAngularResolution, minimumAngleTotalResolution;
     private final double approx_epsilon = 0.001; // If the angle plus epsilon doesn't change  after number of iterations, then it is converged
     private double actAngle = 0;
+    private double actAngularAngle = 0;
+    private double actTotalAngle = 0;
     private Mapper<INode, PointD> positions;
     Optional<SidePanelTab> tab;
 
     private int totalVertecies, totalEdges;
-    private double currentGraphSizeX, currentGraphSizeY, minimumAngle, testAngle;
+    private double currentGraphSizeX, currentGraphSizeY, minimumAngle, testAngle, totalAngle;
 
     public  Experiment (MainFrame mainFrame, String fileName, String inputDir, String outputDir, int boxSize) {
         this.mainFrame = mainFrame;
@@ -81,7 +83,7 @@ public class Experiment {
         sb.append(';');
         sb.append(this.numOfCrossings);
         sb.append(';');
-        sb.append(this.angular_resolution);
+        sb.append(this.actAngularAngle);
         sb.append(';');
         sb.append(this.aspect_ratio);
         sb.append(';');
@@ -133,6 +135,8 @@ public class Experiment {
         sb.append(';');
         sb.append(this.minimumAngleNumOfCrossings);
         sb.append(';');
+        sb.append(this.minimumAngleTotalResolution);
+        sb.append(';');
         sb.append('\n');
 
         Path path = FileSystems.getDefault().getPath(outputDirectory , "\\" + filename);
@@ -166,11 +170,13 @@ public class Experiment {
         sb.append(';');
         sb.append(this.graph.getEdges().size());
         sb.append(';');
-        sb.append(this.angular_resolution);
+        sb.append(this.actAngularAngle);
         sb.append(';');
         sb.append(this.aspect_ratio);
         sb.append(';');
         sb.append(this.numOfCrossings);
+        sb.append(';');
+        sb.append(this.actTotalAngle);
         sb.append(';');
         sb.append('\n');
 
@@ -200,17 +206,19 @@ public class Experiment {
         sb.append(';');
[…]

### 2018-05-31 src/experiment/Experiments.java

@@ -50,8 +50,8 @@ public class Experiments {
     private boolean unconnectedGraphsAllowed = true;
     boolean reached90Deg = false;
     private String[] childernP;
-    private boolean withCrossingRes = true;
-    private boolean withAngularRes = false;
+    private boolean useCrossingRes = true;
+    private boolean useAngularRes = false;
 
 
 
@@ -78,7 +78,8 @@ public class Experiments {
 
     }
 
-    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize, boolean planarGraphsAllowed, boolean unconnectedGraphsAllowed, boolean withCrossingRes, boolean withAngularRes){
+    public Experiments(String inputDirectory, String outputDirectory, long maxCalcTime, int numOfIterationPerStep, int numOfSteps, int boxSize, boolean planarGraphsAllowed,
+                       boolean unconnectedGraphsAllowed, boolean useCrossingRes, boolean useAngularRes){
         this.comp = new GraphComponent();
         this.inputDirectory = inputDirectory;
         this.outputDirectory = outputDirectory;
@@ -88,8 +89,8 @@ public class Experiments {
         this.boxSize = boxSize;
         this.planarGraphsAllowed = planarGraphsAllowed;
         this.unconnectedGraphsAllowed = unconnectedGraphsAllowed;
-        this.withCrossingRes = withCrossingRes;
-        this.withAngularRes = withAngularRes;
+        this.useCrossingRes = useCrossingRes;
+        this.useAngularRes = useAngularRes;
 
     }
 
@@ -217,7 +218,7 @@ public class Experiments {
         frame.setDefaultCloseOperation(WindowConstants.DISPOSE_ON_CLOSE);
 
 
-       // frame.BOX_SIZE = this.boxSize; //TODO unterschiedliche boxsize?
+       // frame.BOX_SIZE = this.boxSize; 
         //String fileName = index +"";
         String fileName = pattern;
         System.out.println("Graph " + fileName + " started.");
@@ -229,6 +230,8 @@ public class Experiments {
 
         experiment.setMaxIterations(this.numOfIteration);
         experiment.setMaxTime(this.maxCalcTime);
+        experiment.getTab().get().setEnableAngularResolution(this.useAngularRes);
+        experiment.getTab().get().setEnableCrossingResolution(this.useCrossingRes);
 
[…]

### 2018-05-31 src/experiment/MainExperiment.java

@@ -15,17 +15,17 @@ public class MainExperiment {
 
       Experiments e = new Experiments(inputDir, outputDir,
               Long.MAX_VALUE, 100, 1000, 10000,
-              false, false, true, false);
+              false, false, true, true);
        e.runOnlyRandom();
 
        e = new Experiments(inputDir, outputDir,
                Long.MAX_VALUE, 100, 1000, 10000,
-               false, false, true, false);
+               false, false, true, true);
         e.runOnlyForce();
 
         e = new Experiments(inputDir, outputDir,
                 Long.MAX_VALUE, 100, 1000, 10000,
-                false, false, true, false);
+                false, false, true, true);
         e.runOnlyTotalResolutionForce();
 
     }

### 2018-05-31 src/main/MinimumAngleMonitor.java

@@ -21,8 +21,10 @@ public class MinimumAngleMonitor {
   private GraphComponent view;
   private double oldMinAngle;
   private double oldAngularRes;
+  private double oldTotalRes;
   private double currentCrossingResolution;
   private double currentAngularResolution;
+  private double currentTotalResolution;
   private boolean useCrossingResolution;
   private boolean useAngularResolution;
 
@@ -44,6 +46,7 @@ public class MinimumAngleMonitor {
     this.oldAngularRes = 0;
     this.currentCrossingResolution = 0;
     this.currentAngularResolution = 0;
+    this.currentTotalResolution = 0;
     this.useCrossingResolution = true;
     this.useAngularResolution = false;
   }
@@ -150,6 +153,39 @@ public class MinimumAngleMonitor {
 
   }
 
+  public double computeTotalResolution() {
+    computeMinimumAngle();
+    computeAngularResolution();
+
+    System.out.println("COMPUTE Total Resolution:  cross "  + this.currentCrossingResolution);
+    System.out.println("COMPUTE Total Resolution:   andular  "  +  this.currentAngularResolution );
+    if (useAngularResolution && useCrossingResolution) {
+      if (currentAngularResolution < currentCrossingResolution) {
+        this.currentTotalResolution = currentAngularResolution;
+      } else {
+        this.currentTotalResolution = currentCrossingResolution;
+      }
+    } else if (useAngularResolution) {
+      this.currentTotalResolution = currentAngularResolution;
+    } else {
+      this.currentTotalResolution = currentCrossingResolution;
+    }
+
+    System.out.println("COMPUTE Total Resolution:   total   "  +  this.currentTotalResolution);
+
+    if(oldTotalRes <= this.currentTotalResolution){
+      this.oldTotalRes = this.currentTotalResolution;
+    }
+
+    System.out.println("COMPUTE Total Resolution:   old crossing"  + this.oldMinAngle);
+    System.out.println("COMPUTE Total Resolution:   old angular"  +  this.oldAngularRes);
+    System.out.println("COMPUTE Total Resolution:   old total"  + this.oldTotalRes );
+
+    return this.currentTotalResolution;
+  }
+
+
+
   public void registerGraphChangedListeners() {
     graph.addNodeLayoutChangedListener(minimumAngleLayoutChangedHandler);
     graph.addEdgeCreatedListener(minimumAngleEdgeCreatedListener);
[…]

### 2018-06-04 src/experiment/Experiment.java
north Test

@@ -323,7 +323,8 @@ public class Experiment {
             System.out.println("Try to open .txt");
             this.mainFrame.openContestFile(this.inputDirectory + this.fileName);
         } else if(this.fileName.endsWith(".graphml")){
-            this.mainFrame.openFile(this.inputDirectory + this.fileName);
+           // this.mainFrame.openFile(this.inputDirectory + this.fileName);
+            this.mainFrame.openSimpleFile(this.inputDirectory + this.fileName);
         }
 
 

### 2018-06-04 src/experiment/Experiments.java

@@ -45,7 +45,7 @@ public class Experiments {
     private int numOfIterationFactor = 1000;
     private long maxCalcTime = Long.MAX_VALUE; //in mili. sec
     private int boxSize = 10000;
-    private int maxNumberOfUnchangedAngle = 2;
+    private int maxNumberOfUnchangedAngle = 5;
     private boolean planarGraphsAllowed = false;
     private boolean unconnectedGraphsAllowed = true;
     boolean reached90Deg = false;
@@ -214,7 +214,7 @@ public class Experiments {
     private void openFrame(String pattern, String algorithmName) {
         MainFrame frame = new MainFrame();
         frame.init();
-        frame.setVisible(false); //TODO
+        frame.setVisible(true); //TODO
         frame.setDefaultCloseOperation(WindowConstants.DISPOSE_ON_CLOSE);
 
 

### 2018-06-04 src/experiment/MainExperiment.java

@@ -6,27 +6,94 @@ package experiment;
 
 public class MainExperiment {
     static String inputDir = "E:\\graph\\rome\\rome\\";
-    static String outputDir = "E:\\graph\\results\\TESTrome290518\\";
+    static String outputDir = "E:\\graph\\results\\rome_Angular_01062018\\";
 
     public static void main(String[] args) {
         //Experiments e = new Experiments();
 
         // Experiments e = new Experiments(("E:\\graph\\graphml\\"), "E:\\graph\\afterRandom\\graphml2\\");
 
+
+
+        //rome
+
       Experiments e = new Experiments(inputDir, outputDir,
               Long.MAX_VALUE, 100, 1000, 10000,
-              false, false, true, true);
-       e.runOnlyRandom();
+              false, false, false, true);
+    //   e.runOnlyRandom();
 
        e = new Experiments(inputDir, outputDir,
                Long.MAX_VALUE, 100, 1000, 10000,
-               false, false, true, true);
+               false, false, false, true);
+      //  e.runOnlyForce();
+
+        e = new Experiments(inputDir, outputDir,
+                Long.MAX_VALUE, 100, 1000, 10000,
+                false, false, false, true);
+        //e.runOnlyTotalResolutionForce();
+
+
+
+        //north
+
+
+         inputDir = "E:\\graph\\north\\north\\north\\";
+         outputDir = "E:\\graph\\results\\north_Crossing_01062018\\";
+
+        e = new Experiments(inputDir, outputDir,
+                Long.MAX_VALUE, 100, 1000, 10000,
+                false, false, true, false);
+        e.runOnlyRandom();
+
+        e = new Experiments(inputDir, outputDir,
+                Long.MAX_VALUE, 100, 1000, 10000,
+                false, false, true, false);
+        e.runOnlyForce();
+
+        e = new Experiments(inputDir, outputDir,
+                Long.MAX_VALUE, 100, 1000, 10000,
+                false, false, true, false);
+        e.runOnlyTotalResolutionForce();
+
+        inputDir = "E:\\graph\\north\\north\\north\\";
+        outputDir = "E:\\graph\\results\\north_Angular_01062018\\";
+
+        e = new Experiments(inputDir, outputDir,
+                Long.MAX_VALUE, 100, 1000, 10000,
+                false, false, false, true);
+        e.runOnlyRandom();
+
+        e = new Experiments(inputDir, outputDir,
[…]

### 2018-06-04 src/experiment/Experiments.java
added aspect ratio to experiments

@@ -52,6 +52,7 @@ public class Experiments {
     private String[] childernP;
     private boolean useCrossingRes = true;
     private boolean useAngularRes = false;
+    private boolean useAspectRaio = true;
 
 
 
@@ -232,6 +233,7 @@ public class Experiments {
         experiment.setMaxTime(this.maxCalcTime);
         experiment.getTab().get().setEnableAngularResolution(this.useAngularRes);
         experiment.getTab().get().setEnableCrossingResolution(this.useCrossingRes);
+        experiment.getTab().get().setEnableAspectRazio(this.useAspectRaio);
 
         YGraphAdapter graphAdapter = new YGraphAdapter(frame.graph);
 
@@ -250,24 +252,27 @@ public class Experiments {
 
             if(algorithmName.equals(this.randomMovementString)){
                 experiment.setAlgorithm(RandomMovementLayout.class);
+                for(int i = 0; i < experiment.getTab().get().configurator.getItems().size(); i++){
+                    System.out.println(i + "   " + experiment.getTab().get().configurator.getItems().get(i).getName());
+                }
                                     /* Set config. parameters */
                 experiment.getTab().get().configurator.getItems().get(0).setValue(10.0); // Minimum step size
                 experiment.getTab().get().configurator.getItems().get(1).setValue(150.0); // Maximum step size
                 experiment.getTab().get().configurator.getItems().get(2).setValue(50); // Failed iteration ... detect loacal maximum
                 experiment.getTab().get().configurator.getItems().get(3).setValue(50); // numbers .. maximum resolving
-                experiment.getTab().get().configurator.getItems().get(4).setValue(false); // Allow decreasing minimum angle at local maximum
-                experiment.getTab().get().configurator.getItems().get(5).setValue(false);  // Only use grid coord.
-                experiment.getTab().get().configurator.getItems().get(6).setValue(true); // focus on critical nodes
-                experiment.getTab().get().configurator.getItems().get(7).setValue(true); // Automatically toggle focusing in critical nodes
-                experiment.getTab().get().configurator.getItems().get(8).setValue(false); // auto increase step size
-                experiment.getTab().get().configurator.getItems().get(9).setValue(useCrossingRes); // use crossing Resolution
-                experiment.getTab().get().configurator.getItems().get(10).setValue(useAngularRes); // use angular Resolution
[…]

### 2018-06-04 src/experiment/MainExperiment.java

@@ -17,25 +17,65 @@ public class MainExperiment {
 
         //rome
 
+     inputDir = "E:\\graph\\rome\\rome\\";
+     outputDir = "E:\\graph\\results\\rome_Crossing_aspectratio_of_5_04062018\\";
+
       Experiments e = new Experiments(inputDir, outputDir,
               Long.MAX_VALUE, 100, 1000, 10000,
               false, false, false, true);
-    //   e.runOnlyRandom();
+       e.runOnlyRandom();
 
        e = new Experiments(inputDir, outputDir,
                Long.MAX_VALUE, 100, 1000, 10000,
                false, false, false, true);
-      //  e.runOnlyForce();
+        //e.runOnlyForce();
 
         e = new Experiments(inputDir, outputDir,
                 Long.MAX_VALUE, 100, 1000, 10000,
                 false, false, false, true);
         //e.runOnlyTotalResolutionForce();
 
+     inputDir = "E:\\graph\\rome\\rome\\";
+     outputDir = "E:\\graph\\results\\rome_Angular_aspectratio_of_5_04062018\\";
+
+    e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
+        e.runOnlyRandom();
+
+     e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
+       //e.runOnlyForce();
+
+     e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
+     //e.runOnlyTotalResolutionForce();
 
 
-        //north
+     inputDir = "E:\\graph\\rome\\rome\\";
+     outputDir = "E:\\graph\\results\\rome_Total_aspectratio_of_5_04062018\\";
 
+     e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
+        e.runOnlyRandom();
+
+     e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
+     //  e.runOnlyForce();
+
+     e = new Experiments(inputDir, outputDir,
+             Long.MAX_VALUE, 100, 1000, 10000,
+             false, false, false, true);
[…]

### 2018-06-04 src/sidepanel/SidePanelTab.java

@@ -387,6 +387,12 @@ public class SidePanelTab {
         return this.enableAngularResolution.isSelected();
     }
 
+    public void setEnableAspectRazio(boolean value) {
+        this.enableAspectRatio.setSelected(value);
+    }
+
+    public boolean getEnableAspectRatio(){return this.enableAspectRatio.isSelected();}
+
     public void setAllowClickCreateNodeEdge(boolean value) {
         this.allowClickCreateNodeEdge.setSelected(value);
     }

### 2018-06-04 paper/paper.tex
Working a bit on the description

@@
 
-In this section, we describe our heuristic algorithm for obtaining drawings with high crossing resolution.
-We assume that the input of our algorithm consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ that is of some crossing resolution $c(\Gamma_0)$.
-We also assume that no two edges of $G$ overlap in $\Gamma_0$, that is, $c(\Gamma_0)>0$.
-A circular drawing of $G$ or a drawing obtained by applying a standard force-directed algorithm on $G$ clearly meets these preconditions.
+In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
+The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
+We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
+A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
 
-Our algorithm is iterative and at each iteration step performs some operations that are mainly based on randomization.
-More precisely, at the $i$-th iteration step, we assume that we have computed a drawing $\Gamma_{i-1}$ of some crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant property of our algorithm, that the crossing resolution cannot be decreased at some iteration step.
+Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
+At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
+%, where $\Gamma_0$ is our initial drawing. 
+In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
 Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
@@
 To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges, say $e$ and $e'$, is called \emph{critical} in drawing $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$.
-The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
+A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
+The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
 
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}.
-The reason is that by slightly changing the location of a critical vertex or the location of one of its neighbours appropriately, one naturally expects to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a weighted random selection procedure, in which the vertices which are at distance $i$ from the vertices of the vertex-pool in the graph have higher weights than the corresponding vertices at distance $j$, when $ \leq <i<j$.
-Hence, if the vertex-pool contains critical vertices, then the closer a vertex is to a critical vertex, the more likely it is to be chosen.
+The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a weighted random selection procedure, in which the vertices at distance~$i$ from the ones of the vertex-pool in the graph have higher weights than the corresponding ones at distance $j$, when $0 \leq i<j$.
+So, if the vertex-pool contains critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
 Otherwise, each vertex of the graph can be chosen with the same probability.
 
-With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices of the drawing, it is highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-Hence, special care is needed in order to avoid these bottlenecks, especially when the graph to be drawn is large.
-We discuss heuristics to avoid such bottlenecks in  Section~\ref{ssec:maxima}.
+%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+However, by focusing exclusively at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
+So, special care is needed to avoid these bottlenecks, especially when the graph to be drawn is large.
+We discuss ways to avoid them in  Section~\ref{ssec:maxima}.
 

### 2018-06-05 src/experiment/Experiments.java
fixed maxNode bug for the upper bound of the resultsfile

@@ -387,7 +387,7 @@ public class Experiments {
         }
 
     private void createFiles(String algoName, int minNodeNum, int maxNodeNum){
-        String filename = algoName+ minNodeNum + "_to_" + maxNodeNum + ".csv";
+        String filename = algoName + minNodeNum + "_to_" + maxNodeNum + ".csv";
         String filenameBest = algoName + "best_" + minNodeNum + "_to_" + maxNodeNum + ".csv";
 
         createFile(filename);
@@ -451,7 +451,7 @@ public class Experiments {
         int nodeNum = g.getNodes().size();
         int maxNodeNum = 0;
         for(maxNodeNum = 0; maxNodeNum <= 100; maxNodeNum+=20){
-            if(maxNodeNum > nodeNum){
+            if(maxNodeNum >= nodeNum){
                 break;
             }
         }

### 2018-06-05 src/main/MinimumAngleMonitor.java

@@ -164,8 +164,6 @@ public class MinimumAngleMonitor {
     computeCrossingResolution();
     computeAngularResolution();
 
-    System.out.println("COMPUTE Total Resolution:  cross "  + this.currentCrossingResolution);
-    System.out.println("COMPUTE Total Resolution:   andular  "  +  this.currentAngularResolution );
     if (useAngularResolution && useCrossingResolution) {
       if (currentAngularResolution < currentCrossingResolution) {
         this.currentTotalResolution = currentAngularResolution;
@@ -178,15 +176,10 @@ public class MinimumAngleMonitor {
       this.currentTotalResolution = currentCrossingResolution;
     }
 
-    System.out.println("COMPUTE Total Resolution:   total   "  +  this.currentTotalResolution);
-
     if(bestTotalResolution <= this.currentTotalResolution){
       this.bestTotalResolution = this.currentTotalResolution;
     }
 
-    System.out.println("COMPUTE Total Resolution:   old crossing"  + this.bestCrossingResolution);
-    System.out.println("COMPUTE Total Resolution:   old angular"  +  this.bestAngularResolution);
-    System.out.println("COMPUTE Total Resolution:   old total"  + this.bestTotalResolution);
 
     return this.currentTotalResolution;
   }

### 2018-06-05 paper/paper.tex
Work on the paper; basically shortening to make space for the evaluation.

@@
 \graphicspath{{figures/}}
+
+\newcommand{\myparagraph}[1]{\medskip\noindent\textbf{#1}.}
 
@@
 Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-
-\begin{itemize}
-\item either all the vertices of drawing $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of drawing $\Gamma_{i-1}$, which we call \emph{critical}.
-\end{itemize}
+%
+%\begin{itemize}
+\begin{inparaenum}[(i)]
+\item either all vertices of $\Gamma_{i-1}$, or
+\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
+\end{inparaenum}
 
@@
 
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations in order to converge to a good solution, because it is simply very unlike to select one of them to move.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a weighted random selection procedure, in which the vertices at distance~$i$ from the ones of the vertex-pool in the graph have higher weights than the corresponding ones at distance $j$, when $0 \leq i<j$.
+The role of critical vertices is central in our algorithm
+%\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations to converge to a good solution, because it is very unlike to select one of them to move.}
+\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a weighted random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool have higher weights than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
 So, if the vertex-pool contains critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, each vertex of the graph can be chosen with the same probability.
+Otherwise, each vertex can be chosen with the same probability.
 
@@
 What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing exclusively at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the graph to be drawn is large.
-We discuss ways to avoid them in  Section~\ref{ssec:maxima}.
+However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
+So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
+We discuss ways to avoid them later in this section.
 
-So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-We have not described, however, how the chosen vertex is moved, i.e., how we compute its new position in the next drawing.
+So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
+Next, we  described how to compute its new position in the next drawing.
 
-Let $v_i$ be the vertex of the drawing $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration of our algorithm.
-Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of vertex $v_i$ in drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from the point $p(v_i)$ of vertex $v_i$ in drawing $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; refer to Fig.~\ref{fig:algo} for an illustration.
-The position of vertex $v_i$ in drawing $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
+Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration. 
+%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
+To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
+These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
+The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
 More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$, and there is no vertex of drawing $\Gamma_{i-1}$ at point $\pi_j$, i.e., we do not introduce any vertex-vertex overlap.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
 
@@
 	\subfloat[\label{fig:algo-rays} {}]{	
-	\includegraphics[page=1]{figures/algorithm}}
+	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
 	\hfil
 	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2]{figures/algorithm}}
+	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
 	\caption{%
@@
 
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the new position of vertex $v_i$ in drawing $\Gamma_i$ is $p(v_i)$, that is, the same as in drawing $\Gamma_{i-1}$, since the crossing resolution of drawing $\Gamma_i$ must be at least as large as the one of $\Gamma_{i-1}$.
-On the other hand, if there is one or more feasible points (among $\pi_0,\pi_1,\ldots,\pi_{\rho-1}$), then one may consider two different approaches to determine the new position of vertex $v_i$ in drawing $\Gamma_i$.
+If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
+If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
 The most natural is to chose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the new position of vertex $v_i$ in drawing $\Gamma_i$.
-In our implementation, we did not observe any significant difference in these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-In Section~\ref{ssec:termination}, we describe conditions under which we determine that our algorithm has converged.
+As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
+We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
 
-\subsection{Avoiding local maxima.}
-\label{ssec:maxima}
+The termination condition of our algorithm is simple and depends on an input parameter~$\tau$.
+More specifically, if the crossing resolution has not been improved during the last $\tau$ iterations, then we assume that the algorithm has converged and we do not proceed any more.
 
-To avoid getting trapped to locally optimal solutions with the aforementioned procedure, we mainly investigated two different approaches.
-Both of them are parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first one mimics the human behaviour.
-What would one do to get away from a locally optimal solution?
-First, she would stop trying to move the endvertices of the edges defining the crossing resolution; she would start moving ``irrelevant'' vertices hoping that by changing the embedding of the graph or its crossing structure, a better solution would be easier to be computed again.
-Our algorithm is mimicking this idea by allowing the vertex-pool to also contain all the vertices of the graph, as follows:
+\myparagraph{Avoiding local maxima}
+%
+To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
+The first mimics the human behaviour.
+What would~one do to escape from a locally optimal solution?
+She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
+%by changing the embedding of the graph or its crossing structure, 
+by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows: 
+%
+\begin{inparaenum}[(i)]
+%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices. 
+\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} containing all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
+\item afterwards, the vertex-pool switches back to the critical vertices.
+\end{inparaenum}
+%
+%From our extensive practical analysis, we quickly realized that, while 
+While this approach turned out to be quite effective for medium-size graphs, for larger graphs, unfortunately, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
 
-\begin{itemize}
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices. 
-\item If during the last $\zeta$ iterations the crossing resolution has not been improved, then the vertex-pool becomes \emph{wider} containing all the vertices of the graph, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations.
-\item After these $\zeta'$ iterations, the vertex-pool switches back to the critical vertices.
-\end{itemize}
+%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
+
+Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
+Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastical'' or ``sharp'' move may help to escape.
+We turned this idea into an algorithmic implementation as follows:
 %
-From our extensive practical analysis, we quickly realized that, while this approach turned out to be quite effective for medium-size graphs, for larger and denser graphs, unfortunately, it was not so efficient.
-What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
+\begin{inparaenum}[(i)]
+\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
+\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
+\end{inparaenum}
+%
+Of course, this approach may lead to drawings with larger area, but this is expected, as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
 
-The second approach that we investigated to avoid locally optimal solutions was based on the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastical'' or ``sharp'' move may help to get untrapped.
-We turned this idea into an algorithmic step as follows:
+\myparagraph{Complexity related issues}
+%
+A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing).
+Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
 
-\begin{itemize}
-\item If during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these parameters for $\zeta'$ iterations.
-\item After these $\zeta'$ iterations, the parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to their initial (input) values.
-\end{itemize}
-%
-Of course, this approach may lead to drawings with larger area, but this is somehow expected, as it turns out that drawings with high crossing resolution may require large drawing area; see, e.g.,~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-\subsection{Conditions for termination.}
-\label{ssec:termination}
-
-The termination condition of our algorithm is simple and depends also on an input parameter $\tau$, such that $\tau \geq \zeta + \zeta'$.
-More precisely, if the crossing resolution during the last $\tau$ iterations has not been improved, then we assume that the algorithm has converged and we do not proceed any more.
-
-\subsection{Complexity Related Issues.}
-\label{ssec:complexity}
-
-A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points among the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing).
-Given  a drawing, a naive approach to compute its crossing points requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; refer, e.g., to~\cite{DBLP:books/lib/BergCKO08}.
-
-However, if the algorithm had to compute all crossing points and the corresponding angles at these points for each candidate position at each iteration step, then our algorithm would not be useful in practice.
+However, if the algorithm had to compute all crossing points and the corresponding angles at these points for each candidate position of each iteration, then it would not be useful in practice.
 Instead, we adopted a different approach, which turned out to be quite efficient in practice.
@@
 Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in drawing $\Gamma_{i-1}$.
+Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
 Let $\phi_i$ be the minimum crossing angle computed with this procedure; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that vertex $v_i$ is at position $\pi_j$.
+Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
 Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, in our context, position $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
+Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
 Note that the complexity of this approach is $O(deg(v_i) \cdot m)$, which in worst-case is $O(nm)$.
 
-\subsection{Grid drawings: An interesting variant.}
-\label{ssec:grid}
+\subsection{Some interesting variants}
 
-The algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
+In the remainder of this section, we discuss some interesting variants of our~algorithm, which are motivated by the following preliminary observations that we made from the computed layouts: Drawings with good crossing resolution tend to have bad aspect ratio (that is, the ratio of the longest to the shortest edge is large) and poor angular resolution (that is, the angles formed by adjacent edges are small).
+The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
+The later can easily become clear by a simple example; if in a drawing all edges are either almost horizontally or almost vertically drawn, then while its crossing resolution is $\epsilon$-close to $90^\circ$, its angular resolution is $\epsilon$-close to~$0^\circ$.
+
+\myparagraph{Aspect ratio}
+%
+Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge.
+Sometimes it is also used as a measure of the area of non-grid drawings.
+It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
+What we simply had to do was to reject candidate positions, which violate this precondition.
+
+\myparagraph{Total resolution} 
+%
+The notion of the total resolution of a drawing was relatively recently introduced with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
+Formally, it is defined as the minimum of these two measures.
+It was not difficult to adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
+In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
+Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
+We do so by simply rejecting candidate positions, which yield a reduced total resolution.
+
+\myparagraph{Angular resolution} 
+%  
+As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
+We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
+
+\myparagraph{Grid drawings}
+%
+Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
 However, it can be easily adjusted to produce such drawings.
-Recall that, at the $i$-th iteration step, our algorithm chooses at random a vertex $v_i$ (based on the content of the vertex-pool), and determines a set of $\rho$ candidate points $\pi_0,\ldots,\pi_{\rho-1}$.
-If we round these points to their closest grid points, say $\overline{\pi}_0,\ldots,\overline{\pi}_{\rho-1}$, and use them as candidates for the next position of $v_i$, then the obtained drawing will be grid, assuming that the starting drawing is grid (the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
+More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
+One can even bound the size of the grid, which simply implies that candidate grid positions outside the bounds must be rejected.
 
-An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
+ %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
+
+%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
 Critical here is not to affect the crossing resolution too much during the conversion.

### 2018-06-05 paper/paper.tex
improving a bit the presentation

@@
 The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-The later can easily become clear by a simple example; if in a drawing all edges are either almost horizontally or almost vertically drawn, then while its crossing resolution is $\epsilon$-close to $90^\circ$, its angular resolution is $\epsilon$-close to~$0^\circ$.
+The later can easily become clear by an example; if in a drawing all edges are either almost horizontally or almost vertically drawn, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
 
@@
 %
-The notion of the total resolution of a drawing was relatively recently introduced with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
+The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
 Formally, it is defined as the minimum of these two measures.

### 2018-06-07 src/experiment/MainExperiment.java
save nodes over 100

@@ -11,7 +11,7 @@ public class MainExperiment {
     //static String inputRomeDir = "E:\\graph\\rome\\smallSet\\"; // testing
    //  static String inputNorthDir = "E:\\graph\\rome\\smallSet\\"; // testing
 
-    static String outputDir = "E:\\graph\\results\\05_06_2018\\";
+    static String outputDir = "E:\\graph\\results\\06_06_2018_with_planar_40-60\\";
 
     static  Experiments e;
 
@@ -20,7 +20,7 @@ public class MainExperiment {
     static int numOfIterationPerStep = 100;
     static int numOfSteps = 1000;
     static int boxSize = 10000;
-    static boolean planarGraphsAllowed = false;
+    static boolean planarGraphsAllowed = true;
     static boolean unconnectedGraphsAllowed = false;
     static boolean useCrossingRes;  // if crossing and angular is enabled, the experiment focus on total resolution
     static boolean useAngularRes;
@@ -29,10 +29,10 @@ public class MainExperiment {
     /** run booleans  **/
     static boolean runRome = true;
     static boolean runNorth = false;
-    static boolean runRandomMovement = false;
+    static boolean runRandomMovement = true;
     static boolean runRandomMovementWithAspectRatio = true;
-    static boolean runForce = false;
-    static boolean runTotalForce = false;
+    static boolean runForce = true;
+    static boolean runTotalForce = true;
 
 
 

### 2018-06-07 paper/paper.tex
save nodes over 100
Starting a bit on the experiments.

@@
 
-% ============================================================
+% ==================================================================
 \begin{document}
@@
 
+% ==================================================================
 \section{Introduction}
 \label{sec:introduction}
+% ==================================================================
 
@@
 
+% ==================================================================
 \section{Related Work}
 \label{sec:relatedwork}
+% ==================================================================
 
@@
 
+% ==================================================================
 \section{Description of our Heuristic Approach}
 \label{sec:algorithm}
+% ==================================================================
 
@@
 
+% ==================================================================
 \myparagraph{Avoiding local maxima}
+% ==================================================================
 %
@@
 
+% ==================================================================
 \myparagraph{Complexity related issues}
+% ==================================================================
 %
@@
 
+% ==================================================================
 \subsection{Some interesting variants}
+\label{ssec:variants}
+% ==================================================================
 
@@
 
+% ==================================================================
 \myparagraph{Aspect ratio}
+% ==================================================================
 %
@@
 
+% ==================================================================
 \myparagraph{Total resolution} 
+% ==================================================================
 %
@@
 
+% ==================================================================
 \myparagraph{Angular resolution} 
+% ==================================================================
 %  
@@
 
+% ==================================================================
 \myparagraph{Grid drawings}
+% ==================================================================
 %
@@
 
+% ==================================================================
 \section{Experimental Evaluation}
 \label{sec:experiments}
+% ==================================================================
+
+In this section, we present the results of our experimental evaluation.
+For comparison purposes, apart from our algorithm, we have also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+The implementations were in Java using the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
+The experiment was performed on a Linux laptop with four cores at 2.5 GHz and 3 GB RAM.
+%
+As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs (commonly used for testing the efficiency of algorithms for drawing graphs).
+
+The experiment was performed as follows.
+Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
+Starting from this layout, every graph was drawn with
+\begin{inparaenum}[(i)]
+\item our algorithm, 
+\item our algorithm adjusted not to violate the aspect ratio of the initial layout, 
+\item the force-directed algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+\item the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+\end{inparaenum}
+%
+We compared the quality of the produced drawings based on the following aesthetic properties:
+%
+\begin{enumerate}[P.1.]
+\item \label{p:cr-res} crossing resolution, 
+\item \label{p:to-res} total resolution, 
+\item \label{p:an-res} angular resolution, 
+\item \label{p:as-rat} aspect ratio, and
+\item \label{p:no-xing} number of crossings.
+\end{enumerate}
+
+The comparisons were on average across dirrerent drawings with same number of vertices.
+For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize the corresponding measures.
+Then, for each of them we reported the average of the corresponding aspect ratios and number of crossings of the produced layouts.
+Note that all algorithms of the experiment can easily be adjusted to maximize only the crossing resolution, or only the angular resolution or both (by maximizing the total resolution).
+In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by appropriately selecting the set of forces to use (as each of them, has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution)\todo{up to here}.
+
+. If both the angular and the
+crossing resolution between two consecutive iterations of each
+algorithm were not improved more than 0.001 degrees, we assumed that
+the algorithm has converged and we did not proceed any more. We note
+that the termination condition is quite strict and demands a large
+number of iterations. The maximum number of iterations that an
+algorithm could perform in order to converge was set to 100.000.
+However, very few graphs required the maximum number of iterations
+in order to converge to a layout. Once an algorithm converged, we
+measured its angular, crossing and total resolution, its average
+edge length, the number of crossings, the running time, and the
+required number of iterations to converge.
 
 \begin{itemize}
-\item Experimental setup.
-\item Benchmarks + degree-3.
-\item Aesthetics:
-
-\begin{itemize}
-\item Number of crossings (\# of crossings).
-\item Average size of crossing angles (angle size).
-\item Average of edge lengths (edge length).
-\item Angular resolution (angular res.).
-\end{itemize}
-
 \item Results.
@@
 
+% ==================================================================
 \section{Conclusions}
 \label{sec:conclusions}
+% ==================================================================
 

### 2018-06-08 src/experiment/MainExperiment.java
Experiment: Big angle force are variable with CR, AR, TR

@@ -11,7 +11,7 @@ public class MainExperiment {
     //static String inputRomeDir = "E:\\graph\\rome\\smallSet\\"; // testing
    //  static String inputNorthDir = "E:\\graph\\rome\\smallSet\\"; // testing
 
-    static String outputDir = "E:\\graph\\results\\06_06_2018_with_planar_40-60\\";
+    static String outputDir = "E:\\graph\\results\\07_06_2018\\";
 
     static  Experiments e;
 
@@ -20,7 +20,7 @@ public class MainExperiment {
     static int numOfIterationPerStep = 100;
     static int numOfSteps = 1000;
     static int boxSize = 10000;
-    static boolean planarGraphsAllowed = true;
+    static boolean planarGraphsAllowed = false;
     static boolean unconnectedGraphsAllowed = false;
     static boolean useCrossingRes;  // if crossing and angular is enabled, the experiment focus on total resolution
     static boolean useAngularRes;
@@ -29,10 +29,10 @@ public class MainExperiment {
     /** run booleans  **/
     static boolean runRome = true;
     static boolean runNorth = false;
-    static boolean runRandomMovement = true;
-    static boolean runRandomMovementWithAspectRatio = true;
-    static boolean runForce = true;
-    static boolean runTotalForce = true;
+    static boolean runRandomMovement = false;
+    static boolean runRandomMovementWithAspectRatio = false;
+    static boolean runForce = true; //big angle force
+    static boolean runTotalForce = false;
 
 
 

### 2018-06-08 paper/paper.tex
Working a bit in the experimental evaluation

@@
 \newcommand{\myparagraph}[1]{\medskip\noindent\textbf{#1}.}
+\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
 
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
 \authorrunning{M.~A.~Bekos et al.}
@@
 
-% ==================================================================
-\myparagraph{Angular resolution} 
+% ==================================================================\myparagraph{Angular resolution} 
 % ==================================================================
@@
 The implementations were in Java using the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.5 GHz and 3 GB RAM.
+The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
 %
@@
 
-The comparisons were on average across dirrerent drawings with same number of vertices.
+The comparisons were on average across different drawings with same number of vertices.
 For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize the corresponding measures.
@@
 Note that all algorithms of the experiment can easily be adjusted to maximize only the crossing resolution, or only the angular resolution or both (by maximizing the total resolution).
-In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by appropriately selecting the set of forces to use (as each of them, has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution)\todo{up to here}.
+In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by switching on only the forces that maximize the corresponding measured properties (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
 
-. If both the angular and the
-crossing resolution between two consecutive iterations of each
-algorithm were not improved more than 0.001 degrees, we assumed that
-the algorithm has converged and we did not proceed any more. We note
-that the termination condition is quite strict and demands a large
-number of iterations. The maximum number of iterations that an
-algorithm could perform in order to converge was set to 100.000.
-However, very few graphs required the maximum number of iterations
-in order to converge to a layout. Once an algorithm converged, we
-measured its angular, crossing and total resolution, its average
-edge length, the number of crossings, the running time, and the
-required number of iterations to converge.
+\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
+It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{unrestricted}, in Fig.~\ref{fig:cr-res-1}.
+The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ar-restricted}, in Fig.~\ref{fig:cr-res-1}.
+Finally, the two spring-embedder algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-black and dotted-gray curves of Fig.~\ref{fig:cr-res-1}.
+
+While our unrestricted algorithm produces drawings of better crossing resolution, this seems to come at a cost of increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+There it seems that the forces due to angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
+Regarding our unrestricted algorithm, its behavior is up to a certain degree expected mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
+On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to decrease, has more or less comparable performace (in terms of the aspect ratio of the computed drawings) as the one of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+
+Regarding the number of crossings of the computed drawings, we observe that the restricted variant of our algorithm and the spring embedder algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our expertiment; refer to Fig.~\ref{fig:cr-res-3}.
+
+A different behavior can be observed in the number of iterations, which are required by the algorithms of our experiments, in order to converge; refer to Fig.~\ref{fig:cr-res-4}.
+Note here that we used different criteria to determine whether our algorithm (and its variant) and whether the two spring embedder algorithms have converged, mainly due to the different nature of these algorithms.
+For each of the two spring embedder algorithms, if the crossing resolution between two consecutive iterations was not improved more than 0.001 degrees, we assumed that the algorithm had converged and we did not proceed any more.
+On the other hand, for both of our algorithms we used a more relaxed criterion, because these algorithms may perform several iterations without changing the drawing (and thus, its crossing resolution), as they based on randomization.
+However, if after 100 consecutive iterations there was no improvement, then we assumed that the algorithm had converged and we did not proceed any more.
+The maximum number of iterations that each of the algorithms of our experiment could perform in order to converge was set to 100.000.
+By Fig.~\ref{fig:cr-res-3}, it is immediate to see that both the unrestricted and the restricted variant of our algorithm require less iterations than the two spring-embedder algorithms.
+
+%. If both the angular and the
+%crossing resolution between two consecutive iterations of each
+%algorithm were not improved more than 0.001 degrees, we assumed that
+%the algorithm has converged and we did not proceed any more. We note
+%that the termination condition is quite strict and demands a large
+%number of iterations. 
+%However, very few graphs required the maximum number of iterations
+%in order to converge to a layout. Once an algorithm converged, we
+%measured its angular, crossing and total resolution, its average
+%edge length, the number of crossings, the running time, and the
+%required number of iterations to converge.
 

### 2018-06-08 paper/paper.tex
Experimental evaluation section.

@@
 In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with significantly better crossing resolution, but this comes at the cost of slightly higher running time, especially when the input graph is large and dense.
+Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
 \end{abstract}
@@
 \myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
+Here, each algorithm was adjusted so to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
 It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{unrestricted}, in Fig.~\ref{fig:cr-res-1}.
 The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ar-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two spring-embedder algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-black and dotted-gray curves of Fig.~\ref{fig:cr-res-1}.
+Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-black and dotted-gray curves of Fig.~\ref{fig:cr-res-1}.
 
 While our unrestricted algorithm produces drawings of better crossing resolution, this seems to come at a cost of increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-There it seems that the forces due to angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Regarding our unrestricted algorithm, its behavior is up to a certain degree expected mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
+For the latter algorithm, it seems that the forces due to angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
+Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected mainly due to the fact that there is no control on the lengths of the edges.
+% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths). 
 On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to decrease, has more or less comparable performace (in terms of the aspect ratio of the computed drawings) as the one of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 
-Regarding the number of crossings of the computed drawings, we observe that the restricted variant of our algorithm and the spring embedder algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our expertiment; refer to Fig.~\ref{fig:cr-res-3}.
+Regarding the number of crossings, we observe that the restricted variant of our algorithm and the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our experiment; refer to Fig.~\ref{fig:cr-res-3}.
 
-A different behavior can be observed in the number of iterations, which are required by the algorithms of our experiments, in order to converge; refer to Fig.~\ref{fig:cr-res-4}.
-Note here that we used different criteria to determine whether our algorithm (and its variant) and whether the two spring embedder algorithms have converged, mainly due to the different nature of these algorithms.
-For each of the two spring embedder algorithms, if the crossing resolution between two consecutive iterations was not improved more than 0.001 degrees, we assumed that the algorithm had converged and we did not proceed any more.
+A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
+We note here that we used different criteria to determine whether our algorithm (and its variant) and whether the two force-directed algorithms have converged, mainly due to the different nature of these algorithms.
+For each of the two force-directed algorithms, if the crossing resolution between two consecutive iterations was not improved more than 0.001 degrees, we assumed that the algorithm had converged and we did not proceed any more.
 On the other hand, for both of our algorithms we used a more relaxed criterion, because these algorithms may perform several iterations without changing the drawing (and thus, its crossing resolution), as they based on randomization.
@@
 The maximum number of iterations that each of the algorithms of our experiment could perform in order to converge was set to 100.000.
-By Fig.~\ref{fig:cr-res-3}, it is immediate to see that both the unrestricted and the restricted variant of our algorithm require less iterations than the two spring-embedder algorithms.
+By Fig.~\ref{fig:cr-res-4}, it is immediate to see that both the unrestricted and the restricted variant of our algorithm require less iterations than the two force-directed algorithms.
 
-%. If both the angular and the
-%crossing resolution between two consecutive iterations of each
-%algorithm were not improved more than 0.001 degrees, we assumed that
-%the algorithm has converged and we did not proceed any more. We note
-%that the termination condition is quite strict and demands a large
-%number of iterations. 
-%However, very few graphs required the maximum number of iterations
-%in order to converge to a layout. Once an algorithm converged, we
-%measured its angular, crossing and total resolution, its average
-%edge length, the number of crossings, the running time, and the
-%required number of iterations to converge.
+\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
+Here, each algorithm was adjusted so to maximize the minimum of the crossing and of the angular resolution.
+We observe that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} tend to produce drawings of the same total resolution.
+Slightly worse is the performance of the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Fig.~\ref{fig:to-res-1}.
+However, the aspect ratio of the drawings of the latter algorithm is better than the corresponding aspect ratios of the remaining algorithms of the experiment; see Fig.~\ref{fig:to-res-2}.
+More concritely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
+Then, the ones produced by the force directed algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} follow.
+However, we again observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
 
-\begin{itemize}
-\item Results.
-\item Discussion.
-\end{itemize}
+Similar observations can be made also for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
+The algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} yields drawings with the least number of crossings.
+Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the restricted variant of our algorithm and by the force-directed algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+Our unrestricted algorithm seems to require the largest number of crossings, which turns out to be notably higher than the corresponding ones of the remaining algorithms of our experiment especially for large graphs.
+On the positive side, both the unrestricted and the restricted variant of our algorithm require less iterations than the two force-directed algorithms, as it can be easily seen in Fig.~\ref{fig:to-res-4}.
+
+\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; refer to Fig.~\ref{fig:an-res}.
+Here, each algorithm was adjusted  to maximize exclusively the angular resolution (i.e., by ignoring the drawing's crossing resolution).
+A notable observation is that the force-directed algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} tends to have the same performance (in term of angular resolution) as our unrestricted algorithm, which is slightly better but mostly for small graphs; see Fig.~\ref{fig:an-res-1}.
+The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
+The force-directed algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} seems to be outperformed by all algorithms of the experiment.
+
+On the other hand, the results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
+This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
 

### 2018-06-08 paper/paper.tex
Work on the experiments
Minor changes... Further work tomorrow.

@@
 
-The comparisons were on average across different drawings with same number of vertices.
-For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize the corresponding measures.
+The reported results are on average across different drawings with same number of vertices.
+For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize exclusively the corresponding measures.
 Then, for each of them we reported the average of the corresponding aspect ratios and number of crossings of the produced layouts.
 Note that all algorithms of the experiment can easily be adjusted to maximize only the crossing resolution, or only the angular resolution or both (by maximizing the total resolution).
-In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by switching on only the forces that maximize the corresponding measured properties (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
 
 \myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted so to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
+Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
 It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{unrestricted}, in Fig.~\ref{fig:cr-res-1}.
 The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ar-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-black and dotted-gray curves of Fig.~\ref{fig:cr-res-1}.
+Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, respectively).
 
-While our unrestricted algorithm produces drawings of better crossing resolution, this seems to come at a cost of increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+While our unrestricted algorithm produces drawings of better crossing resolution, this seems to come at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
 For the latter algorithm, it seems that the forces due to angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
@@
 % (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths). 
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to decrease, has more or less comparable performace (in terms of the aspect ratio of the computed drawings) as the one of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performace (in terms of the aspect ratio of the computed drawings) as the one of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 

### 2018-06-09 paper/paper.tex
Finished a first draft of the main part with Henry; Please if you have time proofread it

--- 2018-06-08 paper/paper.tex
@@
 \usepackage{amsmath,amssymb}
-\usepackage{graphicx,rotating}
+\usepackage{graphicx,pgfplots}
 \usepackage[font=small]{subfig,caption}
@@
 \usepackage{todonotes,lineno,paralist,soul}
-\usepackage{placeins,rotating,hyperref}
+\usepackage{rotating,hyperref,multicol}
 \graphicspath{{figures/}}
 
-\newcommand{\myparagraph}[1]{\medskip\noindent\textbf{#1}.}
+% ==================================================================
+% Format tiks and pgfplots
+\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
+\pgfplotsset{ytick style={draw=none}}
+\renewcommand{\axisdefaultheight}{185pt}
+\pgfkeys{/pgf/number format/.cd,1000 sep={}}
+\pgfplotscreateplotcyclelist{mylist}{%
+{black,thick,smooth},
+{black!50!white,thick,smooth},
+{black,thick,densely dashed,smooth},
+{black!50!white,thick,densely dashed,smooth}
+}
+\pgfplotsset{
+        table/search path={data},
+    }
+\pgfplotsset{
+  /pgfplots/xlabel near ticks/.style={
+     /pgfplots/every axis x label/.style={
+        at={(ticklabel cs:0.5)},anchor=near ticklabel
+     }
+  },
+  /pgfplots/ylabel near ticks/.style={
+     /pgfplots/every axis y label/.style={
+        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
+     }
+}
+% ==================================================================
+
+\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
 \author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
@@
 From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.\.
 Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
@@
 %In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper. 
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\ in Section~\ref{sec:experiments}.
 We conclude in Section~\ref{sec:conclusions} with open problems.
@@
 %
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.\ and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 
@@
 %
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are also based on the force-directed technique.
+%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
 
@@
 
-In the remainder of this section, we discuss some interesting variants of our~algorithm, which are motivated by the following preliminary observations that we made from the computed layouts: Drawings with good crossing resolution tend to have bad aspect ratio (that is, the ratio of the longest to the shortest edge is large) and poor angular resolution (that is, the angles formed by adjacent edges are small).
+In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
+Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observations that we made during our experimental evaluation (see Section~\ref{sec:experiments}): Drawings with good crossing resolution tend to have bad aspect ratio
+%(that is, the ratio of the longest to the shortest edge is large) 
+and poor angular resolution.
+%(that is, the angles formed by adjacent edges are small). 
 The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-The later can easily become clear by an example; if in a drawing all edges are either almost horizontally or almost vertically drawn, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
+For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
 
@@
 Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by simply rejecting candidate positions, which yield a reduced total resolution.
+We do so by rejecting candidate positions, which yield a reduced total~resolution.
 
-% ==================================================================\myparagraph{Angular resolution} 
+% ==================================================================
+\myparagraph{Angular resolution} 
 % ==================================================================
@@
 In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we have also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 The implementations were in Java using the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
@@
 \item our algorithm, 
-\item our algorithm adjusted not to violate the aspect ratio of the initial layout, 
-\item the force-directed algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-\item the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
+\item by Argyriou et al.\ and
+\item by Huang et al.
 \end{inparaenum}
 %
-We compared the quality of the produced drawings based on the following aesthetic properties:
+We compared the quality of the produced drawings based on the following aesthetic~properties:
 %
+\begin{multicols}{2}
 \begin{enumerate}[P.1.]
-\item \label{p:cr-res} crossing resolution, 
-\item \label{p:to-res} total resolution, 
-\item \label{p:an-res} angular resolution, 
-\item \label{p:as-rat} aspect ratio, and
-\item \label{p:no-xing} number of crossings.
+\item \label{p:cr-res} crossing resolution 
+\item \label{p:to-res} total resolution 
+\item \label{p:an-res} angular resolution 
+\columnbreak
+\item \label{p:as-rat} aspect ratio
+\item \label{p:no-xing} number of crossings
 \end{enumerate}
-
-The comparisons were on average across different drawings with same number of vertices.
-For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize the corresponding measures.
-Then, for each of them we reported the average of the corresponding aspect ratios and number of crossings of the produced layouts.
-Note that all algorithms of the experiment can easily be adjusted to maximize only the crossing resolution, or only the angular resolution or both (by maximizing the total resolution).
-In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by switching on only the forces that maximize the corresponding measured properties (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+\end{multicols}
+%
+Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), for~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
+In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+The reported results are on average across different drawings with same number of vertices.
 
 \myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ar-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two spring-embedder algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-black and dotted-gray curves of Fig.~\ref{fig:cr-res-1}.
+Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
+It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
+The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
+Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
 
-While our unrestricted algorithm produces drawings of better crossing resolution, this seems to come at a cost of increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-There it seems that the forces due to angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Regarding our unrestricted algorithm, its behavior is up to a certain degree expected mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to decrease, has more or less comparable performace (in terms of the aspect ratio of the computed drawings) as the one of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+\begin{figure}[t]
+\centering
+\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================
+% Crossing Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
+mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
+\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%===================================
+% Crossing Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
 
-Regarding the number of crossings of the computed drawings, we observe that the restricted variant of our algorithm and the spring embedder algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our expertiment; refer to Fig.~\ref{fig:cr-res-3}.
+\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================================
+% Crossing Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%==================================
+% Crossing Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids] 
+\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Illustration of our experimental results on the crossing resolution.}
+\label{fig:cr-res}
+\end{figure}
 
-A different behavior can be observed in the number of iterations, which are required by the algorithms of our experiments, in order to converge; refer to Fig.~\ref{fig:cr-res-4}.
-Note here that we used different criteria to determine whether our algorithm (and its variant) and whether the two spring embedder algorithms have converged, mainly due to the different nature of these algorithms.
-For each of the two spring embedder algorithms, if the crossing resolution between two consecutive iterations was not improved more than 0.001 degrees, we assumed that the algorithm had converged and we did not proceed any more.
-On the other hand, for both of our algorithms we used a more relaxed criterion, because these algorithms may perform several iterations without changing the drawing (and thus, its crossing resolution), as they based on randomization.
-However, if after 100 consecutive iterations there was no improvement, then we assumed that the algorithm had converged and we did not proceed any more.
-The maximum number of iterations that each of the algorithms of our experiment could perform in order to converge was set to 100.000.
-By Fig.~\ref{fig:cr-res-3}, it is immediate to see that both the unrestricted and the restricted variant of our algorithm require less iterations than the two spring-embedder algorithms.
+While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
+For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
+Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
+% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths). 
+On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
 
-%. If both the angular and the
-%crossing resolution between two consecutive iterations of each
-%algorithm were not improved more than 0.001 degrees, we assumed that
-%the algorithm has converged and we did not proceed any more. We note
-%that the termination condition is quite strict and demands a large
-%number of iterations. 
-%However, very few graphs required the maximum number of iterations
-%in order to converge to a layout. Once an algorithm converged, we
-%measured its angular, crossing and total resolution, its average
-%edge length, the number of crossings, the running time, and the
-%required number of iterations to converge.
+Regarding the number of crossings, we observe that the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our experiment; refer to Fig.~\ref{fig:cr-res-3}.
 
-\begin{itemize}
-\item Results.
-\item Discussion.
-\end{itemize}
+A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
+We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
+For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
+For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
+We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
+In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
+Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
+The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
+We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
+However, for bigger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
+On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
+
+\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
+Here, each algorithm was adjusted to maximize the minimum of the crossing and of the angular resolution.
+For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
+The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
+It is worth noting, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs (even though there seems to be a small difference in our favor).
+
+\begin{figure}[t]
+\centering
+\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================
+% Total Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
+mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
+\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
+\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%===================================
+% Total Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
+\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================================
+% Total Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
+\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%==================================
+% Total Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids] 
+\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
+\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
+\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Illustration of our experimental results on the total resolution.}
+\label{fig:to-res}
+\end{figure}
+
+Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
+More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
+Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
+Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
+
+%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}. 
+The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
+Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
+Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the remaining algorithms of our experiment (especially for large graphs).
+
+On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
+Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
+The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
+We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
+
+\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
+Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
+A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
+For large graphs, the two algorithms tend to have the same performance.
+The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
+The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
+
+\begin{figure}[t]
+\centering
+\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================
+% Total Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
+mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
+\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
+\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%===================================
+% Total Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
+\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================================
+% Total Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
+\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%==================================
+% Total Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids] 
+\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
+\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
+\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Illustration of our experimental results on the angular resolution.}
+\label{fig:an-res}
+\end{figure}
+
+The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
+This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
+The small differences result from the fact that the crossing resolution cannot be entirely neglected.
 
@@
 
+In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
+Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
+As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+
 \bibliographystyle{abbrvurl}

### 2018-06-09 src/experiment/Experiment.java
Experiment setup - gridding

@@ -11,6 +11,7 @@ import graphoperations.Scaling;
 import io.ContestIOHandler;
 import layout.algo.execution.ILayout;
 import layout.algo.randommovement.RandomMovementLayout;
+import layout.algo.utils.LayoutUtils;
 import layout.algo.utils.PositionMap;
 import main.MainFrame;
 import sidepanel.SidePanelTab;
@@ -329,7 +330,7 @@ public class Experiment {
 
         mainFrame.minimumAngleMonitor.updateCrossingResolutionInfoBar();
         this.minimumAngle = Math.round(1000.0 * mainFrame.minimumAngleMonitor.getBestCrossingResolution()) / 1000.0;
-        System.out.println("load Graph " + this.minimumAngle);
+//        System.out.println("load Graph " + this.minimumAngle);
         this.aspect_ratio = GraphOperations.getAspectRatio(this.graph).getValue();
     }
 
@@ -356,8 +357,72 @@ public class Experiment {
     }
 
     void isGridded() {
-        System.out.println("GRIDDED: "+GridGraph.isGridGraph(graph));
+        double threshold = 0.001;
+        double edgeThreshold = 0.001;
+        boolean nodeOverlap = false;
+        boolean nodeEdgeOverlap = false;
+        boolean negativeNodes = false;
+        for (INode u : this.mainFrame.graph.getNodes()) {
 
+            if (u.getPorts().size() == 1) {
+                for (IEdge e : graph.edgesAt(u.getPorts().first())) {
+                    INode src = e.getSourceNode();
+                    INode dst = e.getTargetNode();
+                    for (INode v : this.mainFrame.graph.getNodes()) {
+                        if ((u.hashCode() != v.hashCode()) &&
+                                v.hashCode() != src.hashCode() && v.hashCode() != dst.hashCode()) {
+                            double x0 = v.getLayout().getCenter().getX();
+                            double y0 = v.getLayout().getCenter().getY();
+                            double x1 = src.getLayout().getCenter().getX();
+                            double y1 = src.getLayout().getCenter().getY();
+                            double x2 = dst.getLayout().getCenter().getX();
+                            double y2 = dst.getLayout().getCenter().getY();
+                            double dist;
+                            if (y1 == y2) { //vertical
+                                dist = Math.abs(x0 - x1);
+                            } else if (x1 == x2) {  //horizontal
[…]

### 2018-06-09 src/experiment/Experiments.java

@@ -54,7 +54,7 @@ public class Experiments {
     private String[] childernP;
     private boolean useCrossingRes = true;
     private boolean useAngularRes = false;
-    private boolean useAspectRatio = true;
+    private boolean useAspectRatio = false;
     private boolean useOnlyGrid = true;
     private int gridSize = 1000000;
 
@@ -230,25 +230,27 @@ public class Experiments {
 
             //if(maxNodeNum != 40){frame.dispose();return;}//TODO: Nur zum testen
 
-            experiment.runOrganic();
-            experiment.scaleToBox();
+//            experiment.runOrganic();
+//            experiment.scaleToBox();
 
             //First grid the graph
             if(useOnlyGrid) {
-                experiment.setAlgorithm(CombinedGridder.class);
-                /* Set config. parameters */
-                experiment.getTab().get().configurator.getItems().get(0).setValue(1); // Number of parallel executions
-                experiment.getTab().get().configurator.getItems().get(1).setValue(true); // Respect Minimum Angle
-                experiment.getTab().get().configurator.getItems().get(2).setValue(0.1); // Allow decreasing angle by
-                experiment.getTab().get().configurator.getItems().get(3).setValue(false); // Scale up if necessary
-                experiment.getTab().get().configurator.getItems().get(4).setValue(20); // Iterations until scale-up
-                experiment.getTab().get().configurator.getItems().get(5).setValue(2.0); // Scale by
-                experiment.getTab().get().configurator.getItems().get(6).setValue("");  // Status message
-                experiment.getTab().get().configurator.getItems().get(7).setValue(true); // Force grid after stop was pressed
-
-                //TODO: run gridding
-                experiment.runAlgorithms();
+//                experiment.setAlgorithm(CombinedGridder.class);
+//                /* Set config. parameters */
+//                experiment.getTab().get().configurator.getItems().get(0).setValue(1); // Number of parallel executions
+//                experiment.getTab().get().configurator.getItems().get(1).setValue(true); // Respect Minimum Angle
+//                experiment.getTab().get().configurator.getItems().get(2).setValue(0.1); // Allow decreasing angle by
+//                experiment.getTab().get().configurator.getItems().get(3).setValue(false); // Scale up if necessary
+//                experiment.getTab().get().configurator.getItems().get(4).setValue(20); // Iterations until scale-up
+//                experiment.getTab().get().configurator.getItems().get(5).setValue(2.0); // Scale by
[…]

### 2018-06-09 src/experiment/MainExperiment.java

@@ -56,12 +56,12 @@ static String inputRomeDir = "/home/ch/uni/graph_drawing/rome/";
     }
     public static void run(String mode){
         if(runRome){
-//            int[] boxSizes = {100,1000,10000,1000000};
+//            int[] boxSizes = {100,1000,10000, 1000000};
             int[] boxSizes = {1000000};
             for (int i = 0; i < boxSizes.length; i++) {
                 useAspectRatio = false;
 //            e = new Experiments(inputRomeDir, outputDir + "rome\\"+ mode,
-                e = new Experiments(inputRomeDir, outputDir + "rome/" + mode,
+                e = new Experiments(inputRomeDir, outputDir + "rome-"+String.valueOf(boxSizes[i])+"/" + mode,
                         maxCalcTime, numOfIterationPerStep, numOfSteps, boxSizes[i],
                         planarGraphsAllowed, unconnectedGraphsAllowed, useCrossingRes, useAngularRes, useAspectRatio);
                 if (runRandomMovement)

### 2018-06-09 src/graphoperations/GraphOperations.java

@@ -125,15 +125,19 @@ public class GraphOperations {
   }
 
   public static AspectRatio getAspectRatio(IGraph g){
-    IEdge sEdge = getShortestEdge(g);
-    INode sSource = sEdge.getSourceNode();
-    INode sTarget = sEdge.getTargetNode();
-    double sLength = euclidDist(sSource.getLayout().getCenter().getX(), sSource.getLayout().getCenter().getY(), sTarget.getLayout().getCenter().getX(), sTarget.getLayout().getCenter().getY());
-    IEdge lEdge = getLongestEdge(g);
-    INode lSource = lEdge.getSourceNode();
-    INode lTarget = lEdge.getTargetNode();
-    double lLength = euclidDist(lSource.getLayout().getCenter().getX(), lSource.getLayout().getCenter().getY(), lTarget.getLayout().getCenter().getX(), lTarget.getLayout().getCenter().getY());
-    return new AspectRatio(sLength, sEdge, lLength, lEdge);
+    try {
+      IEdge sEdge = getShortestEdge(g);
+      INode sSource = sEdge.getSourceNode();
+      INode sTarget = sEdge.getTargetNode();
+      double sLength = euclidDist(sSource.getLayout().getCenter().getX(), sSource.getLayout().getCenter().getY(), sTarget.getLayout().getCenter().getX(), sTarget.getLayout().getCenter().getY());
+      IEdge lEdge = getLongestEdge(g);
+      INode lSource = lEdge.getSourceNode();
+      INode lTarget = lEdge.getTargetNode();
+      double lLength = euclidDist(lSource.getLayout().getCenter().getX(), lSource.getLayout().getCenter().getY(), lTarget.getLayout().getCenter().getX(), lTarget.getLayout().getCenter().getY());
+      return new AspectRatio(sLength, sEdge, lLength, lEdge);
+    } catch (NullPointerException e) {
+      return new AspectRatio(0,null, 0, null);
+    }
   }
 
   //returns false if moving the node to the position results in a worse aspect ratio

### 2018-06-09 src/layout/algo/gridding/QuickGridder.java

@@ -50,7 +50,10 @@ public class QuickGridder implements IGridder {
       double allowedAngle = oldAngle - configurator.allowDecreasingBy.getValue();
 
       Stream<PointD> samplePositions = getNeighborGridPositions(oldPosition, iteration).stream()
-          .filter(position -> !reservedPositions.contains(position));
+          .filter(position -> !reservedPositions.contains(position))
+          .filter(position -> LayoutUtils.overlap(position, positions, node, graph))    //nodes not on top of each other
+          .filter(position -> LayoutUtils.overlapNodeEdge(position, positions, graph))   //nodes not on top of edges
+          .filter(position -> position.getX() >= 0 && position.getY() >= 0);      //no negative
 
       if (configurator.respectMinimumAngle.getValue()) {
         samplePositions = samplePositions.filter(position -> {
@@ -60,6 +63,7 @@ public class QuickGridder implements IGridder {
       }
 
       List<PointD> goodPositions = samplePositions.collect(Collectors.toList());
+//      System.out.println("good pos " + goodPositions.size());
 
       if (goodPositions.size() == 0) {
         positions.setValue(node, oldPosition);
@@ -97,6 +101,9 @@ public class QuickGridder implements IGridder {
           .filter(node -> !griddedNodes.contains(node))
           .forEach(node -> positions.setValue(node, getNeighborGridPositions(positions.getValue(node), 1).stream()
               .filter(position -> !reservedPositions.contains(position))
+              .filter(position -> LayoutUtils.overlap(position, positions, node, graph))    //nodes not on top of each other
+              .filter(position -> LayoutUtils.overlapNodeEdge(position, positions, graph))   //nodes not on top of edges
+              .filter(position -> position.getX() >= 0 && position.getY() >= 0)      //no negative
               .findFirst()
               .orElseGet(() -> {
                 configurator.statusMessage.setValue("Force gridding failed!\n(node overlaps with another node)");

### 2018-06-09 src/layout/algo/randommovement/RandomMovementLayout.java

@@ -118,11 +118,26 @@ public class RandomMovementLayout implements ILayout {
 
           if (configurator.onlyGridPositions.getValue()) {
             sample.position = LayoutUtils.round(sample.position);
+//            System.out.println(sample.position);
+//            System.out.println(!LayoutUtils.checkNodeEdgeOverlap(graph, positions) + " and " + !LayoutUtils.overlapNodeEdge());
           }
         })
         .filter(sample -> boundingBox.contains(sample.position))
-        .filter(sample -> configurator.onlyGridPositions.getValue() || LayoutUtils.overlap(sample.position, positions, node, graph))
-//        .filter(sample -> !configurator.useAspectRatio.getValue() || GraphOperations.improvedAspectRatio(sample.position, node, graph, currentAspectRatio, configurator.maxAspectRatio.getValue()) ) //filter out non-ok aspect ratio
+        .filter(sample -> !configurator.onlyGridPositions.getValue() || LayoutUtils.overlap(sample.position, positions, node, graph))
+//        .filter(sample -> !configurator.onlyGridPositions.getValue() || !LayoutUtils.checkNodeEdgeOverlap(graph, positions))
+//            .filter(sample -> LayoutUtils.overlap(sample.position, positions, node, graph))
+            .filter(sample -> {
+              Mapper<INode, PointD> test = PositionMap.copy(positions);
+              test.setValue(node, sample.position);
+//              System.out.println(!LayoutUtils.checkNodeEdgeOverlap(graph, test) + " and " + LayoutUtils.overlapNodeEdge(sample.position, test, graph) + " and " + !LayoutUtils.suckit(sample.position, graph, test));
+//              return !LayoutUtils.checkNodeEdgeOverlap(graph, positions);
+//              return LayoutUtils.overlapNodeEdge(sample.position, test, graph) && !LayoutUtils.checkNodeEdgeOverlap(graph, test);// && !LayoutUtils.suckit(sample.position, graph, test);
+//              return !LayoutUtils.suckit(sample.position, graph, positions);
+              return !LayoutUtils.checkNodeEdgeOverlap(graph, test);
+            })
+
+
+        .filter(sample -> !configurator.useAspectRatio.getValue() || GraphOperations.improvedAspectRatio(sample.position, node, graph, currentAspectRatio, configurator.maxAspectRatio.getValue()) ) //filter out non-ok aspect ratio
         .filter(sample -> !configurator.useAspectRatio.getValue() ||
[…]

### 2018-06-09 src/layout/algo/utils/LayoutUtils.java

@@ -2,6 +2,7 @@ package layout.algo.utils;
 
 import java.util.Iterator;
 import java.util.Map.Entry;
+import java.util.Random;
 
 import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.graph.IEdge;
@@ -10,6 +11,10 @@ import com.yworks.yfiles.graph.INode;
 import com.yworks.yfiles.graph.IPort;
 import com.yworks.yfiles.graph.Mapper;
 import com.yworks.yfiles.utils.IListEnumerable;
+import graphoperations.GraphOperations;
+import util.BoundingBox;
+
+import static graphoperations.GraphOperations.euclidDist;
 
 public class LayoutUtils {
   public static PointD stepInDirection(PointD oldPosition, PointD direction, double stepSize) {
@@ -23,28 +28,219 @@ public class LayoutUtils {
     return new PointD(Math.round(p.getX()), Math.round(p.getY()));
   }
 
-public static Boolean overlap(PointD position, Mapper<INode, PointD> positions, INode node, IGraph graph) {
-	Iterator<Entry<INode, PointD>> It = positions.getEntries().iterator();
-	PointD tmp;
-	while(It.hasNext()){
-		tmp=It.next().getValue();
-		if (position.hits(tmp, 0.5) || position.equals(tmp)) {
-			return false;
+	public static Boolean overlap(PointD position, Mapper<INode, PointD> positions, INode node, IGraph graph) {		//TODO: actually returns NOT overlap
+		Iterator<Entry<INode, PointD>> It = positions.getEntries().iterator();
+		PointD tmp;
+		while(It.hasNext()){
+			tmp=It.next().getValue();
+			if (position.hits(tmp, 0.5) || position.equals(tmp)) {
+				return false;
+			}
 		}
-	}
-	IListEnumerable<IPort> ports= node.getPorts();
-	for(IPort port: ports){
-		for(IEdge edge: graph.edgesAt(port)){
-			It = positions.getEntries().iterator();
-			while(It.hasNext()){
-				tmp=It.next().getValue();
-				if(tmp!=positions.getValue(edge.getSourceNode()) && tmp!=positions.getValue(edge.getTargetNode())
-						&& tmp.hitsLineSegment(positions.getValue(edge.getSourceNode()), positions.getValue(edge.getTargetNode()), 0.5)){
-					return false;
+		IListEnumerable<IPort> ports= node.getPorts();
[…]

### 2018-06-09 src/main/MainFrame.java

@@ -36,7 +36,7 @@ public class MainFrame extends JFrame {
 
     /* Box related issue*/
 
-    public static double BOX_SIZE[] = {1000000, 1000000};
+    public static int BOX_SIZE[] = {100, 100};
 
     /* Graph Drawing related objects */
     public GraphComponent view;

### 2018-06-09 src/sidepanel/SidePanelTab.java

@@ -23,6 +23,7 @@ import layout.algo.NodeSwapper;
 import layout.algo.execution.BasicIGraphLayoutExecutor;
 import layout.algo.execution.ILayout;
 import layout.algo.layoutinterface.ILayoutConfigurator;
+import layout.algo.utils.LayoutUtils;
 import layout.algo.utils.PositionMap;
 import main.MainFrame;
 import util.BoundingBox;
@@ -342,6 +343,13 @@ public class SidePanelTab {
         swapperItem.addActionListener(this::swapperItemActionPerformed);
         custom.add(swapperItem, cCustomPanel);
 
+        JButton randomLayoutItem = new JButton("Random Layout");
+        cCustomPanel.fill = GridBagConstraints.HORIZONTAL;
+        cCustomPanel.gridx = 1;
+        cCustomPanel.gridy = cCustomPanelY;
+        randomLayoutItem.addActionListener(this::randomLayoutItemActionPerformed);
+        custom.add(randomLayoutItem, cCustomPanel);
+
         cSidePanel.fill = GridBagConstraints.HORIZONTAL;
         cSidePanel.gridx = 0;
         cSidePanel.gridy = 0;
@@ -524,11 +532,11 @@ public class SidePanelTab {
         modifyGraph(() -> {
             initSidePanel.removeDefaultListeners();
 
-            IGraph graph = initSidePanel.mainFrame.graph;
-            Mapper<INode, PointD> nodePositions = PositionMap.FromIGraph(graph);
+//            IGraph graph = initSidePanel.mainFrame.graph;
+            Mapper<INode, PointD> nodePositions = PositionMap.FromIGraph(initSidePanel.mainFrame.graph);
             RectD bounds = BoundingBox.from(nodePositions);
             Scaling.scaleBy(Math.min((MainFrame.BOX_SIZE[0] / bounds.getWidth()), (MainFrame.BOX_SIZE[1] / bounds.getHeight())), nodePositions);
-            PositionMap.applyToGraph(graph, nodePositions);
+            PositionMap.applyToGraph(initSidePanel.mainFrame.graph, nodePositions);
             initSidePanel.mainFrame.view.fitGraphBounds();
 
             bounds = BoundingBox.from(nodePositions);
@@ -666,8 +674,8 @@ public class SidePanelTab {
 
 //--------------------------------------------------------------------
     }
-    
[…]

### 2018-06-09 paper/paper.tex
wrote an appendix for the grid drawing experiment.

@@
 {black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth}
+{black!50!white,thick,densely dashed,smooth},
+{black,double,smooth}
 }
@@
 
+\newpage
+\appendix
+
+\section{Experiments on Grid Drawings}
+
+In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to be placed on grid coordinates whilst avoiding vertices to overlap with other vertices or edges.
+Further, we restricted the grid size to \begin{inparaenum}[(i)]
+\item $1,000,000 \times 1,000,000$, denoted by $1M \times 1M$ for short,
+\item $10,000 \times 10,000$, denoted by $10k \times 10k$ for short,
+\item $1,000 \times 1,000$, denoted by $1k \times 1k$ for short, and
+\item $100 \times 100$.
+
+\end{inparaenum}
+While the test set for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, we computed a different initial since our algorithm requires its input drawing to maintain its invariants (that is, vertices must be on the grid).
+More precisely, we computed a random grid layout of each graph where vertices to overlap with other vertices or edges.
+For each grid size, we computed a layout with the variant of our algorithm focusing on crossing resolution.
+Fig.~\ref{fig:experimentsGrid} summarizes the results.
+
+\begin{figure}[!b]
+\centering
+\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================
+% Crossing Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[
+axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
+legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
+width=0.475\textwidth,
+cycle list name=mylist, 
+mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
+xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={10,20,30,40,50,60,70,80,90},
+tick pos=left,ymajorgrids]
+\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
+\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%===================================
+% Crossing Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
+xlabel={Number of Vertices},
+xmin=10,
+xmax=100,
+ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
+\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================================
+% Crossing Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
+xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
+\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%==================================
+% Crossing Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left, 
+ylabel={Iterations},
+xlabel={Number of Vertices},ymajorgrids]
+\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
+\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Illustration of our experimental results on the crossing resolution with grid restriction. The double line shows the results of our unrestricted algorithm from Section~\ref{sec:experiments}.}
+\label{fig:experimentsGrid}
+\end{figure}
+
+Regarding the crossing resulotion, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+More precisely, the grid of size $100 \times 100$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
+Starting from grid size $1,000 \times 1,000$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
+For grid size $1,000 \times 1,000$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
+For grid size $10,000 \times 10,000$, the algorithm performed only about $10^\circ$ worse than the unrestricted version of our algorithm which could only slightly be improved by increasing the grid size to $1,000,000 \times 1,000,000$.
+
+The aspect ratio of computed drawings was more or less the same for all grid size with the exception of size $100 \times 100$ and always about twice the aspect ratio of the unrestricted variant of our algorithm; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+For grid size $100 \times 100$, this may again be explained by the fact, that for the majority of graphs, the initial layout could not be improved.
+
+The number of crossings increased with the restriction on the size of the grid; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
+As with the crossing resolution, there is clear differences between grid sizes $100 \times 100$, $1,000 \times 1,000$ and $10,000 \times 10,000$ whereas there is only a slight improvement from grid size $10,000 \times 10,000$ to $1,000,000 \times 1,000,000$ which still uses twice as many crossings as our unrestricted variant.
+This can partially be explained with the different choice of the initial drawing.
+
+Finally, we observe that the number of iterations needed for convergence for grid size $1,000 \times 1,000$ was already twice as much as for our unrestricted algorithm; see Fig.~\ref{fig:crossingResolutionTimeGrid}.
+From there on, the number of iterations for convergence increases with the grid size.
+Note that grid size $1,000,000 \times 1,000,000$ needed about 1000 to 2000 iterations more than grid size $10,000 \times 10,000$ even though the final layout was only marginally better as discussed prior.
+The curve for $100 \times 100$ highlights that this grid size was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
+
+In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid as long as the grid is not too small.
+However, the computation of grid drawings takes much longer than in the unrestricted version (up to five times as long for grid size $10,000 \times 10,000$ or eight times as long for grid size $1,000,000 \times 1,000,000$).
+In order to improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
+Further experiments in this direction are needed.
+
 \end{document}

### 2018-06-09 paper/paper.tex
Addressing a comment by Jan

@@
 From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.\.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
+The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.\ Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
 Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.

### 2018-06-09 paper/paper.tex
inserted north graph evaluation.

@@
 
+\section{Experiments on the North Graph Testset}
+
+\todo[inline]{Say something about north graph}
+
+\begin{figure}[htb]
+\centering
+\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================
+% Crossing Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[
+axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
+legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
+width=0.475\textwidth,
+cycle list name=mylist, 
+mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
+xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={10,20,30,40,50,60,70,80,90},
+tick pos=left,ymajorgrids]
+\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%===================================
+% Crossing Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
+xlabel={Number of Vertices},
+xmin=10,
+xmax=100,
+ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
+\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================================
+% Crossing Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
+xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
+\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%==================================
+% Crossing Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left, 
+ylabel={Iterations},
+xlabel={Number of Vertices},ymajorgrids]
+\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
+\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
+\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Experimental results for the crossing resolution experiment on the North graph testset.}
+\label{northCrossing}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================
+% Total Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[
+axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
+legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
+width=0.475\textwidth,
+cycle list name=mylist, 
+mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
+xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={10,20,30,40,50,60,70,80,90},
+tick pos=left,ymajorgrids]
+\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
+\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%===================================
+% Total Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
+xlabel={Number of Vertices},
+xmin=10,
+xmax=100,
+ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
+\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================================
+% Total Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
+xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
+\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
+\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%==================================
+% Total Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left, 
+ylabel={Iterations},
+xlabel={Number of Vertices},ymajorgrids]
+\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
+\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
+\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Experimental results for the total resolution experiment on the North graph testset.}
+\label{northTotal}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================
+% Angular Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[
+axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
+legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
+width=0.475\textwidth,
+cycle list name=mylist, 
+mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
+xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={10,20,30,40,50,60,70,80,90},
+tick pos=left,ymajorgrids]
+\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
+\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%===================================
+% Angular Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
+xlabel={Number of Vertices},
+xmin=10,
+xmax=100,
+ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
+\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%======================================
+% Angular Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
+xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
+\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
+\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
+\centering
+%==================================
+% Angular Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left, 
+ylabel={Iterations},
+xlabel={Number of Vertices},ymajorgrids]
+\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
+\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
+\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Experimental results for the angular resolution experiment on the North graph testset.}
+\label{northAngular}
+\end{figure}
+
 \end{document}

### 2018-06-10 paper/paper.tex
wrote introductory sentence for AT&T graphs

@@
 \graphicspath{{figures/}}
-
 % ==================================================================
@@
 Further, we restricted the grid size to \begin{inparaenum}[(i)]
-\item $1,000,000 \times 1,000,000$, denoted by $1M \times 1M$ for short,
-\item $10,000 \times 10,000$, denoted by $10k \times 10k$ for short,
-\item $1,000 \times 1,000$, denoted by $1k \times 1k$ for short, and
-\item $100 \times 100$.
-
-\end{inparaenum}
-While the test set for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, we computed a different initial since our algorithm requires its input drawing to maintain its invariants (that is, vertices must be on the grid).
+\item $10^6 \times 10^6$
+\item $10^4 \times 10^4$
+\item $10^3 \times 10^3$, and
+\item $10^2 \times 10^2$.
+\end{inparaenum} The test set for this experiment was again the set of non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, but we computed a different initial since our algorithm requires its input drawing to maintain its invariants (that is, vertices must be on the grid).
 More precisely, we computed a random grid layout of each graph where vertices to overlap with other vertices or edges.
@@
 \addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}\end{axis}
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
+\end{axis}
 \end{tikzpicture}}}
@@
 \addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
 \end{axis}
@@
 \addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
 \end{axis}
@@
 \addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$1M \times 1M$},{$10k \times 10k$},{$1k \times 1k$},{$100 \times 100$}}
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
 \end{axis}
@@
 Regarding the crossing resulotion, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, the grid of size $100 \times 100$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
-Starting from grid size $1,000 \times 1,000$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-For grid size $1,000 \times 1,000$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10,000 \times 10,000$, the algorithm performed only about $10^\circ$ worse than the unrestricted version of our algorithm which could only slightly be improved by increasing the grid size to $1,000,000 \times 1,000,000$.
-
-The aspect ratio of computed drawings was more or less the same for all grid size with the exception of size $100 \times 100$ and always about twice the aspect ratio of the unrestricted variant of our algorithm; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-For grid size $100 \times 100$, this may again be explained by the fact, that for the majority of graphs, the initial layout could not be improved.
+More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
+Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
+For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
+For grid size $10^4 \times 10^4$, the algorithm performed only about $10^\circ$ worse than the unrestricted version of our algorithm which could only slightly be improved by increasing the grid size to $10^6 \times 10^6$.
+
+The aspect ratio of computed drawings was more or less the same for all grid size with the exception of size $10^2 \times 10^2$ and always about twice the aspect ratio of the unrestricted variant of our algorithm; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+For grid size $10^2 \times 10^2$, this may again be explained by the fact, that for the majority of graphs, the initial layout could not be improved.
 
 The number of crossings increased with the restriction on the size of the grid; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As with the crossing resolution, there is clear differences between grid sizes $100 \times 100$, $1,000 \times 1,000$ and $10,000 \times 10,000$ whereas there is only a slight improvement from grid size $10,000 \times 10,000$ to $1,000,000 \times 1,000,000$ which still uses twice as many crossings as our unrestricted variant.
+As with the crossing resolution, there is clear differences between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$ whereas there is only a slight improvement from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$ which still uses twice as many crossings as our unrestricted variant.
 This can partially be explained with the different choice of the initial drawing.
 
-Finally, we observe that the number of iterations needed for convergence for grid size $1,000 \times 1,000$ was already twice as much as for our unrestricted algorithm; see Fig.~\ref{fig:crossingResolutionTimeGrid}.
+Finally, we observe that the number of iterations needed for convergence for grid size $10^3 \times 10^3$ was already twice as much as for our unrestricted algorithm; see Fig.~\ref{fig:crossingResolutionTimeGrid}.
 From there on, the number of iterations for convergence increases with the grid size.
-Note that grid size $1,000,000 \times 1,000,000$ needed about 1000 to 2000 iterations more than grid size $10,000 \times 10,000$ even though the final layout was only marginally better as discussed prior.
-The curve for $100 \times 100$ highlights that this grid size was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
+Note that grid size $10^6 \times 10^6$ needed about 1000 to 2000 iterations more than grid size $10^4 \times 10^4$ even though the final layout was only marginally better as discussed prior.
+The curve for $10^2 \times 10^2$ highlights that this grid size was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
 
 In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid as long as the grid is not too small.
-However, the computation of grid drawings takes much longer than in the unrestricted version (up to five times as long for grid size $10,000 \times 10,000$ or eight times as long for grid size $1,000,000 \times 1,000,000$).
+However, the computation of grid drawings takes much longer than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$).
 In order to improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
@@
 
-\section{Experiments on the North Graph Testset}
-
-\todo[inline]{Say something about north graph}
+\section{Experiments on the AT\&T  Graph Test Set}
+
+We repeated our experimental evaluation on the crossing resolution, total resolution and angular resolution without grid constraint on a second test set of graphs, the set of non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
+The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
 
@@
 xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
 \addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
@@
 \caption{Experimental results for the crossing resolution experiment on the North graph testset.}
-\label{northCrossing}
+\label{fig:northCrossing}
 \end{figure}
@@
 \caption{Experimental results for the total resolution experiment on the North graph testset.}
-\label{northTotal}
+\label{fig:northTotal}
 \end{figure}
@@
 \caption{Experimental results for the angular resolution experiment on the North graph testset.}
-\label{northAngular}
+\label{fig:northAngular}
 \end{figure}

### 2018-06-10 paper/paper.tex
paper: minor typos + email

@@
 Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{christian-marius.geckeler,jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
+\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
+\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
 }
@@
 
-Formally, the value of the minimum angle formed by any two crossing edges in a drawing is refered to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
+Formally, the value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
 Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; see Figure~\ref{fig:examples}.
@@
 An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
+Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
 COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
@@
 Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastical'' or ``sharp'' move may help to escape.
+Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
 We turned this idea into an algorithmic implementation as follows:
@@
 Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions, which yield a reduced total~resolution.
+We do so by rejecting candidate positions which yield a reduced total~resolution.
 
@@
 
-Regarding the crossing resulotion, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
 More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.

### 2018-06-11 paper/paper.tex
worked on the appendix.

@@
 
-The termination condition of our algorithm is simple and depends on an input parameter~$\tau$.
-More specifically, if the crossing resolution has not been improved during the last $\tau$ iterations, then we assume that the algorithm has converged and we do not proceed any more.
+The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
+More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
 
@@
 % ==================================================================
-\myparagraph{Complexity related issues}
-% ==================================================================
-%
-A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing).
-Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-However, if the algorithm had to compute all crossing points and the corresponding angles at these points for each candidate position of each iteration, then it would not be useful in practice.
+\myparagraph{Complexity issues}
+% ==================================================================
+%
+A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
+Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
+%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
+
+If the algorithm had to compute all crossing points and the corresponding angles for each candidate position of each iteration, then it would not be useful in practice.
 Instead, we adopted a different approach, which turned out to be quite efficient in practice.
@@
 Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed with this procedure; this is our reference angle.
+Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
 Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
@@
 Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i) \cdot m)$, which in worst-case is $O(nm)$.
+Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
 
@@
 More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, which simply implies that candidate grid positions outside the bounds must be rejected.
+One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
 
@@
 
+\section*{Appendix}
+
 \section{Experiments on Grid Drawings}
@@
 The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
+In general we observed that the variance of results is much larger than in the experiments on the Rome graphs.
+This manifests in spikes of large magnitude in the illustrations of results and indicates that the structural properties of graphs in this second test set varies vastly between different graph sizes.
 
@@
 
+For the crossing resolution we observed that both the unrestricted and the aspect ratio restricted variants of our algorithm again outperformed the two evaluated force directed approachs; see Fig.~\ref{fig:northCrossing-1}.
+Remarkable is the synchronous behaviour of all four algorithms with respect to the crossing resolution on different graph sizes such that the curves are nearly parallel.
+By all these results, we can classify the graphs into ``hard'' or ``easy'' graphs with respect to the crossing resolution maximization; in particular graphs between 50 and 70 vertices appear to be harder to improve than graphs between 70 and 80 vertices.
+With respect to the aspect ratio of produced drawings we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force directed algorithms appears to be quite unstable resulting in a large variance.
+Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio which is much lower than the one of our unrestricted algorithm for larger graphs.
+All four algorithms behave nearly the same in terms of number of crossings; see Fig.~\ref{fig:northCrossing-3}.
+In terms of the number of iterations, we observe that surprisingly the algorithm of Argyriou et al.
+converges in the least amout of iterations throughout the test set whereas the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
+
 \begin{figure}[htb]
@@
 
+In the total resolution experiment, for smaller graphs,  we observed similar results as in the experiment on the Rome graphs, that is, our unrestricted algorithm achieves best total resolution followed by the restricted variant and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
+For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
+The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
+performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
+With respect to the number of iterations our two algorithms and the one by Argyriou et al.
+show similar behaviour needing more iterations than the algorithm by Huang et al.; see Fig.~\ref{fig:northTotal-4}.
+Observe that the number of iterations does not seem to correlate with the number of vertices.
+
 \begin{figure}[htb]
@@
 
+In the angular resolution experiments we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
+Only the algorithm by Huang et al.
+seems to be mostly at the last rank.
+Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
+while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
+For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
+achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
+Finally, both our algorithms and the one by Huang et al.
+need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
+
+\section{Graph Drawing Contest 2017 Graphs}
+
+We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}} on the crossing angle maximization and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
+Note that in the contest the teams had only one hour to compute a layout for all 15 contest graphs.
+
+\begin{table}[htb]
+
+\caption{Results for the Graph Drawing Contest Graphs.}
+\centering
+\begin{tabular}{c|c|c|c}
+Graph & CoffeeVM & Our New Approach & TuebingenMidnight \\
+\hline 
+1 & 90 & 89.78 & 77 \\
+2 & 88.23 & 88.7 & 42 \\
+3 & 90 & 89.95 & 89 \\
+4 & 88.97 & 89.05 & 89 \\
+5 & 80.4 & 86.96 &  30 \\
+6 & 90 & 89.72 & 78 \\
+7 & 56.537 & 63.62 & 34 \\
+8 & 84.95 & 89.28 & 61 \\
+9 & 59.885 & 88.2 & 9 \\
+10 & 20.978 & 23.72 & 4 \\
+11 & 46.684 & 72 & 6 \\
+12 & 36.47 & 35.86 & 5 \\
+13 & 25.456 & 33.68 & 4 \\
+14 & 33.52 & 43.08 & 5 \\
+15 & 20.512 & 29.18 & 4 \\
+%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\ 
+%\hline 
+%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\ 
+%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\ 
+\end{tabular} 
+\label{tab:gdContest2017}
+\end{table}
+
+For all the graphs, our results are (sometimes considerably) better than or about the same as the contest winner's.
+
+%\begin{tabular}{c|ccc}
+%Graph & Contest Winner & Our New Approach & T\"ubingen Midnight \\
+%\hline 
+%
+%Graph &  8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\ 
+%Contest Winner & 84.95 & 59.885 & 20.978 & 46.684 & 36.47 & 25.456 & 33.52 & 20.512 \\ 
+%Our New Approach & 89.28 & 88.2 & 23.72 & 72 & 35.86 & 33.68 & 43.08 & 29.18 \\ 
+%\end{tabular}
+
 \end{document}

### 2018-06-12 paper/paper.tex
paper

@@
 We conclude in Section~\ref{sec:conclusions} with open problems.
+In the appendix, we add more experimental results on grid restricted drawings, on more test sets and on the graphs from the graph drawing competition in 2017.
 
@@
 One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
+In Appendix~\ref{app:grid}, we report experimental results on this variant.
 
@@
 %
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs (commonly used for testing the efficiency of algorithms for drawing graphs).
+As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs.
+For another test set (AT\&T graphs) refer to Appendix~\ref{app:north}.
 
@@
 \section{Experiments on Grid Drawings}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to be placed on grid coordinates whilst avoiding vertices to overlap with other vertices or edges.
+\label{app:grid}
+
+In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to be placed on grid coordinates while avoiding vertices to overlap with other vertices or edges.
 Further, we restricted the grid size to \begin{inparaenum}[(i)]
@@
 The number of crossings increased with the restriction on the size of the grid; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As with the crossing resolution, there is clear differences between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$ whereas there is only a slight improvement from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$ which still uses twice as many crossings as our unrestricted variant.
+As with the crossing resolution, there are clear differences between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$ whereas there is only a slight improvement from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$ which still uses twice as many crossings as our unrestricted variant.
 This can partially be explained with the different choice of the initial drawing.
@@
 \section{Experiments on the AT\&T  Graph Test Set}
+\label{app:north}
 
@@
 In general we observed that the variance of results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of results and indicates that the structural properties of graphs in this second test set varies vastly between different graph sizes.
+This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of graphs in this second test set varies vastly between different graph sizes.
 
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution experiment on the North graph testset.}
+\caption{Experimental results for the crossing resolution experiment on the AT\&T graph testset.}
 \label{fig:northCrossing}
@@
 For the crossing resolution we observed that both the unrestricted and the aspect ratio restricted variants of our algorithm again outperformed the two evaluated force directed approachs; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms with respect to the crossing resolution on different graph sizes such that the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' graphs with respect to the crossing resolution maximization; in particular graphs between 50 and 70 vertices appear to be harder to improve than graphs between 70 and 80 vertices.
-With respect to the aspect ratio of produced drawings we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force directed algorithms appears to be quite unstable resulting in a large variance.
+Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution on different graph sizes as the curves are nearly parallel.
+By all these results, we can classify the graphs into ``hard'' or ``easy'' graphs regarding the crossing resolution maximization; in particular graphs between 50 and 70 vertices appear to be harder to improve than graphs between 70 and 80 vertices.
+Regarding the aspect ratio of the produced drawings we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force directed algorithms appears to be quite unstable resulting in a large variance.
 Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio which is much lower than the one of our unrestricted algorithm for larger graphs.
 All four algorithms behave nearly the same in terms of number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that surprisingly the algorithm of Argyriou et al.
+In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
 converges in the least amout of iterations throughout the test set whereas the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the total resolution experiment on the North graph testset.}
+\caption{Experimental results for the total resolution experiment on the AT\&T graph testset.}
 \label{fig:northTotal}
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the angular resolution experiment on the North graph testset.}
+\caption{Experimental results for the angular resolution experiment on the AT\&T graph testset.}
 \label{fig:northAngular}
@@
 
+Summarizing this section, we conclude that compared to the Rome graphs the AT\&T graphs show a much higher variance regarding the various resolution measures.
+The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
+
 \section{Graph Drawing Contest 2017 Graphs}
 
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}} on the crossing angle maximization and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
+Our primary motivation for this work was our participation to the graph drawing contest in 2017\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}} on the maximization of the crossing resolution where we miserably performed.
+Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Huang et al.
+and Argyriou et al.
+
+We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
 Note that in the contest the teams had only one hour to compute a layout for all 15 contest graphs.
@@
 
-For all the graphs, our results are (sometimes considerably) better than or about the same as the contest winner's.
+For all the graphs, our results are (sometimes considerably) better than or about the same as last year's contest winner's.
 

### 2018-06-12 paper/paper.tex
appendix c

@@
 We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute a layout for all 15 contest graphs.
+Note that in the contest the teams had only one hour to compute a layout for all 15 contest graphs, for our algorithm we provide results that were achieved with the same time restriction (see column ``Time restricted'') as well as our best result which were achieved without a strict time restriction (see column ``Our best'').
 
@@
 \centering
-\begin{tabular}{c|c|c|c}
-Graph & CoffeeVM & Our New Approach & TuebingenMidnight \\
+\begin{tabular}{c|c|c|c|c}
+Graph & CoffeeVM & TuebingenMidnight & Time restricted & Our best \\
 \hline 
-1 & 90 & 89.78 & 77 \\
-2 & 88.23 & 88.7 & 42 \\
-3 & 90 & 89.95 & 89 \\
-4 & 88.97 & 89.05 & 89 \\
-5 & 80.4 & 86.96 &  30 \\
-6 & 90 & 89.72 & 78 \\
-7 & 56.537 & 63.62 & 34 \\
-8 & 84.95 & 89.28 & 61 \\
-9 & 59.885 & 88.2 & 9 \\
-10 & 20.978 & 23.72 & 4 \\
-11 & 46.684 & 72 & 6 \\
-12 & 36.47 & 35.86 & 5 \\
-13 & 25.456 & 33.68 & 4 \\
-14 & 33.52 & 43.08 & 5 \\
-15 & 20.512 & 29.18 & 4 \\
+1 & 90  & 77  & 89.99 & 89.99\\
+2 & 88.23  & 42 & 88.21 & 88.7 \\
+3 & 90  & 89 & 87.86 & 89.95 \\
+4 & 88.97  & 89 & 77.13  & 89.05 \\
+5 & 80.4  &  30 & 78.68 & 86.96 \\
+6 & 90  & 78 & 89.96 & 89.96\\
+7 & 56.537  & 34 & 55.77  & 63.62 \\
+8 & 84.95  & 61 & 81.18 & 89.28 \\
+9 & 59.885  & 9 & 54.63 & 88.2 \\
+10 & 20.978  & 4 & 23.60  & 23.72 \\
+11 & 46.684  & 6 & 57.00 & 72.00 \\
+12 & 36.47  & 5 & 26.24 & 35.86 \\
+13 & 25.456  & 4 & 22.43 & 33.68 \\
+14 & 33.52  & 5 & 29.69 & 43.08 \\
+15 & 20.512  & 4 & 13.51 & 29.18 \\
 %Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\ 
@@
 
-For all the graphs, our results are (sometimes considerably) better than or about the same as last year's contest winner's.
-
+We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winners.
+On a few graphs (namely graphs 10 and 11) we even achieve better results.
+With a single exception (graph 4), we also outperform our results from last year.
+If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or about the same as last year's contest winner's.
+We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parametrization will be needed to compute competitive solutions within the provided time.
 %\begin{tabular}{c|ccc}

### 2018-06-12 paper/paper.tex
Final version (hopefully).

--- 2018-06-12 paper/paper.tex
@@
 \usepackage{rotating,hyperref,multicol}
+\usepackage{booktabs,tabularx}
 \graphicspath{{figures/}}
@@
 %In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper. 
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\ in Section~\ref{sec:experiments}.
-We conclude in Section~\ref{sec:conclusions} with open problems.
-In the appendix, we add more experimental results on grid restricted drawings, on more test sets and on the graphs from the graph drawing competition in 2017.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\ in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+In the appendix, we provide more experimental results on grid restricted drawings, on more test sets and on the graphs from the graph drawing competition in 2017.
 
@@
 In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observations that we made during our experimental evaluation (see Section~\ref{sec:experiments}): Drawings with good crossing resolution tend to have bad aspect ratio
+Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
 %(that is, the ratio of the longest to the shortest edge is large) 
@@
 %(that is, the angles formed by adjacent edges are small). 
-The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
+%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
 For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
@@
 %
-Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge.
-Sometimes it is also used as a measure of the area of non-grid drawings.
+%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings. 
 It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
@@
 %
-The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
+%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
 Formally, it is defined as the minimum of these two measures.
-It was not difficult to adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
+It was not difficult to
+Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
 In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
@@
 For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations were in Java using the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
+The implementations were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
 The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
 %
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs.
-For another test set (AT\&T graphs) refer to Appendix~\ref{app:north}.
+As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; in Appendix~\ref{app:north}, we also report on the AT\&T graphs.
 
@@
 \end{tikzpicture}}}
-\caption{Illustration of our experimental results on the crossing resolution.}
+\caption{Experimental results on the crossing resolution for the Rome graphs.}
 \label{fig:cr-res}
@@
 \myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize the minimum of the crossing and of the angular resolution.
+Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
 For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
 The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-It is worth noting, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs (even though there seems to be a small difference in our favor).
+Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
 
@@
 \end{tikzpicture}}}
-\caption{Illustration of our experimental results on the total resolution.}
+\caption{Experimental results on the total resolution for the Rome graphs.}
 \label{fig:to-res}
@@
 Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the remaining algorithms of our experiment (especially for large graphs).
+Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
 
@@
 \end{tikzpicture}}}
-\caption{Illustration of our experimental results on the angular resolution.}
+\caption{Experimental results on the angular resolution for the Rome graphs.}
 \label{fig:an-res}
@@
 % ==================================================================
-\section{Conclusions}
-\label{sec:conclusions}
-% ==================================================================
-
-In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
+\myparagraph{Discussion}
+%
+% ==================================================================
+%
+%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
 Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
 As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+%
+While working on this project, we made some useful observations and  obtained some interesting insights.
+In particular, there is a recent hypothesis (also supported by experiments) that drawings in which the crossing angles are large are easy to read and understand.
+What we observed is that the drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
+In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other (nearly at the same spot).
+Hence, the angular resolution was very often poor, while the aspect ratio could be arbitrarily bad.
+The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped in significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
+
+We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
+As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous (as also evidenced by our experimental evaluation).
+The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing (even although our initial plan was to optimize only the crossing resolution).
 
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution experiment on the AT\&T graph testset.}
+\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
 \label{fig:northCrossing}
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the total resolution experiment on the AT\&T graph testset.}
+\caption{Experimental results for the total resolution on the AT\&T graphs.}
 \label{fig:northTotal}
@@
 \end{tikzpicture}}}
-\caption{Experimental results for the angular resolution experiment on the AT\&T graph testset.}
+\caption{Experimental results for the angular resolution on the AT\&T graphs.}
 \label{fig:northAngular}
@@
 
-\section{Graph Drawing Contest 2017 Graphs}
-
-Our primary motivation for this work was our participation to the graph drawing contest in 2017\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}} on the maximization of the crossing resolution where we miserably performed.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Huang et al.
-and Argyriou et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute a layout for all 15 contest graphs.
-
-\begin{table}[htb]
-
-\caption{Results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c}
-Graph & CoffeeVM & Our New Approach & TuebingenMidnight \\
-\hline 
-1 & 90 & 89.78 & 77 \\
-2 & 88.23 & 88.7 & 42 \\
-3 & 90 & 89.95 & 89 \\
-4 & 88.97 & 89.05 & 89 \\
-5 & 80.4 & 86.96 &  30 \\
-6 & 90 & 89.72 & 78 \\
-7 & 56.537 & 63.62 & 34 \\
-8 & 84.95 & 89.28 & 61 \\
-9 & 59.885 & 88.2 & 9 \\
-10 & 20.978 & 23.72 & 4 \\
-11 & 46.684 & 72 & 6 \\
-12 & 36.47 & 35.86 & 5 \\
-13 & 25.456 & 33.68 & 4 \\
-14 & 33.52 & 43.08 & 5 \\
-15 & 20.512 & 29.18 & 4 \\
+\begin{table}[t!]
+
+\caption{Summary of the results for the Graph Drawing Contest Graphs.}
+\centering
+\begin{tabular}{c|c|c|c|c}
+\toprule
+~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
+\midrule
+1 & 90  & 77  & 89.99 & 89.99\\
+2 & 88.23  & 42 & 88.21 & 88.7 \\
+3 & 90  & 89 & 87.86 & 89.95 \\
+4 & 88.97  & 89 & 77.13  & 89.05 \\
+5 & 80.4  &  30 & 78.68 & 86.96 \\
+6 & 90  & 78 & 89.96 & 89.96\\
+7 & 56.537  & 34 & 55.77  & 63.62 \\
+8 & 84.95  & 61 & 81.18 & 89.28 \\
+9 & 59.885  & 9 & 54.63 & 88.2 \\
+10 & 20.978  & 4 & 23.60  & 23.72 \\
+11 & 46.684  & 6 & 57.00 & 72.00 \\
+12 & 36.47  & 5 & 26.24 & 35.86 \\
+13 & 25.456  & 4 & 22.43 & 33.68 \\
+14 & 33.52  & 5 & 29.69 & 43.08 \\
+15 & 20.512  & 4 & 13.51 & 29.18 \\
+\bottomrule
 %Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\ 
@@
 
-For all the graphs, our results are (sometimes considerably) better than or about the same as last year's contest winner's.
-
-%\begin{tabular}{c|ccc}
-%Graph & Contest Winner & Our New Approach & T\"ubingen Midnight \\
-%\hline 
-%
-%Graph &  8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\ 
-%Contest Winner & 84.95 & 59.885 & 20.978 & 46.684 & 36.47 & 25.456 & 33.52 & 20.512 \\ 
-%Our New Approach & 89.28 & 88.2 & 23.72 & 72 & 35.86 & 33.68 & 43.08 & 29.18 \\ 
-%\end{tabular}
+\section{Graph Drawing Contest 2017 Graphs}
+\label{app:contest}
+
+Our primary motivation for this work was our participation to the graph drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}.
+The topic was exactly the maximization of the crossing resolution.
+Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+
+We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
+Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
+For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
+
+We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
+On a few graphs (namely, graphs 10 and 11), we even achieve better results.
+With a single exception (namely, graph 4), we easily outperformed our results from last year.
+If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
+We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parametrization will be needed to compute competitive solutions within the provided time.
 

### 2018-06-13 paper/paper.tex
Unifying a bit the presentation between the main part and the appendix

@@
 
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to be placed on grid coordinates while avoiding vertices to overlap with other vertices or edges.
-Further, we restricted the grid size to \begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum} The test set for this experiment was again the set of non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, but we computed a different initial since our algorithm requires its input drawing to maintain its invariants (that is, vertices must be on the grid).
-More precisely, we computed a random grid layout of each graph where vertices to overlap with other vertices or edges.
-For each grid size, we computed a layout with the variant of our algorithm focusing on crossing resolution.
-Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
 \begin{figure}[!b]
@@
 \end{tikzpicture}}}
-\caption{Illustration of our experimental results on the crossing resolution with grid restriction. The double line shows the results of our unrestricted algorithm from Section~\ref{sec:experiments}.}
+\caption{Our experimental results on the crossing resolution with different grid restrictions.
+The double line corresponds to our unrestricted algorithm.}
 \label{fig:experimentsGrid}
@@
 
+In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
+In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
+%
+\begin{inparaenum}[(i)]
+\item $10^6 \times 10^6$
+\item $10^4 \times 10^4$
+\item $10^3 \times 10^3$, and
+\item $10^2 \times 10^2$.
+\end{inparaenum}
+%
+The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
+However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
+The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
+
 Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
-Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the algorithm performed only about $10^\circ$ worse than the unrestricted version of our algorithm which could only slightly be improved by increasing the grid size to $10^6 \times 10^6$.
-
-The aspect ratio of computed drawings was more or less the same for all grid size with the exception of size $10^2 \times 10^2$ and always about twice the aspect ratio of the unrestricted variant of our algorithm; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-For grid size $10^2 \times 10^2$, this may again be explained by the fact, that for the majority of graphs, the initial layout could not be improved.
-
-The number of crossings increased with the restriction on the size of the grid; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As with the crossing resolution, there are clear differences between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$ whereas there is only a slight improvement from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$ which still uses twice as many crossings as our unrestricted variant.
-This can partially be explained with the different choice of the initial drawing.
-
-Finally, we observe that the number of iterations needed for convergence for grid size $10^3 \times 10^3$ was already twice as much as for our unrestricted algorithm; see Fig.~\ref{fig:crossingResolutionTimeGrid}.
-From there on, the number of iterations for convergence increases with the grid size.
-Note that grid size $10^6 \times 10^6$ needed about 1000 to 2000 iterations more than grid size $10^4 \times 10^4$ even though the final layout was only marginally better as discussed prior.
-The curve for $10^2 \times 10^2$ highlights that this grid size was too restrictive for the vast majority of graphs which often resulted in the initial layout to be accepted as the best solution for larger graphs.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid as long as the grid is not too small.
-However, the computation of grid drawings takes much longer than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$).
-In order to improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
+More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
+As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
+Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when grid size was set to $10^3 \times 10^3$.
+%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
+%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
+For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
+
+The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+The fact that aspect ratio of these drawings was worse can also explained of course by the fact that in most cases an improved drawing could not be reported.
+
+As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
+As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
+It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
+Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
+%This can partially be explained with the different choice of the initial drawing.
+%
+On the other hand, the number of iterations needed for convergence increases with the grid size; see Fig.~\ref{fig:crossingResolutionTimeGrid}, with the exception of the grid of size $10^2 \times 10^2$, which clearly indicates our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+
+In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
+However, the computation of a grid drawing takes longer, which is of course expected.
+%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
+Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
+%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
 Further experiments in this direction are needed.
@@
 
-We repeated our experimental evaluation on the crossing resolution, total resolution and angular resolution without grid constraint on a second test set of graphs, the set of non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
+In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
+Note that we did not impose any  grid constraints our algorithms.
 The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
 In general we observed that the variance of results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of graphs in this second test set varies vastly between different graph sizes.
+This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
 
@@
 
-For the crossing resolution we observed that both the unrestricted and the aspect ratio restricted variants of our algorithm again outperformed the two evaluated force directed approachs; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution on different graph sizes as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' graphs regarding the crossing resolution maximization; in particular graphs between 50 and 70 vertices appear to be harder to improve than graphs between 70 and 80 vertices.
-Regarding the aspect ratio of the produced drawings we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force directed algorithms appears to be quite unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of number of crossings; see Fig.~\ref{fig:northCrossing-3}.
+For the crossing resolution we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
+Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
+By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
+In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
+Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
+Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
+All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
 In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amout of iterations throughout the test set whereas the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[htb]
+converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
+
+\begin{figure}[tb]
 \centering
@@
 
-In the total resolution experiment, for smaller graphs,  we observed similar results as in the experiment on the Rome graphs, that is, our unrestricted algorithm achieves best total resolution followed by the restricted variant and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
+In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
 For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
@@
 
-In the angular resolution experiments we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
+In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
 Only the algorithm by Huang et al.
@@
 
-Summarizing this section, we conclude that compared to the Rome graphs the AT\&T graphs show a much higher variance regarding the various resolution measures.
-The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
+Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
+%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
 
@@
 If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parametrization will be needed to compute competitive solutions within the provided time.
+We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
 

### 2018-06-13 paper/paper.tex
Unifying a bit the presentation between the main part and the appendix

@@
 performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations our two algorithms and the one by Argyriou et al.
-show similar behaviour needing more iterations than the algorithm by Huang et al.; see Fig.~\ref{fig:northTotal-4}.
-Observe that the number of iterations does not seem to correlate with the number of vertices.
+With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
+show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
+%Observe that the number of iterations does not seem to correlate with the number of vertices.
 

### 2018-06-23 src/layout/algo/gridding/QuickGridder.java
Gridding: only positive and no overlap

@@ -50,7 +50,9 @@ public class QuickGridder implements IGridder {
       double allowedAngle = oldAngle - configurator.allowDecreasingBy.getValue();
 
       Stream<PointD> samplePositions = getNeighborGridPositions(oldPosition, iteration).stream()
-          .filter(position -> !reservedPositions.contains(position));
+          .filter(position -> !reservedPositions.contains(position))
+          .filter(position -> LayoutUtils.overlap(position, positions, node, graph))
+          .filter(position -> position.x > 0 && position.y > 0);
 
       if (configurator.respectMinimumAngle.getValue()) {
         samplePositions = samplePositions.filter(position -> {

### 2018-06-23 src/algorithms/graphs/GridGraph.java
Scale Gridding with zoom

@@ -11,7 +11,8 @@ public final class GridGraph {
     return graph.getNodes().stream()
         .allMatch(node -> {
           PointD center = node.getLayout().getCenter();
-          return Util.isInteger(center.getX()) && Util.isInteger(center.getY());
+//          return Util.isInteger(center.getX()) && Util.isInteger(center.getY());
+            return Util.isAlmostInteger(center.getX()) && Util.isAlmostInteger(center.getY());  //TODO: fix with yFiles bug
         });
   }
 }

### 2018-06-23 src/graphoperations/GraphOperations.java

@@ -125,6 +125,9 @@ public class GraphOperations {
   }
 
   public static AspectRatio getAspectRatio(IGraph g){
+    if (g.getNodes().size() <= 0) {
+      return null;
+    }
     IEdge sEdge = getShortestEdge(g);
     INode sSource = sEdge.getSourceNode();
     INode sTarget = sEdge.getTargetNode();

### 2018-06-23 src/main/MainFrame.java

@@ -1,6 +1,7 @@
 package main;
 
 import com.sun.istack.internal.Nullable;
+import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.geometry.SizeD;
 import com.yworks.yfiles.graph.*;
 import com.yworks.yfiles.graph.styles.DefaultLabelStyle;
@@ -37,6 +38,7 @@ public class MainFrame extends JFrame {
     /* Box related issue*/
 
     public static final double BOX_SIZE[] = {1000000, 1000000};
+    private static int gridSize = 20;
 
     /* Graph Drawing related objects */
     public GraphComponent view;
@@ -164,6 +166,9 @@ public class MainFrame extends JFrame {
 
         this.view.addZoomChangedListener((o, zoomItemEventArgs) -> {
             boolean removedListeners = this.initSidePanel.removeDefaultListeners();
+            int spacing = (int)(gridSize/view.getZoom()) > 0 ? (int)(gridSize/view.getZoom()) : 1;
+            gridVisualCreator.getGridInfo().setHorizontalSpacing(spacing);
+            gridVisualCreator.getGridInfo().setVerticalSpacing(spacing);
             Scaling.scaleNodeSizes(view);
             if (this.graph.getNodes().size() > 100) {
                 try {
@@ -181,8 +186,9 @@ public class MainFrame extends JFrame {
         /* Add two listeners two the graph */
         this.graphSnapContext = new GraphSnapContext();
         this.graphEditorInputMode.setSnapContext(this.graphSnapContext);
-        GridInfo gridInfo = new GridInfo();
+        GridInfo gridInfo = new GridInfo(gridSize,gridSize, new PointD(0,0));
         this.gridVisualCreator = new GridVisualCreator(gridInfo);
+        this.gridVisualCreator.setVisibilityThreshold(5);
         this.view.getBackgroundGroup().addChild(this.gridVisualCreator, ICanvasObjectDescriptor.ALWAYS_DIRTY_INSTANCE);
         this.graphSnapContext.setGridSnapType(GridSnapTypes.GRID_POINTS);
         this.graphSnapContext.setNodeGridConstraintProvider(new GridConstraintProvider<>(gridInfo));

### 2018-06-23 src/main/MinimumAngleMonitor.java

@@ -7,6 +7,7 @@ import com.yworks.yfiles.graph.*;
 import com.yworks.yfiles.utils.IEventListener;
 import com.yworks.yfiles.utils.ItemEventArgs;
 import com.yworks.yfiles.view.GraphComponent;
+import graphoperations.GraphOperations;
 import layout.algo.utils.BestSolutionMonitor;
 import layout.algo.utils.PositionMap;
 import util.DisplayMessagesGui;
@@ -129,6 +130,19 @@ public class MinimumAngleMonitor {
     infoLabel.setText(labText.orElse("Graph has no crossings."));
   }
 
+  public void showAspectRatio(IGraph graph, GraphComponent view, JLabel infoLabel, boolean viewCenter) {
+    Optional<GraphOperations.AspectRatio> aspectRatio = Optional.ofNullable(GraphOperations.getAspectRatio(graph));
+    Optional<String> labText = aspectRatio.map(cr -> {
+      String text = DisplayMessagesGui.createAspectRatioMsg(graph, aspectRatio.get());
+
+      view.updateUI();
+
+      return text;
+    });
+
+    infoLabel.setText(labText.orElse("Graph has no crossings."));
+  }
+
 
   public void updateCrossingResolutionInfoBar() {
     showCrossingResolution(graph, view, infoLabel, false);
@@ -142,9 +156,8 @@ public class MinimumAngleMonitor {
     computeCrossingResolution();
     computeAngularResolution();
     if (useAspectRatio) {   //TODO
-        String text = DisplayMessagesGui.createAspectRatioMsg(graph);
-        infoLabel.setText(text);
-        return;
+      showAspectRatio(graph, view, infoLabel, false);
+      return;
     }
     if (useAngularResolution && useCrossingResolution) {
       if (currentAngularResolution < currentCrossingResolution) {

### 2018-06-23 src/util/DisplayMessagesGui.java

@@ -1,7 +1,6 @@
 package util;
 
 import com.yworks.yfiles.graph.IGraph;
-
 import graphoperations.GraphOperations;
 import layout.algo.utils.BestSolutionMonitor;
 import util.graph2d.Intersection;
@@ -44,8 +43,7 @@ public class DisplayMessagesGui {
         return text;
     }
 
-    public static String createAspectRatioMsg(IGraph graph) {
-        GraphOperations.AspectRatio aspectRatio = GraphOperations.getAspectRatio(graph);
+    public static String createAspectRatioMsg(IGraph graph, GraphOperations.AspectRatio aspectRatio) {
         return "Aspect Ratio: " + aspectRatio.getValue() +
                 " | Min: " + (int)aspectRatio.getShortestEdgeLength() +
                 " | Max.: " + (int)aspectRatio.getLongestEdgeLength();

### 2018-06-23 src/util/Util.java

@@ -71,4 +71,8 @@ public abstract class Util {
   public static boolean isInteger(double d) {
     return d % 1 == 0;
   }
+
+  public static boolean isAlmostInteger(double d) {
+    return Math.abs(Math.min(d%1, 1-d%1)) < 0.0000001;  //TODO: fix, yFiles bug where moving a node by hand results in 0.0000000000001 or 0.9999999999999
+  }
 }

### 2018-06-23 paper/paper.tex

@@
 For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
-Notably, RAC graph are sparse (they contain at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, where $n$ denotes the number of  vertices), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+Notably, RAC graphs are sparse (they contain at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, where $n$ denotes the number of  vertices), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
 
@@
 From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.\ Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
+The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
 Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
@@
 %In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\ in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-In the appendix, we provide more experimental results on grid restricted drawings, on more test sets and on the graphs from the graph drawing competition in 2017.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+In the appendix, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
 
@@
 %
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.\ and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 
@@
 So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We discuss ways to avoid them later in this section.
+We will discuss ways to avoid them later in this section.
 
@@
 The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
+More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
 For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
@@
 If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to chose the feasible point that maximizes the crossing resolution of the obtained drawing.
+The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
 As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
@@
 %
-Of course, this approach may lead to drawings with larger area, but this is expected, as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
+This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
 
@@
 
-If the algorithm had to compute all crossing points and the corresponding angles for each candidate position of each iteration, then it would not be useful in practice.
+If the algorithm had to compute all crossing points and the corresponding angles for each candidate position of each iteration, then it would not be useful.
 Instead, we adopted a different approach, which turned out to be quite efficient in practice.
@@
 It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
+Similarl as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
 In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
@@
 We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for bigger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
+However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
 On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
@@
 While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings in which the crossing angles are large are easy to read and understand.
+In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
 What we observed is that the drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
 In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other (nearly at the same spot).
-Hence, the angular resolution was very often poor, while the aspect ratio could be arbitrarily bad.
+Hence, the angular resolution and the aspect ratio were very often poor.
 The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped in significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
@@
 As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous (as also evidenced by our experimental evaluation).
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing (even although our initial plan was to optimize only the crossing resolution).
+The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing (even though our initial plan was to optimize only the crossing resolution).
 
@@
 As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when grid size was set to $10^3 \times 10^3$.
+Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
 %Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
@@
 The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that aspect ratio of these drawings was worse can also explained of course by the fact that in most cases an improved drawing could not be reported.
+The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
 
@@
 %
-On the other hand, the number of iterations needed for convergence increases with the grid size; see Fig.~\ref{fig:crossingResolutionTimeGrid}, with the exception of the grid of size $10^2 \times 10^2$, which clearly indicates our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
 
@@
 In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any  grid constraints our algorithms.
+Note that we did not impose any grid constraint on our algorithms.
 The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general we observed that the variance of results is much larger than in the experiments on the Rome graphs.
+In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
 This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
@@
 
-For the crossing resolution we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
+For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
 Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
@@
 
-Our primary motivation for this work was our participation to the graph drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}.
-The topic was exactly the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+Our primary motivation for this work was our participation to the graph drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
+Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
+and Huang et al.
 

### 2018-08-03 paper/comments/comments.tex
started revision of the paper.

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage[a4paper, total={6in, 9in}]{geometry}
\usepackage{soul}
\usepackage{enumerate}
\usepackage{xcolor,url}

\newcommand{\rcomment}[1]{\vspace{0.3cm} \item \textbf{Reviewer's Comment:} {\em #1}}
\newcommand{\tcomment}[1]{\vspace{0.3cm} {\color{red} \item \textbf{Reviewer's Comment:} {\em #1}}}
\newcommand{\jcomment}[1]{\vspace{0.3cm} {\color{blue} \item \textbf{Reviewer's Comment:} {\em #1}}}

\newcommand{\response}{\vspace{0.2cm} \textbf{Response: }}
\newcommand{\jresponse}[1]{\vspace{0.2cm} \textbf{Response: }{\color{blue} {\em #1}}}

\begin{document}

% ==================================================================================
\section*{Reviewer 1}
% ==================================================================================

\st{----------- Overall evaluation -----------

This paper presents an algorithm for maximising the angle of edge crossings in a graph drawing. The authors then compare the performance of their algorithm with two others, with respect to some graph drawing aesthetic criteria, demonstrating the worth of their approach.

This is a good idea, which clearly gives improved performance.}

\begin{itemize}

\tcomment{However the approach is not novel (although it has been presented as such) - and so the paper comes across as preliminary work that needs to be rooted in a wider context in order to strengthen it.}

\response{}

\tcomment{The process of choosing whether to move a node ("If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position") is an application of a simple hill-climbing algorithm (with the same advantages and disadvantages) - this needs to be acknowledged and clarified.}

\response{}

\jcomment{Or maybe the process might better be related to simulated annealing (with a temperature parameter)?}

\jresponse{If we make a journal version, we might consider to investigate the effect of the random movements gradually becoming smaller.}

\jcomment{The authors should consider the importance and relevance of the extensive prior work done in state-space problem solving when describing their heuristic approach. If they were to do this, then they might see the value of a composite evaluation function that would be able to take into account the other criteria discussed in section 3.1.}

\jresponse{This would require more extensive experiments and I would suggest to consider this approach again for a journal version.}

\end{itemize}

\st{The effort taken to implement the algorithms from two other published works is acknowledged: this is a good idea (and often not easy to do). The use of standard sets of graphs (Rome and AT\&T) is a very good approach in such quantitative evaluations.}

\begin{itemize}

\tcomment{Five aesthetic properties (P1-P5) are clearly defined, but never referred to again. There are only four comparative line charts in each of fig 3 and fig 4 (one of which - iterations - is not an aesthetic property).  P2 and P3 do not seem to have been measured or plotted.}

\response{}

\tcomment{Despite these omissions, the evaluation process appears sound: I would have liked to have seen some examples of drawings - everything is very abstract: what did these graphs actually look like? What does a graph with a high crossing resolution look like in comparison with one with a low crossing resolution?}

\response{}

\tcomment{The authors say that they made some interesting observations about the resulting drawings (e.g. "common to have adjacent edges to run almost in parallel") - some actual examples of actual graphs (before and after versions) would help elucidate these observations.}

\response{}

\end{itemize}

\st{In summary, this is a good idea that needs to be presented within the broader context of state-space search, and with clear examples to demonstrate its success anecdotally - the quantitative evaluation is fine, but a bit dry without some concrete examples.}
% ==================================================================================
\newpage
\section*{Reviewer 2}
% ==================================================================================

\st{----------- Overall evaluation -----------

Motivated by an unsuccessful performance at last year's automated GD contest, the authors designed, implemented and experimentally evaluated heuristic algorithms for optimizing crossing, angular and total (combined) resolution. Thus, the value being optimized is the size of a smallest angle (the bigger the better) created by an edge crossing and a pair of consecutive edges in the rotation at a vertex, resp., in straight-line drawings of graphs. The authors compared the performance of their algorithm on Rome graphs [11] and AT\&T graphs with the two other approaches by Argyriou et al. [5] and Huang et al. [32], both of which are force-directed unlike the heuristic proposed in the paper.

Due to its origin, the studied problem is, of course, a good fit for GD conference.}

\begin{itemize}

\jcomment{Although it perhaps makes more sense to choose the parameter being optimized in a more robust way, e.g., by taking also the average and variance of angle crossings into account.}

\jresponse{This might be an interesting direction to study if we are going to try making a journal version.}

\end{itemize}

\st{Starting from an arbitrary straight-line drawing, the heuristic is iteratively relocating vertices in the drawing one by one while improving upon the required resolution (crossing, angular, total). The choice of the vertex in every iteration is randomized as well as the choice of its new location. In particular, a vertex is chosen uniformly at random according to the distance to the vertices in the so-called vertex pool. The vertex pool consists either of the vertices whose adjacent edge(s) are participating in a crossing with the smallest angle, or of the whole vertex set. Then several new possible locations for the picked vertex, also chosen randomly, are evaluated w.r.t. to the resolution out of which at most one is chosen. In order to escape from a local maximum, the algorithms occasionally increase the size of the vertex pool to include all the vertices.

The proposed heuristic, in spite of its simplicity, looks reasonable. According to the experimental evaluation in the case of the crossing and total resolution, the heuristic outperformed by a considerable margin the other two algorithms by Argyriou et al. and Huang et al. on the graphs in the set. In the case of the angular resolution, the proposed heuristic was considerably better than both other algorithms only on medium-size graphs in the set, but much worse than Argyriou et al. for small graphs (up to 20 vertices), and comparable to its performance for big graphs (about 100 vertices) in the set.

As for the presentation, the paper is well-written and reads smoothly, I could spot only a couple of typos. One downside is that the authors do not discuss how to choose the parameters of the heuristic (see comments below), which makes the obtained result hard to reproduce; and the description of the algorithm is also not very clear at times.}

\begin{itemize}

\rcomment{I would also suggest to the authors to make their code publicly available.}

\response{Added a footnote to the github repo.}

\end{itemize}

\st{Due to the aforementioned lack of clarity I propose weak accept.}

---------------------------------------------------------------------------------------------------------

\begin{itemize}

\tcomment{l.103: The result of Dujmovic et al. looks like folklore. Split the edge set of the drawn graph into $\pi/\alpha$ parts each of which must be planar. If this is indeed correct, you can perhaps mention it.}

\response{}

\tcomment{l.141: "... weighted random selection procedure ..." Could you be more specific?}

\response{}

\rcomment{l.143: Which distance do you have in mind? Graph theoretical, Euclidean, Manhattan?}

\response{Added the following comment: (that is the shortest path to any vertex in the vertex-pool has length~$i$)}

\rcomment{l.143-145: "If the vertex pool contains ONLY critical vertices ... Otherwise, THE VERTEX POOL CONTAINS ALL THE VERTICES and each vertex can be chosen with the same probability." }

\response{Changed.}

\rcomment{l.155: " Next, we describeD ...  "(remove D)}

\response{Removed.}

\tcomment{l.158: How do you choose $\rho$?}

\response{}

\tcomment{l.165: You should discuss the choice of the parameters $\delta_{\min}$ and $\delta_{\max} $.}

\response{}

\rcomment{l.190: " ... becomes wider containing all the vertices" What do you mean?  That the vertex pool becomes wider unless it is not containng all the vertices already, or that you would include all the vertices into the pool.}

\response{Changed "containing" to "by including".}

\tcomment{l.193: How big is a medium-size graph?}

\response{}

\rcomment{l.207: na\"ive}

\response{Done.}

\rcomment{l.235: SimilarlY}

\response{Done.}

\rcomment{l.394: " ... helped IN significantly ... " (remove IN)}

\response{Removed.}

\end{itemize}
% ==================================================================================
\newpage
\section*{Reviewer 3}
% ==================================================================================
\st{----------- Overall evaluation -----------

This paper presents a novel heuristic algorithm for maximizing the crossing resolution of a graph layout, which is defined to be the maximum angle of all the minimum angles formed by any two crossing edges. The idea here is to randomly displace the end vertices of the crossing edges that define the crossing resolution of the entire graph layout for exploring the better solutions that increase the crossing resolution. The algorithm first collects critical edges that define the crossing resolution. It then seeks the better positions of their end vertices by randomly moving it along a set of rays that emanate from their original positions, where the angles of the rays from the x-axis are obtained by uniformly dividing 2*PI radians (360 degrees). The proposed algorithm was tested through the comparison with conventional algorithms based on variants of the force-directed algorithm. The experiments were conducted exhaustively, and the associated results are convincing. Overall, th!e paper has been well written and easy to understand. Interesting variants of the proposed formulation were discussed, as well as possible problems including unwanted traps by local minima and computational complexity issues. Thus, the paper is ready for presentation in the conference after an appropriate revision is carried out.

The authors are encouraged to address technical concerns as follows:}

\begin{itemize}

\tcomment{Improve the explanation at the end of the section "Complexity issues" on page 6.

The last paragraph on page 6 describes the computational complexity of the proposed algorithm for computing edge crossings. This is basically acceptable, while still a bit hard to understand. I believe that the description can be further improved by inserting some visual figures.}

\response{}

\tcomment{Categorization of aesthetic properties for evaluating the present algorithm

Section 4 describes that the present algorithm was compared with existing ones in terms of five aesthetic properties: (P1) crossing resolution, (P2) total resolution, (P3) angular resolution, (P4) aspect ratio, and (P5) number of crossings. However, the first three were employed as the measure for adjusting the algorithms, while the last two were evaluated in the respective experiments. Thus, just listing these five properties looks unnatural. The authors are encouraged to categorize important evaluation properties appropriately from the beginning.}

\response{}

\jcomment{Aspect ratio control in the proposed approach

I understand that the present algorithm may significantly distort the aspect ratio of the drawing areas if we do not impose any specific constraints over the displacement of graph vertices. However, just restricting the positions of vertices within the permitted range of the aspect ratio is quite naive and may be excessively restrictive. The authors are encouraged to discuss possible strategies to effectively control the aspect ratio of the drawing area in the context of the proposed algorithm.}

\jresponse{I guess that goes in the direction of composite evaluation function again (see comment of reviewer 1).}

\end{itemize}

\end{document}

### 2018-08-03 paper/paper.tex
started revision of the paper.

@@
 \footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a weighted random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool have higher weights than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, each vertex can be chosen with the same probability.
+We turned this observation into an algorithmic implementation through a weighted random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool (that is the shortest path to any vertex in the vertex-pool has length~$i$)  have higher weights than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
+So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
+Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
 
@@
 So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we  described how to compute its new position in the next drawing.
+Next, we  describe how to compute its new position in the next drawing.
 
@@
 %\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} containing all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
+\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
 \item afterwards, the vertex-pool switches back to the critical vertices.
@@
 A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
+Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
 %; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
@@
 It was not difficult to
-Similarl as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
+Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
 In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
@@
 For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
+The implementations\footnote{Our implementation is publicly available at \url{https://github.com/michaelbekos/crossingresolution}} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
 The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
@@
 Hence, the angular resolution and the aspect ratio were very often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped in significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
+The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
 

### 2018-08-15 src/sidepanel/SidePanelTab.java
Checkboxes alignment

@@ -282,6 +282,9 @@ public class SidePanelTab {
         cDefaultPanel.weightx = 0.5;
         cDefaultPanel.weighty = 0;
         enableCrossingResolution.addItemListener(this::enableCrossingResolutionActionPerformed);
+        if (!MainFrame.CONTEST_MODE) {
+            defaultPanel.add(enableCrossingResolution, cDefaultPanel);
+        }
         enableCrossingResolution.setSelected(false);
 
         enableAngularResolution = new JCheckBox("Angular Resolution");
@@ -291,6 +294,9 @@ public class SidePanelTab {
         cDefaultPanel.weightx = 0.5;
         cDefaultPanel.weighty = 0;
         enableAngularResolution.addItemListener(this::enableAngularResolutionActionPerformed);
+        if (!MainFrame.CONTEST_MODE) {
+            defaultPanel.add(enableAngularResolution, cDefaultPanel);
+        }
         enableAngularResolution.setSelected(false);
 
         enableAspectRatio = new JCheckBox("Aspect Ratio");
@@ -300,13 +306,11 @@ public class SidePanelTab {
         cDefaultPanel.weightx = 0.5;
         cDefaultPanel.weighty = 0;
         enableAspectRatio.addItemListener(this::enableAspectRatioActionPerformed);
-        enableAspectRatio.setSelected(false);
-
         if (!MainFrame.CONTEST_MODE) {
-            defaultPanel.add(enableCrossingResolution, cDefaultPanel);
-            defaultPanel.add(enableAngularResolution, cDefaultPanel);
             defaultPanel.add(enableAspectRatio, cDefaultPanel);
         }
+        enableAspectRatio.setSelected(false);
+
 
         outputTextArea.setLineWrap(true);
         outputTextArea.setRows(10);

### 2018-08-15 paper/comments/comments.tex

@@
 
-\tcomment{However the approach is not novel (although it has been presented as such) - and so the paper comes across as preliminary work that needs to be rooted in a wider context in order to strengthen it.}
+\rcomment{However the approach is not novel (although it has been presented as such) - and so the paper comes across as preliminary work that needs to be rooted in a wider context in order to strengthen it.}
 
-\response{}
+\response{See below.}
 
-\tcomment{The process of choosing whether to move a node ("If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position") is an application of a simple hill-climbing algorithm (with the same advantages and disadvantages) - this needs to be acknowledged and clarified.}
+\rcomment{The process of choosing whether to move a node ("If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position") is an application of a simple hill-climbing algorithm (with the same advantages and disadvantages) - this needs to be acknowledged and clarified.}
 
-\response{}
+\response{Added a note stating that our approach essentially resembles a hill climbing method.}
 
@@
 
-\tcomment{Five aesthetic properties (P1-P5) are clearly defined, but never referred to again.
+\rcomment{Five aesthetic properties (P1-P5) are clearly defined, but never referred to again.
 There are only four comparative line charts in each of fig 3 and fig 4 (one of which - iterations - is not an aesthetic property).
@@
 
-\response{}
+\response{Removed the notation with P.}
 
@@
 
-\tcomment{l.103: The result of Dujmovic et al.
+\rcomment{l.103: The result of Dujmovic et al.
 looks like folklore.
@@
 
-\tcomment{l.141: "... weighted random selection procedure ..." Could you be more specific?}
+\rcomment{l.141: "... weighted random selection procedure ..." Could you be more specific?}
 
-\response{}
+\response{Changed to probabilistic.}
 
@@
 
-\tcomment{l.193: How big is a medium-size graph?}
+\rcomment{l.193: How big is a medium-size graph?}
 
-\response{}
+\response{Made a note that we consider large to be more than 100 vertices.}
 
@@
 
-\tcomment{Improve the explanation at the end of the section "Complexity issues" on page 6.
+\rcomment{Improve the explanation at the end of the section "Complexity issues" on page 6.
 
@@
 
-\response{}
+\response{Refined the text a bit. No new figure.}
 
-\tcomment{Categorization of aesthetic properties for evaluating the present algorithm
+\rcomment{Categorization of aesthetic properties for evaluating the present algorithm
 
@@
 
-\response{}
+\response{Pointed out that P4 and P5 are additionally measured.}
 

### 2018-08-15 paper/paper.tex

@@
 
-Formally, the value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} graph or \emph{RAC} graph, for short; see Figure~\ref{fig:examples}.
-For these graphs, several results, mostly of theoretical nature, are known (refer to Section~\ref{sec:relatedwork} for a short overview).
-Notably, RAC graphs are sparse (they contain at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, where $n$ denotes the number of  vertices), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
+Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
 
@@
 
-The role of critical vertices is central in our algorithm
-%\footnote{If the focus is not on the critical vertices for a large graph, then an algorithm that is based on randomly selecting a vertex to move, so to improve the crossing resolution, will need a large number of iterations to converge to a good solution, because it is very unlike to select one of them to move.}
-\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a weighted random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool (that is the shortest path to any vertex in the vertex-pool has length~$i$)  have higher weights than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
+The role of critical vertices is central in our algorithmwiki\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool (that is the shortest path to any vertex in the vertex-pool has length~$i$)  have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
 So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
@@
 So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
+Note that this method resembles probabilistic hill climbing approaches.
 Next, we  describe how to compute its new position in the next drawing.
@@
 We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-
 The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
@@
 %From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be quite effective for medium-size graphs, for larger graphs, unfortunately, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
+While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
 
@@
 
-If the algorithm had to compute all crossing points and the corresponding angles for each candidate position of each iteration, then it would not be useful.
-Instead, we adopted a different approach, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate points to move $v_i$.
+Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
+Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
 Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
@@
 For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is publicly available at \url{https://github.com/michaelbekos/crossingresolution}} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
+The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
 The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
@@
 \end{inparaenum}
-%
-We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-\begin{multicols}{2}
-\begin{enumerate}[P.1.]
-\item \label{p:cr-res} crossing resolution
-\item \label{p:to-res} total resolution
-\item \label{p:an-res} angular resolution
-\columnbreak
-\item \label{p:as-rat} aspect ratio
-\item \label{p:no-xing} number of crossings
-\end{enumerate}
-\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), for~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
+
+We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
+Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
+
+All algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (that is,~the total resolution); see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
 In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+Additionally, we evaluate the produced drawings regarding the aspect ratio and the number of crossings.
 The reported results are on average across different drawings with same number of vertices.

### 2018-08-15 paper/comments/comments.tex
included some sample drawings and mentioned parameter settings in the text

@@
 
-\tcomment{l.158: How do you choose $\rho$?}
+\rcomment{l.158: How do you choose $\rho$?}
 
-\response{}
+\response{Added the following comment: For our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$ (where $w$ and $h$ are the width and the height of the initial drawing, respectively), $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$.\footnote{In order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.} }
 
-\tcomment{l.165: You should discuss the choice of the parameters $\delta_{\min}$ and $\delta_{\max} $.}
+\rcomment{l.165: You should discuss the choice of the parameters $\delta_{\min}$ and $\delta_{\max} $.}
 
-\response{}
+\response{See above.}
 

### 2018-08-15 paper/paper.tex
included some sample drawings and mentioned parameter settings in the text

@@
 
+%conf for conference version, arxiv for arxiv version
+\newcommand{\ver}{arxiv}
+\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
+
 % ==================================================================
@@
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-In the appendix, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
+\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
 
@@
 One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-In Appendix~\ref{app:grid}, we report experimental results on this variant.
+\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
 
@@
 %
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; in Appendix~\ref{app:north}, we also report on the AT\&T graphs.
+As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
 
@@
 \end{inparaenum}
+For our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$ (where $w$ and $h$ are the width and the height of the initial drawing, respectively), $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$.\footnote{In order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.}
 
@@
 
-Regarding the number of crossings, we observe that the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms of our experiment; refer to Fig.~\ref{fig:cr-res-3}.
+We observe that the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms; refer to Fig.~\ref{fig:cr-res-3}.
 
@@
 In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-What we observed is that the drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other (nearly at the same spot).
-Hence, the angular resolution and the aspect ratio were very often poor.
+We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
+In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
+Hence, angular resolution and aspect ratio were often poor.
 The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
 
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous (as also evidenced by our experimental evaluation).
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing (even though our initial plan was to optimize only the crossing resolution).
+We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
+As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
+The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
 
@@
 
-\newpage
+\arxapp{\newpage
 \appendix
@@
 
+\newpage
+
+\section{Sample Drawings}
+\label{app:samples}
+
+\begin{figure}[htbp]
+\centering
+%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
+%\centering
+%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+%}
+
+\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
+}
+\hfill
+\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-huang-total}
+}
+\caption{Drawings of graph 5 from last years graph drawing contest.  
+%(a)~Organic Layout by \emph{yFiles}, 
+(a)--(c)~Our results, 
+(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
+(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+\label{fig:graph5}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
+%\centering
+%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+%}
+
+\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
+}
+\hfill
+\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
+}
+\hfill
+\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\hfill
+\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-huang-total}
+}
+\caption{Drawings of graph 9 from last years graph drawing contest. 
+%(a)~Organic Layout by \emph{yFiles}, 
+(a)--(c)~Our results, 
+(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
+(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+\label{fig:graph9}
+\end{figure}}{}
+
 \end{document}

### 2018-08-29 paper/paper.tex
Several minor changes after Henry's pass. Ready from my side.

@@
 \graphicspath{{figures/}}
+
 % ==================================================================
@@
 % ==================================================================
-
 \newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
+
+%conf for conference version, arxiv for arxiv version
+\newcommand{\ver}{conf}
+\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
+
 \author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
@@
 \institute{
-Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
+Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
 \texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
@@
 
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{arxiv}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
 % ==================================================================
@@
 \maketitle
-\linenumbers
+%\linenumbers
 % ==================================================================
@@
 
-In Graph Drawing, there exists a really rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
 However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
@@
 
-The role of critical vertices is central in our algorithmwiki\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at distance~$i$ from the ones of the vertex-pool (that is the shortest path to any vertex in the vertex-pool has length~$i$)  have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
+The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
 So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
@@
 So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Note that this method resembles probabilistic hill climbing approaches.
 Next, we describe how to compute its new position in the next drawing.
+Note that our method resembles probabilistic hill climbing approaches.
 
@@
 \end{inparaenum}
-For our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$ (where $w$ and $h$ are the width and the height of the initial drawing, respectively), $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$.\footnote{In order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.}
-
-We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
+%We compared the quality of the produced drawings based on the following aesthetic~properties:
+%
+%\begin{multicols}{2}
+%\begin{enumerate}[P.1.]
+%\item \label{p:cr-res} crossing resolution
+%\item \label{p:to-res} total resolution
+%\item \label{p:an-res} angular resolution
+%\columnbreak
+%\item \label{p:as-rat} aspect ratio
+%\item \label{p:no-xing} number of crossings
+%\end{enumerate}
+%\end{multicols}
+%
+Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
+In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+The reported results are on average across different drawings with same number of vertices.
+Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
+
+%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
 Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-All algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (that is,~the total resolution); see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-Additionally, we evaluate the produced drawings regarding the aspect ratio and the number of crossings.
-The reported results are on average across different drawings with same number of vertices.
 
@@
 
-We observe that the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the corresponding number of crossings produced by the two other algorithms; refer to Fig.~\ref{fig:cr-res-3}.
+Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
 
@@
 
-Our primary motivation for this work was our participation to the graph drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
+Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
 Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
@@
 
-\newpage
+%\newpage
 
@@
 
-\begin{figure}[htbp]
+In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
+Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
+
+\begin{figure}[h!]
 \centering
@@
 }
-\caption{Drawings of graph 5 from last years graph drawing contest.  
+\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
 %(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~Our results, 
-(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
-(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+(a)--(c)~our algorithm, 
+(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
 \label{fig:graph5}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[p]
 \centering
@@
 }
-\caption{Drawings of graph 9 from last years graph drawing contest. 
+\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
 %(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~Our results, 
-(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
-(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+(a)--(c)~our algorithm, 
+(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
 \label{fig:graph9}

### 2018-08-29 paper/paper.tex
finished sample drawings

--- 2018-08-15 paper/paper.tex
@@
 
+In the following, we showcase drawings computed by our two algorithms and the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+For each algorithm, we present drawings attempting to optimize each of crossing, angular and total resolution.
+The two graphs shown here are taken from last year's graph drawing competition.
+As an initial drawing, we used for all algorithms the SmartOrganic layouts shown in Figure~\ref{fig:graph5And9-yfiles} computed with the yfiles library~\cite{DBLP:books/sp/04/WieseE004}.
+The drawings computed by each of the algorithms are shown in Figures~\ref{fig:graph5} and~\ref{fig:graph9}.
+
+\begin{figure}[htbp]
+\centering
+\subfloat[\label{fig:graph5-yfiles}{Graph 5}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+}
+\hfil
+\subfloat[\label{fig:graph9-yfiles}{Graph 9}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\caption{SmartOrganic layouts by yFiles~\cite{DBLP:books/sp/04/WieseE004} of graphs 5 and 9 from last year's graph drawing contest.}
+\label{fig:graph5And9-yfiles}
+\end{figure}
+
 \begin{figure}[htbp]
@@
 \hfill
+\subfloat[\label{fig:graph5-ourRestricted-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourRestricted-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-angular}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourRestricted-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-total}
+}
+\hfill
 \subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+\includegraphics[width=0.3\textwidth]{graph5-huang-crossing}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+\includegraphics[width=0.3\textwidth]{graph5-huang-angular}
 }
@@
 }
-\caption{Drawings of graph 5 from last years graph drawing contest.  
+\caption{Drawings of graph 5 from last year's graph drawing contest. 
 %(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~Our results, 
-(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
-(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+(a)--(c)~Our best results,
+(d)--(f)~Results of our algorithm restricted to maintain the aspect ratio 
+(g)--(i)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
+(j)--(l)~Results computed with Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.}
 \label{fig:graph5}
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+\includegraphics[width=0.3\textwidth]{graph9-ourBest-angular}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+\includegraphics[width=0.3\textwidth]{graph9-ourBest-total}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourRestricted-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourRestricted-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-angular}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourRestricted-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-total}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+\includegraphics[width=0.3\textwidth]{graph9-huang-crossing}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+\includegraphics[width=0.3\textwidth]{graph9-huang-angular}
 }
@@
 }
-\caption{Drawings of graph 9 from last years graph drawing contest. 
+\caption{Drawings of graph 9 from last year's graph drawing contest. 
 %(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~Our results, 
-(d)--(f)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
-(g)--(i)~Results computed with Huang et al~\cite{DBLP:journals/vlc/HuangEHL13}.}
+(a)--(c)~Our best results,
+(d)--(f)~Results of our algorithm restricted to maintain the aspect ratio 
+(g)--(i)~Results computed with Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, 
+(j)--(l)~Results computed with Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.}
 \label{fig:graph9}

### 2018-08-30 paper/paper.tex

--- 2018-08-29 paper/paper.tex
@@
 %conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
+\newcommand{\ver}{arxiv}
 \newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
@@
 Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
+As an initial drawing, we used for all algorithms the SmartOrganic layouts shown in Figure~\ref{fig:graph5And9-yfiles} computed with the yfiles library~\cite{DBLP:books/sp/04/WieseE004}.
+
+\begin{figure}[htbp]
+\centering
+\subfloat[\label{fig:graph5-yfiles}{Graph 5}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+}
+\hfil
+\subfloat[\label{fig:graph9-yfiles}{Graph 9}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\caption{SmartOrganic layouts by yFiles~\cite{DBLP:books/sp/04/WieseE004} of graphs 5 and 9 from last year's graph drawing contest.}
+\label{fig:graph5And9-yfiles}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
 \subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
@@
 }
+
 \caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
+(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
+(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
@@
 
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
+\begin{figure}[htbp]
+\centering
 \subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
+\includegraphics[width=0.3\textwidth]{graph9-ourBest-total}
+}
+
+\subfloat[\label{fig:graph9-ourRestricted-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourRestricted-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-angular}
+}
+\hfill
+\subfloat[\label{fig:graph9-ourRestricted-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-total}
+}
+
 \subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
@@
 }
-\hfill
+
 \subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+\includegraphics[width=0.3\textwidth]{graph9-huang-crossing}
 }
@@
 }
+
 \caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
+(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
+(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
 \label{fig:graph9}
-\end{figure}}{}
+\end{figure}
+
+}{}
 

### 2018-08-30 src/graphoperations/Scaling.java
updated: Centering and FPP
Centering:  Rotate the graph, so that his max side is parallel to the max side of the box
FPP: removed stretching und added planarity check output to the infobox

@@ -11,6 +11,7 @@ import com.yworks.yfiles.graph.styles.IEdgeStyle;
 import com.yworks.yfiles.graph.styles.PolylineEdgeStyle;
 import com.yworks.yfiles.view.GraphComponent;
 import com.yworks.yfiles.view.Pen;
+import main.MainFrame;
 import util.BoundingBox;
 
 import java.awt.*;
@@ -29,11 +30,13 @@ public final class Scaling {
      * Move all node coordinates to positive values and multiply them by a factor
      */
   public static void scaleBy(double factor, Mapper<INode, PointD> nodePositions) {
-    RectD bounds = BoundingBox.from(nodePositions);
-    scaleByCore(factor, nodePositions, bounds);
-  }
+        RectD bounds = BoundingBox.from(nodePositions);
+        scaleByCore(factor, nodePositions, bounds);
+    }
+
 
   private static void scaleByCore(double factor, Mapper<INode, PointD> nodePositions, RectD bounds) {
+
     for (Map.Entry<INode, PointD> entry : nodePositions.getEntries()) {
       INode node = entry.getKey();
       PointD center = entry.getValue();
@@ -46,7 +49,8 @@ public final class Scaling {
 
   }
 
-  public static void scaleNodeSizes(GraphComponent view) {
+
+    public static void scaleNodeSizes(GraphComponent view) {
       IGraph graph = view.getGraph();
       scaleNodeSizes(graph, view.getZoom());
   }

### 2018-08-30 src/sidepanel/SidePanelTab.java

@@ -39,6 +39,7 @@ import java.beans.PropertyChangeEvent;
 import java.time.Duration;
 import java.util.ArrayList;
 import java.util.LinkedList;
+import java.util.Map;
 import java.util.Optional;
 
 public class SidePanelTab {
@@ -581,15 +582,20 @@ public class SidePanelTab {
             IGraph graph = initSidePanel.mainFrame.graph;
             Mapper<INode, PointD> nodePositions = PositionMap.FromIGraph(graph);
             RectD bounds = BoundingBox.from(nodePositions);
-            Scaling.scaleBy(Math.max(1, Math.min((int) (MainFrame.BOX_SIZE[0] / bounds.getWidth() * scaleFactorWhenCentering),
-             (int) (MainFrame.BOX_SIZE[1] / bounds.getHeight()*scaleFactorWhenCentering))), nodePositions);
-//            Scaling.scaleBy(Math.min(MainFrame.BOX_SIZE[0] / bounds.getWidth() * scaleFactorWhenCentering,
-//                   MainFrame.BOX_SIZE[1] / bounds.getHeight()*scaleFactorWhenCentering), nodePositions);
-            Centering.moveToCenter(MainFrame.BOX_SIZE[0], MainFrame.BOX_SIZE[1], nodePositions);
+            //Rotate the graph, that his max side is parallel to the max side of the box
+            if(MainFrame.BOX_SIZE[0] >= MainFrame.BOX_SIZE[1] && bounds.width < bounds.height || MainFrame.BOX_SIZE[0] < MainFrame.BOX_SIZE[1] && bounds.width > bounds.height){
+                for (Map.Entry<INode, PointD> entry : nodePositions.getEntries()) {
+                    INode node = entry.getKey();
+                    PointD center = entry.getValue();
+                    nodePositions.setValue(node, new PointD(center.getY(),center.getX()));
+                }
+            }
+            bounds = BoundingBox.from(nodePositions);
+             Scaling.scaleBy(Math.max(1, Math.min((int) (MainFrame.BOX_SIZE[0] / bounds.getWidth() * scaleFactorWhenCentering),
+              (int) (MainFrame.BOX_SIZE[1] / bounds.getHeight()*scaleFactorWhenCentering))), nodePositions);
+
             PositionMap.applyToGraph(graph, nodePositions);
             initSidePanel.mainFrame.view.fitGraphBounds();
-
-
 
             bounds = BoundingBox.from(nodePositions);
[…]

### 2018-08-30 src/util/BoundingBox.java

@@ -19,7 +19,8 @@ import main.MainFrame;
 
 
 public class BoundingBox {
-	  public static RectD from(Mapper<INode, PointD> nodes) {
+
+	public static RectD from(Mapper<INode, PointD> nodes) {
 	    double minX = Double.POSITIVE_INFINITY, maxX = Double.NEGATIVE_INFINITY;
 	    double minY = Double.POSITIVE_INFINITY, maxY = Double.NEGATIVE_INFINITY;
 

### 2018-08-30 paper/paper.tex

@@
 The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
+Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=10$.
 
@@
 The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
+
+\paragraph{Acknowlegdments.} This project was supported by DFG grant KA812/18-1.
+The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
 

### 2018-08-30 paper/paper.tex
minor change

--- 2018-08-30 paper/paper.tex
@@
 %conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{arxiv}
+\newcommand{\ver}{conf}
 \newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
@@
 
-The latter result is already an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
+The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
 Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
@@
 Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-As an initial drawing, we used for all algorithms the SmartOrganic layouts shown in Figure~\ref{fig:graph5And9-yfiles} computed with the yfiles library~\cite{DBLP:books/sp/04/WieseE004}.
-
-\begin{figure}[htbp]
-\centering
-\subfloat[\label{fig:graph5-yfiles}{Graph 5}]{
+
+\begin{figure}[h!]
+\centering
+%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
+%\centering
+%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
+%}
+
+\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
+}
+\hfill
+\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
+}
+\hfill
+\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
+\centering
+\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
+}
+\hfill
+\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
 \centering
@@
 }
-\hfil
-\subfloat[\label{fig:graph9-yfiles}{Graph 9}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\caption{SmartOrganic layouts by yFiles~\cite{DBLP:books/sp/04/WieseE004} of graphs 5 and 9 from last year's graph drawing contest.}
-\label{fig:graph5And9-yfiles}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-crossing}
-}
 \hfill
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-angular}
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}
 }
@@
 }
-
 \caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
-(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+%(a)~Organic Layout by \emph{yFiles}, 
+(a)--(c)~our algorithm, 
+(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
@@
 
-\begin{figure}[htbp]
-\centering
+\begin{figure}[p]
+\centering
+%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
+%\centering
+%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+%}
+
 \subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-angular}
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-total}
-}
-
-\subfloat[\label{fig:graph9-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-total}
-}
-
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
+}
+\hfill
 \subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
@@
 }
-
+\hfill
 \subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-crossing}
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
 }
@@
 \centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-angular}
+\includegraphics[width=0.3\textwidth]{graph9-yfiles}
 }
@@
 }
-
 \caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
-(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+%(a)~Organic Layout by \emph{yFiles}, 
+(a)--(c)~our algorithm, 
+(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
+(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
 \label{fig:graph9}
-\end{figure}
-
-}{}
+\end{figure}}{}
 

### 2018-08-30 paper/paper.tex
minor change for consistency reasons.

--- 2018-08-30 paper/paper.tex
@@
 Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-As an initial drawing, we used for all algorithms the SmartOrganic layouts shown in Figure~\ref{fig:graph5And9-yfiles} computed with the yfiles library~\cite{DBLP:books/sp/04/WieseE004}.
+As initial drawings for all algorithms, we used the layouts shown in Figure~\ref{fig:graph5And9-yfiles} that we computed with the SmartOrganic layouter of the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
 
@@
 \centering
-\subfloat[\label{fig:graph5-yfiles}{Graph 5}]{
+\subfloat[\label{fig:graph5-yfiles}{}]{
 \centering
@@
 \hfil
-\subfloat[\label{fig:graph9-yfiles}{Graph 9}]{
+\subfloat[\label{fig:graph9-yfiles}{}]{
 \centering
@@
 }
-\caption{SmartOrganic layouts by yFiles~\cite{DBLP:books/sp/04/WieseE004} of graphs 5 and 9 from last year's graph drawing contest.}
+\caption{Illustration of 
+(a)~the 5th, and of
+(b)~the 9th graph of the Graph Drawing contest 2017.}
 \label{fig:graph5And9-yfiles}

### 2018-08-30 src/graphoperations/Scaling.java
Reinserte centering, change default node pen style

@@ -60,11 +60,12 @@ public final class Scaling {
         double height = graph.getNodeDefaults().getSize().height * scaleValue;
         for (INode u : graph.getNodes()) {
             graph.setNodeLayout(u, new RectD(
-                    u.getLayout().getCenter().getX() - 0.5*width,
-                    u.getLayout().getCenter().getY() - 0.5*height,
+                    u.getLayout().getCenter().getX() - 0.5 * width,
+                    u.getLayout().getCenter().getY() - 0.5 * height,
                     width,
                     height
             ));
+
         }
     }
 
@@ -75,7 +76,6 @@ public final class Scaling {
 
     public static void scaleEdgeSizes(IGraph graph, double zoomValue){
         double scaleValue = 1 / zoomValue; // scale reinserted nodes
-        PolylineEdgeStyle edgeStyle = new PolylineEdgeStyle();
         for(IEdge e : graph.getEdges()){
             IEdgeStyle s = e.getStyle();
             if(s instanceof PolylineEdgeStyle) {
@@ -83,7 +83,6 @@ public final class Scaling {
             } else {
                 System.out.println(s.getClass());
             }
-          //  graph.setStyle(e, edgeStyle);
         }
 
 

### 2018-08-30 src/sidepanel/SidePanelTab.java

@@ -594,6 +594,7 @@ public class SidePanelTab {
              Scaling.scaleBy(Math.max(1, Math.min((int) (MainFrame.BOX_SIZE[0] / bounds.getWidth() * scaleFactorWhenCentering),
               (int) (MainFrame.BOX_SIZE[1] / bounds.getHeight()*scaleFactorWhenCentering))), nodePositions);
 
+            Centering.moveToCenter(MainFrame.BOX_SIZE[0], MainFrame.BOX_SIZE[1], nodePositions);
             PositionMap.applyToGraph(graph, nodePositions);
             initSidePanel.mainFrame.view.fitGraphBounds();
 
@@ -817,9 +818,11 @@ public class SidePanelTab {
 
     private void fppItemActionPerformed(@SuppressWarnings("unused") ActionEvent evt) {
         YGraphAdapter graphAdapter = new YGraphAdapter(initSidePanel.mainFrame.graph);
-        String outString = "Graph is planar: " + GraphChecker.isPlanar(graphAdapter.getYGraph());
+        String outString = "planar: " + GraphChecker.isPlanar(graphAdapter.getYGraph()) +
+                "\nloop free: " + GraphChecker.isSelfLoopFree(graphAdapter.getYGraph()) +
+                "\nmultiple edge free: " + GraphChecker.isMultipleEdgeFree(graphAdapter.getYGraph());
         if( com.yworks.yfiles.algorithms.GraphChecker.isBiconnected(graphAdapter.getYGraph())) {
-            outputTextArea.setText("Graph is biconnected.\n" + outString );
+            outputTextArea.setText("biconnected: true\n" + outString );
             initSidePanel.removeDefaultListeners();
             ICompoundEdit compoundEdit = initSidePanel.mainFrame.graph.beginEdit("Undo layout", "Redo layout");
             FraysseixPachPollack.FPPSettings fppSettings = new FraysseixPachPollack.FPPSettings();
@@ -830,7 +833,7 @@ public class SidePanelTab {
             compoundEdit.commit();
             initSidePanel.addDefaultListeners();
         }else{
-            outputTextArea.setText("Graph is not biconnected!" + outString);
+            outputTextArea.setText("biconnected: false \n" + outString);
         }
 
     }

### 2018-09-12 src/main/ContestWindow.java
fix timer, add automatic option, save all on close, other misc

@@ -17,7 +17,6 @@ import java.util.stream.Stream;
 import static main.GraphFrame.createSeparator;
 
 public class ContestWindow {
-    private static volatile int openFrames;
     private JFrame contestWindow;
     private GridBagConstraints cFrame;
 
@@ -45,8 +44,9 @@ public class ContestWindow {
         int endIndex = startIndex;
         String folderPath = "contest-2018";
         String pattern = "$$$.json";
+        boolean automatic = false;
 
-        if (args.length == 1) {
+        if (args.length == 1) {     //load all files in folder, automate first 8
             folderPath = args[0];
             try (Stream<Path> paths = Files.walk(Paths.get(folderPath))) {
                 List<Integer> list = paths
@@ -56,6 +56,7 @@ public class ContestWindow {
                         .collect(Collectors.toList());
                 startIndex = list.get(0);
                 endIndex = list.get(list.size() - 1);
+                automatic = true;
 
                 System.out.println(startIndex +  " " + endIndex);
             } catch (IOException e) {
@@ -63,19 +64,40 @@ public class ContestWindow {
             }
         }
 
-        try {
-            startIndex = Integer.parseInt(args[0]);
-            if (args.length >= 2) {
-                endIndex = Integer.parseInt(args[1]);
-            } else {
-                endIndex = startIndex;
+        if (args.length == 2) {     //load all files, no automation
+            folderPath = args[0];
+//            try (Stream<Path> paths = Files.walk(Paths.get(folderPath))) {
+            try (Stream<Path> paths = Files.list(Paths.get(folderPath))) { //non-recursive
+                List<Integer> list = paths
+                        .filter(Files::isRegularFile)
+//                        .map(path -> path.getFileName().toString().replaceAll("[^0-9]","")) //TODO: rename files so only number.json
+//                        .map(path -> Integer.parseInt(path.replaceFirst("[.][^.]+$", "")))
+                        .map(path -> Integer.parseInt(path.getFileName().toString().replaceFirst("[.][^.]+$", "")))
+                        .sorted()
+                        .collect(Collectors.toList());
+                startIndex = list.get(0);
+                endIndex = list.get(list.size() - 1);
+                automatic = false;
+
+                System.out.println(startIndex +  " " + endIndex);
+            } catch (IOException e) {
+                e.printStackTrace();
             }
-        } catch (NumberFormatException e) {
[…]

### 2018-09-12 src/main/GraphFrame.java

@@ -1,5 +1,6 @@
 package main;
 
+import algorithms.graphs.GridGraph;
 import com.yworks.yfiles.graph.LayoutUtilities;
 import com.yworks.yfiles.layout.organic.OrganicLayout;
 import layout.algo.randommovement.RandomMovementLayout;
@@ -29,8 +30,7 @@ public class GraphFrame {
     private JLabel minAngleLabel;
     private JLabel timeLabel;
     private JCheckBox allowDecreasing;
-    private long runningTime;
-    private long startTime;
+    private double runningTime;
     private long prevTime;
     private long currTime;
 
@@ -45,9 +45,10 @@ public class GraphFrame {
             directory.mkdir();
         }
         this.fileName = pattern.replace("$$$", Integer.toString(id));
-        initFrame();
-        this.startTime = System.nanoTime();
+        this.runningTime = 0;
+        this.prevTime = System.nanoTime();
         initMainFrame();
+        initFrame();
         updateButtons(getSidePanel());
 
         mainFrame.minimumAngleMonitor.addPropertyChangeListener(this::angleChangedPropertyChanged);
@@ -59,9 +60,7 @@ public class GraphFrame {
         mainFrame.setVisible(false);
         mainFrame.setDefaultCloseOperation(WindowConstants.DISPOSE_ON_CLOSE);
 
-//        addClosingListener(mainFrame);
         loadGraph(mainFrame, folderPath, fileName);
-        System.out.println(folderPath +  "  " +fileName);
         runAlgorithms(mainFrame, automatic);
     }
 
@@ -82,8 +81,8 @@ public class GraphFrame {
 
 
 
-//        JLabel minAngle = new JLabel(Double.toString(mainFrame.minimumAngleMonitor.getBestCrossingResolution()));
-        minAngleLabel = new JLabel("-1");
+        minAngleLabel = new JLabel(String.format("%.5f",mainFrame.minimumAngleMonitor.getCurrentCrossingResolution()));
+//        minAngleLabel = new JLabel("-1");
         cFramePanel.gridx += 1;
         panel.add(minAngleLabel, cFramePanel);
         cFramePanel.gridx += 1;
@@ -96,7 +95,7 @@ public class GraphFrame {
         cFramePanel.gridx += 1;
         panel.add(createSeparator(), cFramePanel);
 
-        allowDecreasing = new JCheckBox("Allow Decreasing");
[…]

### 2018-09-12 src/main/MainFrame.java

@@ -6,7 +6,6 @@ import com.yworks.yfiles.geometry.PointD;
 import com.yworks.yfiles.geometry.SizeD;
 import com.yworks.yfiles.graph.*;
 import com.yworks.yfiles.graph.styles.DefaultLabelStyle;
-import com.yworks.yfiles.graph.styles.IArrow;
 import com.yworks.yfiles.graph.styles.PolylineEdgeStyle;
 import com.yworks.yfiles.graph.styles.ShinyPlateNodeStyle;
 import com.yworks.yfiles.layout.organic.OrganicLayout;

### 2018-09-12 paper/journal.tex
preliminary work

--- 2018-08-30 paper/paper.tex
@@
 
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{arxiv}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
 \author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
@@
 \maketitle
-%\linenumbers
+\linenumbers
 % ==================================================================
@@
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
+In the appendix, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
 
@@
 % ==================================================================
-\subsection{Some interesting variants}
+\subsection{Some Interesting Variants}
 \label{ssec:variants}
@@
 One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
+In Appendix~\ref{app:grid}, we report experimental results on this variant.
 
@@
 %
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
+As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; in Appendix~\ref{app:north}, we also report on the AT\&T graphs.
 
@@
 
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
+We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
 As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
@@
 
-\paragraph{Acknowlegdments.} This project was supported by DFG grant KA812/18-1.
+\paragraph{Acknowledgments.} This project was supported by DFG grant KA812/18-1.
 The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
@@
 
-\arxapp{\newpage
+\newpage
 \appendix
@@
 
-}{}
-
 \end{document}

### 2018-09-13 paper/journal.tex
Creating a first draft of the jounral version

@@
-\documentclass[runningheads]{llncs}
+\documentclass{comjnl}
 \usepackage{amsmath,amssymb}
@@
 \usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
+%\usepackage{algorithm2e}
 \usepackage{todonotes,lineno,paralist,soul}
@@
 \newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
+% ==================================================================
+
+\begin{document}
+
+\author{Michael~A.~Bekos} 
+\author{Henry~F\"orster}
+\author{Christian~Geckeler} 
+\author{Lukas~Holl\"ander} 
+\author{Michael~Kaufmann}
+\author{Amad\"aus~M.~Spallek} 
+\author{Jan~Splett}
+
+\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
+
+\affiliation{
 Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-\linenumbers
+}
+
+\email{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de\\\{geckeler,jan-lukas.hollaender,amadaeus.spallek,jan.splett\}@student.uni-tuebingen.de}
+
+\received{13 September 2018}
+
+\shortauthors{M.~A.~Bekos et al.}
+
+\keywords{Graph Drawing, Angular, Crossing and Total Resolution.}
+
+%\copyrightyear{2018} \vol{00} \issue{0} \DOI{000}
 % ==================================================================
@@
 The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
+Recent experiments suggest that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
+However, maximizing the crossing resolution turns out to be an NP-hard problem in general, and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
 
@@
 % ==================================================================
+\maketitle
+% ==================================================================
+
+% ==================================================================
 \section{Introduction}
@@
 
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
+In Graph Drawing, there exists a rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
+However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a quite difficult and very challenging task, mainly due to the edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
+As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview we point the reader to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
+
+In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence (through eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
+In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing~\cite{DBLP:conf/gd/Purchase97,DBLP:journals/iwc/Purchase00}, it turned out that the types of edge crossings also matter.
 As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
 
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+Formally, the value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution\footnote{Also, refered to as \emph{crossing angle resolution} in the literature.}}.
+Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
+Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (or \emph{RAC}, for short) graph; see Figure~\ref{fig:examples} for an illustration.
+Notably, RAC graphs are sparse (an $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
 
@@
 	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
+	\includegraphics[page=1]{figures/examples}}
 	\hfil
 	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
+	\includegraphics[page=2]{figures/examples}}
+	\caption{Illustration of: %
+	(a)~a RAC drawing of the complete graph $K_5$, and
 	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
@@
 The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
+It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
 
 From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
+The first one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, while the~second one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
 Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
@@
 
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
+In this work, we approach the crossing resolution maximization problem from a different perspective.
+We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
 How would one increase the crossing resolution of a given drawing?
@@
 Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
+If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position.
 
 In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
+Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself),
 we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
+We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms.
+Our evaluation indicates that our method significantly outperforms the force-directed algorithms~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
+Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
 
@@
 The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
 Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
@@
 \medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
+The remainder of this paper is structured as follows.
+Section~\ref{sec:relatedwork} overviews related works.
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-In the appendix, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
 
@@
 The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; in Appendix~\ref{app:north}, we also report on the AT\&T graphs.
+As test sets for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13} and the AT\&T graphs, which are two collections of around 8.100 and 424 benchmark graphs, respectively.
 
 The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
+Initially, each graph of our expertiment was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
 Starting from this layout, every graph was drawn with
@@
 \item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
+\item our algorithm restricted not to violate the aspect ratio of the initial layout, and the force-directed algorithms
 \item by Argyriou et al.\ and
@@
 \end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
+%We compared the quality of the produced drawings based on the following aesthetic properties:
+%
 %\begin{enumerate}[P.1.]
@@
 %\item \label{p:an-res} angular resolution
-%\columnbreak
 %\item \label{p:as-rat} aspect ratio
@@
 %\end{enumerate}
-%\end{multicols}
-%
+
 Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
@@
 
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
+\subsection{The Rome Graph Test Set}
+
+In this subsection, we report the results of our expertimental evaluation for the non-planar Rome graphs, which form a collection of 424 benchmark graphs.
+Note that we did not impose any grid constraint on our algorithms.
+Our results are summarized in Figs.~\ref{fig:cr-res}, \ref{fig:an-res} and~\ref{fig:to-res}.
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
@@
 \label{fig:cr-res}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:to-res}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:an-res}
-\end{figure}
+\end{figure*}
 
@@
 
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\paragraph{Acknowledgments.} This project was supported by DFG grant KA812/18-1.
-The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
+\subsection{The AT\&T  Graph Test Set}
+\label{app:north}
+
+In this subsection, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
+Note that we did not impose any grid constraint on our algorithms.
+The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
+In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
+This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
+
+\begin{figure*}[htb]
+\centering
+\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
+\centering
+\scalebox{.95}{
 \centering
@@
 tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
 \addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
@@
 \label{fig:northCrossing}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
 \centering
@@
 \label{fig:northTotal}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[htb]
+\begin{figure*}[htb]
 \centering
@@
 \label{fig:northAngular}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}[t!]
+\subsection{Graph Drawing Contest 2017}
+\label{app:contest}
+
+Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
+Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
+and Huang et al.
+
+We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
+Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
+For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
+
+We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
+On a few graphs (namely, graphs 10 and 11), we even achieve better results.
+With a single exception (namely, graph 4), we easily outperformed our results from last year.
+If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
+We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
+
+\begin{table*}[t!]
 
@@
 \bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
 \end{tabular}
 \label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
+\end{table*}
+
+\subsection{Grid Restricted Drawings}
+\label{app:grid}
+
+\begin{figure*}[!b]
+\centering
+\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================
+% Crossing Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[
+axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
+legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
+width=0.475\textwidth,
+cycle list name=mylist,
+mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
+xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={10,20,30,40,50,60,70,80,90},
+tick pos=left,ymajorgrids]
+\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%===================================
+% Crossing Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
+xlabel={Number of Vertices},
+xmin=10,
+xmax=100,
+ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
+\end{axis}
+\end{tikzpicture}}}
+
+\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%======================================
+% Crossing Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
+xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
+\end{axis}
+\end{tikzpicture}}}
+\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
+\centering
+\scalebox{0.95}{
+\centering
+%==================================
+% Crossing Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
+anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
+xmin=10,
+xmax=100,
+ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
+ylabel={Iterations},
+xlabel={Number of Vertices},ymajorgrids]
+\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
+\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
+\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
+\end{axis}
+\end{tikzpicture}}}
+\caption{Our experimental results on the crossing resolution with different grid restrictions.
+The double line corresponds to our unrestricted algorithm.}
+\label{fig:experimentsGrid}
+\end{figure*}
+
+In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
+In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
+%
+\begin{inparaenum}[(i)]
+\item $10^6 \times 10^6$
+\item $10^4 \times 10^4$
+\item $10^3 \times 10^3$, and
+\item $10^2 \times 10^2$.
+\end{inparaenum}
+%
+The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
+However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
+The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
+
+Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
+As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
+Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
+%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
+%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
+For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
+
+The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
+
+As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
+As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
+It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
+Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
+%This can partially be explained with the different choice of the initial drawing.
+%
+On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+
+In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
+However, the computation of a grid drawing takes longer, which is of course expected.
+%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
+Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
+%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
+Further experiments in this direction are needed.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
@@
 \label{fig:graph5And9-yfiles}
-\end{figure}
-
-\begin{figure}[p]
+\end{figure*}
+
+\begin{figure*}[p]
 \centering
@@
 \label{fig:graph5}
-\end{figure}
-
-\begin{figure}[htbp]
+\end{figure*}
+
+\begin{figure*}[htbp]
 \centering
@@
 \label{fig:graph9}
-\end{figure}
+\end{figure*}
+
+% ==================================================================
+\section{Conclusions}
+%
+% ==================================================================
+%
+%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
+Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
+As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+%
+While working on this project, we made some useful observations and  obtained some interesting insights.
+In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
+We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
+In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
+Hence, angular resolution and aspect ratio were often poor.
+The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
+
+We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
+As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
+The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
+
+\paragraph{Acknowledgments.} This project was supported by DFG grant KA812/18-1.
+The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
+
+\bibliographystyle{compj}
+\bibliography{references}
 

### 2018-09-14 paper/cr-res-north.tex
Work on the journal version.

--- 2018-09-13 paper/journal.tex
@@
-\documentclass{comjnl}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-%\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-% ==================================================================
-
-\begin{document}
-
-\author{Michael~A.~Bekos} 
-\author{Henry~F\"orster}
-\author{Christian~Geckeler} 
-\author{Lukas~Holl\"ander} 
-\author{Michael~Kaufmann}
-\author{Amad\"aus~M.~Spallek} 
-\author{Jan~Splett}
-
-\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-
-\affiliation{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-}
-
-\email{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de\\\{geckeler,jan-lukas.hollaender,amadaeus.spallek,jan.splett\}@student.uni-tuebingen.de}
-
-\received{13 September 2018}
-
-\shortauthors{M.~A.~Bekos et al.}
-
-\keywords{Graph Drawing, Angular, Crossing and Total Resolution.}
-
-%\copyrightyear{2018} \vol{00} \issue{0} \DOI{000}
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments suggest that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general, and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\maketitle
-% ==================================================================
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a quite difficult and very challenging task, mainly due to the edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics~\cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview we point the reader to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence (through eye-tracking experiments), that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing~\cite{DBLP:conf/gd/Purchase97,DBLP:journals/iwc/Purchase00}, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-Formally, the value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution\footnote{Also, refered to as \emph{crossing angle resolution} in the literature.}}.
-Analogously, the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (or \emph{RAC}, for short) graph; see Figure~\ref{fig:examples} for an illustration.
-Notably, RAC graphs are sparse (an $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}), while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2]{figures/examples}}
-	\caption{Illustration of: %
-	(a)~a RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, while the~second one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective.
-We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed algorithms~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows.
-Section~\ref{sec:relatedwork} overviews related works.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some Interesting Variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-In Appendix~\ref{app:grid}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-As test sets for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13} and the AT\&T graphs, which are two collections of around 8.100 and 424 benchmark graphs, respectively.
-
-The experiment was performed as follows.
-Initially, each graph of our expertiment was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic properties:
-%
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=10$.
-
-\subsection{The Rome Graph Test Set}
-
-In this subsection, we report the results of our expertimental evaluation for the non-planar Rome graphs, which form a collection of 424 benchmark graphs.
-Note that we did not impose any grid constraint on our algorithms.
-Our results are summarized in Figs.~\ref{fig:cr-res}, \ref{fig:an-res} and~\ref{fig:to-res}.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure*}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure*}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure*}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-\subsection{The AT\&T  Graph Test Set}
-\label{app:north}
-
-In this subsection, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure*}[htb]
+\begin{figure*}[t!]
 \centering
 \subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 
 \subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \caption{Experimental results for the crossing resolution on the AT\&T graphs.}
@@
 \end{figure*}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure*}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure*}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure*}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure*}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\subsection{Graph Drawing Contest 2017}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-\begin{table*}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table*}
-
-\subsection{Grid Restricted Drawings}
-\label{app:grid}
-
-\begin{figure*}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure*}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-As initial drawings for all algorithms, we used the layouts shown in Figure~\ref{fig:graph5And9-yfiles} that we computed with the SmartOrganic layouter of the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
-
-\begin{figure*}[htbp]
-\centering
-\subfloat[\label{fig:graph5-yfiles}{}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfil
-\subfloat[\label{fig:graph9-yfiles}{}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\caption{Illustration of 
-(a)~the 5th, and of
-(b)~the 9th graph of the Graph Drawing contest 2017.}
-\label{fig:graph5And9-yfiles}
-\end{figure*}
-
-\begin{figure*}[p]
-\centering
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
-(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure*}
-
-\begin{figure*}[htbp]
-\centering
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-total}
-}
-
-\subfloat[\label{fig:graph9-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-total}
-}
-
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
-(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(j)--(l)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure*}
-
-% ==================================================================
-\section{Conclusions}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\paragraph{Acknowledgments.} This project was supported by DFG grant KA812/18-1.
-The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
-
-\bibliographystyle{compj}
-\bibliography{references}
-
-\end{document}

### 2018-09-14 paper/journal.tex
Work on the journal version.

@@
 \documentclass{comjnl}
+
+%% These two lines are needed to get the correct paper size
+%% in TeX Live 2016
+\let\pdfpageheight\paperheight
+\let\pdfpagewidth\paperwidth
+
 \usepackage{amsmath,amssymb}
@@
 % ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
+\newcommand{\myparagraph}[1]{\medskip\noindent\textbf{#1}.}
 % ==================================================================
@@
 
+% ==================================================================
 \author{Michael~A.~Bekos} 
@@
 
-\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
+\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution%
+\thanks{A preliminary version of this work has appeared in the Proceedings of the 26th International Symposium on Graph Drawing and Network Visualization (GD 2018).
+This work is supported by the DFG grant Ka812/17-1}}
 
@@
 
-%\copyrightyear{2018} \vol{00} \issue{0} \DOI{000}
+\volumeyear{2018} 
+\volumenumber{--} 
+\issuenumber{-} 
 % ==================================================================
@@
 
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed algorithms~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
+Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs of general graphs.
+Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we followed a more practical approach.
+We implemented our algorithm and the force-directed ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms.
+Our evaluation indicates that our method significantly outperforms the aforementioned force-directed algorithms in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.
+Analogous results are obtained, when our algorithm and the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
+
+\paragraph{Preliminaries:}
 Unless otherwise specified, in this paper we consider simple undirected graphs.
@@
 Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
+
+\paragraph{Structure of the paper:}
 The remainder of this paper is structured as follows.
 Section~\ref{sec:relatedwork} overviews related works.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.
 
@@
 
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
+As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
+Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
 The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
+Actually, several relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
+Deciding, however, whether a $1$-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
 Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
+
 The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
+To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
 More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
@@
 
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
+An immediate observation emerging from the above overview is that the focus has been primarily on theoretical aspects of the problem.
+Most of the approaches that could be useful in practice are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants we point the reader to~\cite{DBLP:books/ph/BattistaETT99}.
+
+More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA},  to support conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
+%
+In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
+%
+However, the obtained drawings are not straight-line.
 For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
+However, its usefulness is only limited to small graphs and to a specific kind of drawings.
+
+Of more general scope are, of course, the already mentioned works by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, which are also based on the force-directed technique.
 
@@
 The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
+We assume that no two edges of graph $G$ overlap in drawing $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
+A circular drawing or a drawing obtained by applying a force-directed algorithm on graph $G$ clearly meets this precondition.
 
 Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
+At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of graph $G$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
+In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration of our algorithm (so, in a sense, our algorithm resembles probabilistic hill climbing approaches).
+We proceed by choosing a vertex of $\Gamma_{i-1}$ arbitrarily at random based on the content of the so-called \emph{vertex-pool}, which may contain:
+
+\begin{itemize}
+\item[--] either a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}, or
+\item[--] all vertices of $\Gamma_{i-1}$.
+\end{itemize}
 
@@
 To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
+A pair of edges $e$ and $e'$ is called \emph{critical} in drawing $\Gamma_{i-1}$ if and only if edges $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to the crossing resolution of drawing $\Gamma_{i-1}$ (that is, to $c(\Gamma_{i-1})$).
+The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
+
+The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a solution with good crossing resolution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex that is in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
+We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at graph-distance~$j$  in $G$, when $0 \leq i<j$.
+In other words, the closer a vertex is to the critical vertices, the more likely it is to be chosen.
+On the other hand, if the vertex-pool contains all vertices of the graph, then each vertex can be chosen with the same probability.
+
 What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
+However, by focusing only at the critical vertices, it is highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
+Therefore, special care is needed to avoid these bottlenecks, especially when the input graph is large.
+We will discuss ways to avoid local maxima later in this section (refer to subsection~\ref{ssec:bottlenecks}).
 
 So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
+Next, we describe how to compute the new position of the chosen vertex in the next drawing $\Gamma_i$.
+Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
+
+Let $v_i$ be the vertex of drawing $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
+To compute the position of vertex $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
+These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo} for an illustration.
 The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
 More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
+For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; we call point $\pi_j$ \emph{feasible} if and only if there is no vertex of drawing $\Gamma_{i-1}$ at point $\pi_j$ and the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of the graph in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$.
 
@@
 	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
+	\includegraphics[page=1]{figures/algorithm}}
 	\hfil
 	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
+	\includegraphics[page=2]{figures/algorithm}}
 	\caption{%
 	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
+	(a)~the chosen vertex is the white one;
 	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
 	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
+	among them, $\pi_4$ yields the best solution, and
+	(b)~the resulting drawing after moving the vertex at position $\pi_2$.}
 	\label{fig:algo}
@@
 
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
+If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., the same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
+If there is one or more feasible points, then one may consider two different approaches to determine the position of vertex $v_i$ in $\Gamma_i$.
 The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
+As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of vertex $v_i$ in $\Gamma_i$.
 We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
+The termination condition of our algorithm is simple and depends on an input parameter $\tau$.
+More specifically, if the crossing resolution has not improved during the last $\tau$ iterations, we assume that the algorithm has converged and we stop.
+
+% ==================================================================
+\subsection{Avoiding local maxima}
+\label{ssec:bottlenecks}
 % ==================================================================
@@
 The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
+What would one do to escape from a locally optimal solution?
+She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that by doing so a better solution will be easier to be computed afterwards.
+Our algorithm is mimicking this idea as follows:
+%
+\begin{itemize}
+\item[--] if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
+\item[--] afterwards, the vertex-pool switches back to the critical vertices.
+\end{itemize}
+
+From our extensive practical analysis, we quickly realized that, while this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm to proceed.
 
@@
 %
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
+\begin{itemize}
+\item[--] if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
+\item[--] afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
+\end{itemize}
+
+Note, however, that this approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
+
+% ==================================================================
+\subsection{Complexity issues}
+\label{ssec:complexity}
+% ==================================================================
+
 A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
+Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
 
@@
 Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
+Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i$ denotes the degree of vertex $v_i$.
+Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$ we compute the crossings and the corresponding crossing angles of edge $e_k$ with all other edges in $\Gamma_{i-1}$.
 Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
+Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that vertex $v_i$ is at position $\pi_j$.
+Let $\chi_j$ be the minimum crossing angle computed with this approach, when vertex $v_i$ is at position $\pi_j$.
+Clearly, $\pi_j$ is feasible, only if $\chi_j \geq \phi_i$.
+Note that the complexity of this approach is $O(\deg(v_i)m) = O(nm)$.
 
@@
 In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
+Towards this direction, we discuss in this section interesting variants of our algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:conclusions}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio (that is, the ratio of the longest to the shortest edge is large) and poor angular resolution (that is, the angles formed by adjacent edges are small).
+The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
+For the latter, observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
 
@@
 %
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
+Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge.
+Sometimes it is also used as a measure of the area of non-grid drawings.
 It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
@@
 %
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
+The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
 Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
+It was not difficult to adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
 In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
@@
 %
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
+As it is the case with the force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, our algorithm can also be restricted to maximize only the angular resolution (by neglecting its crossing resolution).
 We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
@@
 More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-In Appendix~\ref{app:grid}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
+One can even bound the size of the grid, by rejecting candidate grid positions that lie outside the permitted bounds.
+
+Note that an alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
+However, critical in this step is not to affect the crossing resolution of the computed drawing too much during its conversion to grid.
+It is worth noting that we tried to apply such a technique (using again randomization).
+Our idea was to choose uniformly at random, for each vertex of the non-grid starting drawing, one of the grid points from its \emph{neighbourhood}, and move it to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of the vertex was initially set to its four closest grid points.
+If after a number of iterations a grid drawing without vertex-vertex overlaps and of similar crossing resolution as the one of the starting drawing could not be reported, then we had to augment the neighbourhood of each vertex so to contain more grid points (hoping that by this augmentation the computation of an acceptable grid drawing will be become feasible).
+
+However, we faced two main difficulties when adopting this alternative approach.
+First, we could not always guarantee that the crossing resolution of the final grid drawing will be similar to the one of the starting drawing without increasing the drawing area too much.
+The second one is that the conversion itself is an extra step in the algorithm, which some times was more demanding than the main step of the algorithm, which was the actual computation of the drawing with high crossing resolution.
+So, we decided not to adopt this alternative approach.
 
@@
 In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
+For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
+The implementations\footnote{All implementations are available on request by the authors.} were in Java using the yFiles graph drawing library~\cite{DBLP:books/sp/04/WieseE004}.
+
+% ==================================================================
+\subsection{Experiment's Setup}
+\label{ssec:setup}
+% ==================================================================
+
+As test sets for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13} and the non-planar AT\&T graphs (also known as Graph Catalog or North graphs; available at \texttt{http://graphdrawing.org/data}), which are two collections of around 8.100 and 424 benchmark graphs, respectively, that are widely used in Graph Drawing for the evaluation of different drawing algorithms.
+
 The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-As test sets for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13} and the AT\&T graphs, which are two collections of around 8.100 and 424 benchmark graphs, respectively.
-
 The experiment was performed as follows.
 Initially, each graph of our expertiment was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic properties:
-%
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=10$.
-
+Starting from this layout, every graph was drawn with:
+%
+\begin{itemize}
+\item[--] our algorithm with $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=10$,
+\item[--] our algorithm restricted not to violate the aspect ratio of the initial layout, 
+\item[--] the force-directed algorithm by Argyriou et al., and
+\item[--] the force-directed algorithm by Huang et al.
+\end{itemize}
+
+Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of them to maximize exclusively the corresponding measures.
+In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we discussed in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
+The reported results, illustrated in Figures~\ref{fig:cr-res}--\ref{fig:northAngular}, are on average across different drawings with same number of vertices.
+
+% ==================================================================
 \subsection{The Rome Graph Test Set}
-
-In this subsection, we report the results of our expertimental evaluation for the non-planar Rome graphs, which form a collection of 424 benchmark graphs.
+\label{ssec:rome}
+% ==================================================================
+
+In this subsection, we report the results of our expertimental evaluation for the non-planar Rome graphs.
 Note that we did not impose any grid constraint on our algorithms.
-Our results are summarized in Figs.~\ref{fig:cr-res}, \ref{fig:an-res} and~\ref{fig:to-res}.
+Our results are summarized in Figures~\ref{fig:cr-res}, \ref{fig:an-res} and~\ref{fig:to-res}.
 
@@
 
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure*}
+% ==================================================================
+\input{cr-res-rome.tex}
+% ==================================================================
+
+% ==================================================================
+\input{to-res-rome.tex}
+% ==================================================================
+
+% ==================================================================
+\input{an-res-rome.tex}
+% ==================================================================
+
+% ==================================================================
+\input{cr-res-north.tex}
+% ==================================================================
+
+% ==================================================================
+\input{to-res-north.tex}
+% ==================================================================
+
+% ==================================================================
+\input{an-res-north.tex}
+% ==================================================================
+
+% ==================================================================
+\input{grid-rome.tex}
+% ==================================================================
 
@@
 For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
+Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
+On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) as the one of Huang et al.
 
@@
 We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
+In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
+Observe that even under this different convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
+
+The maximum number of iterations that each of the algorithms of the experiment could perform in order to converge was set to 100.000, but notably this limit was never reached.
+
+We observe that both force-directed algorithms by Argyriou et al.\ and by Huang et al.\ seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is more likely to be computed.
+However, that for larger graphs the algorithm by Huang et al.\ requires the least amount of iterations.
+Regarding the unrestricted and the restricted variants of our algorithm, it is clear that both require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
 
 \myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
+Here, each algorithm was adjusted to maximize both the crossing and the angular resolution, simultaneously.
 For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield drawings with better total resolution than the corresponding ones by Argyriou et al.
@@
 
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure*}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
+Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of this measure) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
+%More concretely, the 
+The drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
 Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
@@
 
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
+Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
+The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings.
 Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
+Our unrestricted algorithm requires the largest number of crossings, which turns out to be notably higher than the corresponding ones of the other three algorithms.
+
+On the negative side, both the unrestricted and the restricted variants of our algorithm require more iterations than the force-directed algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:to-res-4}.
 Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
@@
 
-\begin{figure*}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure*}
-
 The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
@@
 
+% ==================================================================
 \subsection{The AT\&T  Graph Test Set}
-\label{app:north}
-
-In this subsection, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
+\label{ssec:north}
+% ==================================================================
+
+In this subsection, we report the results of our experimental evaluation for the non-planar AT\&T graphs.
+Note that again we did not impose any grid constraint on our algorithms.
 The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
 In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure*}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure*}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
+This manifests in spikes of large magnitude in the illustrations of the results indicates that the structural properties of the graphs in this second test set varies vastly between graphs of different sizes.
+
+\myparagraph{Crossing resolution} As with the Rome graphs, we observed again that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
 Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
+The results actually indicate that one can classify these graphs into ``hard'' and ``easy'' ones, when maximizing their crossing resolution.
+In particular, the graphs with 50 to 70 vertices appear to be harder than the graphs with 70 to 80 vertices.
+As expected, the aspect ratio of the drawings produced by both the unrestricted and the restricted variants of our algorithm depends vastly on the number of vertices of the graphs.
+However, this natural behaviour cannot be observed in the resulting drawings of the two force-directed algorithms, which appear to be quite unstable resulting in a large variance.
 Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
@@
 In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure*}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure*}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
+converges using the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
+
+\myparagraph{Total resolution} For small graphs, we observed similar results as in the experiment on the Rome graphs, that is, our unrestricted algorithm outperforms (in terms of the total resolution) the other three algorithms of the experiment, while the restricted variant of our algorithm is slightly better than the algorithms by Argyriou et al.\ and by Huang et al.; see Fig.~\ref{fig:northTotal-1}.
+For larger graphs, however, the two variants of our algorithm achieve similar results while still outperforming the algorithm by Huang et al.
+Regarding the aspect ratio and the number of crossings of the produced drawings, we observe that all algorithms have more or less the same behaviour as when adjusted
+
+The results for the aspect ratio and number of crossings are similar to those observed  the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
 performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
@@
 %Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure*}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure*}
 
@@
 
+% ==================================================================
+\subsection{Grid Restricted Drawings}
+\label{ssec:grid}
+% ==================================================================
+
+In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
+In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
+%
+\begin{inparaenum}[(i)]
+\item $10^6 \times 10^6$
+\item $10^4 \times 10^4$
+\item $10^3 \times 10^3$, and
+\item $10^2 \times 10^2$.
+\end{inparaenum}
+%
+The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
+However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
+The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
+
+Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
+As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
+Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
+%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
+%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
+For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
+
+The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
+
+As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
+As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
+It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
+Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
+%This can partially be explained with the different choice of the initial drawing.
+%
+On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+
+In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
+However, the computation of a grid drawing takes longer, which is of course expected.
+%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
+Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
+%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
+Further experiments in this direction are needed.
+
+% ==================================================================
 \subsection{Graph Drawing Contest 2017}
-\label{app:contest}
+\label{ssec:contest17}
+% ==================================================================
 
@@
 
-\subsection{Grid Restricted Drawings}
-\label{app:grid}
-
-\begin{figure*}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure*}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
 \section{Sample Drawings}
@@
 \section{Conclusions}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
+\label{sec:conclusions}
+% ==================================================================
+
+%
+In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
 Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
+%As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+
 While working on this project, we made some useful observations and  obtained some interesting insights.
@@
 
-\paragraph{Acknowledgments.} This project was supported by DFG grant KA812/18-1.
-The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
+\paragraph{Acknowledgments.} The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
 

### 2018-09-16 paper/journal.tex
Working a bit more on the journal version

@@
 To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
+The first mimics the human behavior.
 What would one do to escape from a locally optimal solution?
@@
 For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
+Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
 On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) as the one of Huang et al.
@@
 
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
+A different behavior can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
 We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
@@
 \myparagraph{Crossing resolution} As with the Rome graphs, we observed again that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
+Remarkable is the synchronous behavior of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
 The results actually indicate that one can classify these graphs into ``hard'' and ``easy'' ones, when maximizing their crossing resolution.
@@
 As expected, the aspect ratio of the drawings produced by both the unrestricted and the restricted variants of our algorithm depends vastly on the number of vertices of the graphs.
-However, this natural behaviour cannot be observed in the resulting drawings of the two force-directed algorithms, which appear to be quite unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
+However, this natural behavior cannot be observed in the resulting drawings of the two force-directed algorithms, which appear to be quite unstable resulting in a large variance.
+Again the restricted variant of our algorithm and the two force-directed algorithms produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
 All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges using the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
+In terms of the number of iterations need for convergence, we observe that somewhat surprisingly the algorithm of Argyriou et al.\ converges using the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
 
 \myparagraph{Total resolution} For small graphs, we observed similar results as in the experiment on the Rome graphs, that is, our unrestricted algorithm outperforms (in terms of the total resolution) the other three algorithms of the experiment, while the restricted variant of our algorithm is slightly better than the algorithms by Argyriou et al.\ and by Huang et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, the two variants of our algorithm achieve similar results while still outperforming the algorithm by Huang et al.
-Regarding the aspect ratio and the number of crossings of the produced drawings, we observe that all algorithms have more or less the same behaviour as when adjusted
+For larger graphs, however, the two variants of our algorithm achieve similar results, while still outperforming the algorithm by Huang et al.
+Regarding the aspect ratio and the number of crossings of the produced drawings, we observe that all algorithms have more or less the same behavior as when adjusted for maximizing the crossing resolution only.
+The only exception can be observed in the algorithm by Huang et al., which performs more stable; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
+With respect to the number of iterations required for convergence, the two variants of our algorithms and the force-directed algorithm by Argyriou et al.\ show similar behavior.
+The algorithm by Huang et al., on the other hand, needs less iterations for convergence; see Fig.~\ref{fig:northTotal-4}.
+Note, however, that there is no immediate correlation with the number of vertices of the graph.
 
-The results for the aspect ratio and number of crossings are similar to those observed  the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
+\myparagraph{Angular resolution} Here, the picture is not so clear concerning the ranking of the algorithms; see Fig.~\ref{fig:northAngular-1}.
+Only the algorithm by Huang et al.\ seems to be mostly at last in the rank.
+Concerning the aspect ratio, we see very good behavior for the restricted variant of our algorithm and the algorithm by Huang et al., while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
+Regarding the number of crossings, we again observe that all algorithms have similar behavior.
+Note, however, that the unrestricted variant of our algorithm and the algorithm by Argyriou et al.\ yield drawings with more crossing, especially for graphs with large number of vertices; see Fig.~\ref{fig:northAngular-3}.
+Finally, both variants of our algorithm and the force-directed algorithm by Huang et al.\ need comparable number of iterations for convergence, which most of the times is lower than the corresponding one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
 
@@
 
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
+In addition to the experiments that we have described so far, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
+In particular, we were interesting to see how the different quality measures that we have evaluated so far are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
+
+\begin{enumerate}[(i)]
 \item $10^6 \times 10^6$
@@
 \item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
+\end{enumerate}
 
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
+The test suite for this experiment was the non-planar Rome graphs.
+The algorithm to be evaluated was the variant of our algorithm that optimizes the drawing's crossing resolution.
+However, since this algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
+Fig.~\ref{fig:experimentsGrid} summarizes the results.
+
+Regarding the crossing resolution, we natural observe that the bigger the grid is, the easier it is to compute a drawing with higher crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
 As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
+Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the underlying grid was of size $10^3 \times 10^3$.
+The crossing resolution of the drawings produced by our algorithm, when the underlying grid was of size $10^4 \times 10^4$, were on average worse by only $10^\circ$ than those produced by our algorithm, when imposing no restriction on the size of the underlying grid (refer to the double line in Fig.~\ref{fig:crossingResolutionGrid}).
+As expected, the gap was more narrow, when the underlying grid was of size $10^6 \times 10^6$.
 
@@
 As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+As a result, the drawings produced by our algorithm, when imposing no restriction on the size of the underlying grid, are clearly better (in terms of the number of crossings) than those produced by our algorithm, when the size of the underlying grid is limited.
+As a matter of fact, the larger the underlying grid is, the fewer crossing the reported drawings have, and the differences are quite clear.
+Exceptional are the grids of sizes $10^4 \times 10^4$ and $10^6 \times 10^6$, where the improvement (in the number of crossings) from the latter to the former is small on average; this can partially be explained by the different choices of the initial drawings.
 
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
+In general, the number of iterations needed for convergence increases with the size of the underlying grid; see Fig.~\ref{fig:crossingResolutionTimeGrid}.
+Exceptional seems to be the $10^2 \times 10^2$ grid, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
+
+In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution, when restricted to a grid, as long as the grid is not too small.
 However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
+Finally, the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution), but also with respect to the number of iterations needed for convergence.
 

### 2018-09-17 src/main/ContestWindow.java
Add scrollpane, visualize button, move best button to only display angle

@@ -18,6 +18,8 @@ import static main.GraphFrame.createSeparator;
 
 public class ContestWindow {
     private JFrame contestWindow;
+    private JPanel contestWindowPanel;
+    private JScrollPane scrollPane;
     private GridBagConstraints cFrame;
 
     private ArrayList<GraphFrame> graphFrames;
@@ -139,7 +141,6 @@ public class ContestWindow {
 
     public ContestWindow() {
         this.contestWindow = new JFrame("Contest Window");
-        this.contestWindow.setLayout(new GridBagLayout());
         this.contestWindow.setDefaultCloseOperation(JFrame.DO_NOTHING_ON_CLOSE);
         this.contestWindow.addWindowListener(new WindowAdapter() {
             @Override
@@ -152,9 +153,18 @@ public class ContestWindow {
                 System.exit(0); //maybe nicer way
             }
         });
-        this.contestWindow.setMinimumSize(new Dimension(700,700));
+        this.contestWindow.setSize(new Dimension(900,600));
         this.contestWindow.setVisible(true);
-//        this.contestWindow.pack();
+        this.contestWindow.revalidate();
+        this.contestWindow.repaint();
+
+        this.contestWindowPanel = new JPanel();
+        this.contestWindowPanel.setLayout(new GridBagLayout());
+        this.scrollPane = new JScrollPane(contestWindowPanel);
+        this.scrollPane.getVerticalScrollBar().setUnitIncrement(16);
+        this.contestWindow.getContentPane().add(scrollPane, BorderLayout.CENTER);
+
+
 
         this.cFrame = new GridBagConstraints();
         this.cFrame.gridx = 0;
@@ -164,7 +174,7 @@ public class ContestWindow {
 
 
         //TODO align columns
-        this.contestWindow.add(createLegend(), this.cFrame);
+        this.contestWindowPanel.add(createLegend(), this.cFrame);
 
         graphFrames = new ArrayList<>();
 
@@ -204,7 +214,13 @@ public class ContestWindow {
         cFramePanel.gridx += 1;
         panel.add(createSeparator(), cFramePanel);
 
-        JLabel timeElapsed = new JLabel("Time Elapsed (s)");
+        JLabel showBest = new JLabel("Best Angle");
+        cFramePanel.gridx += 1;
+        panel.add(showBest, cFramePanel);
+        cFramePanel.gridx += 1;
[…]

### 2018-09-17 src/main/GraphFrame.java

@@ -21,6 +21,7 @@ public class GraphFrame {
     public JPanel panel;
     public MainFrame mainFrame;
     public double minAngle;
+    public double bestAngle;
     public String fileName;
     public String folderPath;
     public String savePath;
@@ -28,6 +29,7 @@ public class GraphFrame {
     private JButton startPauseButton;
     private JButton stopButton;
     private JLabel minAngleLabel;
+    private JLabel bestAngleLabel;
     private JLabel timeLabel;
     private JCheckBox allowDecreasing;
     private double runningTime;
@@ -88,6 +90,11 @@ public class GraphFrame {
         cFramePanel.gridx += 1;
         panel.add(createSeparator(), cFramePanel);
 
+        bestAngleLabel = new JLabel("-1");
+        cFramePanel.gridx += 1;
+        panel.add(bestAngleLabel, cFramePanel);
+        cFramePanel.gridx += 1;
+        panel.add(createSeparator(), cFramePanel);
 
         timeLabel = new JLabel("-1");
         cFramePanel.gridx += 1;
@@ -103,12 +110,6 @@ public class GraphFrame {
         cFramePanel.gridx += 1;
         panel.add(createSeparator(), cFramePanel);
 
-        JButton showBest = new JButton("Best");
-        cFramePanel.gridx += 1;
-        panel.add(showBest, cFramePanel);
-        showBest.addActionListener(this::showBestActionPerformed);
-        cFramePanel.gridx += 1;
-        panel.add(createSeparator(), cFramePanel);
 
         startPauseButton = new JButton("Start");
         cFramePanel.gridx += 1;
@@ -124,6 +125,16 @@ public class GraphFrame {
         cFramePanel.gridx += 1;
         saveButton.addActionListener(this::saveGraphActionPerformed);
         panel.add(saveButton, cFramePanel);
+
+        cFramePanel.gridx += 1;
+        panel.add(createSeparator(), cFramePanel);
+
+        JCheckBox allowDecreasing = new JCheckBox("Visualize");
+        cFramePanel.gridx += 1;
+        panel.add(allowDecreasing, cFramePanel);
+        allowDecreasing.addItemListener(this::visualizeEnabled);
+        allowDecreasing.setSelected(false);
+
     }
 
     public  static JComponent createSeparator() {
@@ -236,13 +247,17 @@ public class GraphFrame {
     }
 
     public void updateTime() {
[…]

### 2018-09-17 paper/an-res-north.tex

--- 2018-08-30 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-%\linenumbers
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some interesting variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
-
-The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\columnbreak
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-%\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
-
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\arxapp{\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure}[htb]
+\begin{figure*}[t!]
 \centering
@@
 \label{fig:northAngular}
-\end{figure}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\begin{table}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure}
-
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure}}{}
-
-\end{document}
+\end{figure*}

### 2018-09-17 paper/an-res-rome.tex

--- 2018-08-30 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-%\linenumbers
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some interesting variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
-
-The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\columnbreak
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-%\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
-
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
 \subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 
 \subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \caption{Experimental results on the angular resolution for the Rome graphs.}
 \label{fig:an-res}
-\end{figure}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\arxapp{\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\begin{table}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure}
-
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure}}{}
-
-\end{document}
+\end{figure*}

### 2018-09-17 paper/cr-res-rome.tex

--- 2018-08-30 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-%\linenumbers
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some interesting variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
-
-The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\columnbreak
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-%\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
-
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
 \subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 
 \subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \caption{Experimental results on the crossing resolution for the Rome graphs.}
 \label{fig:cr-res}
-\end{figure}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\arxapp{\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\begin{table}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure}
-
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure}}{}
-
-\end{document}
+\end{figure*}

### 2018-09-17 paper/to-res-north.tex

--- 2018-08-30 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-%\linenumbers
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some interesting variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
-
-The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\columnbreak
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-%\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
-
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Total Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Total Resolution rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Total Resolution rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Total Resolution fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-totalResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {totalResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{totalResolution.csv};
-\addplot table [x=Nodes, y=Mixed, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the total resolution for the Rome graphs.}
-\label{fig:to-res}
-\end{figure}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\arxapp{\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[tb]
+\begin{figure*}[t!]
 \centering
 \subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 
 \subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \caption{Experimental results for the total resolution on the AT\&T graphs.}
 \label{fig:northTotal}
-\end{figure}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\begin{table}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure}
-
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure}}{}
-
-\end{document}
+\end{figure*}

### 2018-09-17 paper/to-res-rome.tex

--- 2018-08-30 paper/paper.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
-\usepackage[font=small]{subfig,caption}
-\usepackage{algorithm2e}
-\usepackage{todonotes,lineno,paralist,soul}
-\usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
-\graphicspath{{figures/}}
-
-% ==================================================================
-% Format tiks and pgfplots
-\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]
-\pgfplotsset{ytick style={draw=none}}
-\renewcommand{\axisdefaultheight}{185pt}
-\pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
-{black,double,smooth}
-}
-\pgfplotsset{
-        table/search path={data},
-    }
-\pgfplotsset{
-  /pgfplots/xlabel near ticks/.style={
-     /pgfplots/every axis x label/.style={
-        at={(ticklabel cs:0.5)},anchor=near ticklabel
-     }
-  },
-  /pgfplots/ylabel near ticks/.style={
-     /pgfplots/every axis y label/.style={
-        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
-     }
-}
-% ==================================================================
-\newcommand{\myparagraph}[1]{\smallskip\noindent\textbf{#1}.}
-
-%conf for conference version, arxiv for arxiv version
-\newcommand{\ver}{conf}
-\newcommand{\arxapp}[2]{\ifthenelse{\equal{\ver}{conf}}{#2}{#1}}
-
-\author{Michael~A.~Bekos, Henry~F\"orster, Christian~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~M.~Spallek, Jan~Splett}
-
-\authorrunning{M.~A.~Bekos et al.}
-\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
-\titlerunning{Drawings of Graphs with High Crossing Resolution}
-
-\institute{
-Institut f\"ur Informatik, Universit\"at T\"ubingen, T\"ubingen, Germany\\
-\texttt{\{bekos,foersth,geckeler,mk\}@informatik.uni-tuebingen.de}\\
-\texttt{\{jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
-}
-
-% ==================================================================
-\begin{document}
-\maketitle
-%\linenumbers
-% ==================================================================
-
-\begin{abstract}
-The \emph{crossing resolution} of a non-planar drawing of a graph is the value of the minimum angle formed by any pair of crossing edges.
-Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph.
-However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.
-
-In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
-\end{abstract}
-
-% ==================================================================
-\section{Introduction}
-\label{sec:introduction}
-% ==================================================================
-
-In Graph Drawing, there exists a rich literature and a wide range of~techniques for drawing planar graphs; see, e.g.,~\cite{DBLP:journals/combinatorica/FraysseixPP90,DBLP:conf/gd/GutwengerM98,DBLP:journals/algorithmica/Kant96}.
-However, drawing a non-planar graph, and in particular when it does not have some special structure (e.g., degree restriction), is a difficult and challenging task, mainly due to the~edge crossings that negatively affect the drawing's quality~\cite{DBLP:journals/iwc/Purchase00}.
-As a result, the established techniques are significantly fewer (e.g., crossing minimization heuristics \cite{DBLP:journals/algorithmica/EadesW94,DBLP:journals/tsmc/SugiyamaTT81}, energy-based layout algorithms~\cite{DBLP:journals/congnum/Eades84,DBLP:journals/spe/FruchtermanR91}); for an overview refer to~\cite{DBLP:books/ph/BattistaETT99,DBLP:conf/dagstuhl/1999dg,DBLP:reference/crc/2013gd}.
-
-In this context, Huang et al.~\cite{DBLP:conf/apvis/Huang07,DBLP:journals/vlc/HuangEH14} a decade ago introduced some important experimental evidence,
-% (through eye-tracking experiments),
-that edge crossings may not negatively affect the drawing's quality too much (and hence the human's ability to read and interpret it), when the angles formed by the crossing edges are large.
-In other words, while prior to these experiments it was commonly accepted that mainly the number of crossings is the most important parameter for judging the quality of a non-planar graph drawing, it turned out that the types of edge crossings also matter.
-As a result, a new and prominent research direction was initiated, recognized under the term ``beyond planarity''~\cite{Shonan2016,Dagstuhl2016,SoCG2017}, which focuses on graphs and their properties, when different constraints on the types of edges crossings are imposed; refer to~\cite{DBLP:journals/corr/abs-1804-07257} for a recent survey.
-
-The value of the minimum angle formed by any two crossing edges in a drawing is referred to as its \emph{crossing resolution}; the crossing resolution of a graph is defined as the maximum crossing resolution over all its drawings.
-Clearly, the crossing resolution of a non-planar graph is at most $90^\circ$, while a graph that admits a drawing with crossing resolution $90^\circ$ is called \emph{right-angle-crossing} (\emph{RAC}) graph; see Figure~\ref{fig:examples}.
-Notably, RAC graphs are sparse with at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:k5} {}]{
-	\includegraphics[page=1,scale=0.85]{figures/examples}}
-	\hfil
-	\subfloat[\label{fig:k6} {}]{
-	\includegraphics[page=2,scale=0.85]{figures/examples}}
-	\caption{%
-	(a)~A RAC drawing of the complete graph $K_5$, and
-	(b)~a drawing of the complete graph $K_6$, whose crossing resolution is arbitrarily close to $90^\circ$.}
-	\label{fig:examples}
-\end{figure}
-
-The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-Also, the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
-
-From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}.
-The first one is due to Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, while the~second one is due to Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges.
-Each of them uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea of both is the same: the smaller the crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.
-
-In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method,
-% for computing drawings with high crossing resolution,
-which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing.
-How would one increase the crossing resolution of a given drawing?
-First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase.
-Of course, we cannot consider all possible positions for the vertex to be moved.
-Instead, we consider a small set of randomly generated ones.
-If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this~position.
-
-In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs.
-Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings,
-%(mainly due to the nature of the problem itself),
-we followed a more practical approach.
-We implemented our algorithm and the force-directed ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs.
-% (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms.
-Our evaluation indicates that our method significantly outperforms the force-directed ones~\cite{DBLP:journals/cj/ArgyriouBS13,DBLP:journals/vlc/HuangEHL13} in terms of crossing resolution, but this comes at the cost of slightly worse running time
-%, especially
-for large and dense graphs.
-Analogous results are obtained, when our algorithm and the ones of~\cite{DBLP:journals/cj/ArgyriouBS13} and~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize  the \emph{angular resolution} (i.e., the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (i.e., the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).
-
-\medskip\noindent{\it Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
-Let $G=(V,E)$ be such a graph.
-The degree of vertex $u\in V$ of $G$ is denoted by $d(u)$.
-The degree $d(G)$ of  graph $G$ is defined as the maximum degree of its vertices, i.e., $d(G)=\max_{u\in V}d(u)$.
-%
-Given a drawing $\Gamma(G)$ of $G$, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$.
-%For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.
-
-\medskip\noindent{\it Structure of the paper:}
-The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works.
-%In Section~\ref{sec:preliminaries}, we introduce preliminary notions and the notation used throughout the paper.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-\arxapp{In the appendix}{In~\cite{arxivVersion}}, we provide experimental results on grid restricted drawings, on more test sets and on the graphs from the Graph Drawing Competition in 2017.
-
-% ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
-% ==================================================================
-
-As already mentioned, the study of the crossing resolution maximization problem has mainly focused on its optimal case, i.e., on the study of RAC graphs.
-An $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}.
-The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge.
-Actually, several~relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}.
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.
-The results are fewer, when the right-angle constraint is relaxed.
-Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph with crossing resolution at least $\alpha$ radians, has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results are also known in the presence of bends~\cite{DBLP:journals/siamdm/AckermanFT12,DBLP:journals/mst/GiacomoDLM11}.
-
-An immediate observation emerging from the above overview is that the~focus has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practice are based on force-directed techniques~\cite{DBLP:books/ph/BattistaETT99,DBLP:journals/congnum/Eades84}.
-COWA is a system that supports conceptual web site traffic analysis~\cite{DBLP:conf/apvis/DidimoLR10}; its algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe topology-driven force-directed heuristics to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency; the obtained drawings, however, are not straight-line.
-%
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-program to increase the crossing angles of circular drawings.
-%
-Of more general scope are the already mentioned force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-
-%As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
-Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges.
-Angelini et al.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11} presented results for two interesting variants, in which the input graph is either an acyclic digraph or a graph of bounded vertex degree.
-Argyriou et al.~\cite{DBLP:journals/jgaa/ArgyriouBS12} showed that the problem of deciding whether a graph is RAC is NP-hard.
-Angelini et al.~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11} observed that planar graphs of bounded degree admit straight-line RAC drawings in subquadratic area.
-Didimo et al.~\cite{DBLP:journals/ipl/DidimoEL10} characterized the complete bipartite RAC graphs.
-Di~Giacomo et al.~\cite{DBLP:journals/algorithmica/GiacomoDEL14}, and Hong and Nagamochi~\cite{DBLP:conf/wg/HongN15} studied variants, in which the vertices of the input graph are restricted on two parallel lines and on the circumference of a circle, respectively.
-Eades and Liotta~\cite{DBLP:journals/dam/EadesL13} showed that the optimal RAC graphs (i.e., those of maximum density) are 1-planar (i.e., they admit drawings in which each edge is crossed at most once).
-Deciding, however, whether a 1-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}.
-Other relationships between the class of RAC graphs and subclasses of 1-planar graphs have also been studied by Bachmaier et al.~\cite{DBLP:journals/dam/BachmaierBHNR17} and Brandenburg et al.~\cite{DBLP:journals/tcs/BrandenburgDEKL16}.
-Note that the problem of finding RAC drawings has been also studied in the presence of bends; see e.g.~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11}.
-
-%To the best of our knowledge, there is only one work, by Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11}, which studies the crossing resolution maximization problem by relaxing the right-angle constraint on the crossing angles of the computed drawings.
-More precisely, in their work, Dujmovic et al.~\cite{DBLP:journals/cjtcs/DujmovicGMW11} proved that an $n$-vertex graph, whose crossing resolution is at least $\alpha$ (in radians) has at most $(3n-6)\pi/\alpha$ edges.
-Corresponding density results for the case, in which few bends are allowed along each edge, are known by Ackerman et al.~\cite{DBLP:journals/siamdm/AckermanFT12} and Di Giacomo et al.~\cite{DBLP:journals/mst/GiacomoDLM11}.
-
-%An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem.
-Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}.
-%Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.
-%
-%More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution.
-%
-%In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency.
-%
-%However, the obtained drawings are not straight-line.
-For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings.
-%
-%Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.\, which are also based on the force-directed technique.
-
-% ==================================================================
-\section{Description of our Heuristic Approach}
-\label{sec:algorithm}
-% ==================================================================
-
-In this section, we describe our heuristic for obtaining drawings with high crossing resolution.
-The input of our heuristic consists of a graph $G$ and an initial drawing $\Gamma_0$ of $G$ with crossing resolution $c(\Gamma_0)$.
-We assume that no two edges of~$G$ overlap in $\Gamma_0$, i.e., $c(\Gamma_0)>0$.
-A circular drawing or a drawing obtained by applying a force-directed algorithm on $G$ clearly meets this precondition.
-
-Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-%, where $\Gamma_0$ is our initial drawing.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration.
-Then, a vertex of $\Gamma_{i-1}$ is chosen arbitrarily at random based on the so-called \emph{vertex-pool}, which may contain:
-%
-%\begin{itemize}
-\begin{inparaenum}[(i)]
-\item either all vertices of $\Gamma_{i-1}$, or
-\item a prespecified subset of the vertices of $\Gamma_{i-1}$, called \emph{critical}.
-\end{inparaenum}
-
-Intuitively, the critical vertices are the endpoints of the edges that define the crossing resolution of drawing $\Gamma_{i-1}$.
-To formally define them, we first need to introduce the notion of critical edge-pairs.
-A pair of edges $e$ and $e'$ is called \emph{critical} in $\Gamma_{i-1}$, if $e$ and $e'$ cross in $\Gamma_{i-1}$ and the minimum angle that is formed at their crossing point is equal to $c(\Gamma_{i-1})$.
-The set of critical vertices of $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
-
-The role of critical vertices is central in our algorithm\footnote{If the focus is not on the critical vertices for a large graph, then our algorithm will need a large number of iterations to converge to a good solution, because it is simply very unlikely to select to move one of the vertices that define the crossing resolution.}: By appropriately changing the location of a critical vertex or of a vertex in the neighbourhood of the critical vertices, we naturally expect to improve the crossing resolution of the current drawing.
-We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at distance $j$  in the graph, when $0 \leq i<j$.
-So, if the vertex-pool contains only critical vertices, then the closer a vertex is to the critical vertices, the more likely it is to be chosen.
-Otherwise, the vertex-pool contains all vertices and each vertex can be chosen with the same probability.
-
-%With this procedure, what we quickly realized from our experimental evaluation, is that by appropriately changing the location of a critical vertex at each iteration of our algorithm, the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-What we quickly realized from our practical analysis, is that the crossing resolution of the initial drawing improves rapidly during the first iterations of the algorithm.
-However, by focusing only at the critical vertices, it is~highly possible that the algorithm will get trapped to some local maxima after a number of iterations.
-So, special care is needed to avoid these bottlenecks, especially when the input graph is large.
-We will discuss ways to avoid them later in this section.
-
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
-Next, we describe how to compute its new position in the next drawing.
-Note that our method resembles probabilistic hill climbing approaches.
-
-Let $v_i$ be the vertex of $\Gamma_{i-1}$ that has been chosen to be moved at the $i$-th iteration.
-%Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$.
-To compute the position of $v_i$ in the next drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from $p(v_i)$ in $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm.
-These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$; see Fig.~\ref{fig:algo}.
-The position of vertex $v_i$ in $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$.
-More precisely, for each ray $r_i$ we choose a distance value $\delta_i$ uniformly at random from the interval $[\delta_{min},\delta_{max}]$, where $\delta_{min}$ and $\delta_{max}$ are two positive parameters of the algorithm.
-For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ is obtained by translating $p(u)$ along $r_j$ by a distance $\delta_j$; point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at $\pi_j$ and by keeping all other vertices of $G$ in their positions in $\Gamma_{i-1}$ is at least as large as the crossing resolution of $\Gamma_{i-1}$, and there is no vertex of $\Gamma_{i-1}$ at $\pi_j$.
-
-\begin{figure}[t!]
-	\centering
-	\subfloat[\label{fig:algo-rays} {}]{
-	\includegraphics[page=1, scale=0.9]{figures/algorithm}}
-	\hfil
-	\subfloat[\label{fig:algo-move} {}]{
-	\includegraphics[page=2, scale=0.9]{figures/algorithm}}
-	\caption{%
-	Illustration of an iteration step of our algorithm:
-	(a)~The chosen vertex is the white one;
-	the computed rays $r_0,\ldots,r_7$ have been rotated by $8^\circ$;
-	the black-colored points along these rays are points $\pi_0,\ldots,\pi_7$;
-	among them, $\pi_4$ yields the best solution.
-	(b)~The resulting drawing after moving the vertex at position $\pi_2$.}
-	\label{fig:algo}
-\end{figure}
-
-If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the position of $v_i$ in $\Gamma_i$ is $p(v_i)$, i.e., same as in $\Gamma_{i-1}$, since $c(\Gamma_i) \geq c(\Gamma_{i-1})$ must hold.
-If there is one or more feasible points, then one may consider two different approaches to determine the position of $v_i$ in $\Gamma_i$.
-The most natural is to choose the feasible point that maximizes the crossing resolution of the obtained drawing.
-As an alternative, one may rely again on randomization and chose uniformly at random one of the feasible points as the position of $v_i$ in $\Gamma_i$.
-We note that we did not observe any significant difference between these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one.
-The termination condition of our algorithm is simple and depends on an in\-put parameter~$\tau$.
-More specifically, if the crossing resolution has not improved~during the last $\tau$ iterations, we assume that the algorithm has converged and we~stop.
-
-% ==================================================================
-\myparagraph{Avoiding local maxima}
-% ==================================================================
-%
-To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$.
-The first mimics the human behaviour.
-What would~one do to escape from a locally optimal solution?
-She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that
-%by changing the embedding of the graph or its crossing structure,
-by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows:
-%
-\begin{inparaenum}[(i)]
-%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices.
-\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} by including all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
-\item afterwards, the vertex-pool switches back to the critical vertices.
-\end{inparaenum}
-%
-%From our extensive practical analysis, we quickly realized that, while
-While this approach turned out to be effective for smaller graphs, for graphs with more than 100 vertices, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.
-
-%What we observed is that in larger graphs it was unlikely to change the embedding of the graph in a beneficial way for the algorithm to proceed, mainly because of the random choice of the vertices to be moved; in most iterations with the wider vertex-pool, the vertices chosen to be moved were not in the part of the drawing where the critical vertices resided.
-
-Our second approach is based on parameters $\rho$, $\delta_{min}$ and $\delta_{max}$ of the algorithm.
-Our idea was that if the algorithm gets trapped to a locally optimal solution, then a ``drastic'' or ``sharp'' move may help to escape.
-We turned this idea into an algorithmic implementation as follows:
-%
-\begin{inparaenum}[(i)]
-\item if during the last $\zeta$ iterations the crossing resolution has not been improved, we double the values of $\rho$, $\delta_{min}$ and $\delta_{max}$, and the algorithm is executed with these values for $\zeta'$ iterations;
-\item afterwards, $\rho$, $\delta_{min}$ and $\delta_{max}$ switch back to this initial value.
-\end{inparaenum}
-%
-This approach may lead to drawings with larger area, but this is ``expected'', as it turns out that drawings with high crossing resolution may require large area~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/tcs/BrandenburgDEKL16}.
-
-% ==================================================================
-\myparagraph{Complexity issues}
-% ==================================================================
-%
-A factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges and the corresponding angles at these points.
-Given  a drawing, a na\"ive approach to compute its crossings requires $O(m^2)$ time, which can be improved by a plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ denote the number of edges and crossings.
-%; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.
-
-Instead of computing all crossing points and the corresponding angles for each candidate position of each iteration, we adopted a different approach for determining the set of feasible candidate positions, which turned out to be quite efficient in practice.
-Recall that we denoted by $v_i$ the vertex chosen at the $i$-th iteration step, and by $\pi_0,\ldots,\pi_{\rho-1}$ the candidate positions to move $v_i$.
-Let $e_0,\ldots,e_{d_i-1}$ be the edges incident to $v_i$, where $d_i=deg(v_i)$.
-Next, for each edge $e_k$ with $k=0,\ldots,d_i-1$  we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges in $\Gamma_{i-1}$.
-Let $\phi_i$ be the minimum crossing angle computed; this is our reference angle.
-Also, for each candidate position $\pi_j$ with $j=0,\ldots,\rho-1$, and for each edge $e_k$ with $k=0,\ldots,d_i-1$, we compute the crossings and the corresponding crossing angles of $e_k$ with all other edges of the drawing, assuming that $v_i$ is at $\pi_j$.
-Let $\chi_j$ be the minimum crossing angle computed with this approach, when $v_i$ is at position $\pi_j$.
-Clearly, $\pi_j$ is feasible only if $\chi_j \geq \phi_i$.
-Note that the complexity of this approach is $O(deg(v_i)m) = O(nm)$.
-
-% ==================================================================
-\subsection{Some interesting variants}
-\label{ssec:variants}
-% ==================================================================
-
-In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our~algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:experiments}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio
-%(that is, the ratio of the longest to the shortest edge is large)
-and poor angular resolution.
-%(that is, the angles formed by adjacent edges are small).
-%The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
-For the latter observe that if in a drawing all edges are either almost horizontally or almost vertically drawn, such that only ``horizontal'' and ``vertical'' edges cross, then the crossing resolution of this drawing is arbitrarily close to $90^\circ$, while its angular resolution is arbitrarily close to~$0^\circ$.
-
-% ==================================================================
-\myparagraph{Aspect ratio}
-% ==================================================================
-%
-%Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
-What we simply had to do was to reject candidate positions, which violate this precondition.
-
-% ==================================================================
-\myparagraph{Total resolution}
-% ==================================================================
-%
-%The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}.
-Formally, it is defined as the minimum of these two measures.
-It was not difficult to
-Similarly as above, we could adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing.
-In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution.
-Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease.
-We do so by rejecting candidate positions which yield a reduced total~resolution.
-
-% ==================================================================
-\myparagraph{Angular resolution}
-% ==================================================================
-%
-As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution).
-We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).
-
-% ==================================================================
-\myparagraph{Grid drawings}
-% ==================================================================
-%
-Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates.
-However, it can be easily adjusted to produce such drawings.
-More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid).
-One can even bound the size of the grid, by rejecting candidate grid positions outside the bounds.
-\arxapp{In Appendix~\ref{app:grid}}{In~\cite{arxivVersion}}, we report experimental results on this variant.
-
- %(the latter requirement is not difficult to be achieved, e.g., by slightly adjusting the positions of the vertices of a circular layout, whose underlying circle has large radius, to be at grid points).
-
-%An alternative approach is to let our algorithm compute a (non-grid) drawing and then apply some technique to convert it to grid.
-Critical here is not to affect the crossing resolution too much during the conversion.
-Of course, if the crossing resolution of the obtained grid drawing is at least as large as the crossing resolution of the non-grid starting drawing, then we considered the computed grid drawing \emph{acceptable}.
-In our approach, we also accepted drawings whose crossing resolution is at least as large as the crossing resolution of the non-grid starting drawing minus an input parameter $\epsilon$, which we used as \emph{slack} for tolerating slightly worse solutions.
-For converting a non-grid drawing to grid, we repeated the following procedure, that is also based on randomization, until we compute an acceptable grid drawing: For each vertex of the non-grid starting drawing, chose uniformly at random one of the grid points from the so-called \emph{neighbourhood} of the vertex, and move the vertex to this point (assuming of course that there is no other vertex at this point), where the neighbourhood of a vertex was initially set to its four closest grid points.
-If after a number of iterations an acceptable grid drawing without vertex-vertex overlaps could not be reported, then we augmented the neighbourhood of all vertices to contain more grid points, in particular, the grid points that are in horizontal and vertical distance~$1$ from the grid points of the current neighbourhood\todo{Not forget to mention that we may have to double the area.}.
-
-% ==================================================================
-\section{Experimental Evaluation}
-\label{sec:experiments}
-% ==================================================================
-
-In this section, we present the results of our experimental evaluation.
-For comparison purposes, apart from our algorithm, we also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The implementations\footnote{Our implementation is available on request from the authors.} were in Java using yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-The experiment was performed on a Linux laptop with four cores at 2.4 GHz and 8 GB RAM.
-%
-As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs; \arxapp{in Appendix~\ref{app:north}}{in~\cite{arxivVersion}}, we also report on the AT\&T graphs.
-
-The experiment was performed as follows.
-Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}.
-Starting from this layout, every graph was drawn with
-\begin{inparaenum}[(i)]
-\item our algorithm,
-\item our algorithm restricted~not to violate the aspect ratio of the initial layout, and the force-directed algorithms
-\item by Argyriou et al.\ and
-\item by Huang et al.
-\end{inparaenum}
-%We compared the quality of the produced drawings based on the following aesthetic~properties:
-%
-%\begin{multicols}{2}
-%\begin{enumerate}[P.1.]
-%\item \label{p:cr-res} crossing resolution
-%\item \label{p:to-res} total resolution
-%\item \label{p:an-res} angular resolution
-%\columnbreak
-%\item \label{p:as-rat} aspect ratio
-%\item \label{p:no-xing} number of crossings
-%\end{enumerate}
-%\end{multicols}
-%
-Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of the algorithms to maximize exclusively the corresponding measures; see Figs.~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
-In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results are on average across different drawings with same number of vertices.
-Finally, we mention that for our algorithm, we chose $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=50$; also, in order to further decrease the time needed for an iteration, we only computed a candidate position for 10 of the 50 rays.
-
-%We compared the quality of the produced drawings based on crossing resolution, total resolution, and angular resolution.
-Additionally, we consider the aspect ratio, and the number of crossings of the corresponding drawings.
-
-\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}.
-Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
-Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
-\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the crossing resolution for the Rome graphs.}
-\label{fig:cr-res}
-\end{figure}
-
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
-For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behaviour is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges.
-% (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) with the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
-
-A different behaviour can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}.
-We note here that we used different criteria to determine whether the algorithms of our experiment had converged.
-For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees.
-For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations.
-We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment.
-In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis\, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this more restricted convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
-The maximum number of iterations that each of the algorithms could perform in order to converge was set to 100.000, but that limit was never reached.
-We observe that both force-directed algorithms seem to require a great amount of iterations to converge for small graphs, where a drawing with really good crossing resolution is possible.
-However, for larger graphs  the algorithm by Huang et al.\ requires the least amount of iterations.
-On the other hand, both the unrestricted and the restricted variant of our algorithm require comparable number of iterations to converge, but clearly more than the ones of the algorithm by Huang et al.
-
-\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}.
-Here, each algorithm was adjusted to maximize both the crossing and the angular resolution.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield  drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
-
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
 \subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 
 \subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
 \centering
@@
 \end{axis}
-\end{tikzpicture}}}
+\end{tikzpicture}}
 \caption{Experimental results on the total resolution for the Rome graphs.}
 \label{fig:to-res}
-\end{figure}
-
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of aspect ratio) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
-More concretely, the drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
-Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
-
-%Analogous observations can be made for the number of crossings of the produced layouts; see Fig.~\ref{fig:to-res-3}.
-The restricted variant of our algorithm and the algorithm by Huang et al.\ yield drawings with the least number of crossings; see Fig.~\ref{fig:to-res-3}.
-Comparable but slightly worse (in terms of the number of crossings) are the drawings produced by the force-directed algorithm by Argyriou et al.
-Our unrestricted algorithm seems to require the largest number of crossings, which turn out to be notably higher than the corresponding ones of the other three algorithms.
-
-On the negative side, both the unrestricted and the restricted variant of our algorithm require more iterations than the  force-directed algorithm by Huang et al.; see Fig.~\ref{fig:to-res-4}.
-Recall, however, that the latter algorithm is clearly outperformed by both our variants in term of total resolution.
-The algorithm by Argyriou et al.\ clearly requires the highest number of iterations (especially for large graphs).
-We note that the convergence criterion was the same as for the crossing resolution; however, the measured quality was (not the crossing but) the total resolution.
-
-\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the  results for the angular resolution; see Fig.~\ref{fig:an-res}.
-Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
-The restricted variant of our algorithm yields drawings with slightly worse angular resolution.
-The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-\begin{figure}[t]
-\centering
-\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
-mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Angular Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
-\addplot table [x=n, y=Angular resolution rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Angular resolution rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Angular resolution fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-angularResolution.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-aspectRatio.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-crossings.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
-\addplot table [x=n, y=Iterations rm, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {angularResolution.csv};
-\addplot table [x=n, y=Iterations fa, col sep=semicolon]{angularResolution.csv};
-\addplot table [x=Nodes, y=Angular Only, col sep=semicolon]{abs-iterations.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results on the angular resolution for the Rome graphs.}
-\label{fig:an-res}
-\end{figure}
-
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
-The small differences result from the fact that the crossing resolution cannot be entirely neglected.
-
-% ==================================================================
-\myparagraph{Discussion}
-%
-% ==================================================================
-%
-%In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-%
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details \arxapp{see Appendix~\ref{app:contest}}{see~\cite{arxivVersion}}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\bibliographystyle{abbrvurl}
-\bibliography{references}
-
-\arxapp{\newpage
-\appendix
-
-\section*{Appendix}
-
-\section{Experiments on Grid Drawings}
-\label{app:grid}
-
-\begin{figure}[!b]
-\centering
-\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=n, y=crossing resolution rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=n, y=Aspect ratio rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=n, y=Crossing number rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Crossing number rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:crossingResolutionTimeGrid}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{0.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=n, y=Iterations rm1M, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm10k, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm1000, col sep=semicolon]{gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm100, col sep=semicolon] {gridExperiment.csv};
-\addplot table [x=n, y=Iterations rm, col sep=semicolon] {gridExperiment.csv};
-\legend{{$10^6 \times 10^6$},{$10^4 \times 10^4$},{$10^3 \times 10^3$},{$10^2 \times 10^2$}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Our experimental results on the crossing resolution with different grid restrictions.
-The double line corresponds to our unrestricted algorithm.}
-\label{fig:experimentsGrid}
-\end{figure}
-
-In addition to the experiments described in Section~\ref{sec:experiments}, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we evaluated in Section~\ref{sec:experiments} are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
-%
-\begin{inparaenum}[(i)]
-\item $10^6 \times 10^6$
-\item $10^4 \times 10^4$
-\item $10^3 \times 10^3$, and
-\item $10^2 \times 10^2$.
-\end{inparaenum}
-%
-The test suite for this experiment was again the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}.
-However, since our algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
-The layouts for each of the different sizes of the grid were computed with the variant of our algorithm that optimizes the crossing resolution; Fig.~\ref{fig:experimentsGrid} summarizes the results.
-
-Regarding the crossing resolution, we can observe that with increasing grid size, we could achieve better crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
-More precisely, a grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
-As a result, the reported drawings were often the initial ones (as our algorithm could not improve them), especially for large graphs.
-Significantly fewer were the graphs for which our algorithm could not report an improved drawing, when the grid size was set to $10^3 \times 10^3$.
-%Starting from grid size $10^3 \times 10^3$, we could observe, for small graphs, the performance was close to the performance of the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}).
-%For grid size $10^3 \times 10^3$, the performance declined for the larger graphs in the test set and we again could observe that for some of them the algorithm was not able to improve the initial layout.
-For grid size $10^4 \times 10^4$, the drawings produced by our algorithm were on average by only $10^\circ$ worse than those produced by the unrestricted version of our algorithm (double line in Fig.~\ref{fig:crossingResolutionGrid}), while the gap was closer for grid size $10^6 \times 10^6$.
-
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
-
-As expected, the smaller the underlying grid is, the more crossings the computed drawings contain; see Fig.~\ref{fig:crossingResolutionCrossingNumberGrid}.
-As a result, the unrestricted variant of our algorithm clearly outperforms all other ones.
-It is worth noting that the differences are clear between grid sizes $10^2 \times 10^2$, $10^3 \times 10^3$ and $10^4 \times 10^4$.
-Notably, there is only a slight improvement (in terms of the number of crossings) from grid size $10^4 \times 10^4$ to $10^6 \times 10^6$.
-%This can partially be explained with the different choice of the initial drawing.
-%
-On the other hand, the number of iterations needed for convergence increases with the grid size (see Fig.~\ref{fig:crossingResolutionTimeGrid}), with the exception of the grid of size $10^2 \times 10^2$, which verifies our previous observation that for the vast majority of the graphs an improved drawing could not be reported.
-
-In conclusion, we can state that our algorithm is still able to compute drawings with high crossing resolution when restricted to a grid, as long as the grid is not too small.
-However, the computation of a grid drawing takes longer, which is of course expected.
-%than in the unrestricted version (up to five times as long for grid size $10^4 \times 10^4$ or eight times as long for grid size $10^6 \times 10^6$). In order
-Finally, note that the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution) but also with respect to the number of iterations needed to converge.
-%To improve the performance of our algorithm in this restricted variant, it may be interesting to see how much the choice of the initial drawing affects the resulting crossing resolution.
-Further experiments in this direction are needed.
-
-\section{Experiments on the AT\&T  Graph Test Set}
-\label{app:north}
-
-In this section, we report the results of our expertimental evaluation (on the crossing, total and angular resolutions) for the non-planar AT\&T graphs, which form a collection of 424 benchmark graphs (also known as Graph Catalog and North graphs; available at \texttt{http://graphdrawing.org/data}).
-Note that we did not impose any grid constraint on our algorithms.
-The corresponding results are illustrated in Figs.~\ref{fig:northCrossing}, \ref{fig:northTotal} and~\ref{fig:northAngular}.
-In general, we observed that the variance of the results is much larger than in the experiments on the Rome graphs.
-This manifests in spikes of large magnitude in the illustrations of the results and indicates that the structural properties of the graphs in this second test set varies vastly between different graph sizes.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Crossing Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=cr rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cr tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200,500,1000,2000},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northCrossing-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{crossingResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {crossingResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
-\label{fig:northCrossing}
-\end{figure}
-
-For the crossing resolution, we observed that both variants of our algorithm again outperformed the two force-directed algorithms; see Fig.~\ref{fig:northCrossing-1}.
-Remarkable is the synchronous behaviour of all four algorithms regarding the crossing resolution, as the curves are nearly parallel.
-By all these results, we can classify the graphs into ``hard'' or ``easy'' when maximizing their crossing resolution.
-In particular, graphs with 50 to 70 vertices appear to be harder to improve than graphs with 70 to 80 vertices.
-Regarding the aspect ratio of the produced drawings, we observe that while our algorithms show a slight increase with the number of vertices, the behaviour for both force-directed algorithms appears to be unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force directed approaches produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
-All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
-In terms of the number of iterations, we observe that somewhat surprisingly the algorithm of Argyriou et al.
-converges in the least amount of iterations, while the remaining three algorithms behave nearly the same; see Fig.~\ref{fig:northCrossing-4}.
-
-\begin{figure}[tb]
-\centering
-\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Total Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Total Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=tr rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=tr tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Total Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Total Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northTotal-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Total Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {totalResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{totalResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {totalResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the total resolution on the AT\&T graphs.}
-\label{fig:northTotal}
-\end{figure}
-
-In the total resolution experiment, we observed similar results as in the experiment on the Rome graphs for small graphs, that is, our unrestricted algorithm outperforms the other three ones, while the restricted variant of our algorithm yields drawings and then the algorithm by Argyriou et al.; see Fig.~\ref{fig:northTotal-1}.
-For larger graphs, however, these three algorithms achieve similar results while still outperforming the algorithm by Huang et al.
-The results for the aspect ratio and number of crossings are similar to those of the crossing resolution experiment, with the exception of the fact that the algorithm of Huang et al.
-performs more stable with respect to the aspect ratio; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
-With respect to the number of iterations, our two algorithms and the one by Argyriou et al.
-show similar behavior needing more iterations than the algorithm by Huang et al.\ in order to converge; see Fig.~\ref{fig:northTotal-4}.
-%Observe that the number of iterations does not seem to correlate with the number of vertices.
-
-\begin{figure}[htb]
-\centering
-\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Angular Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,
-legend style={at={(0.425,1.35)},anchor=north,,legend columns=2,draw=none},
-width=0.475\textwidth,
-cycle list name=mylist,
-mark repeat={5},xlabel={Number of Vertices},ylabel={Angular Resolution},
-xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={10,20,30,40,50,60,70,80,90},
-tick pos=left,ymajorgrids]
-\addplot table [x=avg, y=an rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=an fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=an tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Angular Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none},ylabel style={yshift=-0.2cm}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={0,10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio},
-xlabel={Number of Vertices},
-xmin=10,
-xmax=100,
-ytick={1,2,5,10,20,50,100,200},ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
-\addplot table [x=avg, y=ar rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=ar tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Angular Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,250,500,750,1000,1250,1500,1750},ylabel={Number of Crossings},
-xlabel={Number of Vertices},tick pos=left,ylabel style={yshift=-0.2cm}, ymajorgrids]
-\addplot table [x=avg, y=cn rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=cn tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\subfloat[\label{fig:northAngular-4}{Iterations vs no.~of vertices}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Angular Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},ylabel style={yshift=-0.2cm},axis y line*=left,legend style={at={(0.425,1.35)},
-anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},
-xmin=10,
-xmax=100,
-ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},tick pos=left,
-ylabel={Iterations},
-xlabel={Number of Vertices},ymajorgrids]
-\addplot table [x=avg, y=it rm, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it rm-1, col sep=semicolon] {angularResolutionNorth.csv};
-\addplot table [x=avg, y=it fa, col sep=semicolon]{angularResolutionNorth.csv};
-\addplot table [x=avg, y=it tr, col sep=semicolon] {angularResolutionNorth.csv};
-\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
-\end{axis}
-\end{tikzpicture}}}
-\caption{Experimental results for the angular resolution on the AT\&T graphs.}
-\label{fig:northAngular}
-\end{figure}
-
-In the angular resolution experiment, we again obtain a not-so-clear picture concerning the ranking of the algorithms, especially for higher number of vertices the ranking varies; see Fig.~\ref{fig:northAngular-1}.
-Only the algorithm by Huang et al.
-seems to be mostly at the last rank.
-Concerning the aspect ratio we see very good behaviour for our restricted variant and the algorithm by Huang et al.
-while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
-For the number of crossings, we again observe that all algorithms achieve similar values, however, our unrestricted algorithm and the algorithm by Argyriou et al.
-achieve slightly higher values for larger graphs; see Fig.~\ref{fig:northAngular-3}.
-Finally, both our algorithms and the one by Huang et al.
-need a similar number of iterations for convergence which is lower than the one by Argyriou et al.; see Fig.~\ref{fig:northAngular-4}.
-
-Summarizing, we conclude that compared to the Rome graphs, the AT\&T graphs show a much higher variance regarding the various resolution measures.
-%The crossing resolution results of our method still outperform the other approaches by far but especially for larger graphs as well as angular and total resolution the methods tend to perform similarly.
-
-\begin{table}[t!]
-
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-%Graph & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
-%\hline
-%Contest Winner & 90 & 88.23 & 90 & 88.97 & 80.4 & 90 & 56.537  \\
-%Our New Approach & 89.78 & 88.7 & 89.95 & 89.05 & 86.96 & 89.72 & 63.62  \\
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table}
-
-\section{Graph Drawing Contest 2017 Graphs}
-\label{app:contest}
-
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
-
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
-
-%\newpage
-
-\section{Sample Drawings}
-\label{app:samples}
-
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
-Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-
-\begin{figure}[h!]
-\centering
-%\subfloat[\label{fig:graph5-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-%}
-
-\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-The aesthetic criterion optimized by each variant is reported in the caption of its subfigure.}
-\label{fig:graph5}
-\end{figure}
-
-\begin{figure}[p]
-\centering
-%\subfloat[\label{fig:graph9-yfiles}{Organic Layout}]{
-%\centering
-%\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-%}
-
-\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
-\hfill
-\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
-\hfill
-\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
-%(a)~Organic Layout by \emph{yFiles}, 
-(a)--(c)~our algorithm, 
-(d)--(f)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
-(g)--(i)~the algorithm by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
-Each variant was obtained by optimizing a different aesthetic criterion, which is named in the caption of each subfigure.}
-\label{fig:graph9}
-\end{figure}}{}
-
-\end{document}
+\end{figure*}

### 2018-09-23 paper/journal.tex
The pass is finished; only the results from the GD contest 2018 are to be added.

@@
 \volumenumber{--} 
-\issuenumber{-} 
+\issuenumber{--} 
 % ==================================================================
@@
 
+\begin{table*}[t!]
+\caption{Summary of the results for the Graph Drawing Contest 2017.}
+\centering
+\begin{tabular}{c|c|c|c|c}
+\toprule
+~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
+\midrule
+1 & 90  & 77  & 89.99 & 89.99\\
+2 & 88.23  & 42 & 88.21 & 88.7 \\
+3 & 90  & 89 & 87.86 & 89.95 \\
+4 & 88.97  & 89 & 77.13  & 89.05 \\
+5 & 80.4  &  30 & 78.68 & 86.96 \\
+6 & 90  & 78 & 89.96 & 89.96\\
+7 & 56.537  & 34 & 55.77  & 63.62 \\
+8 & 84.95  & 61 & 81.18 & 89.28 \\
+9 & 59.885  & 9 & 54.63 & 88.2 \\
+10 & 20.978  & 4 & 23.60  & 23.72 \\
+11 & 46.684  & 6 & 57.00 & 72.00 \\
+12 & 36.47  & 5 & 26.24 & 35.86 \\
+13 & 25.456  & 4 & 22.43 & 33.68 \\
+14 & 33.52  & 5 & 29.69 & 43.08 \\
+15 & 20.512  & 4 & 13.51 & 29.18 \\
+\bottomrule
+\end{tabular}
+\label{tab:gdContest2017}
+\end{table*}
+
 While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
@@
 In addition to the experiments that we have described so far, we also evaluated how our algorithm performs, if we restrict its vertices to integer grid coordinates.
-In particular, we were interesting to see how the different quality measures that we have evaluated so far are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
+In particular, we were interesting to see how the different quality measures that we have investigated so far are affected by the restrictions imposed on having the vertices of the graph on integer grids of different sizes:
 
@@
 The test suite for this experiment was the non-planar Rome graphs.
-The algorithm to be evaluated was the variant of our algorithm that optimizes the drawing's crossing resolution.
-However, since this algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the grid, ensuring that neither two vertices nor two edges overlap.
+The algorithm to be evaluated was the variant of our algorithm that optimizes the drawing's crossing resolution (without any restriction on the aspect ratio).
+However, since this algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the underlying grid, ensuring that neither two vertices nor two edges overlap.
 Fig.~\ref{fig:experimentsGrid} summarizes the results.
 
-Regarding the crossing resolution, we natural observe that the bigger the grid is, the easier it is to compute a drawing with higher crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+Regarding the crossing resolution, we natural observe that the bigger the grid is, the easier it is to compute a drawing with high crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
 More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
@@
 
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the grid of size $10^2 \times 10^2$; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the $10^2 \times 10^2$ grid; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
 The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
@@
 % ==================================================================
-\subsection{Graph Drawing Contest 2017}
-\label{ssec:contest17}
+\subsection{Graph Drawing Contests}
+\label{ssec:contest}
 % ==================================================================
 
-Our primary motivation for this work was our participation to the Graph Drawing contest in 2017, where we miserably performed\footnote{\texttt{http://www.graphdrawing.de/contest2017/results.html}}; the topic was the maximization of the crossing resolution.
-Our approach for the contest in 2017 was a mixture of the two force directed algorithms by Argyriou et al.
-and Huang et al.
+Every year, the Graph Drawing conference organizes a contest, in which the task is to visualize a set of challenge graphs in a predefined drawing style.
+The teams participating to the contest receive a set of challenge graphs, and after one hour they must submit their final drawings for each of them.
+The team with the highest cumulative score wins.
+In years 2017 and 2018, the task of the contest was the crossing resolution maximization problem, which actually formed our primary motivation for this work.
 
-We give a comparison of our new approach to the performances of the clear winner ``CoffeeVM'' of last year's graph drawing contest and our previous team ``TuebingenMidnight'' in Table~\ref{tab:gdContest2017}.
-Note that in the contest the teams had only one hour to compute layouts for all 15 contest graphs.
-For our algorithm, we provide results that were achieved with the same time limit (see column ``Time restricted''), as well as our best results which were achieved without a strict time limit (see column ``Our best'').
+\myparagraph{Graph Drawing Contest 2017} In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
+%
+Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.
 
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the last year's contest winner.
-On a few graphs (namely, graphs 10 and 11), we even achieve better results.
-With a single exception (namely, graph 4), we easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as last year's contest winner.
-We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
+It is far from clear that the wining team was the team  ``CoffeeVM''.
+Our team ``TuebingenMidnight''  miserably performed.
+As already mentioned, each team had only one hour to compute a layout for each of the $16$ challenge graphs.
+In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved with our algorithm with the same time limit (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
 
-\begin{table*}[t!]
+We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the contest's winner when restricted within the one hour.
+On a few graphs (namely, graphs 10 and 11), we even achieved better results.
+With a single exception (namely, graph 4), we could easily outperformed our results from last year.
+If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as the contest's winner.
 
-\caption{Summary of the results for the Graph Drawing Contest Graphs.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42 & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
-4 & 88.97  & 89 & 77.13  & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
-8 & 84.95  & 61 & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
-12 & 36.47  & 5 & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
-14 & 33.52  & 5 & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
-\bottomrule
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table*}
+%We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
+
+\myparagraph{Graph Drawing Contest 2018} In 2018,  
+\vspace{10cm}
 
@@
 
-In this section, we present drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
+In this section, we present sample drawings of the 5th and of the 9th graph given in the Graph Drawing contest 2017 that are produced by different variants of our algorithm and of the algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}; see Figures~\ref{fig:graph5} and~\ref{fig:graph9}, respectively.
 Each variant was obtained by adjusting each of the aforementioned algorithms to optimize the crossing resolution, the angular resolution and the total resolution, respectively; the aesthetic criterion optimized by each variant is reported in the caption of its corresponding subfigure.
-As initial drawings for all algorithms, we used the layouts shown in Figure~\ref{fig:graph5And9-yfiles} that we computed with the SmartOrganic layouter of the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
+As initial drawings for all algorithms, we used the layouts shown in Fig.~\ref{fig:graph5And9-yfiles} that we computed with the SmartOrganic layouter of the yFiles library~\cite{DBLP:books/sp/04/WieseE004}.
 
-\begin{figure*}[htbp]
+\begin{figure}[t]
 \centering
 \subfloat[\label{fig:graph5-yfiles}{}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}
-}
-\hfil
+\includegraphics[width=0.3\textwidth]{graph5-yfiles}}
+
 \subfloat[\label{fig:graph9-yfiles}{}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-yfiles}
-}
+\includegraphics[width=0.27\textwidth]{graph9-yfiles}}
 \caption{Illustration of 
@@
 \label{fig:graph5And9-yfiles}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[p]
+% ==================================================================
+\section{Conclusions}
+\label{sec:conclusions}
+% ==================================================================
+
+%
+In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
+Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
+%As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+
+It also worth noting that while working on this project, we made some useful observations and obtained some interesting insights.
+In particular, there is a recent observation (also supported by experiments) that drawings, in which the crossing angles are large, are easy to read and understand.
+We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
+In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other (see, e.g., Fig.~\ref{fig:graph5-argyriou-cr}).
+Hence, their angular resolution and aspect ratio were often poor.
+The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped in significantly improving the readability of the produced drawings, without loosing too much of their quality in terms of the crossing resolution.
+
+We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm.
+As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely better as also evidenced by our experiments.
+The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
+
+\paragraph{Acknowledgments.} The authors would also like to thank Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
+
+\bibliographystyle{compj}
+\bibliography{references}
+
+\clearpage
+
+\begin{figure*}[h!]
 	\centering
 	\subfloat[\label{fig:graph5-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourBest-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph5-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourBest-angular}}
 	\hfill
 	\subfloat[\label{fig:graph5-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourBest-total}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourBest-total}}
 	\hfill
 	\subfloat[\label{fig:graph5-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourRestricted-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph5-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourRestricted-angular}}
 	\hfill
 	\subfloat[\label{fig:graph5-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-ourRestricted-total}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-ourRestricted-total}}
 	\hfill
 	\subfloat[\label{fig:graph5-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-argyriou-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph5-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-argyriou-angular}}
 	\hfill
 	\subfloat[\label{fig:graph5-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-argyriou-total}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-argyriou-total}}
 	\hfill
 	\subfloat[\label{fig:graph5-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-huang-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph5-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph5-huang-angular}}
 	\hfill
 	\subfloat[\label{fig:graph5-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph5-huang-total}
-}
-
-\caption{Different drawings of the 5th graph given in the Graph Drawing 2017 contest produced by different variants of
+	\includegraphics[width=0.27\textwidth]{graph5-huang-total}}
+	\caption{Different drawings of the 5th challenge graph given in the Graph Drawing 2017 contest produced by different variants of
 	(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
@@
 	\subfloat[\label{fig:graph9-ourBest-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-ourBest-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph9-ourBest-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-ourBest-angular}}
 	\hfill
 	\subfloat[\label{fig:graph9-ourBest-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourBest-total}
-}
-
+	\includegraphics[width=0.27\textwidth]{graph9-ourBest-total}}
+	\hfill	
 	\subfloat[\label{fig:graph9-ourRestricted-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-ourRestricted-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph9-ourRestricted-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-ourRestricted-angular}}
 	\hfill
 	\subfloat[\label{fig:graph9-ourRestricted-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-ourRestricted-total}
-}
-
+	\includegraphics[width=0.27\textwidth]{graph9-ourRestricted-total}}	
+	\hfill	
 	\subfloat[\label{fig:graph9-argyriou-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-argyriou-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph9-argyriou-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-argyriou-angular}}
 	\hfill
 	\subfloat[\label{fig:graph9-argyriou-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-argyriou-total}
-}
-
+	\includegraphics[width=0.27\textwidth]{graph9-argyriou-total}}	
+	\hfill
 	\subfloat[\label{fig:graph9-huang-cr}{Crossing Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-crossing}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-huang-crossing}}
 	\hfill
 	\subfloat[\label{fig:graph9-huang-ar}{Angular Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-angular}
-}
+	\includegraphics[width=0.27\textwidth]{graph9-huang-angular}}
 	\hfill
 	\subfloat[\label{fig:graph9-huang-tr}{Total Resolution}]{
-\centering
-\includegraphics[width=0.3\textwidth]{graph9-huang-total}
-}
-
-\caption{Different drawings of the 9th graph given in the Graph Drawing 2017 contest produced by different variants of
+	\includegraphics[width=0.27\textwidth]{graph9-huang-total}}
+	\caption{Different drawings of the 9th challenge graph given in the Graph Drawing 2017 contest produced by different variants of
 	(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
@@
 
-% ==================================================================
-\section{Conclusions}
-\label{sec:conclusions}
-% ==================================================================
-
-%
-In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-%As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
-
-While working on this project, we made some useful observations and  obtained some interesting insights.
-In particular, there is a recent hypothesis (also supported by experiments) that drawings, in which the crossing angles, are large are easy to read and understand.
-We observed that drawings that are optimized only in terms of the crossing angles might be arbitrarily bad and may have several undesired properties.
-In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other.
-Hence, angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped significantly improving the readability of the drawings, without loosing too much of their quality in terms of the crossing resolution.
-
-We conclude by noting that our motivation to work with this problem was our participation to GD2017 contest, where we performed miserably using a force-directed algorithm; for details see Appendix~\ref{app:contest}.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely more promising than our previous as evidenced by our experiments.
-The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
-
-\paragraph{Acknowledgments.} The authors would also like to acknowledge Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
-
-\bibliographystyle{compj}
-\bibliography{references}
-
 \end{document}

### 2018-09-25 paper/journal.tex
new figs

@@
 % ==================================================================
-\input{cr-res-rome.tex}
+%\input{cr-res-rome.tex}
 % ==================================================================
+\begin{figure*}[t]
+	\centering
+	\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=1]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=2]{figures/rome}}
+
+	\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=3]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=4]{figures/rome}}
+	\caption{Experimental results on the crossing resolution for the Rome graphs.}
+	\label{fig:cr-res}
+\end{figure*}
 
 % ==================================================================
-\input{to-res-rome.tex}
+%\input{to-res-rome.tex}
 % ==================================================================
+\begin{figure*}[t]
+	\centering
+	\subfloat[\label{fig:to-res-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=5]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=6]{figures/rome}}
+
+	\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=7]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=8]{figures/rome}}
+	\caption{Experimental results on the total resolution for the Rome graphs.}
+	\label{fig:to-res}
+\end{figure*}
 
 % ==================================================================
-\input{an-res-rome.tex}
+%\input{an-res-rome.tex}
 % ==================================================================
+\begin{figure*}[t]
+	\centering
+	\subfloat[\label{fig:an-res-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=9]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=10]{figures/rome}}
+
+	\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=11]{figures/rome}}
+	\hfil
+	\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=12]{figures/rome}}
+	\caption{Experimental results on the angular resolution for the Rome graphs.}
+	\label{fig:an-res}
+\end{figure*}
 
 % ==================================================================
-\input{cr-res-north.tex}
+%\input{cr-res-north.tex}
 % ==================================================================
+\begin{figure*}[t!]
+	\centering
+	\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=1]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=2]{figures/north}}
+
+	\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=3]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northCrossing-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=4]{figures/north}}
+	\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
+	\label{fig:northCrossing}
+\end{figure*}
 
 % ==================================================================
-\input{to-res-north.tex}
+%\input{to-res-north.tex}
 % ==================================================================
+\begin{figure*}[t!]
+	\centering
+	\subfloat[\label{fig:northTotal-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=5]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=6]{figures/north}}
+
+	\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=7]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northTotal-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=8]{figures/north}}
+	\caption{Experimental results for the total resolution on the AT\&T graphs.}
+	\label{fig:northTotal}
+\end{figure*}
 
 % ==================================================================
-\input{an-res-north.tex}
+%\input{an-res-north.tex}
 % ==================================================================
+\begin{figure*}[t!]
+	\centering
+	\subfloat[\label{fig:northAngular-1}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=9]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=10]{figures/north}}
+
+	\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=11]{figures/north}}
+	\hfil
+	\subfloat[\label{fig:northAngular-4}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=12]{figures/north}}
+	\caption{Experimental results for the angular resolution on the AT\&T graphs.}
+	\label{fig:northAngular}
+\end{figure*}
 
@@
 
-\begin{table*}[t!]
+\begin{table*}
 \caption{Summary of the results for the Graph Drawing Contest 2017.}
@@
 \midrule
-1 & 90  & 77  & 89.99 & 89.99\\
+1 & 90.00  & 77  & 89.99 & 89.99\\
 2 & 88.23  & 42  & 88.21 & 88.7 \\
-3 & 90  & 89 & 87.86 & 89.95 \\
+3 & 90.00  & 89  & 87.86 & 89.95 \\
 4 & 88.97  & 89  & 77.13 & 89.05 \\
-5 & 80.4  &  30 & 78.68 & 86.96 \\
-6 & 90  & 78 & 89.96 & 89.96\\
-7 & 56.537  & 34 & 55.77  & 63.62 \\
+5 & 80.40  & 30  & 78.68 & 86.96 \\
+6 & 90.00  & 78  & 89.96 & 89.96\\
+7 & 56.53  & 34  & 55.77 & 63.62 \\
 8 & 84.95  & 61  & 81.18 & 89.28 \\
-9 & 59.885  & 9 & 54.63 & 88.2 \\
-10 & 20.978  & 4 & 23.60  & 23.72 \\
-11 & 46.684  & 6 & 57.00 & 72.00 \\
+9 & 59.88  & 9   & 54.63 & 88.2 \\
+10 & 20.97  & 4  & 23.60 & 23.72 \\
+11 & 46.68  & 6  & 57.00 & 72.00 \\
 12 & 36.47  & 5  & 26.24 & 35.86 \\
-13 & 25.456  & 4 & 22.43 & 33.68 \\
+13 & 25.45  & 4  & 22.43 & 33.68 \\
 14 & 33.52  & 5  & 29.69 & 43.08 \\
-15 & 20.512  & 4 & 13.51 & 29.18 \\
+15 & 20.51  & 4  & 13.51 & 29.18 \\
+16 & 19.83  & 3  & 12.13 & 26.12 \\
 \bottomrule
@@
 \subfloat[\label{fig:graph5-yfiles}{}]{
-\includegraphics[width=0.3\textwidth]{graph5-yfiles}}
+\includegraphics[width=0.4\textwidth]{graph5-yfiles}}
 
 \subfloat[\label{fig:graph9-yfiles}{}]{
-\includegraphics[width=0.27\textwidth]{graph9-yfiles}}
+\includegraphics[width=0.37\textwidth]{graph9-yfiles}}
 \caption{Illustration of 
@@
 In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
-%As a future direction, we plan to evaluate the performance of variants of our heuristic that compromise between even more aesthetic criteria.
+Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
 
@@
 The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.
+Thus, as a future direction, we plan to further extend and evaluate our technique, e.g., by using a multi-criteria objective function (i.e., that is not limited to the crossing angles) that takes into account more aesthetic criteria.
 

### 2018-09-29 paper/journal.tex
Continue working.

@@
 \midrule
-1 & 90.00  & 77  & 89.99 & 89.99\\
-2 & 88.23  & 42  & 88.21 & 88.7 \\
-3 & 90.00  & 89  & 87.86 & 89.95 \\
-4 & 88.97  & 89  & 77.13 & 89.05 \\
-5 & 80.40  & 30  & 78.68 & 86.96 \\
-6 & 90.00  & 78  & 89.96 & 89.96\\
-7 & 56.53  & 34  & 55.77 & 63.62 \\
-8 & 84.95  & 61  & 81.18 & 89.28 \\
-9 & 59.88  & 9   & 54.63 & 88.2 \\
-10 & 20.97  & 4  & 23.60 & 23.72 \\
-11 & 46.68  & 6  & 57.00 & 72.00 \\
-12 & 36.47  & 5  & 26.24 & 35.86 \\
-13 & 25.45  & 4  & 22.43 & 33.68 \\
-14 & 33.52  & 5  & 29.69 & 43.08 \\
-15 & 20.51  & 4  & 13.51 & 29.18 \\
-16 & 19.83  & 3  & 12.13 & 26.12 \\
+1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$ & 89.99$^\circ$\\
+2 & 88.23$^\circ$  & 42$^\circ$ & 88.21$^\circ$ & 88.70$^\circ$\\
+3 & 90.00$^\circ$  & 89$^\circ$ & 87.86$^\circ$ & 89.95$^\circ$\\
+4 & 88.97$^\circ$  & 89$^\circ$ & 77.13$^\circ$ & 89.05$^\circ$\\
+5 & 80.40$^\circ$  & 30$^\circ$ & 78.68$^\circ$ & 86.96$^\circ$\\
+6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$ & 89.96$^\circ$\\
+7 & 56.53$^\circ$  & 34$^\circ$ & 55.77$^\circ$ & 63.62$^\circ$\\
+8 & 84.95$^\circ$  & 61$^\circ$ & 81.18$^\circ$ & 89.28$^\circ$\\
+9 & 59.88$^\circ$  & 9$^\circ$  & 54.63$^\circ$ & 88.20$^\circ$\\
+10 & 20.97$^\circ$ & 4$^\circ$  & 23.60$^\circ$ & 23.72$^\circ$\\
+11 & 46.68$^\circ$ & 6$^\circ$  & 57.00$^\circ$ & 72.00$^\circ$\\
+12 & 36.47$^\circ$ & 5$^\circ$  & 26.24$^\circ$ & 35.86$^\circ$\\
+13 & 25.45$^\circ$ & 4$^\circ$  & 22.43$^\circ$ & 33.68$^\circ$\\
+14 & 33.52$^\circ$ & 5$^\circ$  & 29.69$^\circ$ & 43.08$^\circ$\\
+15 & 20.51$^\circ$ & 4$^\circ$  & 13.51$^\circ$ & 29.18$^\circ$\\
+16 & 19.83$^\circ$ & 3$^\circ$  & 12.13$^\circ$ & 26.12$^\circ$\\
 \bottomrule
@@
 \label{tab:gdContest2017}
+\end{table*}
+
+\begin{table*}
+\caption{Summary of the results for the Graph Drawing Contest 2018.}
+\centering
+\begin{tabular}{c|c|c|c}
+\toprule
+~~Graph~~ & ~~TuebingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
+\midrule
+1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$\\
+2 & 88.23$^\circ$  & 42$^\circ$ & 88.70$^\circ$\\
+3 & 90.00$^\circ$  & 89$^\circ$ & 89.95$^\circ$\\
+4 & 88.97$^\circ$  & 89$^\circ$ & 89.05$^\circ$\\
+5 & 80.40$^\circ$  & 30$^\circ$ & 86.96$^\circ$\\
+6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$\\
+7 & 56.53$^\circ$  & 34$^\circ$ & 63.62$^\circ$\\
+8 & 84.95$^\circ$  & 61$^\circ$ & 89.28$^\circ$\\
+9 & 59.88$^\circ$  & 9$^\circ$  & 88.20$^\circ$\\
+10 & 20.97$^\circ$ & 4$^\circ$  & 23.72$^\circ$\\
+11 & 46.68$^\circ$ & 6$^\circ$  & 72.00$^\circ$\\
+12 & 36.47$^\circ$ & 5$^\circ$  & 35.86$^\circ$\\
+13 & 25.45$^\circ$ & 4$^\circ$  & 33.68$^\circ$\\
+14 & 33.52$^\circ$ & 5$^\circ$  & 43.08$^\circ$\\
+\bottomrule
+\end{tabular}
+\label{tab:gdContest2018}
 \end{table*}
@@
 % ==================================================================
-\subsection{Graph Drawing Contests}
+\subsection{Graph Drawing Contests 2018 and 2019}
 \label{ssec:contest}
@@
 The team with the highest cumulative score wins.
-In years 2017 and 2018, the task of the contest was the crossing resolution maximization problem, which actually formed our primary motivation for this work.
+In years 2017 and 2018, the task of the contest was to produce a drawing for each of the given challenge graphs with as high crossing resolution as possible (which actually also formed our primary motivation for this work).
 
@@
 
-It is far from clear that the wining team was the team  ``CoffeeVM''.
-Our team ``TuebingenMidnight''  miserably performed.
-As already mentioned, each team had only one hour to compute a layout for each of the $16$ challenge graphs.
-In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved with our algorithm with the same time limit (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
+It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''.
+Our team ``\texttt{TuebingenMidnight}''  miserably performed; we actually took the third position out of a total of four teams that participated.
+As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$.
+In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
 
@@
 On a few graphs (namely, graphs 10 and 11), we even achieved better results.
-With a single exception (namely, graph 4), we could easily outperformed our results from last year.
-If we neglect the time restriction, for all the graphs, the results  are (sometimes considerably) better than or at least about the same as the contest's winner.
+With a single exception (namely, graph 4), we could easily outperform the results that we achieved with the spring-embedder algorithm.
+If we neglect the time restriction, then for all the graphs, the results  are (sometimes considerably) better than or at least about the same as the contest's winner.
 
-%We can conclude that our new approach has good potential for the application in this year's graph drawing contest, however, we also note that more careful graph dependent parameterization will be needed to compute competitive solutions within the provided time.
+\myparagraph{Graph Drawing Contest 2018} In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper).
+The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''.
+Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.
 
-\myparagraph{Graph Drawing Contest 2018} In 2018,  
-\vspace{10cm}
+It is eye-catching that our team was the winning one; team \texttt{CoffeeVM+} took the second position of the contest (out of a total of four teams that participated to the contest).
+It is worth noting that for almost all $14$ challenge graphs our team achieved significantly better results than those of the second team \texttt{CoffeeVM+}, which clearly indicates that our new approach is very competitive (and definitely better than our previous,  as also evidenced by the experimental evaluation).
 
@@
 We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm.
-As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account; our new approach is definitely better as also evidenced by our experiments.
+As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account.
+
 The framework that we developed seems to be quite adaptable to optimize or to take into account also other desired aesthetic properties of a drawing.

### 2018-10-04 paper/journal.tex
some more work on the journal
Table 2 -> update with Henry
current version

@@
 \usepackage{amsmath,amssymb}
-\usepackage{graphicx,pgfplots}
+\usepackage{graphicx,color,pgfplots}
 \usepackage[font=small]{subfig,caption}
-%\usepackage{algorithm2e}
 \usepackage{todonotes,lineno,paralist,soul}
 \usepackage{rotating,hyperref,multicol}
-\usepackage{booktabs,tabularx}
+\usepackage{booktabs,tabularx,colortbl}
 \graphicspath{{figures/}}
@@
 The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$.
-It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.
+It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced from our related work section.
 
@@
 
-\paragraph{Structure of the paper:}
-The remainder of this paper is structured as follows.
-Section~\ref{sec:relatedwork} overviews related works.
-Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
-We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.
+\begin{table*}
+\caption{Summary of the results for the Graph Drawing Contest 2017.}
+\centering
+\begin{tabular}{c|c|c|c|c}
+\toprule
+\rowcolor{gray!30!}~~Graph~~ & ~~CoffeeVM~~ & ~~T\"ubingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
+\midrule
+\rowcolor{gray!30!} 1  & 90$^\circ$  & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!00!} 2  & 88$^\circ$  & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 3  & 90$^\circ$  & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!00!} 4  & 89$^\circ$  & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 5  & 80$^\circ$  & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
+\rowcolor{gray!00!} 6  & 90$^\circ$  & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!30!} 7  & 57$^\circ$  & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
+\rowcolor{gray!00!} 8  & 85$^\circ$  & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 9  & 60$^\circ$  & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
+\rowcolor{gray!00!} 10 & 21$^\circ$  & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
+\rowcolor{gray!30!} 11 & 47$^\circ$  & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
+\rowcolor{gray!00!} 12 & 36$^\circ$  & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
+\rowcolor{gray!30!} 13 & 25$^\circ$  & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
+\rowcolor{gray!00!} 14 & 34$^\circ$  & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
+\rowcolor{gray!30!} 15 & 21$^\circ$  & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
+\rowcolor{gray!00!} 16 & 20$^\circ$  & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
+\bottomrule
+\end{tabular}
+\label{tab:gdContest2017}
+\end{table*}
 
 % ==================================================================
-\section{Related Work}
-\label{sec:relatedwork}
+\paragraph{Related Work:}
+%\label{sec:relatedwork}
 % ==================================================================
-
 As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem.
@@
 Of more general scope are, of course, the already mentioned works by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, which are also based on the force-directed technique.
+
+% ==================================================================
+\paragraph{Motivation:}
+%\label{sec:motivation}
+% ==================================================================
+
+Every year, the Graph Drawing conference organizes a contest, in which the task is to visualize a set of challenge graphs in a predefined drawing style.
+The teams participating to the contest receive a set of challenge graphs, and after one hour they must submit their final drawings for each of them.
+The team with the highest cumulative score wins.
+In years 2017 and 2018, the task of the contest was to produce a drawing for each of the given challenge graphs with as high crossing resolution as possible (which actually also formed our primary motivation for this work).
+
+%\myparagraph{Graph Drawing Contest 2017} 
+In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
+%
+Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.
+
+It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''.
+Our team ``\texttt{TuebingenMidnight}''  miserably performed; we actually took the third position out of a total of four teams that participated.
+As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$.
+In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
+
+We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the contest's winner when restricted within the one hour.
+On a few graphs (namely, graphs 10 and 11), we even achieved better results.
+With a single exception (namely, graph 4), we could easily outperform the results that we achieved with the spring-embedder algorithm.
+If we neglect the time restriction, then for all the graphs, the results  are (sometimes considerably) better than or at least about the same as the contest's winner.
+
+\begin{table*}
+\caption{Summary of the results for the Graph Drawing Contest 2018.}
+\centering
+\begin{tabular}{c|c|c|c|c}
+\toprule
+\rowcolor{gray!30!} ~~Graph~~ & ~~Grid size~~ & ~~T\"ubingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
+\midrule
+\rowcolor{gray!30!} 1  & 100 $\times$ 100             & 87$^\circ$ & 58$^\circ$ & 89.99$^\circ$\\
+\rowcolor{gray!00!} 2  & 100 $\times$ 100             & 90$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!30!} 3  & 150 $\times$ 150             & 88$^\circ$ & 85$^\circ$ & 89.95$^\circ$\\
+\rowcolor{gray!00!} 4  & 200 $\times$ 200             & 88$^\circ$ & 60$^\circ$ & 89.05$^\circ$\\
+\rowcolor{gray!30!} 5  & 200 $\times$ 200             & 71$^\circ$ & 48$^\circ$ & 78$^\circ$\\
+\rowcolor{gray!00!} 6  & 200 $\times$ 200             & 89$^\circ$ & 71$^\circ$ & 89.96$^\circ$\\
+\rowcolor{gray!30!} 7  & 250 $\times$ 250             & 80$^\circ$ & 43$^\circ$ & 63.62$^\circ$\\
+\rowcolor{gray!00!} 8  & 10,000 $\times$ 10,000       & 34$^\circ$ & 27$^\circ$ & 43$^\circ$\\
+\rowcolor{gray!30!} 9  & 10,000 $\times$ 10,000       & 77$^\circ$ & 60$^\circ$ & 83$^\circ$\\
+\rowcolor{gray!00!} 10 & 30,000 $\times$ 30,000       & 13$^\circ$ & 19$^\circ$ & 24$^\circ$\\
+\rowcolor{gray!30!} 11 & 50,000 $\times$ 50,000       & 11$^\circ$ & 22$^\circ$ & 72.00$^\circ$\\
+\rowcolor{gray!00!} 12 & 100,000 $\times$ 100,000     & 74$^\circ$ &  9$^\circ$ & 80$^\circ$\\
+\rowcolor{gray!30!} 13 & 1,000,000 $\times$ 1,000,000 &  2$^\circ$ &  2$^\circ$ & 33.68$^\circ$\\
+\rowcolor{gray!00!} 14 & 1,000,000 $\times$ 1,000,000 &  3$^\circ$ & 11$^\circ$ & 11$^\circ$\\
+\bottomrule
+\end{tabular}
+\label{tab:gdContest2018}
+\end{table*}
+
+%\myparagraph{Graph Drawing Contest 2018} 
+In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper).
+The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''.
+Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.
+
+Note that in year 2018, the organizers of the contest posed one additional constraint, namely, the drawing area was restricted to a certain size, which was different for each graph; see column ``Grid size'' in Table~\ref{tab:gdContest2018}.
+It is eye-catching that our team was the winning one; team \texttt{CoffeeVM+} took the second position of the contest (out of a total of four teams that participated to the contest).
+It is worth noting that for almost all $14$ challenge graphs our team achieved significantly better results than those of the second team \texttt{CoffeeVM+}, which clearly indicates that our new approach is very competitive (and definitely better than our previous,  as also evidenced by the experimental evaluation).
+
+\paragraph{Structure of the paper:}
+The remainder of this paper is structured as follows. 
+%Section~\ref{sec:relatedwork} overviews related works. 
+Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.
 
@@
 
-\begin{table*}
-\caption{Summary of the results for the Graph Drawing Contest 2017.}
-\centering
-\begin{tabular}{c|c|c|c|c}
-\toprule
-~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
-\midrule
-1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$ & 89.99$^\circ$\\
-2 & 88.23$^\circ$  & 42$^\circ$ & 88.21$^\circ$ & 88.70$^\circ$\\
-3 & 90.00$^\circ$  & 89$^\circ$ & 87.86$^\circ$ & 89.95$^\circ$\\
-4 & 88.97$^\circ$  & 89$^\circ$ & 77.13$^\circ$ & 89.05$^\circ$\\
-5 & 80.40$^\circ$  & 30$^\circ$ & 78.68$^\circ$ & 86.96$^\circ$\\
-6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$ & 89.96$^\circ$\\
-7 & 56.53$^\circ$  & 34$^\circ$ & 55.77$^\circ$ & 63.62$^\circ$\\
-8 & 84.95$^\circ$  & 61$^\circ$ & 81.18$^\circ$ & 89.28$^\circ$\\
-9 & 59.88$^\circ$  & 9$^\circ$  & 54.63$^\circ$ & 88.20$^\circ$\\
-10 & 20.97$^\circ$ & 4$^\circ$  & 23.60$^\circ$ & 23.72$^\circ$\\
-11 & 46.68$^\circ$ & 6$^\circ$  & 57.00$^\circ$ & 72.00$^\circ$\\
-12 & 36.47$^\circ$ & 5$^\circ$  & 26.24$^\circ$ & 35.86$^\circ$\\
-13 & 25.45$^\circ$ & 4$^\circ$  & 22.43$^\circ$ & 33.68$^\circ$\\
-14 & 33.52$^\circ$ & 5$^\circ$  & 29.69$^\circ$ & 43.08$^\circ$\\
-15 & 20.51$^\circ$ & 4$^\circ$  & 13.51$^\circ$ & 29.18$^\circ$\\
-16 & 19.83$^\circ$ & 3$^\circ$  & 12.13$^\circ$ & 26.12$^\circ$\\
-\bottomrule
-\end{tabular}
-\label{tab:gdContest2017}
-\end{table*}
-
-\begin{table*}
-\caption{Summary of the results for the Graph Drawing Contest 2018.}
-\centering
-\begin{tabular}{c|c|c|c}
-\toprule
-~~Graph~~ & ~~TuebingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
-\midrule
-1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$\\
-2 & 88.23$^\circ$  & 42$^\circ$ & 88.70$^\circ$\\
-3 & 90.00$^\circ$  & 89$^\circ$ & 89.95$^\circ$\\
-4 & 88.97$^\circ$  & 89$^\circ$ & 89.05$^\circ$\\
-5 & 80.40$^\circ$  & 30$^\circ$ & 86.96$^\circ$\\
-6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$\\
-7 & 56.53$^\circ$  & 34$^\circ$ & 63.62$^\circ$\\
-8 & 84.95$^\circ$  & 61$^\circ$ & 89.28$^\circ$\\
-9 & 59.88$^\circ$  & 9$^\circ$  & 88.20$^\circ$\\
-10 & 20.97$^\circ$ & 4$^\circ$  & 23.72$^\circ$\\
-11 & 46.68$^\circ$ & 6$^\circ$  & 72.00$^\circ$\\
-12 & 36.47$^\circ$ & 5$^\circ$  & 35.86$^\circ$\\
-13 & 25.45$^\circ$ & 4$^\circ$  & 33.68$^\circ$\\
-14 & 33.52$^\circ$ & 5$^\circ$  & 43.08$^\circ$\\
-\bottomrule
-\end{tabular}
-\label{tab:gdContest2018}
-\end{table*}
-
 While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
@@
 Finally, the choice of the initial grid drawing seems to affect the performance of our algorithm, both with respect to the quality of the produced drawings (counted here in terms of the crossing resolution), but also with respect to the number of iterations needed for convergence.
-
-% ==================================================================
-\subsection{Graph Drawing Contests 2018 and 2019}
-\label{ssec:contest}
-% ==================================================================
-
-Every year, the Graph Drawing conference organizes a contest, in which the task is to visualize a set of challenge graphs in a predefined drawing style.
-The teams participating to the contest receive a set of challenge graphs, and after one hour they must submit their final drawings for each of them.
-The team with the highest cumulative score wins.
-In years 2017 and 2018, the task of the contest was to produce a drawing for each of the given challenge graphs with as high crossing resolution as possible (which actually also formed our primary motivation for this work).
-
-\myparagraph{Graph Drawing Contest 2017} In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
-%
-Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.
-
-It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''.
-Our team ``\texttt{TuebingenMidnight}''  miserably performed; we actually took the third position out of a total of four teams that participated.
-As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$.
-In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
-
-We can observe that for almost all graphs, our new approach achieves only slightly worse results than the ones of the contest's winner when restricted within the one hour.
-On a few graphs (namely, graphs 10 and 11), we even achieved better results.
-With a single exception (namely, graph 4), we could easily outperform the results that we achieved with the spring-embedder algorithm.
-If we neglect the time restriction, then for all the graphs, the results  are (sometimes considerably) better than or at least about the same as the contest's winner.
-
-\myparagraph{Graph Drawing Contest 2018} In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper).
-The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''.
-Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.
-
-It is eye-catching that our team was the winning one; team \texttt{CoffeeVM+} took the second position of the contest (out of a total of four teams that participated to the contest).
-It is worth noting that for almost all $14$ challenge graphs our team achieved significantly better results than those of the second team \texttt{CoffeeVM+}, which clearly indicates that our new approach is very competitive (and definitely better than our previous,  as also evidenced by the experimental evaluation).
 

### 2018-10-05 paper/journal.tex
added our best results for new contest graphs
Updates; adding a reference to Tamara's paper

@@
 \centering
-\begin{tabular}{c|c|c|c|c}
+\begin{tabular}{c|c|c|c|c|c|c}
 \toprule
-\rowcolor{gray!30!}~~Graph~~ & ~~CoffeeVM~~ & ~~T\"ubingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
+\rowcolor{gray!30!}~Graph~ & ~no. of vertices~ & ~no. of edges~ & ~CoffeeVM~ & ~T\"ubingenMidnight~ & ~Time restricted~ & ~Our best~ \\
 \midrule
-\rowcolor{gray!30!} 1  & 90$^\circ$  & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
-\rowcolor{gray!00!} 2  & 88$^\circ$  & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
-\rowcolor{gray!30!} 3  & 90$^\circ$  & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
-\rowcolor{gray!00!} 4  & 89$^\circ$  & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
-\rowcolor{gray!30!} 5  & 80$^\circ$  & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
-\rowcolor{gray!00!} 6  & 90$^\circ$  & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
-\rowcolor{gray!30!} 7  & 57$^\circ$  & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
-\rowcolor{gray!00!} 8  & 85$^\circ$  & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
-\rowcolor{gray!30!} 9  & 60$^\circ$  & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
-\rowcolor{gray!00!} 10 & 21$^\circ$  & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
-\rowcolor{gray!30!} 11 & 47$^\circ$  & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
-\rowcolor{gray!00!} 12 & 36$^\circ$  & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
-\rowcolor{gray!30!} 13 & 25$^\circ$  & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
-\rowcolor{gray!00!} 14 & 34$^\circ$  & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
-\rowcolor{gray!30!} 15 & 21$^\circ$  & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
-\rowcolor{gray!00!} 16 & 20$^\circ$  & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
+\rowcolor{gray!30!} 1  &  9  & 18    & 90$^\circ$ & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!00!} 2  & 12  & 36    & 88$^\circ$ & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 3  & 16  & 42    & 90$^\circ$ & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!00!} 4  & 17  & 54    & 89$^\circ$ & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 5  & 30  & 95    & 80$^\circ$ & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
+\rowcolor{gray!00!} 6  & 40  & 75    & 90$^\circ$ & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!30!} 7  & 80  & 254   & 57$^\circ$ & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
+\rowcolor{gray!00!} 8  & 85  & 136   & 85$^\circ$ & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 9  & 149 & 258   & 60$^\circ$ & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
+\rowcolor{gray!00!} 10 & 174 & 344   & 21$^\circ$ & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
+\rowcolor{gray!30!} 11 & 250 & 400   & 47$^\circ$ & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
+\rowcolor{gray!00!} 12 & 300 & 987   & 36$^\circ$ & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
+\rowcolor{gray!30!} 13 & 392 & 562   & 25$^\circ$ & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
+\rowcolor{gray!00!} 14 & 510 & 820   & 34$^\circ$ & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
+\rowcolor{gray!30!} 15 & 650 & 2,173 & 21$^\circ$ & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
+\rowcolor{gray!00!} 16 & 998 & 1,596 & 20$^\circ$ & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
 \bottomrule
@@
 
-Of more general scope are, of course, the already mentioned works by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, which are also based on the force-directed technique.
+Of more general scope are, of course, the already mentioned works by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, which are based on the force-directed technique.
 
@@
 
-Every year, the Graph Drawing conference organizes a contest, in which the task is to visualize a set of challenge graphs in a predefined drawing style.
+Every year, the Graph Drawing conference organizes a contest, in which the task is to visualize a set of challenge graphs in a prescribed drawing style.
 The teams participating to the contest receive a set of challenge graphs, and after one hour they must submit their final drawings for each of them.
 The team with the highest cumulative score wins.
-In years 2017 and 2018, the task of the contest was to produce a drawing for each of the given challenge graphs with as high crossing resolution as possible (which actually also formed our primary motivation for this work).
+%
+In years 2017 and 2018, the task of the contest was to produce a drawing for each of the given challenge graphs with as high crossing resolution as possible (i.e., ignoring other aesthetic measures on the produced drawings).
+This task was actually our primary motivation for this work.
 
 %\myparagraph{Graph Drawing Contest 2017} 
-In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
+In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}).
 %
@@
 
-It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''.
+It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''; details for their approach can be found in~\cite{DBLP:journals/corr/abs-1807-09483}.
 Our team ``\texttt{TuebingenMidnight}''  miserably performed; we actually took the third position out of a total of four teams that participated.
@@
 If we neglect the time restriction, then for all the graphs, the results are (sometimes considerably) better than or at least about the same as the contest's winner.
+Nevertheless, what we realized is that the spring-embedder algorithm was not efficient enough for participating to the contest.
 
@@
 \centering
-\begin{tabular}{c|c|c|c|c}
+\begin{tabular}{c|c|c|c|c|c|c}
 \toprule
-\rowcolor{gray!30!} ~~Graph~~ & ~~Grid size~~ & ~~T\"ubingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
+\rowcolor{gray!30!} Graph & no. of vertices & no. of edges  & Grid size & T\"ubingenColdShower & CoffeeVM+ & Our best \\
 \midrule
-\rowcolor{gray!30!} 1  & 100 $\times$ 100             & 87$^\circ$ & 58$^\circ$ & 89.99$^\circ$\\
-\rowcolor{gray!00!} 2  & 100 $\times$ 100             & 90$^\circ$ & 90$^\circ$ & 90$^\circ$\\
-\rowcolor{gray!30!} 3  & 150 $\times$ 150             & 88$^\circ$ & 85$^\circ$ & 89.95$^\circ$\\
-\rowcolor{gray!00!} 4  & 200 $\times$ 200             & 88$^\circ$ & 60$^\circ$ & 89.05$^\circ$\\
-\rowcolor{gray!30!} 5  & 200 $\times$ 200             & 71$^\circ$ & 48$^\circ$ & 78$^\circ$\\
-\rowcolor{gray!00!} 6  & 200 $\times$ 200             & 89$^\circ$ & 71$^\circ$ & 89.96$^\circ$\\
-\rowcolor{gray!30!} 7  & 250 $\times$ 250             & 80$^\circ$ & 43$^\circ$ & 63.62$^\circ$\\
-\rowcolor{gray!00!} 8  & 10,000 $\times$ 10,000       & 34$^\circ$ & 27$^\circ$ & 43$^\circ$\\
-\rowcolor{gray!30!} 9  & 10,000 $\times$ 10,000       & 77$^\circ$ & 60$^\circ$ & 83$^\circ$\\
-\rowcolor{gray!00!} 10 & 30,000 $\times$ 30,000       & 13$^\circ$ & 19$^\circ$ & 24$^\circ$\\
-\rowcolor{gray!30!} 11 & 50,000 $\times$ 50,000       & 11$^\circ$ & 22$^\circ$ & 72.00$^\circ$\\
-\rowcolor{gray!00!} 12 & 100,000 $\times$ 100,000     & 74$^\circ$ &  9$^\circ$ & 80$^\circ$\\
-\rowcolor{gray!30!} 13 & 1,000,000 $\times$ 1,000,000 &  2$^\circ$ &  2$^\circ$ & 33.68$^\circ$\\
-\rowcolor{gray!00!} 14 & 1,000,000 $\times$ 1,000,000 &  3$^\circ$ & 11$^\circ$ & 11$^\circ$\\
+\rowcolor{gray!30!} 1  & 10    & 32    & 100 $\times$ 100             & 87$^\circ$ & 58$^\circ$ & 88$^\circ$\\
+\rowcolor{gray!00!} 2  & 12    & 24    & 100 $\times$ 100             & 90$^\circ$ & 90$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!30!} 3  & 18    & 49    & 150 $\times$ 150             & 88$^\circ$ & 85$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!00!} 4  & 23    & 46    & 200 $\times$ 200             & 88$^\circ$ & 60$^\circ$ & 89$^\circ$\\
+\rowcolor{gray!30!} 5  & 30    & 91    & 200 $\times$ 200             & 71$^\circ$ & 48$^\circ$ & 78$^\circ$\\
+\rowcolor{gray!00!} 6  & 40    & 68    & 200 $\times$ 200             & 89$^\circ$ & 71$^\circ$ & 90$^\circ$\\
+\rowcolor{gray!30!} 7  & 50    & 94    & 250 $\times$ 250             & 80$^\circ$ & 43$^\circ$ & 83$^\circ$\\
+\rowcolor{gray!00!} 8  & 102   & 400   & 10,000 $\times$ 10,000       & 34$^\circ$ & 27$^\circ$ & 43$^\circ$\\
+\rowcolor{gray!30!} 9  & 128   & 256   & 10,000 $\times$ 10,000       & 77$^\circ$ & 60$^\circ$ & 83$^\circ$\\
+\rowcolor{gray!00!} 10 & 500   & 1,534 & 30,000 $\times$ 30,000       & 13$^\circ$ & 19$^\circ$ & 24$^\circ$\\
+\rowcolor{gray!30!} 11 & 709   & 1,602 & 50,000 $\times$ 50,000       & 11$^\circ$ & 22$^\circ$ & 18$^\circ$\\
+\rowcolor{gray!00!} 12 & 1,800 & 6,961 & 100,000 $\times$ 100,000     & 74$^\circ$ &  9$^\circ$ & 80$^\circ$\\
+\rowcolor{gray!30!} 13 & 1,500 & 4,467 & 1,000,000 $\times$ 1,000,000 &  2$^\circ$ &  2$^\circ$ & 4$^\circ$\\
+\rowcolor{gray!00!} 14 & 3,000 & 5,899 & 1,000,000 $\times$ 1,000,000 &  3$^\circ$ & 11$^\circ$ & 11$^\circ$\\
 \bottomrule

### 2018-10-05 paper/journal.tex
Revising a bit the conclusion

@@
 
-%
 In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms, even when restricted to a given maximum aspect ratio.
+Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms (even when restricted to a given maximum aspect ratio).
+For the latter algorithms, our evaluation indicates that their performance is good, only when several aesthetic criteria are taken into account.
 
@@
 
-We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm.
+%We conclude by noting that our motivation to work with this problem was our participation to the Graph Drawing contest 2017, where we performed miserably using a force-directed algorithm.
 As our evaluation shows, the performance of such algorithms is good, only when several aesthetic criteria are taken into account.

### 2018-10-05 paper/journal.tex
Final version

@@
 
-\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution%
-\thanks{A preliminary version of this work has appeared in the Proceedings of the 26th International Symposium on Graph Drawing and Network Visualization (GD 2018).
-This work is supported by the DFG grant Ka812/17-1}}
+\title[Drawings of Graphs with High Crossing Resolution]{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}
 
@@
 
-\paragraph{Acknowledgments.} The authors would also like to thank Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
+\paragraph{Acknowledgments.} The authors would like to thank Simon Wegendt and Jessica Wolz for implementing the first version of our prototype.
+This work is supported by the DFG grant Ka812/17-1.
+A preliminary version of this work has appeared in the Proceedings of the 26th International Symposium on Graph Drawing and Network Visualization (GD 2018).
 

### 2019-07-25 paper/comments/reply-letter.tex
Starting a reply letter and addressing few comments

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage[a4paper, total={6in, 9in}]{geometry}
\usepackage{soul}
\usepackage{enumerate}
\usepackage{xcolor,url}

\newcommand{\rcomment}[1]{\vspace{0.3cm} \item \textbf{Reviewer's Comment:} {\em #1}}
\newcommand{\tcomment}[1]{\vspace{0.3cm} {\color{red} \item \textbf{Reviewer's Comment:} {\em #1}}}
\newcommand{\jcomment}[1]{\vspace{0.3cm} {\color{blue} \item \textbf{Reviewer's Comment:} {\em #1}}}

\newcommand{\response}{\vspace{0.2cm} \textbf{Response: }}
\newcommand{\jresponse}[1]{\vspace{0.2cm} \textbf{Response: }{\color{blue} {\em #1}}}

\begin{document}

% ==================================================================================
\section*{Reviewer 1}
% ==================================================================================

The authors present a heuristic that aims to maximize the crossing resolution of graph. Their heuristic can be adopted to also aim towards maximizing angular and total resolution. The authors present an extensive experimental analysis which demonstrates that their heuristic
outperforms the previously known approaches ([21,22], which are both ``spring-embedder'' like heuristics). The authors also demonstrate that their approach is efficient, resulting results in reasonable
amounts of time. On the practical side they also produce grid drawings of restricted size, which is something previously known algorithms did not address.

The paper is well written. The heuristic is presented in a clear and simple way so that it is possible to be programmed by other researchers/developers. Sufficients details are also provided
on how to run each step of the heuristic efficiently. Given the limited applied previous work on the topic (most of the previous work focuses on the case of RAC graphs), i believe that their contribution is significant and it is definetely worth publishing. I suggest ACCEPTANCE subject to addressing the following

MINOR issues:

\begin{itemize}
\rcomment{pg 2, ln 19, right. Remove ``Unless otherwise specified''. No directed or non-simple graphs are considered.}

\response{Done.}

\rcomment{pg 3, ln 30 right. change to ''... performed miserably''}

\response{Done.}

\rcomment{Table 1: Indicate in the caption or in the column headings that the last 2 columns correspond to the results presented in this paper.}

\response{Done.}

\rcomment{Table 2: Use a better heading than "our best" (for example, "no-time restriction"?)

\response{Revised as suggested.}

\rcomment{pg 4, ln 41, left: Mention Section 4}

\response{Done.}

\rcomment{pg 4, ln 60, left: The text followin "In other words.." is not implied by the formula. The formula should be: $c(\Gamma_{k}) \geq c(\Gamma_{k-1}), ~~0 \leq k \leq i-1$ or similar.}

\response{Done.}

7. pg 4, ln 46, right: "or of a vertex that is in the neighborhood of the critical vertex..."
Why is this true in general? The edges incident to vertices in the neighborhood are not related
to the
crossing considered.

8. pg 4, ln 52, right: If possible, provide an intuition why you are using graph-distance instead of
geometric-distance.
9. pg 5, ln 17, left: "chooses uniformly at random". This is not accurate. You are not using the
uniform distribution.
In the previous page you assigned probalilities to the vertices wrt their graph-distance from the
critical
vertex. Are you applying this rule. Please clarify.
10. pg 5. ln 30 left: Mention that $\rho$ is also a parameter of the algorithm.
11. General comment for section 2. Mention early in the section that your heuristic, as described,
produces real coordinates
and that you explain in Sect 2.3 how to produce a grid drawing. (Having seen Table 2, a reader
expects to get an algorithm that
produces grid drawing at the first place).
12. General comment for section 2. Have a discussion regarding the real-number computations
involved. How many decimal points used
for crossing resolution, vertex position, etc.
13. Pg 6, ln 11, right: "Aspect ratio" This is not the traditional definition. Aspect ratio is
typically defined as the ration of
largest to smaller side-length of the enclosing renctange of the drawing. The smaller edge to
largest edge ratio is more of a
"resolution" type measurement.
14. Figures 4,5,7,8. In caption (a) you state that it is the crossing resolution. In these figures it is
either angular of total
resolution.
15. Figure 9. Clearly state that the double line refers to an algorithm that is unrestricted wrt the
grid size. (So far, you have
used the term "unrestricted" wrt the time devoted to each run; the user then expectes to have 4
such lines, one per grid size).
16. pg 16, line 38, left: "naturalLY observe".
17. General comment: It should be also interesting to compare the crossing resolution between
the "non-grid" and the "grid"
versions of your algorithm

% ==================================================================================
\newpage
\section*{Reviewer 2}
% ==================================================================================

\st{In graph drawing, book embeddings represent an intuitive way to draw graphs. Vertices are ordered along a /spine/, and edges are embedded as arcs on /pages/ without crossings. Book embeddings have been thoroughly studied, however several important questions still remain. In the present paper, the authors study a more restrictive variant of book embeddings, which requires each page to have no edges incident to the same vertex. Such an embedding is called a /dispersable/ book embedding.}

\st{The authors conduct a thorough study of this variant, centered on disproving a long-standing open conjecture from the original paper on dispersable book embedings by Bernhart and Kainen from 1979: that any $k$-regular bipartite graph is /dispersable/ (i.e., it has a dispersable book embedding on $k$ pages). The present authors successfully disprove the conjecture for $k=3$ and $k=4$; however, it is still unknown for $k>4$. They further show that all 3-connected 3-regular planar bipartite graphs are dispersable, and conjecture that all 3-regular planar bipartite graphs are dispersable.}

\st{This represents a major step in furthering research on dispersable book embeddings, and has the potential to spur on further research in the area. Although a majority of the paper is dedicated to a heavy (and somewhat tedious) case analysis, the result represents a significant contribution to graph drawing and, by extension, combinatorics. Lastly, aside from some minor grammatical errors, the paper is written rather well.}

\st{Contributions, Techniques and Novelty}

\st{The paper can be broken down into 3 main results, which I will address in turn:}

\begin{enumerate}

\item \st{First, the authors disprove the conjecture for $k=3$. That is, they prove that not all 4-regular bipartite graphs are dispersable. They prove this by finding one 4-regular bipartite graph, namely the Folkman graph, that is not dispersable. Their proof that this graph is not dispersable is a nontrivial, case-heavy analysis. The proof relies on the fact that a graph has a $k$-page book embedding if and only if it also has a circular embedding with a $k$-edge coloring, such that no two edges with the same color cross. And furthermore there is a simple transformation from one to the other. The authors argue using the structure of such a circular layout (and the fact that crossing edges do not have the same color), and show that forbidden patterns in the ordering of vertices around the circular layout emerge as a result. The end result of these forbidden patterns is that a $4$-page dispersable book embedding does not exist.}

\item \st{Next, the authors briefly discuss a computer-assisted proof that the conjecture fails also for $k=3$. The authors investigate the /Gray/ graph, and use an extension of a recently introduced SAT formulation to check that it does not have a $3$-page dispersable book embedding.}

\item \st{Finally, and perhaps most interestingly, the authors show that the conjecture is true for $k=3$ when the graph is additionally 3-connected and planar. This argument relies on the existence of 3-coloring of the edges, and  a 3-coloring of the vertices in the dual graph (that is, a coloring of the faces), such that faces have a color different from any of the edges incident to the face. This proof is a refreshing departure from the heavy case analysis given for the first result, and in my opinion, is also more appealing than the computer-aided proof.}
\end{enumerate}

Minor Comments

\begin{itemize}
\rcomment{Line 6: ``equals to the maximum'' $\rightarrow$ ``equals the maximum''}

\response{Done.}

\rcomment{Lines 41/42: ``The book thickness is known to be bounded only for bounded genus graphs.'' $\rightarrow$ Suggest ``...only known to be bounded for bounded genus graphs.'' Otherwise, the former is stating that it is known that only bounded genus graphs have bounded book thickness. Is this intentional?}

\response{Added e.g. instead.}

\rcomment{Lines 132/133 ``there is no twins'' $\rightarrow$ ``there are no twins''}

\response{Done.}

\rcomment{Line 136: ``twin pairs'' $\rightarrow$ Is this different from ``twins'' or are they equivalent? I notice several uses of both, and I’m not sure if the meanings are distinct.}

\response{Changed a bit to make it more clear.}

\rcomment{Line 153: I suggest preceding $(A_2,ab)$ with ``edges'' since these also resemble open intervals.}

\response{Done, but only in the journal version due to space limitation.}

\rcomment{Line 169 (and others): for each declaration of a forbidden pair, I was initially confused by the center dots in the notation. Are they meant to match a single connector vertex? Please mention this explicitly somewhere.}

\response{Done in the journal version.}

\rcomment{Line 179 ``there are not exactly two same twin vertices'' $\rightarrow$ I found this confusing, and the same can be said of several of the forbidden pair definitions. Do you mean ``there is not a single twin pair'' or even ``there is not a single twin''. Perhaps clarifying the usage of the word twin would be helpful.}

\response{Done.}

\rcomment{For all forbidden pairs, A and B are used; however they apply without loss of generality to any distinct twins. That should perhaps be stated somewhere.}

\response{Come on... that should be clear.}

\rcomment{Line 206: ``no a A’s connector'' $\rightarrow$ ``no connector of A''. There are many uses like this, I suggest carefully searching the manuscript to find all occurrences. (See further, 207 ``a A’s connectors'', 208 ``the B’s connectors'', footnote 4: ``there are C’s connectors'', and others.)}

\response{Done.}

\rcomment{Line 247: ``Each of X,Y and U,V are different.'' The exact meaning is not clear to me. Is it that X and Y are not twins? and U,V are not twins?}

\response{Done.}

\rcomment{Line 285: It took me a while to realize that $d(A_1,A_2)$ includes $A_1$ and $A_2$, which results in $d(A)$ counts.}

\response{Added a note in the journal version.}

\rcomment{Line 298: ``By Lemma 3 applied for A and C'' $\rightarrow$ ``Applying Lemma 3 for A and C''}

\response{Done.}

\rcomment{Line 306: ``We study the book thickness of'' $\rightarrow$ ``We study the dispersable book thickness of''}

\response{Done.}

\rcomment{Line 308/309: I highly suggesting fitting image 9 in the paper; it is very helpful to visualize the Gray graph}

\response{Come on... that's impossible.}

\rcomment{Line 314: ``From this formulation'' $\rightarrow$ Based on the previous sentence, it is not clear if you are referring to the original SAT formulation, or your modified SAT formulation.}

\response{Done.}

\rcomment{Lines 326-331: It is worth mentioning how the computer actually assisted; what I think happened is that you ran a SAT solver, and it did not find $3$-page dispersable book embedding, and therefore there does not exist one. This deserves to be explicitly stated. Did it take long to verify?}

\response{Revised in the journal version.}

\rcomment{Line 333: I believe there should be a Corollary here as a result of Theorem 2, indicating that the conjecture is false for $k=3$.}

\response{It's exactly like this in the journal version.}

\rcomment{Line 334: Capitalize connected and regular}

\response{Done.}

\rcomment{References: At least references 9, 11 and 12 have minor errors. in [9], the journal is called ``SIAM Journal of Algebraic and Discrete Methods''. In [11] and [12] The same author is abbreviated differently in each, and in [11], ``Geomentry'' $\rightarrow$ ``Geometry''. Please thoroughly inspect the references.}

\response{Done.}
\end{itemize}

% ==================================================================================
\newpage
\section*{Reviewer 3}
% ==================================================================================

\st{The paper deals with dispersable book embeddings, i.e. classical book embeddings (the vertices are ordered along a line l (the spine), the edges are drawn crossing-free at different half-planes (pages) bounded by l), with the additional property that the graphs induced by the edges of each page are 1-regular. It is clear that the number of pages needed for a dispersable book embedding of a graph G is at least the maximum degree of G. If this number of pages suffices, the graph G is called dispersable. A conjecture of Barnhart and Kainen from 1979 states that each k-regular bipartite graph is dispersable
(for k=1,2 this is clearly true).}

\st{In the paper the following three things are shown:}
\begin{enumerate}[(a)]
\item \st{The conjecture of Barnhart and Kainen is disproven for the case k=3 with a computer-aided proof.}
\item \st{The conjecture of Barnhart and Kainen is disproven for the case k=4 with a hand-written proof.}
\item \st{It is proven that each 3-connected 3-regular bipartite planar graph is dispersable.}
\end{enumerate}

\begin{enumerate}[(a)]
\item \st{There is a known reduction of the problem of finding a book embedding with a given number of pages to SAT.
   The authors extend this reduction to the case of dispersable book embeddings what is really simple and straight-forward.
   Then they say that computer experiments showed that there is no 3-page dispersable book embedding of the 47-vertex
   Gray graph. So the main contribution in this part is that the authors have found a graph for which the computer
   program gave a negative answer, and not the computer program itself.}
\item \st{It is proven that there is no 4-page dispersable book embedding of the 20-vertex Folkman graph.
   The proof works as follows: By proving forbidden patterns for the order of the vertices on the spine
   of the book embedding gradually all permutations of the vertices are excluded as orders on the spine.
   This leads to a lot of case distinctions over more than 13 pages. I read the part of the proof in the main part,
   but not the remaining part in the appendix. There are some inaccuracies in the proof, especially in the part where
   it is shown that the forbidden patterns indeed forbid every permutation. In the end, I did not find any mistakes which cannot be fixed easily, but it is hard to check its correctness as part of a converence review because it is very
   technical.}
\item \st{First the authors show that 3-connected 3-regular bipartite planar graphs have a coloring of the edges
   and faces with certain properties. Afterwards this coloring is used for a recursive geometric construction
   of the 3-page dispersable book embedding of the graph.}
\end{enumerate}

\st{Evaluation:}

\st{With results A and B the authors make a good progress in the area of book embeddings. These results are interesting
and should be published. Unfortunately, these results are not suitable for a presentation at a conference
since result A is proven by the computer and result B is proven by a long series of case distinctions without
good structural insight for other graphs than the considered Folkman graph.
After disproving the conjecture of Barnhart and Kainen, result C is an appropriate weaker variant to consider.
The proof seems to be correct and interesting for at least parts of the WG community.
Since the focus of the paper does not lie on result C (it is only a small part of the paper) and result C on its own would be weak for a WG
contribution, I recommend that this paper should be published in a journal and not be presented at a
conference.}

\st{Comments for the authors:}

\begin{itemize}
\rcomment{l. 19: in early $\rightarrow$ to the early}

\response{Done.}

\rcomment{l. 39: $\rightarrow$ while all graphs with ... treewidth [11] have sublinear book thickness}

\response{Done.}

\rcomment{l. 81: avoid abbreviations at the beginning of a sentence: Figs. $\rightarrow$ Figures}

\response{Done.}

\rcomment{l. 88: later you call this construction a subdivision. why not here?}

\response{Done in the journal version.}

\rcomment{l. 118: $\rightarrow$ be the two intervals defined by the twins in a dispersable order.}

\response{Done.}

\rcomment{l. 136: $\rightarrow$ of pairs of twin pairs}

\response{Done.}

\rcomment{l. 139: pairs of non-crossing twins $\rightarrow$ non-crossing twin pairs}

\response{Done.}

\rcomment{l. 143, 145: twins $\rightarrow$ twin vertices}

\response{Done.}

\rcomment{l. 148: a $\rightarrow$ another}

\response{Done in the journal version.}

\rcomment{l. 151 ff: add reference to Figure 3}

\response{Done.}

\rcomment{l. 161: no comma after $B_2$}

\response{Done.}

\rcomment{l. 161: pairs of crossing twins $\rightarrow$ crossing pairs of twins}

\response{Done.}

\rcomment{l. 167: twins $\rightarrow$ twin vertices}

\response{Done.}

\rcomment{l. 179: including ab: add reference to Lemma 3}

\response{Done.}

\rcomment{l. 197: overfull box}

\response{Ignored}

\rcomment{l. 206: remove a}

\response{Done.}

\rcomment{l. 206: $\rightarrow$ Lemma 4 applied for}

\response{Done.}

\rcomment{l. 207: a $\rightarrow$ an}

\response{Done.}

\rcomment{l. 211: then $\rightarrow$ the edge}

\response{Done.}

\rcomment{l. 216: $[C_2,A_2]$ $\rightarrow$ $[C_2,A_1]$}

\response{Done.}

\rcomment{l. 221: no comma after that}

\response{Done.}

\rcomment{l. 224: twins $\rightarrow$ twin vertices}

\response{Done.}

\rcomment{l. 225: $\leq$ $\rightarrow$ $=$}

\response{Not done.}

\rcomment{l. 228: twin $\rightarrow$ twin vertex}

\response{Done.}

\rcomment{l. 238: later $\rightarrow$ latter}

\response{Done.}

\rcomment{l. 240: $(C_1,ac)$ $\rightarrow$ $(A_1,ac)$}

\response{Done.}

\rcomment{l. 242: overfull box}

\response{Ignored.}

\rcomment{l. 245: no comma}

\response{Done.}

\rcomment{l. 245: $\rightarrow$ there are at least two twin vertices}

\response{Done.}

\rcomment{l. 247: different $\rightarrow$ no pair of twin vertices}

\response{Done.}

\rcomment{l. 250, 251: twins $\rightarrow$ twin vertices}

\response{Done.}

\rcomment{l. 258: no comma before is}

\response{Done.}

\rcomment{l. 261: at $\rightarrow$ in}

\response{Done.}

\rcomment{l. 284 ff: before you used the notation delta and counted strictly inside. why do you change the notation?}

\response{The notation changed.}

\rcomment{l. 287: two twins are $\rightarrow$ any pair of twins is}

\response{Done.}

\rcomment{l. 289: 4 $\rightarrow$ 6}

\response{Done.}

\rcomment{l. 291: Figs. $\rightarrow$ Figures}

\response{Done.}

\rcomment{l. 292: are $\rightarrow$ is}

\response{Done.}

\rcomment{l. 292: a $\rightarrow$ an}

\response{Done.}

\rcomment{l. 294: later $\rightarrow$ latter}

\response{Done.}

\rcomment{l. 296: the order of the vertices does not coincide with Figure 4e}

\response{Done.}

\rcomment{l. 301: $y = ab.$ Here some arguments are missing. My suggestion: (a) analogously to $z \neq ab$ we get $x \neq ab$, (b) $u,w \neq ab$ since $u$ is D's connector and $w$ is E's connector, (c) $v \neq ab$ because the following 5 pairwise crossing edges: $A_1 ab$, $A_2 ab$, $B_1 ab$, $B_2 ab$, $C_1 ce$}

\response{There is no reason to add this discussion. Symmetry implies the properties.}

\rcomment{l. 321: An $\rightarrow$ A set of}

\response{Done.}

\rcomment{l. 323: add "we" after "details"}

\response{Done.}

\rcomment{l. 384: $\rightarrow$ and its right side}

\response{Done.}

\rcomment{l. 428: exist $\rightarrow$ exists}

\response{Done.}

\rcomment{l. 436: $\rightarrow$ study the dispersable}

\response{Done.}

\rcomment{Fig. 5: the dotted gray is hard to see}

\response{Done.}

\end{itemize}

% ==================================================================================
\newpage
\section*{Reviewer 4}
% ==================================================================================

\st{In this paper, the authors tackle an older open question:  Given a k-regular bipartite graph, can we split its edges into k matchings and find a vertex order such that each matching forms a planar graph when placing the vertices in this order around a cycle?}

\st{It was known that this is true for k=1,2 and conjectured that it is also true for larger k.  But, the authors disprove it for k=3 and k=4.  The proof for k=4 is combinatorial, while the proof for k=3 was done by phrasing the problem as a SAT problem and running experiments.  The authors also show that the conjecture is true for k=3 if additionally the graph is 3-connected and planar.}

\st{On the whole, this is a nice paper.  However, I was very lost in the combinatorial proof.  The authors state a crucial Lemma 2, and then use it to classify vertices as either far or near.  But I don't see how that far/near follows from the lemma.  In consequence, I could not follow the rest of the proof.  As such, I'm a bit unsure whether this main result is true.  (My guess is that it is true---the authors probably checked that with their SAT solver?---but this proof does not convince me.) That's why I ranked the paper somewhat lower.}

\st{Detailed comments to the authors:}

\begin{itemize}
\rcomment{Lemma 2 classifies a pair of twins A1,A2, and the connectors that relates to them (ab,ac,ad,ae).  But then in the next paragraph, you claim that the first configuration implies that the twins A1,A2 have to lie next to each other, with no other twin between them.  I don't see why this is true.  For example, what's wrong with the order A1,ac,B1,bc,A2,ab,...,ad....ae.  This satisfies all the conclusions for Lemma 2 and has A in the 1-3-configuration, but A1 and A2 are not close.  Neither are they far (because the way you define far means that they have to be in the 2-2-configuration).    So this is an entirely different configuration that you are omitting in all the later cases. You need to either convince me that this order is impossible, or you need to expand all the later analysis to include this possibility.}

\response{We agreed that this comment is not correct.}

\rcomment{Lemma 3 also has some issues.  Statement iii just says that the four twins are separated by A's connectors.  But then the explanatory side-phrase says something stronger, namely, that the connector betwee A1 and A2 *must* be ab.    This should be said clearly in the main statement of iii.  (Same thing for iv.)}

\response{We agreed that this comment is not correct.}
\end{itemize}

Further minor comments:

\begin{itemize}
\rcomment{p2, l40: "while sublinear book thickness have all graphs". This is not proper English, reword.  Maybe "while sublinear book thickness can be shown for all graphs..."}

\response{Revised based on some earlier comment of another reviewer.}

\rcomment{p2 (and throughout the paper):  You seem to use the term "1-regular" in a very strange sense.  Normally d-regular means that every vertex has degree *exactly* d, neither more nor less.  But you seem to equate "1-regular" with "is a matching", i.e., every vertex has degree *at*most* 1.  At the very minimum you should make this very clear.  Better even, avoid the term 1-regular and replace it with "maximum degree 1" everywhere.}

\response{MB: I believe that this is a wrong comment, but Martin suggested an easy hack which we applied.}

\rcomment{p2, l51: define "circular embedding"}

\response{Come on... We have references and also a figure thas illustrate this obvious thing.}

\rcomment{p4, l125: "with three or four A's connectors" insert "of" before "A's" (and same thing on the next line)}

\response{Done.}
\end{itemize}

% ==================================================================================
\newpage
\section*{Reviewer 5}
% ==================================================================================

\st{This is a deep paper with solutions to two long standing open questions involving Dispersable Book Embeddings (DBE).  The minimum number of pages needed for a DBE is called the dispersible book thickness (dbt) and it is clear that $dbt(G) >= max-degree-of-G$ and it was conjectured by Bernhart and Kainen that equality holds for every k-regular bipartite graph.  They showed that the conjecture holds for k =1, 2.  In this paper, counter-examples are presented for $k = 3, 4$.}

\st{The paper is well written with only a few typos}

\begin{itemize}
\rcomment{one common (see for example pg 6, line 172) is the absence of "of" between "two" and "A's connectors".}

\response{Done.}

\rcomment{Also on page 10, line 348, I suggest including a statement of the 3-color theorem.}

\response{Added a note in the journal version.}
\end{itemize}

% ==================================================================================
\newpage
\section*{Reviewer 6}
% ==================================================================================

\st{This is a nice one. The authors disprove a conjecture from the late 70's by means of a computer proof (and a somewhat weaker proof by hand). They also show that the desired property (being dispersable) is held by Barnette graphs.}

\st{The writing is very nice and the math seems sound. The only drawback of the paper might be that it fits into the scope of WG only to some extend, as no deeper connection of the result to concepts in computer science is given. This seems like a minor issue to me, though.}

\end{document}

### 2019-07-25 paper/journal.tex
Almost done with reviewer 1. One more comment to be addressed.

@@
 
-\usepackage{amsmath,amssymb}
+\usepackage{amsmath,amssymb,adjustbox}
 \usepackage{graphicx,color,pgfplots}
@@
 In this paper, we propose a new heuristic algorithm for the crossing resolution maximization problem and we experimentally compare it against the known approaches from the literature.
-Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher aspect ratio, especially when the input graph is large.
+Our experimental evaluation indicates that the new heuristic produces drawings with better crossing resolution, but this comes at the cost of slightly higher edge-length ratio, especially when the input graph is large.
 \end{abstract}
@@
 \paragraph{Preliminaries:}
-Unless otherwise specified, in this paper we consider simple undirected graphs.
+In this paper we consider simple undirected graphs.
 Let $G=(V,E)$ be such a graph.
@@
 \begin{table*}
-\caption{Summary of the results for the Graph Drawing Contest 2017.}
+\caption{Summary of the results for the Graph Drawing Contest 2017; the last two columns correspond to results presented in this paper (with and without time limitations, respectively).}
 \centering
+\begin{adjustbox}{width=1\textwidth}
 \begin{tabular}{c|c|c|c|c|c|c}
 \toprule
-\rowcolor{gray!30!}~Graph~ & ~no. of vertices~ & ~no. of edges~ & ~CoffeeVM~ & ~T\"ubingenMidnight~ & ~Time restricted~ & ~Our best~ \\
+\rowcolor{gray!30!}~Graph~ & ~no.~of vertices~ & ~no.~of edges~ & CoffeeVM & T\"ubingenMidnight & Time restricted & No-time restriction \\
 \midrule
@@
 \end{tabular}
+\end{adjustbox}
 \label{tab:gdContest2017}
@@
 It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''; details for their approach can be found in~\cite{DBLP:journals/corr/abs-1807-09483}.
-Our team ``\texttt{TuebingenMidnight}''  miserably performed; we actually took the third position out of a total of four teams that participated.
+Our team ``\texttt{TuebingenMidnight}''  performed miserably; we actually took the third position out of a total of four teams that participated.
 As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$.
-In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as our best results, which were achieved without any time limitations (see column ``Our best'').
+In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as s results, which were achieved without any time limitations (see column ``No-time restriction'').
 
@@
 \centering
+\begin{adjustbox}{width=1\textwidth}
 \begin{tabular}{c|c|c|c|c|c|c}
 \toprule
-\rowcolor{gray!30!} Graph & no. of vertices & no. of edges  & Grid size & T\"ubingenColdShower & CoffeeVM+ & Our best \\
+\rowcolor{gray!30!} Graph & no.~of vertices & no.~of edges  & Grid size & T\"ubingenColdShower & CoffeeVM+ & No-time restriction \\
 \midrule
@@
 \end{tabular}
+\end{adjustbox}
 \label{tab:gdContest2018}
@@
 Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project.
+In Section~\ref{sec:samples} we present sample drawings produced by our algorithm.
 We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.
@@
 A circular drawing or a drawing obtained by applying a force-directed algorithm on graph $G$ clearly meets this precondition.
+The output of our heuristic will be a (not necessarily grid) drawing $\Gamma$ with crossing resolution $c(\Gamma)$, such that $c(\Gamma) \geq c(\Gamma_0)$.
+In Section~\ref{ssec:variants}, we discuss how one can easily modify our heuristic so that $\Gamma$ is necessarily a grid drawing, if  the input drawing $\Gamma_0$ is a grid drawing, as well.
 
 Our algorithm is iterative and at each iteration performs some operations that are mainly based on randomization.
-At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of graph $G$ of crossing resolution $c(\Gamma_{i-1}) \geq c(\Gamma_0)$.
-In other words, we assume, as an invariant for our algorithm, that the crossing resolution cannot be decreased at some iteration of our algorithm (so, in a sense, our algorithm resembles probabilistic hill climbing approaches).
+At the $i$-th iteration, we assume that we have computed a drawing $\Gamma_{i-1}$ of graph $G$ with crossing resolution $c(\Gamma_{i-1})$.
+As an invariant for our algorithm, we further assume that for all values of $k$ with $0 \leq k \leq i-1$ the following condition holds %
+%
+$$c(\Gamma_{k-1}) \leq c(\Gamma_k).$$ 
+%
+In other words, we assume that the crossing resolution cannot be decreased at some iteration of our algorithm (so, in a sense, our algorithm resembles probabilistic hill climbing approaches), which also implies that $c(\Gamma_{i-1}) \geq c(\Gamma_{0})$.
 We proceed by choosing a vertex of $\Gamma_{i-1}$ arbitrarily at random based on the content of the so-called \emph{vertex-pool}, which may contain:
@@
 The set of critical vertices of drawing $\Gamma_{i-1}$ is then defined by the four endvertices of each critical edge-pair.
+It is worth noting at this point that several well-known problems related to precision in real-number computations may occur, while computing the set of critical vertices of drawing $\Gamma_{i-1}$.
+Here, we assume that this set has been computed correctly up to the reliability of computations in the used machine.
 
@@
 We turned this observation into an algorithmic implementation through a probabilistic random selection procedure, so that the vertices at graph-distance~$i$ from the ones of the vertex-pool have higher probability for selection than the corresponding ones at graph-distance~$j$  in $G$, when $0 \leq i<j$.
-In other words, the closer a vertex is to the critical vertices, the more likely it is to be chosen.
+In other words, the closer a vertex is to the critical vertices, the more likely it is to be chosen\footnote{Note that in approach we have chosen the graph-distance as the criterion for choosing the vertices to be moved, mainly for reasons of efficiency.
+Alternatively, one could use the Euclidean distance for this purpose.}.
 On the other hand, if the vertex-pool contains all vertices of the graph, then each vertex can be chosen with the same probability.
@@
 
-So far, we have described the main idea of our algorithm, which at each iteration chooses uniformly at random a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
+So far, we have described the main idea of our algorithm, which at each iteration chooses  a vertex of the current drawing to move (based on the content of the vertex-pool), so to improve the crossing resolution.
 Next, we describe how to compute the new position of the chosen vertex in the next drawing $\Gamma_i$.
@@
 In general, aesthetically pleasant drawings of graphs are usually the result of compromising between different aesthetic criteria.
-Towards this direction, we discuss in this section interesting variants of our algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:conclusions}): Drawings that are optimised only in terms of the crossing resolution tend to have bad aspect ratio (that is, the ratio of the longest to the shortest edge is large) and poor angular resolution (that is, the angles formed by adjacent edges are small).
+Towards this direction, we discuss in this section interesting variants of our algorithm, which are motivated by the following observation that we made while working on this project (see Section~\ref{sec:conclusions}): Drawings that are optimised only in terms of the crossing resolution tend to have bad edge-length ratio (that is, the ratio of the longest to the shortest edge is large) and poor angular resolution (that is, the angles formed by adjacent edges are small).
 The former seems to be a consequence of the fact that drawings with good crossing resolution tend to be quite demanding in area.
@@
 % ==================================================================
-\myparagraph{Aspect ratio}
+\myparagraph{Edge-length ratio}
 % ==================================================================
 %
-Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge.
+Formally, the edge-length ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge; see, e.g.,~\cite{DBLP:journals/tcs/LazardLL19}.
 Sometimes it is also used as a measure of the area of non-grid drawings.
-It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value.
+It was easy to instruct our algorithm to prevent producing drawings with edge-length ratio either higher than the one of the starting layout or higher than a given input value.
 What we simply had to do was to reject candidate positions, which violate this precondition.
@@
 \item[--] our algorithm with $\delta_{max}=\frac{1}{2}\max\{w,h\}$, where $w$ and $h$ are the width and the height of the initial drawing, respectively, $\delta_{min}=\frac{1}{100}\delta_{max}$ and $\rho=10$,
-\item[--] our algorithm restricted not to violate the aspect ratio of the initial layout, 
+\item[--] our algorithm restricted not to violate the edge-length ratio of the initial layout, 
 \item[--] the force-directed algorithm by Argyriou et al., and
@@
 Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the aspect ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
-The variant of our algorithm, which does not violate the aspect ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{AR-restricted}, in Fig.~\ref{fig:cr-res-1}.
+It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the edge-length ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
+The variant of our algorithm, which does not violate the edge-length ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ELR-restricted}, in Fig.~\ref{fig:cr-res-1}.
 Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).
@@
 	\hfil
-	\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:cr-res-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=2]{figures/rome}}
@@
 	\centering
-	\subfloat[\label{fig:to-res-1}{Crossing resolution vs no.~of vertices}]{
+	\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=5]{figures/rome}}
 	\hfil
-	\subfloat[\label{fig:to-res-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:to-res-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=6]{figures/rome}}
@@
 	\centering
-	\subfloat[\label{fig:an-res-1}{Crossing resolution vs no.~of vertices}]{
+	\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=9]{figures/rome}}
 	\hfil
-	\subfloat[\label{fig:an-res-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:an-res-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=10]{figures/rome}}
@@
 	\hfil
-	\subfloat[\label{fig:northCrossing-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:northCrossing-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=2]{figures/north}}
@@
 	\centering
-	\subfloat[\label{fig:northTotal-1}{Crossing resolution vs no.~of vertices}]{
+	\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=5]{figures/north}}
 	\hfil
-	\subfloat[\label{fig:northTotal-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:northTotal-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=6]{figures/north}}
@@
 	\centering
-	\subfloat[\label{fig:northAngular-1}{Crossing resolution vs no.~of vertices}]{
+	\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=9]{figures/north}}
 	\hfil
-	\subfloat[\label{fig:northAngular-2}{Aspect ratio vs no.~of vertices}]{
+	\subfloat[\label{fig:northAngular-2}{Edge-length ratio vs no.~of vertices}]{
 	\includegraphics[scale=0.99,page=10]{figures/north}}
@@
 
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased aspect ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding aspect ratio of the drawings produced by the algorithm of Argyriou et al.
+While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased edge-length ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding edge-length ratio of the drawings produced by the algorithm of Argyriou et al.
 For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
 Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the aspect ratio to increase, has more or less comparable performance (in terms of aspect ratio) as the one of Huang et al.
+On the other hand, the restricted variant of our algorithm, which does not allow the edge-length ratio to increase, has more or less comparable performance (in terms of edge-length ratio) as the one of Huang et al.
 
@@
 
-Contrary to the results for the total resolution, the results for the aspect ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of this measure) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
+Contrary to the results for the total resolution, the results for the edge-length ratio show that the drawings produced by the algorithm by Huang et al.\ are better (in terms of this measure) than the drawings produced by remaining algorithms; see Fig.~\ref{fig:to-res-2}.
 %More concretely, the 
-The drawings produced by the restricted variant of our algorithm have slightly worse aspect ratios.
+The drawings produced by the restricted variant of our algorithm have slightly worse edge-length ratios.
 Then, the ones produced by the force-directed algorithm by Argyriou et al.\ follow.
-Again, we observe that our unrestricted algorithm leads to drawings with very high aspect ratio.
+Again, we observe that our unrestricted algorithm leads to drawings with very high edge-length ratio.
 
@@
 
-The results for the aspect ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
+The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
 This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
@@
 In particular, the graphs with 50 to 70 vertices appear to be harder than the graphs with 70 to 80 vertices.
-As expected, the aspect ratio of the drawings produced by both the unrestricted and the restricted variants of our algorithm depends vastly on the number of vertices of the graphs.
+As expected, the edge-length ratio of the drawings produced by both the unrestricted and the restricted variants of our algorithm depends vastly on the number of vertices of the graphs.
 However, this natural behavior cannot be observed in the resulting drawings of the two force-directed algorithms, which appear to be quite unstable resulting in a large variance.
-Again the restricted variant of our algorithm and the two force-directed algorithms produce drawings with similar aspect ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
+Again the restricted variant of our algorithm and the two force-directed algorithms produce drawings with similar edge-length ratio, which is much lower than the one of our unrestricted algorithm for larger graphs.
 All four algorithms behave nearly the same in terms of the number of crossings; see Fig.~\ref{fig:northCrossing-3}.
@@
 For larger graphs, however, the two variants of our algorithm achieve similar results, while still outperforming the algorithm by Huang et al.
-Regarding the aspect ratio and the number of crossings of the produced drawings, we observe that all algorithms have more or less the same behavior as when adjusted for maximizing the crossing resolution only.
+Regarding the edge-length ratio and the number of crossings of the produced drawings, we observe that all algorithms have more or less the same behavior as when adjusted for maximizing the crossing resolution only.
 The only exception can be observed in the algorithm by Huang et al., which performs more stable; see Figs.~\ref{fig:northTotal-2} and~\ref{fig:northTotal-3}.
@@
 Only the algorithm by Huang et al.\ seems to be mostly at last in the rank.
-Concerning the aspect ratio, we see very good behavior for the restricted variant of our algorithm and the algorithm by Huang et al., while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
+Concerning the edge-length ratio, we see very good behavior for the restricted variant of our algorithm and the algorithm by Huang et al., while the remaining two algorithms show large variance and much worse values; see Fig.~\ref{fig:northAngular-2}.
 Regarding the number of crossings, we again observe that all algorithms have similar behavior.
@@
 The test suite for this experiment was the non-planar Rome graphs.
-The algorithm to be evaluated was the variant of our algorithm that optimizes the drawing's crossing resolution (without any restriction on the aspect ratio).
+The algorithm to be evaluated was the variant of our algorithm that optimizes the drawing's crossing resolution (without any restriction on the edge-length ratio).
 However, since this algorithm is guaranteed to produce a grid drawing, only if its initial drawing is grid, each of the Rome graphs was initially laid out by randomly placing its vertices on the underlying grid, ensuring that neither two vertices nor two edges overlap.
@@
 
-Regarding the crossing resolution, we natural observe that the bigger the grid is, the easier it is to compute a drawing with high crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
+Regarding the crossing resolution, we naturally observe that the bigger the grid is, the easier it is to compute a drawing with high crossing resolution; see Fig.~\ref{fig:crossingResolutionGrid}.
 More precisely, the grid of size $10^2 \times 10^2$ was too restrictive for the vast majority of the graphs.
@@
 
-The aspect ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the $10^2 \times 10^2$ grid; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
-The fact that the aspect ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
+The edge-length ratio of the computed drawings was more or less the same regardless of the size of the underlying grid, with the exception of the drawings computed on the $10^2 \times 10^2$ grid; see Fig.~\ref{fig:crossingResolutionAspectRatioGrid}.
+The fact that the edge-length ratio of these drawings was worse can be explained of course by the fact that in most cases an improved drawing could not be reported.
 
@@
 \section{Sample Drawings}
-\label{app:samples}
+\label{sec:samples}
 
@@
 In this paper, we introduced a new heuristic aiming to produce drawings of high crossing resolution, for which we also presented variants that take into account other common aesthetic criteria in Graph Drawing.
-Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms (even when restricted to a given maximum aspect ratio).
+Our experimental evaluation indicates that the new heuristic is competitive to the state of the art force-directed algorithms (even when restricted to a given maximum edge-length ratio).
 For the latter algorithms, our evaluation indicates that their performance is good, only when several aesthetic criteria are taken into account.
@@
 In particular, in these drawings it was very common to have adjacent edges to run almost in parallel and vertices to be very close to each other (see, e.g., Fig.~\ref{fig:graph5-argyriou-cr}).
-Hence, their angular resolution and aspect ratio were often poor.
-The additional restrictions that we imposed regarding the angular resolution and the aspect ratio helped in significantly improving the readability of the produced drawings, without loosing too much of their quality in terms of the crossing resolution.
+Hence, their angular resolution and edge-length ratio were often poor.
+The additional restrictions that we imposed regarding the angular resolution and the edge-length ratio helped in significantly improving the readability of the produced drawings, without loosing too much of their quality in terms of the crossing resolution.
 
@@
 	\caption{Different drawings of the 5th challenge graph given in the Graph Drawing 2017 contest produced by different variants of
-	(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-	(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
+	(a)--(c)~the variant of our algorithm without restrictions on the edge-length ratio, 
+	(d)--(f)~the variant of our algorithm forced to maintain the input edge-length ratio,
 	(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
@@
 	\caption{Different drawings of the 9th challenge graph given in the Graph Drawing 2017 contest produced by different variants of
-	(a)--(c)~the variant of our algorithm without restrictions on the aspect ratio, 
-	(d)--(f)~the variant of our algorithm forced to maintain the input aspect ratio,
+	(a)--(c)~the variant of our algorithm without restrictions on the edge-length ratio, 
+	(d)--(f)~the variant of our algorithm forced to maintain the input edge-length ratio,
 	(g)--(i)~the algorithm by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and

### 2019-07-25 paper/comments/reply-letter.tex
Adding a cover page to the reply letter.

@@
 
+\title{{\normalsize Revision Report for the Computer Journal submission}\\
+{\Large\em ``A Heuristic Approach towards Drawings of Graphs with\\High Crossing Resolution''}}
+\author{}
+\date{}
+
+%\author{Patrizio Angelini, Michael A. Bekos, Giuseppe Liotta, Fabrizio Montecchiani}
+
+\author{M.~Bekos, H.~F\"orster, C.~Geckeler, L.~Holl\"ander, M.~Kaufmann, A.~Spallek, J.~Splett}
+
 \begin{document}
+
+\maketitle
+
+\noindent We thank all reviewers for their very helpful comments and suggestions.
+We have addressed all the raised issues.
+Details are provided below, where we list all reviews together with a detailed response to each suggestion by the reviewers.
+
+\newpage
 
@@
 Their heuristic can be adopted to also aim towards maximizing angular and total resolution.
-The authors present an extensive experimental analysis which demonstrates that their heuristic
-outperforms the previously known approaches ([21,22], which are both ``spring-embedder'' like heuristics).
-The authors also demonstrate that their approach is efficient, resulting results in reasonable
-amounts of time. On the practical side they also produce grid drawings of restricted size, which is something previously known algorithms did not address.
+The authors present an extensive experimental analysis which demonstrates that their heuristic outperforms the previously known approaches ([21,22], which are both ``spring-embedder'' like heuristics).
+The authors also demonstrate that their approach is efficient, resulting results in reasonable amounts of time.
+On the practical side they also produce grid drawings of restricted size, which is something previously known algorithms did not address.
 
-The paper is well written. The heuristic is presented in a clear and simple way so that it is possible to be programmed by other researchers/developers. Sufficients details are also provided
-on how to run each step of the heuristic efficiently.
-Given the limited applied previous work on the topic (most of the previous work focuses on the case of RAC graphs), i believe that their contribution is significant and it is definetely worth publishing.
+The paper is well written.
+The heuristic is presented in a clear and simple way so that it is possible to be programmed by other researchers/developers.
+Sufficient details are also provided on how to run each step of the heuristic efficiently.
+Given the limited applied previous work on the topic (most of the previous work focuses on the case of RAC graphs), i believe that their contribution is significant and it is definitely worth publishing.
 I suggest ACCEPTANCE subject to addressing the following
@@
 
-\rcomment{Table 2: Use a better heading than "our best" (for example, "no-time restriction"?)
+\rcomment{Table 2: Use a better heading than "our best" (for example, "no-time restriction"?)}
 
@@
 
-7. pg 4, ln 46, right: "or of a vertex that is in the neighborhood of the critical vertex..."
-Why is this true in general? The edges incident to vertices in the neighborhood are not related
-to the
-crossing considered.
+\rcomment{pg 4, ln 46, right: ``or of a vertex that is in the neighborhood of the critical vertex...'' Why is this true in general?
+The edges incident to vertices in the neighborhood are not related to the crossing considered.}
 
-8. pg 4, ln 52, right: If possible, provide an intuition why you are using graph-distance instead of
-geometric-distance.
-9. pg 5, ln 17, left: "chooses uniformly at random". This is not accurate. You are not using the
-uniform distribution.
-In the previous page you assigned probalilities to the vertices wrt their graph-distance from the
-critical
-vertex. Are you applying this rule. Please clarify.
-10. pg 5. ln 30 left: Mention that $\rho$ is also a parameter of the algorithm.
-11. General comment for section 2. Mention early in the section that your heuristic, as described,
-produces real coordinates
-and that you explain in Sect 2.3 how to produce a grid drawing. (Having seen Table 2, a reader
-expects to get an algorithm that
-produces grid drawing at the first place).
-12. General comment for section 2. Have a discussion regarding the real-number computations
-involved. How many decimal points used
-for crossing resolution, vertex position, etc.
-13. Pg 6, ln 11, right: "Aspect ratio" This is not the traditional definition. Aspect ratio is
-typically defined as the ration of
-largest to smaller side-length of the enclosing renctange of the drawing. The smaller edge to
-largest edge ratio is more of a
-"resolution" type measurement.
-14. Figures 4,5,7,8. In caption (a) you state that it is the crossing resolution. In these figures it is
-either angular of total
-resolution.
-15. Figure 9. Clearly state that the double line refers to an algorithm that is unrestricted wrt the
-grid size. (So far, you have
-used the term "unrestricted" wrt the time devoted to each run; the user then expectes to have 4
-such lines, one per grid size).
-16. pg 16, line 38, left: "naturalLY observe".
-17. General comment: It should be also interesting to compare the crossing resolution between
-the "non-grid" and the "grid"
-versions of your algorithm
+\response{This is not easy to answer precisely.
+Indeed, the vertices in the neighborhood are not related to the crossing resolution of the current drawing.
+However, by also considering these vertices (and in particular by moving them), we have observed that in several cases it is possible to gain the necessary space for the actual critical vertices, i.e., those that define the drawing's crossing resolution, to move and thus to improve the drawing's crossing resolution.
+This is the idea underneath.}
+
+\rcomment{pg 4, ln 52, right: If possible, provide an intuition why you are using graph-distance instead of geometric-distance.}
+
+\response{Of course, one could naturally use the Euclidean distance as an alternative criterion for selecting the vertices to be moved.
+Here, we have chosen the graph-distance instead, mainly for reasons of efficiency.
+In the revised version, we have added a corresponding note discussing this observation.}
+
+\rcomment{pg 5, ln 17, left: ``chooses uniformly at random''.
+This is not accurate.
+You are not using the uniform distribution.
+In the previous page you assigned probalilities to the vertices wrt their graph-distance from the critical vertex.
+Are you applying this rule.
+Please clarify.}
+
+\response{Indeed, this statement was not accurate. The quoted text has been removed in the revised version.}
+
+\rcomment{pg 5. ln 30 left: Mention that $\rho$ is also a parameter of the algorithm.}
+
+\response{In the revised version, we mention that $\rho$ is a parameter of the algorithm as follows: ``where $\rho > 0$ is an integer parameter of the algorithm''.}
+
+\rcomment{General comment for section 2.
+Mention early in the section that your heuristic, as described, produces real coordinates and that you explain in Sect 2.3 how to produce a grid drawing.
+(Having seen Table 2, a reader expects to get an algorithm that produces grid drawing at the first place).}
+
+\response{Revised according to the reviewer's suggestion.}
+
+\rcomment{General comment for section 2. Have a discussion regarding the real-number computations involved. How many decimal points used for crossing resolution, vertex position, etc.}
+
+\response{We have made a corresponding note about this well-known issue in Section~2; in particular, at the point in which we describe how to compute the set of critical vertices.
+Note that in our approach we do not adopt any special data structure from Computational Geometry to increase the precision in our computations.
+Hence, the precision depends on the used machine.}
+
+\rcomment{pg 6, ln 11, right: "Aspect ratio" This is not the traditional definition.
+Aspect ratio is typically defined as the ratio of largest to smaller side-length of the enclosing renctange of the drawing.
+The smaller edge to largest edge ratio is more of a
+"resolution" type measurement.}
+
+\response{Indeed, we were misusing the term ``aspect ratio''.
+The correct term is ``edge-length ratio''.
+In the revised version, this problem has been resolved.
+We have also added a reference to a recent paper introducing and studying this measure for outerplanar graphs (Sylvain Lazard, William J.
+Lenhart, Giuseppe Liotta: On the edge-length ratio of outerplanar graphs.
+Theor.
+Comput.
+Sci.
+770: 88-94 (2019)).}
+
+\rcomment{Figures 4,5,7,8. In caption (a) you state that it is the crossing resolution. In these figures it is either angular of total resolution.}
+
+\response{We have resolved this problem.}
+
+\rcomment{Figure 9.
+Clearly state that the double line refers to an algorithm that is unrestricted wrt the grid size.
+(So far, you have used the term "unrestricted" wrt the time devoted to each run; the user then expectes to have 4 such lines, one per grid size).}
+
+\response{In the revised version, we have revised the caption of Figure~9 to make this clear.}
+
+\rcomment{pg 16, line 38, left: "naturalLY observe".}
+
+\response{Done.}
+
+\tcomment{General comment: It should be also interesting to compare the crossing resolution between the "non-grid" and the "grid" versions of your algorithm}
+
+\response{\textcolor{red}{Which version of the algorithm do we use in the experiments?}}
+\end{itemize}
 
@@
 
-\st{In graph drawing, book embeddings represent an intuitive way to draw graphs.
-Vertices are ordered along a /spine/, and edges are embedded as arcs on /pages/ without crossings.
-Book embeddings have been thoroughly studied, however several important questions still remain.
-In the present paper, the authors study a more restrictive variant of book embeddings, which requires each page to have no edges incident to the same vertex.
-Such an embedding is called a /dispersable/ book embedding.}
+The paper studies how to compute straight-line drawings of graphs such that the angle at which two edges cross is as large as possible.
+Deciding whether a graph has a drawing where each pair of crossing edges cross at a 90-degree angle is an NP-hard problem.
+The literature contains three heuristic approaches that aim to optimize the crossing angle, two force-directed approaches (Argyriou et al, Huang et al.) and a randomized approach (Demel et al., aka CoffeeVM).
+The randomized approach has been published at the same conference as the preliminary version of the paper at hand.
+The authors participated with their approach at the annual Graph Drawing Contest 2018.
+Their approach outperformed the winner (CoffeeVM) of the 2018 edition of the contest [0,1].
 
-\st{The authors conduct a thorough study of this variant, centered on disproving a long-standing open conjecture from the original paper on dispersable book embedings by Bernhart and Kainen from 1979: that any $k$-regular bipartite graph is /dispersable/ (i.e., it has a dispersable book embedding on $k$ pages).
-The present authors successfully disprove the conjecture for $k=3$ and $k=4$; however, it is still unknown for $k>4$.
-They further show that all 3-connected 3-regular planar bipartite graphs are dispersable, and conjecture that all 3-regular planar bipartite graphs are dispersable.}
+The paper introduces a heuristic based on local modifications.
+The approach iteratively moves a vertex v to a better position.
+Since it is unclear how to efficiently compute a good position of a vertex, the authors follow a randomized approach.
+They compute a random sample S following certain restrictions and move to v to the position in S that maximizes the crossing angle of the edges incident to v.
+In order to escape local maxima, the authors introduce some parameters to control the sample space and the vertex candidates.
+The approach is evaluated in an experimental evaluation.
+First, the authors show that the approach computes larger crossing angles compared to the force-directed approaches.
+Moreover, they provide insights on the interaction between large crossing angle affects other metrics, in particular, the number of crossing and the aspect ratio of the drawing.
 
-\st{This represents a major step in furthering research on dispersable book embeddings, and has the potential to spur on further research in the area.
-Although a majority of the paper is dedicated to a heavy (and somewhat tedious) case analysis, the result represents a significant contribution to graph drawing and, by extension, combinatorics.
-Lastly, aside from some minor grammatical errors, the paper is written rather well.}
-
-\st{Contributions, Techniques and Novelty}
-
-\st{The paper can be broken down into 3 main results, which I will address in turn:}
-
-\begin{enumerate}
-
-\item \st{First, the authors disprove the conjecture for $k=3$.
-That is, they prove that not all 4-regular bipartite graphs are dispersable.
-They prove this by finding one 4-regular bipartite graph, namely the Folkman graph, that is not dispersable.
-Their proof that this graph is not dispersable is a nontrivial, case-heavy analysis.
-The proof relies on the fact that a graph has a $k$-page book embedding if and only if it also has a circular embedding with a $k$-edge coloring, such that no two edges with the same color cross.
-And furthermore there is a simple transformation from one to the other.
-The authors argue using the structure of such a circular layout (and the fact that crossing edges do not have the same color), and show that forbidden patterns in the ordering of vertices around the circular layout emerge as a result.
-The end result of these forbidden patterns is that a $4$-page dispersable book embedding does not exist.}
-
-\item \st{Next, the authors briefly discuss a computer-assisted proof that the conjecture fails also for $k=3$.
-The authors investigate the /Gray/ graph, and use an extension of a recently introduced SAT formulation to check that it does not have a $3$-page dispersable book embedding.}
-
-\item \st{Finally, and perhaps most interestingly, the authors show that the conjecture is true for $k=3$ when the graph is additionally 3-connected and planar.
-This argument relies on the existence of 3-coloring of the edges, and  a 3-coloring of the vertices in the dual graph (that is, a coloring of the faces), such that faces have a color different from any of the edges incident to the face.
-This proof is a refreshing departure from the heavy case analysis given for the first result, and in my opinion, is also more appealing than the computer-aided proof.}
-\end{enumerate}
-
-Minor Comments
+In contrast to the conference version, the full journal contains a significant amount of new material. Nevertheless, before publishing, I would like to see the following two concerns addressed.
 
 \begin{itemize}
-\rcomment{Line 6: ``equals to the maximum'' $\rightarrow$ ``equals the maximum''}
+\tcomment{First, the authors claim (Page 2, L20) to be aware of only two approaches (Argyriou et al.
+and Huang et al.) that optimize the crossing angle.
+On the other hand, the introduction contains a comparison to a third approach (CoffeeVM).
+The abstract in [1] suggests that CoffeeVM follows a similar local randomized approach.
+In my opinion, it is only fair to the authors of CoffeeVM to discuss similarities and differences between the two approaches.
+Moreover, the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems.
+I propose to add [1-6] as relevant related work.}
 
-\response{Done.}
+\response{}
 
-\rcomment{Lines 41/42: ``The book thickness is known to be bounded only for bounded genus graphs.'' $\rightarrow$ Suggest ``...only known to be bounded for bounded genus graphs.'' Otherwise, the former is stating that it is known that only bounded genus graphs have bounded book thickness.
-Is this intentional?}
+\tcomment{Second, I think that the structure of the evaluation can be improved by formulating a set of research questions. For example:
 
-\response{Added e.g. instead.}
+I) Does algorithm A compute larger crossing angles than algorithm B?
 
-\rcomment{Lines 132/133 ``there is no twins'' $\rightarrow$ ``there are no twins''}
+II) What is the angular resolution in drawings with small crossing angles?
 
-\response{Done.}
+III) How many iterations need the algorithms to converge?
 
-\rcomment{Line 136: ``twin pairs'' $\rightarrow$ Is this different from ``twins'' or are they equivalent? I notice several uses of both, and I’m not sure if the meanings are distinct.}
+Once the research questions are stated, one can think about proper setup and tools to investigate these questions.
+Of course with an experimental evaluation, it is unlikely that we are able to verify the research questions in their entirety.
+But we can collect indications that the statements are likely to be true (on specific data sets).
+One indicator that can be compared is the mean of two datasets, that is is the sole statistic used by the authors.
+}
 
-\response{Changed a bit to make it more clear.}
+\response{}
 
-\rcomment{Line 153: I suggest preceding $(A_2,ab)$ with ``edges'' since these also resemble open intervals.}
+\tcomment{In my opinion, the mean should be accompanied by more statistics (at least the standard deviation), since the mean is sensitive to outliers.
+Moreover, I have the impression that the writing of the experimental evaluation lacks precision.
+A more formal style of writing reduces the possibility that statements can be misunderstood.
+For example, the word /significant/ has a defined meaning, it implies that a statement is very unlikely and this has been confirmed with a proper statistical test.
+But comparing the means of two sequences of observations A and B does not imply significance.
+Instead of writing "A has significantly smaller crossing angles than B", I propose to write "The mean of A is (considerably) smaller than the mean of B", or "the mean of A is X degrees smaller than the mean of B".
+In the following, I give an incomplete list of statements that should be revised.}
 
-\response{Done, but only in the journal version due to space limitation.}
+\response{}
 
-\rcomment{Line 169 (and others): for each declaration of a forbidden pair, I was initially confused by the center dots in the notation.
-Are they meant to match a single connector vertex?
-Please mention this explicitly somewhere.}
+\tcomment{p7,c2, ~l42) 'It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution[...]'.
 
-\response{Done in the journal version.}
+The statement suggests an irrevocable truth but the only conclusion that you can draw from the plots is the following 'the mean crossing resolution of drawings obtained by your algorithm is considerably larger than the mean crossing resolutions of the remaining algorithms'.
+Alternatively, (globally) define the meaning of 'outperforms' "A outperforms B if the mean of A is X degrees smaller than the mean of B"}
 
-\rcomment{Line 179 ``there are not exactly two same twin vertices'' $\rightarrow$ I found this confusing, and the same can be said of several of the forbidden pair definitions.
-Do you mean ``there is not a single twin pair'' or even ``there is not a single twin''.
-Perhaps clarifying the usage of the word twin would be helpful.}
+\response{}
 
-\response{Done.}
+\tcomment{p7, c2, l54: While our unrestricted algorithm produces drawings with better crossing resolution no need to say /better/, it is /larger/}
 
-\rcomment{For all forbidden pairs, A and B are used; however they apply without loss of generality to any distinct twins. That should perhaps be stated somewhere.}
+\response{}
 
-\response{Come on... that should be clear.}
+\tcomment{p8, c1, l56: ... has more or less comparable performance.... what does 'more or less' mean? Does the mean crossing resolution differ by at most x degrees?}
 
-\rcomment{Line 206: ``no a A’s connector'' $\rightarrow$ ``no connector of A''.
-There are many uses like this, I suggest carefully searching the manuscript to find all occurrences.
-(See further, 207 ``a A’s connectors'', 208 ``the B’s connectors'', footnote 4: ``there are C’s connectors'', and others.)}
+\response{}
 
-\response{Done.}
+\tcomment{p8, c1/2: ...which at the same time is significantly smaller;}
 
-\rcomment{Line 247: ``Each of X,Y and U,V are different.'' The exact meaning is not clear to me. Is it that X and Y are not twins? and U,V are not twins?}
+\response{}
 
-\response{Done.}
+\tcomment{p9, c1, l55: ... significant... I do not see why you are able to claim significance.}
 
-\rcomment{Line 285: It took me a while to realize that $d(A_1,A_2)$ includes $A_1$ and $A_2$, which results in $d(A)$ counts.}
+\response{}
 
-\response{Added a note in the journal version.}
+\tcomment{p10 c1, l47: For the vast majority of the graphs in the experiments; I am not sure whether you are still comparing the mean or the number of graphs that have a smaller angle.
+In the latter case there is no need to be vague: e.g.
+'On 90\% of the graphs our algorithms compute drawings with a smaller crossing than drawings obtained by B'}
 
-\rcomment{Line 298: ``By Lemma 3 applied for A and C'' $\rightarrow$ ``Applying Lemma 3 for A and C''}
+\response{}
 
-\response{Done.}
+\tcomment{p10, c1, l57: ...tend to produce drawings of the same total resolution on larger graphs
 
-\rcomment{Line 306: ``We study the book thickness of'' $\rightarrow$ ``We study the dispersable book thickness of''}
+What does 'tend to the same' mean? Either they are equal or they are not.
 
-\response{Done.}
+How many vertices do the larger graphs have?
 
-\rcomment{Line 308/309: I highly suggesting fitting image 9 in the paper; it is very helpful to visualize the Gray graph}
+I guess what you would like to say is along the following lines:
 
-\response{Come on... that's impossible.}
+The mean values for the total resolution of drawings obtained by our algorithm and by Argyriou et al. differ by a most 5(?) degrees for graphs with more than 90 vertices.}
 
-\rcomment{Line 314: ``From this formulation'' $\rightarrow$ Based on the previous sentence, it is not clear if you are referring to the original SAT formulation, or your modified SAT formulation.}
+\response{}
 
-\response{Done.}
+\tcomment{p11, c2, ~l51: ...small graphs.... medium-size graphs
 
-\rcomment{Lines 326-331: It is worth mentioning how the computer actually assisted; what I think happened is that you ran a SAT solver, and it did not find $3$-page dispersable book embedding, and therefore there does not exist one.
-This deserves to be explicitly stated.
-Did it take long to verify?}
+As before, it is not clear when a graph is small, medium-sized or large.
+In case that you would like to use these terms frequently, I propose to introduce graphs classes: S, M, L that partition the rome graphs based on their number of vertices.}
 
-\response{Revised in the journal version.}
+\response{}
 
-\rcomment{Line 333: I believe there should be a Corollary here as a result of Theorem 2, indicating that the conjecture is false for $k=3$.}
+\tcomment{p12, c1, l48: the angular resolution dominates the crossing resolution
 
-\response{It's exactly like this in the journal version.}
+I have problems understanding this statement.
+What does dominate mean?
+Do you want to investigate the number of graphs whose crossing resolution is smaller than the total resolution?
+If so, I do not see how you can observe this based on the plots (even ignoring that my statement is on the number of graphs).
+Moreover, it seems strange to refactor this information from your plots when you could simply get this information from your data.}
 
-\rcomment{Line 334: Capitalize connected and regular}
-
-\response{Done.}
-
-\rcomment{References: At least references 9, 11 and 12 have minor errors.
-in [9], the journal is called ``SIAM Journal of Algebraic and Discrete Methods''.
-In [11] and [12] The same author is abbreviated differently in each, and in [11], ``Geomentry'' $\rightarrow$ ``Geometry''.
-Please thoroughly inspect the references.}
-
-\response{Done.}
-\end{itemize}
-
-% ==================================================================================
-\newpage
-\section*{Reviewer 3}
-% ==================================================================================
-
-\st{The paper deals with dispersable book embeddings, i.e.
-classical book embeddings (the vertices are ordered along a line l (the spine), the edges are drawn crossing-free at different half-planes (pages) bounded by l), with the additional property that the graphs induced by the edges of each page are 1-regular.
-It is clear that the number of pages needed for a dispersable book embedding of a graph G is at least the maximum degree of G.
-If this number of pages suffices, the graph G is called dispersable.
-A conjecture of Barnhart and Kainen from 1979 states that each k-regular bipartite graph is dispersable
-(for k=1,2 this is clearly true).}
-
-\st{In the paper the following three things are shown:}
-\begin{enumerate}[(a)]
-\item \st{The conjecture of Barnhart and Kainen is disproven for the case k=3 with a computer-aided proof.}
-\item \st{The conjecture of Barnhart and Kainen is disproven for the case k=4 with a hand-written proof.}
-\item \st{It is proven that each 3-connected 3-regular bipartite planar graph is dispersable.}
-\end{enumerate}
-
-\begin{enumerate}[(a)]
-\item \st{There is a known reduction of the problem of finding a book embedding with a given number of pages to SAT.
-   The authors extend this reduction to the case of dispersable book embeddings what is really simple and straight-forward.
-   Then they say that computer experiments showed that there is no 3-page dispersable book embedding of the 47-vertex
-   Gray graph. So the main contribution in this part is that the authors have found a graph for which the computer
-   program gave a negative answer, and not the computer program itself.}
-\item \st{It is proven that there is no 4-page dispersable book embedding of the 20-vertex Folkman graph.
-   The proof works as follows: By proving forbidden patterns for the order of the vertices on the spine
-   of the book embedding gradually all permutations of the vertices are excluded as orders on the spine.
-   This leads to a lot of case distinctions over more than 13 pages. I read the part of the proof in the main part,
-   but not the remaining part in the appendix. There are some inaccuracies in the proof, especially in the part where
-   it is shown that the forbidden patterns indeed forbid every permutation.
-In the end, I did not find any mistakes which cannot be fixed easily, but it is hard to check its correctness as part of a converence review because it is very
-   technical.}
-\item \st{First the authors show that 3-connected 3-regular bipartite planar graphs have a coloring of the edges
-   and faces with certain properties. Afterwards this coloring is used for a recursive geometric construction
-   of the 3-page dispersable book embedding of the graph.}
-\end{enumerate}
-
-\st{Evaluation:}
-
-\st{With results A and B the authors make a good progress in the area of book embeddings. These results are interesting
-and should be published. Unfortunately, these results are not suitable for a presentation at a conference
-since result A is proven by the computer and result B is proven by a long series of case distinctions without
-good structural insight for other graphs than the considered Folkman graph.
-After disproving the conjecture of Barnhart and Kainen, result C is an appropriate weaker variant to consider.
-The proof seems to be correct and interesting for at least parts of the WG community.
-Since the focus of the paper does not lie on result C (it is only a small part of the paper) and result C on its own would be weak for a WG
-contribution, I recommend that this paper should be published in a journal and not be presented at a
-conference.}
-
-\st{Comments for the authors:}
-
-\begin{itemize}
-\rcomment{l. 19: in early $\rightarrow$ to the early}
-
-\response{Done.}
-
-\rcomment{l. 39: $\rightarrow$ while all graphs with ... treewidth [11] have sublinear book thickness}
-
-\response{Done.}
-
-\rcomment{l. 81: avoid abbreviations at the beginning of a sentence: Figs. $\rightarrow$ Figures}
-
-\response{Done.}
-
-\rcomment{l. 88: later you call this construction a subdivision. why not here?}
-
-\response{Done in the journal version.}
-
-\rcomment{l. 118: $\rightarrow$ be the two intervals defined by the twins in a dispersable order.}
-
-\response{Done.}
-
-\rcomment{l. 136: $\rightarrow$ of pairs of twin pairs}
-
-\response{Done.}
-
-\rcomment{l. 139: pairs of non-crossing twins $\rightarrow$ non-crossing twin pairs}
-
-\response{Done.}
-
-\rcomment{l. 143, 145: twins $\rightarrow$ twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 148: a $\rightarrow$ another}
-
-\response{Done in the journal version.}
-
-\rcomment{l. 151 ff: add reference to Figure 3}
-
-\response{Done.}
-
-\rcomment{l. 161: no comma after $B_2$}
-
-\response{Done.}
-
-\rcomment{l. 161: pairs of crossing twins $\rightarrow$ crossing pairs of twins}
-
-\response{Done.}
-
-\rcomment{l. 167: twins $\rightarrow$ twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 179: including ab: add reference to Lemma 3}
-
-\response{Done.}
-
-\rcomment{l. 197: overfull box}
-
-\response{Ignored}
-
-\rcomment{l. 206: remove a}
-
-\response{Done.}
-
-\rcomment{l. 206: $\rightarrow$ Lemma 4 applied for}
-
-\response{Done.}
-
-\rcomment{l. 207: a $\rightarrow$ an}
-
-\response{Done.}
-
-\rcomment{l. 211: then $\rightarrow$ the edge}
-
-\response{Done.}
-
-\rcomment{l. 216: $[C_2,A_2]$ $\rightarrow$ $[C_2,A_1]$}
-
-\response{Done.}
-
-\rcomment{l. 221: no comma after that}
-
-\response{Done.}
-
-\rcomment{l. 224: twins $\rightarrow$ twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 225: $\leq$ $\rightarrow$ $=$}
-
-\response{Not done.}
-
-\rcomment{l. 228: twin $\rightarrow$ twin vertex}
-
-\response{Done.}
-
-\rcomment{l. 238: later $\rightarrow$ latter}
-
-\response{Done.}
-
-\rcomment{l. 240: $(C_1,ac)$ $\rightarrow$ $(A_1,ac)$}
-
-\response{Done.}
-
-\rcomment{l. 242: overfull box}
-
-\response{Ignored.}
-
-\rcomment{l. 245: no comma}
-
-\response{Done.}
-
-\rcomment{l. 245: $\rightarrow$ there are at least two twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 247: different $\rightarrow$ no pair of twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 250, 251: twins $\rightarrow$ twin vertices}
-
-\response{Done.}
-
-\rcomment{l. 258: no comma before is}
-
-\response{Done.}
-
-\rcomment{l. 261: at $\rightarrow$ in}
-
-\response{Done.}
-
-\rcomment{l. 284 ff: before you used the notation delta and counted strictly inside. why do you change the notation?}
-
-\response{The notation changed.}
-
-\rcomment{l. 287: two twins are $\rightarrow$ any pair of twins is}
-
-\response{Done.}
-
-\rcomment{l. 289: 4 $\rightarrow$ 6}
-
-\response{Done.}
-
-\rcomment{l. 291: Figs. $\rightarrow$ Figures}
-
-\response{Done.}
-
-\rcomment{l. 292: are $\rightarrow$ is}
-
-\response{Done.}
-
-\rcomment{l. 292: a $\rightarrow$ an}
-
-\response{Done.}
-
-\rcomment{l. 294: later $\rightarrow$ latter}
-
-\response{Done.}
-
-\rcomment{l. 296: the order of the vertices does not coincide with Figure 4e}
-
-\response{Done.}
-
-\rcomment{l.
-301: $y = ab.$ Here some arguments are missing.
-My suggestion: (a) analogously to $z \neq ab$ we get $x \neq ab$, (b) $u,w \neq ab$ since $u$ is D's connector and $w$ is E's connector, (c) $v \neq ab$ because the following 5 pairwise crossing edges: $A_1 ab$, $A_2 ab$, $B_1 ab$, $B_2 ab$, $C_1 ce$}
-
-\response{There is no reason to add this discussion. Symmetry implies the properties.}
-
-\rcomment{l. 321: An $\rightarrow$ A set of}
-
-\response{Done.}
-
-\rcomment{l. 323: add "we" after "details"}
-
-\response{Done.}
-
-\rcomment{l. 384: $\rightarrow$ and its right side}
-
-\response{Done.}
-
-\rcomment{l. 428: exist $\rightarrow$ exists}
-
-\response{Done.}
-
-\rcomment{l. 436: $\rightarrow$ study the dispersable}
-
-\response{Done.}
-
-\rcomment{Fig. 5: the dotted gray is hard to see}
-
-\response{Done.}
+\response{}
 
@@
 
-% ==================================================================================
-\newpage
-\section*{Reviewer 4}
-% ==================================================================================
+\textbf{References:}
 
-\st{In this paper, the authors tackle an older open question:  Given a k-regular bipartite graph, can we split its edges into k matchings and find a vertex order such that each matching forms a planar graph when placing the vertices in this order around a cycle?}
+[0] Devanny W., Kindermann P., Löffler M., Rutter I.
+(2018) Graph Drawing Contest Report.
+In: Frati F., Ma KL.
+(eds) Graph Drawing and Network Visualization.
+GD 2017.
+Lecture Notes in Computer Science, vol 10692.
+Springer, Cham
 
-\st{It was known that this is true for k=1,2 and conjectured that it is also true for larger k.
-But, the authors disprove it for k=3 and k=4.
-The proof for k=4 is combinatorial, while the proof for k=3 was done by phrasing the problem as a SAT problem and running experiments.
-The authors also show that the conjecture is true for k=3 if additionally the graph is 3-connected and planar.}
+[1] Demel A., Dürrschnabel D., Mchedlidze T., Radermacher M., Wulf L.
+(2018) A Greedy Heuristic for Crossing-Angle Maximization.
+In: Biedl T., Kerren A.
+(eds) Graph Drawing and Network Visualization.
+GD 2018.
+Lecture Notes in Computer Science, vol 11282.
+Springer, Cham
 
-\st{On the whole, this is a nice paper.
-However, I was very lost in the combinatorial proof.
-The authors state a crucial Lemma 2, and then use it to classify vertices as either far or near.
-But I don't see how that far/near follows from the lemma.
-In consequence, I could not follow the rest of the proof.
-As such, I'm a bit unsure whether this main result is true.
-(My guess is that it is true---the authors probably checked that with their SAT solver?---but this proof does not convince me.) That's why I ranked the paper somewhat lower.}
+[2] Johan Ugander and Lars Backstrom.
+2013.
+Balanced label propagation for partitioning massive graphs.
+In Proceedings of the sixth ACM international conference on Web search and data mining (WSDM '13).
+ACM, New York, NY, USA, 507-516.
+DOI: https://doi.org/10.1145/2433396.2433461
 
-\st{Detailed comments to the authors:}
+[3] C. M. Fiduccia and R. M. Mattheyses, "A Linear-Time Heuristic for Improving Network Partitions," 19th Design Automation Conference, Las Vegas, NV, USA, 1982, pp. 175-181.
+doi: 10.1109/DAC.1982.1585498
 
-\begin{itemize}
-\rcomment{Lemma 2 classifies a pair of twins A1,A2, and the connectors that relates to them (ab,ac,ad,ae).
-But then in the next paragraph, you claim that the first configuration implies that the twins A1,A2 have to lie next to each other, with no other twin between them.
-I don't see why this is true.
-For example, what's wrong with the order A1,ac,B1,bc,A2,ab,...,ad....ae.
-This satisfies all the conclusions for Lemma 2 and has A in the 1-3-configuration, but A1 and A2 are not close.
-Neither are they far (because the way you define far means that they have to be in the 2-2-configuration).
-So this is an entirely different configuration that you are omitting in all the later cases.
-You need to either convince me that this order is impossible, or you need to expand all the later analysis to include this possibility.}
+[4] Kernighan, B. W. and Lin, S. (1970), An Efficient Heuristic Procedure for Partitioning Graphs. Bell System Technical Journal, 49: 291-307. doi:10.1002/j.1538-7305.1970.tb01770.x
 
-\response{We agreed that this comment is not correct.}
+[5]  A Geometric Heuristic for Rectilinear Crossing Minimization
+Marcel Radermacher, Klara Reichard, Ignaz Rutter, and Dorothea Wagner
+2018 Proceedings of the Twentieth Workshop on Algorithm Engineering and Experiments (ALENEX). 2018, 129-138
 
-\rcomment{Lemma 3 also has some issues.
-Statement iii just says that the four twins are separated by A's connectors.
-But then the explanatory side-phrase says something stronger, namely, that the connector betwee A1 and A2 *must* be ab.
-This should be said clearly in the main statement of iii.
-(Same thing for iv.)}
-
-\response{We agreed that this comment is not correct.}
-\end{itemize}
-
-Further minor comments:
-
-\begin{itemize}
-\rcomment{p2, l40: "while sublinear book thickness have all graphs". This is not proper English, reword.  Maybe "while sublinear book thickness can be shown for all graphs..."}
-
-\response{Revised based on some earlier comment of another reviewer.}
-
-\rcomment{p2 (and throughout the paper):  You seem to use the term "1-regular" in a very strange sense.
-Normally d-regular means that every vertex has degree *exactly* d, neither more nor less.
-But you seem to equate "1-regular" with "is a matching", i.e., every vertex has degree *at*most* 1.
-At the very minimum you should make this very clear.
-Better even, avoid the term 1-regular and replace it with "maximum degree 1" everywhere.}
-
-\response{MB: I believe that this is a wrong comment, but Martin suggested an easy hack which we applied.}
-
-\rcomment{p2, l51: define "circular embedding"}
-
-\response{Come on... We have references and also a figure thas illustrate this obvious thing.}
-
-\rcomment{p4, l125: "with three or four A's connectors" insert "of" before "A's" (and same thing on the next line)}
-
-\response{Done.}
-\end{itemize}
-
-% ==================================================================================
-\newpage
-\section*{Reviewer 5}
-% ==================================================================================
-
-\st{This is a deep paper with solutions to two long standing open questions involving Dispersable Book Embeddings (DBE).
-The minimum number of pages needed for a DBE is called the dispersible book thickness (dbt) and it is clear that $dbt(G) >= max-degree-of-G$ and it was conjectured by Bernhart and Kainen that equality holds for every k-regular bipartite graph.
-They showed that the conjecture holds for k =1, 2.
-In this paper, counter-examples are presented for $k = 3, 4$.}
-
-\st{The paper is well written with only a few typos}
-
-\begin{itemize}
-\rcomment{one common (see for example pg 6, line 172) is the absence of "of" between "two" and "A's connectors".}
-
-\response{Done.}
-
-\rcomment{Also on page 10, line 348, I suggest including a statement of the 3-color theorem.}
-
-\response{Added a note in the journal version.}
-\end{itemize}
-
-% ==================================================================================
-\newpage
-\section*{Reviewer 6}
-% ==================================================================================
-
-\st{This is a nice one.
-The authors disprove a conjecture from the late 70's by means of a computer proof (and a somewhat weaker proof by hand).
-They also show that the desired property (being dispersable) is held by Barnette graphs.}
-
-\st{The writing is very nice and the math seems sound.
-The only drawback of the paper might be that it fits into the scope of WG only to some extend, as no deeper connection of the result to concepts in computer science is given.
-This seems like a minor issue to me, though.}
+[6] Bläsius T., Radermacher M., Rutter I.
+(2017) How to Draw a Planarization.
+In: Steffen B., Baier C., van den Brand M., Eder J., Hinchey M., Margaria T.
+(eds) SOFSEM 2017: Theory and Practice of Computer Science.
+SOFSEM 2017.
+Lecture Notes in Computer Science, vol 10139.
+Springer, Cham
 

### 2019-07-26 paper/comments/reply-letter.tex
Most of the comments have been addressed.

@@
 
-%\author{Patrizio Angelini, Michael A. Bekos, Giuseppe Liotta, Fabrizio Montecchiani}
-
-\author{M.~Bekos, H.~F\"orster, C.~Geckeler, L.~Holl\"ander, M.~Kaufmann, A.~Spallek, J.~Splett}
+\author{M.~A.~Bekos, H.~F\"orster, C.~Geckeler, L.~Holl\"ander,\\M.~Kaufmann, A.~Spallek, J.~Splett}
 
@@
 Given the limited applied previous work on the topic (most of the previous work focuses on the case of RAC graphs), i believe that their contribution is significant and it is definitely worth publishing.
-I suggest ACCEPTANCE subject to addressing the following
-
-MINOR issues:
+I suggest ACCEPTANCE subject to addressing the following MINOR issues:
 
@@
 
-\rcomment{pg 4, ln 60, left: The text followin "In other words.." is not implied by the formula. The formula should be: $c(\Gamma_{k}) \geq c(\Gamma_{k-1}), ~~0 \leq k \leq i-1$ or similar.}
+\rcomment{pg 4, ln 60, left: The text following "In other words.." is not implied by the formula. The formula should be: $c(\Gamma_{k}) \geq c(\Gamma_{k-1}), ~~0 \leq k \leq i-1$ or similar.}
 
-\response{Done.}
+\response{We agree with the reviewer. We have revised this part in the paper to make it clear.}
 
@@
 Indeed, the vertices in the neighborhood are not related to the crossing resolution of the current drawing.
-However, by also considering these vertices (and in particular by moving them), we have observed that in several cases it is possible to gain the necessary space for the actual critical vertices, i.e., those that define the drawing's crossing resolution, to move and thus to improve the drawing's crossing resolution.
+However, by also considering these vertices (and in particular by moving them), we have observed that in several cases it is possible to gain the necessary space for the actual critical vertices (i.e., those that define the drawing's crossing resolution) to move and thus to improve the drawing's crossing resolution.
 This is the idea underneath.}
@@
 
-\response{We have resolved this problem.}
+\response{In the revised version, we have changed the caption of the subfigures.}
 
@@
 
-\response{In the revised version, we have revised the caption of Figure~9 to make this clear.}
+\response{In the revised version, we have changed the caption of Figure~9 to make this clear.}
 
@@
 \begin{itemize}
-\tcomment{First, the authors claim (Page 2, L20) to be aware of only two approaches (Argyriou et al.
+\rcomment{First, the authors claim (Page 2, L20) to be aware of only two approaches (Argyriou et al.
 and Huang et al.) that optimize the crossing angle.
@@
 In my opinion, it is only fair to the authors of CoffeeVM to discuss similarities and differences between the two approaches.
-Moreover, the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems.
-I propose to add [1-6] as relevant related work.}
-
-\response{}
-
-\tcomment{Second, I think that the structure of the evaluation can be improved by formulating a set of research questions. For example:
-
-I) Does algorithm A compute larger crossing angles than algorithm B?
-
-II) What is the angular resolution in drawings with small crossing angles?
-
-III) How many iterations need the algorithms to converge?
-
-Once the research questions are stated, one can think about proper setup and tools to investigate these questions.
-Of course with an experimental evaluation, it is unlikely that we are able to verify the research questions in their entirety.
-But we can collect indications that the statements are likely to be true (on specific data sets).
-One indicator that can be compared is the mean of two datasets, that is is the sole statistic used by the authors.
-}
-
-\response{}
-
-\tcomment{In my opinion, the mean should be accompanied by more statistics (at least the standard deviation), since the mean is sensitive to outliers.
-Moreover, I have the impression that the writing of the experimental evaluation lacks precision.
-A more formal style of writing reduces the possibility that statements can be misunderstood.
-For example, the word /significant/ has a defined meaning, it implies that a statement is very unlikely and this has been confirmed with a proper statistical test.
-But comparing the means of two sequences of observations A and B does not imply significance.
-Instead of writing "A has significantly smaller crossing angles than B", I propose to write "The mean of A is (considerably) smaller than the mean of B", or "the mean of A is X degrees smaller than the mean of B".
-In the following, I give an incomplete list of statements that should be revised.}
-
-\response{}
-
-\tcomment{p7,c2, ~l42) 'It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution[...]'.
-
-The statement suggests an irrevocable truth but the only conclusion that you can draw from the plots is the following 'the mean crossing resolution of drawings obtained by your algorithm is considerably larger than the mean crossing resolutions of the remaining algorithms'.
-Alternatively, (globally) define the meaning of 'outperforms' "A outperforms B if the mean of A is X degrees smaller than the mean of B"}
-
-\response{}
-
-\tcomment{p7, c2, l54: While our unrestricted algorithm produces drawings with better crossing resolution no need to say /better/, it is /larger/}
-
-\response{}
-
-\tcomment{p8, c1, l56: ... has more or less comparable performance.... what does 'more or less' mean? Does the mean crossing resolution differ by at most x degrees?}
-
-\response{}
-
-\tcomment{p8, c1/2: ...which at the same time is significantly smaller;}
-
-\response{}
-
-\tcomment{p9, c1, l55: ... significant... I do not see why you are able to claim significance.}
-
-\response{}
-
-\tcomment{p10 c1, l47: For the vast majority of the graphs in the experiments; I am not sure whether you are still comparing the mean or the number of graphs that have a smaller angle.
-In the latter case there is no need to be vague: e.g.
-'On 90\% of the graphs our algorithms compute drawings with a smaller crossing than drawings obtained by B'}
-
-\response{}
-
-\tcomment{p10, c1, l57: ...tend to produce drawings of the same total resolution on larger graphs
-
-What does 'tend to the same' mean? Either they are equal or they are not.
-
-How many vertices do the larger graphs have?
-
-I guess what you would like to say is along the following lines:
-
-The mean values for the total resolution of drawings obtained by our algorithm and by Argyriou et al. differ by a most 5(?) degrees for graphs with more than 90 vertices.}
-
-\response{}
-
-\tcomment{p11, c2, ~l51: ...small graphs.... medium-size graphs
-
-As before, it is not clear when a graph is small, medium-sized or large.
-In case that you would like to use these terms frequently, I propose to introduce graphs classes: S, M, L that partition the rome graphs based on their number of vertices.}
-
-\response{}
-
-\tcomment{p12, c1, l48: the angular resolution dominates the crossing resolution
-
-I have problems understanding this statement.
-What does dominate mean?
-Do you want to investigate the number of graphs whose crossing resolution is smaller than the total resolution?
-If so, I do not see how you can observe this based on the plots (even ignoring that my statement is on the number of graphs).
-Moreover, it seems strange to refactor this information from your plots when you could simply get this information from your data.}
-
-\response{}
-
-\end{itemize}
-
-\textbf{References:}
-
-[0] Devanny W., Kindermann P., Löffler M., Rutter I.
-(2018) Graph Drawing Contest Report.
-In: Frati F., Ma KL.
-(eds) Graph Drawing and Network Visualization.
-GD 2017.
-Lecture Notes in Computer Science, vol 10692.
-Springer, Cham
 
@@
 Springer, Cham
+}
+
+\response{We added a paragraph discussing the general ideas of the approach in [1]}
+
+\rcomment{Moreover, the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems. I propose to add [1-6] as relevant related work.
 
@@
 In Proceedings of the sixth ACM international conference on Web search and data mining (WSDM '13).
-ACM, New York, NY, USA, 507-516.
-DOI: https://doi.org/10.1145/2433396.2433461
+ACM, New York, NY, USA, 507-516.\\DOI: https://doi.org/10.1145/2433396.2433461
 
@@
 Lecture Notes in Computer Science, vol 10139.
-Springer, Cham
+Springer, Cham}
+
+\response{References [1-6] have been added as related work.}
+
+\rcomment{Second, I think that the structure of the evaluation can be improved by formulating a set of research questions. For example:
+
+I) Does algorithm A compute larger crossing angles than algorithm B?
+
+II) What is the angular resolution in drawings with small crossing angles?
+
+III) How many iterations need the algorithms to converge?
+
+Once the research questions are stated, one can think about proper setup and tools to investigate these questions.
+Of course with an experimental evaluation, it is unlikely that we are able to verify the research questions in their entirety.
+But we can collect indications that the statements are likely to be true (on specific data sets).
+One indicator that can be compared is the mean of two datasets, that is is the sole statistic used by the authors.}
+
+\response{We preferred not to pose such exact questions, because we simply cannot provide answers to any of them.
+The reason is that an experimental evaluation can only give some indications.
+Instead, we now explicitly mention which are the parameters that are interested~in.}
+
+\rcomment{In my opinion, the mean should be accompanied by more statistics (at least the standard deviation), since the mean is sensitive to outliers.}
+
+\response{As mentioned in the previous reply, our experimental evaluation aims in giving some indications.
+The mean is a good metric in this direction, even though it is sensitive to outliers.
+\textcolor{red}{For completeness, we have added a diagram illustrating the standard deviation as suggested by the reviewer.}}
+
+\rcomment{Moreover, I have the impression that the writing of the experimental evaluation lacks precision.
+A more formal style of writing reduces the possibility that statements can be misunderstood.
+For example, the word /significant/ has a defined meaning, it implies that a statement is very unlikely and this has been confirmed with a proper statistical test.
+But comparing the means of two sequences of observations A and B does not imply significance.
+Instead of writing "A has significantly smaller crossing angles than B", I propose to write "The mean of A is (considerably) smaller than the mean of B", or "the mean of A is X degrees smaller than the mean of B".
+In the following, I give an incomplete list of statements that should be revised.}
+
+\response{We agree with the reviewer that some of our statements might be questionable.
+So, in the following, we have addressed each of the reviewer's specific comments in order to improve the presentation.}
+
+\rcomment{p7,c2, ~l42) 'It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution[...]'.
+The statement suggests an irrevocable truth but the only conclusion that you can draw from the plots is the following 'the mean crossing resolution of drawings obtained by your algorithm is considerably larger than the mean crossing resolutions of the remaining algorithms'.
+Alternatively, (globally) define the meaning of 'outperforms' "A outperforms B if the mean of A is X degrees smaller than the mean of B"}
+
+\response{We revised according to the first suggestion of the reviewer.}
+
+\rcomment{p7, c2, l54: While our unrestricted algorithm produces drawings with better crossing resolution no need to say /better/, it is /larger/}
+
+\response{Done.}
+
+\rcomment{p8, c1, l56: ... has more or less comparable performance.... what does 'more or less' mean? Does the mean crossing resolution differ by at most x degrees?}
+
+\response{The phrase 'more or less' was describing the fact that the corresponding two curve in Fig.3b coincide. We have revised the statement to make it more clear.}
+
+\rcomment{p8, c1/2: ...which at the same time is significantly smaller;}
+
+\response{Done.}
+
+\rcomment{p9, c1, l55: ... significant... I do not see why you are able to claim significance.}
+
+\response{Indeed, the claim was overestimating the situation. We have revised accordingly.}
+
+\rcomment{p10 c1, l47: For the vast majority of the graphs in the experiments; I am not sure whether you are still comparing the mean or the number of graphs that have a smaller angle.
+In the latter case there is no need to be vague: e.g.
+'On 90\% of the graphs our algorithms compute drawings with a smaller crossing than drawings obtained by B'}
+
+\response{We are always comparing mean numbers.
+This is explicitly stated at the beginning of Section 3.
+Here, we intended to distinguish between some graphs with small number of vertices ($<20$) and the remaining ones of the experiment.
+In the revised version, we are more precise in our description by explicitly mentioning this.
+We have revised accordingly.}
+
+\rcomment{p10, c1, l57: ...tend to produce drawings of the same total resolution on larger graphs.
+%
+What does 'tend to the same' mean? Either they are equal or they are not. 
+%
+How many vertices do the larger graphs have?    
+%
+I guess what you would like to say is along the following lines:
+%
+The mean values for the total resolution of drawings obtained by our algorithm and by Argyriou et al. differ by a most 5(?) degrees for graphs with more than 90 vertices.}
+
+\response{No, this was not what we intended to describe at this point.
+We really wanted to describe the tendency of the lines in the plot.
+In particular, the fact that they seem to converge to the same values after a point on.
+Nevertheless, to avoid confusion we removed this note in the revised version.}
+
+\rcomment{p11, c2, ~l51: ...small graphs.... medium-size graphs:
+%
+As before, it is not clear when a graph is small, medium-sized or large.
+In case that you would like to use these terms frequently, I propose to introduce graphs classes: S, M, L that partition the rome graphs based on their number of vertices.}
+
+\response{In the revised version, we use neither the term ``small graphs'' nor the term ``medium-size graphs''.}
+
+\rcomment{p12, c1, l48: the angular resolution dominates the crossing resolution:
+%
+I have problems understanding this statement.
+What does dominate mean?
+Do you want to investigate the number of graphs whose crossing resolution is smaller than the total resolution?
+If so, I do not see how you can observe this based on the plots (even ignoring that my statement is on the number of graphs).
+Moreover, it seems strange to refactor this information from your plots when you could simply get this information from your data.}
+
+\response{We agree with the reviewer.
+This observation was questionable, as it was not supported by numbers in our experiment.
+It was simply our intuition but we somehow failed to transfer this information.
+In the revised version, the observation has been completely removed to avoid confusion.}
+
+\end{itemize}
 

### 2019-07-26 paper/journal.tex
Most of the comments have been addressed.

@@
 If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position.
+Note at that in general the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.
 
-Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs of general graphs.
+On the other hand, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs of general graphs.
 Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we followed a more practical approach.
@@
 \begin{table*}
-\caption{Summary of the results for the Graph Drawing Contest 2017; the last two columns correspond to results presented in this paper (with and without time limitations, respectively).}
+\caption{Summary of the results for the Graph Drawing Contest 2017; the last two columns correspond to results obtained by the implementation of the algorithm presented in this paper (with and without time limitations, respectively).}
 \centering
@@
 %\myparagraph{Graph Drawing Contest 2017} 
-In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}).
+In 2017, we participated to the contest~\cite{DBLP:conf/gd/DevannyKLR17} (by implementing a mixture of the two force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}).
 %
@@
 
-It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''; details for their approach can be found in~\cite{DBLP:journals/corr/abs-1807-09483}.
+It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''.
+Notably,  their approach is in principle very similar to the one that we present in this paper.
+In particular, they also adopt the vertex movement paradigm on a set of a randomly chosen samples.
+Once they find a sample position, which improves the crossing resolution of the drawing, then they focus on a region close to this position.
+In this region, they are examining more sample positions, hoping that they will find one that will further improve the crossing resolution of the drawing; for details refer to~\cite{DBLP:conf/gd/DemelDMRW18}.
+
 Our team ``\texttt{TuebingenMidnight}''  performed miserably; we actually took the third position out of a total of four teams that participated.
 As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$.
-In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as s results, which were achieved without any time limitations (see column ``No-time restriction'').
+In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as  results, which were achieved without any time limitations (see column ``No-time restriction'').
 
@@
 %\myparagraph{Graph Drawing Contest 2018} 
-In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper).
+In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}''~\cite{DBLP:conf/gd/DevannyKLR18} (using an implementation of the algorithm presented in this paper).
 The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''.
@@
 In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we discussed in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution).
-The reported results, illustrated in Figures~\ref{fig:cr-res}--\ref{fig:northAngular}, are on average across different drawings with same number of vertices.
+The reported results, illustrated in Figures~\ref{fig:cr-res}--\ref{fig:northAngular}, are on average across different drawings with same number of vertices on the following set of parameters:
+%
+\begin{enumerate}[P.1.]
+\item \label{p:cr-res} crossing resolution
+\item \label{p:to-res} total resolution
+\item \label{p:an-res} angular resolution
+\item \label{p:as-rat} edge-length ratio
+\item \label{p:no-xing} number of crossings
+\end{enumerate}
 
@@
 Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution).
-It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the edge-length ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
+It is immediate to see that on average the crossing resolution of the drawings obtained by your algorithm, which does not impose any restriction on the edge-length ratio of the computed drawings, is considerably larger than the crossing resolutions of the remaining algorithms; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}.
 The variant of our algorithm, which does not violate the edge-length ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ELR-restricted}, in Fig.~\ref{fig:cr-res-1}.
@@
 
-While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased edge-length ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding edge-length ratio of the drawings produced by the algorithm of Argyriou et al.
+While our unrestricted algorithm produces drawings with larger crossing resolution on average, this comes at a cost of drastically increased edge-length ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding edge-length ratio of the drawings produced by the algorithm of Argyriou et al.
 For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short.
-Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths).
-On the other hand, the restricted variant of our algorithm, which does not allow the edge-length ratio to increase, has more or less comparable performance (in terms of edge-length ratio) as the one of Huang et al.
-
-Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.
+Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose spring forces impose strong restrictions on the edge lengths).
+On the other hand, the edge-length ratio of the drawings obtained by the restricted variant of our algorithm, which does not allow the edge-length ratio to increase, is very close to the edge-length ratio of the drawings obtained by the algorithm of Huang et al.
+Note that the latter algorithm is the one yielding drawings that on average have the lowest edge-length ratio.
+Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings.
+On the other hand, the two other algorithms of our experiment produce drawings that on average have more crossings; see Fig.~\ref{fig:cr-res-3}.
 
@@
 In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees.
-Observe that even under this different convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
+Observe that even under this different convergence criterion, the algorithm needs more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.
 
@@
 Here, each algorithm was adjusted to maximize both the crossing and the angular resolution, simultaneously.
-For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield drawings with better total resolution than the corresponding ones by Argyriou et al.
-The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}.
-Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
+For graphs with more 20 vertices, both our unrestricted algorithm and its restricted variant yield drawings with better total resolution than the corresponding ones by Argyriou et al.
+The drawings produced by the algorithm by Huang et al.\ seem to have worse total resolution; see Fig.~\ref{fig:to-res-1}.%Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.
 
@@
 Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
+A notable observation is that, for graphs with less than $30$ vertices the best results are achieved by the algorithm by Argyriou et al., while for the remaining graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
 For large graphs, the two algorithms tend to have the same performance.
@@
 The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
-
-The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.
-This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
+%
+The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.% This observation suggests that, for most of the graphs of our experiment, the angular resolution is greater than the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results.
 The small differences result from the fact that the crossing resolution cannot be entirely neglected.

### 2019-07-26 paper/data/plotsWithTikz.tex
first impression of the new plots

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}
\usepackage{amssymb}
\usepackage{subfig}

%\pgfplotscreateplotcyclelist{mylist}{%
%{blue!80!black,mark options={mark size=1.5pt,fill=white},mark=*},
%{red!80!black,densely dashdotted, mark options={mark size=1.5pt,solid,fill=white},mark=square*},
%{green!50!black,mark options={mark size=1.5pt,solid,fill=white},densely dashed,mark=*},
%{yellow!40!black,densely dashdotdotted,mark options={solid,fill=white},mark=triangle*},
%{densely dotted,mark options={solid,fill=white},mark=diamond*}
%}
\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]

%\definecolor{dark-blue}{rgb}{0,0,0.545}
%\definecolor{dark-red}{rgb}{0.545,0,0}
%\definecolor{dark-green}{rgb}{0,0.392,0}
%\definecolor{dark-magenta}{rgb}{0.545,0,0.545}
%\pgfplotsset{ytick style={draw=o}}
%\renewcommand{\axisdefaultwidth}{300pt}
\renewcommand{\axisdefaultheight}{185pt}

\pgfplotscreateplotcyclelist{mylist}{%
{black,thick,smooth,o},
{black!50!white,thick,smooth,o},
{black,thick,densely dashed,smooth,o},
{black!50!white,thick,densely dashed,smooth,o},
{black,double,smooth,o}
}

\pgfplotscreateplotcyclelist{fourlist}{%
{black,thick,smooth,mark=x},
{black!50!white,thick,smooth,mark=x},
{black,thick,densely dashed,smooth,mark=x},
{black!50!white,thick,densely dashed,smooth,mark=x}
}

\pgfplotsset{
  /pgfplots/xlabel near ticks/.style={
     /pgfplots/every axis x label/.style={
        at={(ticklabel cs:0.5)},anchor=near ticklabel
     }
  },
  /pgfplots/ylabel near ticks/.style={
     /pgfplots/every axis y label/.style={
        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
     }
  }

\begin{document}
\pgfkeys{/pgf/number format/.cd,1000 sep={}}
\begin{figure}
\centering
\subfloat[\label{fig:crossingResolution}{}]{
\centering
\scalebox{.95}{
\centering
%======================
% Crossing Resolution
%======================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,10,20,30,40,50,60,70,80,90},
			 ymajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Crossing Resolution},
			 xlabel={Number of Vertices},
			 ylabel style={yshift=-0.2cm}]
\addplot+[boxplot] table [y=cr-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\subfloat[\label{fig:crossingResolution-aspectRatio}{}]{
\centering
\scalebox{.95}{
\centering
%===================================
% Crossing Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={1,2.5,10,25,100,250,1000,2500,10000},
			 ymode=log, log ticks with fixed point,
			 ymajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel style={yshift=-0.2cm},
			 ylabel={Aspect Ratio},
			 xlabel={Number of Vertices}]
\addplot+[boxplot] table [y=asp-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\end{figure}

\begin{figure}
\centering

\subfloat[\label{fig:crossingResolution-crossings}{}]{
\centering
\scalebox{.95}{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,25,50,75,100,125,150,175,200,225,250,275,300},
			 ymajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Number of Crossings},
			 xlabel={Number of Vertices},
			 ylabel style={yshift=-0.2cm},
			]
\addplot+[boxplot] table [y=cn-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\subfloat[\label{fig:crossingResolution-iterations}{}]{
\centering
\scalebox{.95}{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},
			 ymajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	    xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Iterations},
     	     ylabel style={yshift=-0.2cm},
			 xlabel={Number of Vertices}]
\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[mark = x, boxplot] table [y=iterations-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=iterations-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\end{figure}

\end{document}

### 2019-07-29 paper/data/plotsWithTikz.tex
first try on standard deviation

@@
 \usepgfplotslibrary{statistics}
+\usepgfplotslibrary{fillbetween}
+\usetikzlibrary{patterns}
 \usepackage{amssymb}
@@
 \pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth,o},
-{black!50!white,thick,smooth,o},
-{black,thick,densely dashed,smooth,o},
-{black!50!white,thick,densely dashed,smooth,o},
-{black,double,smooth,o}
+{black,thick,smooth},
+{black!50!white,thick,smooth},
+{black,thick,densely dashed,smooth},
+{black!50!white,thick,densely dashed,smooth},
+{black,double,smooth}
+}
+
+\pgfplotscreateplotcyclelist{mylist-filled}{%
+{black,thick,smooth},
+{black,smooth},
+{black,smooth},
+{black!50!white,thick,smooth},
+{black!50!white,smooth},
+{black!50!white,smooth},
+{black,thick,densely dashed,smooth},
+{black,densely dashed,smooth},
+{black,densely dashed,smooth},
+{black!50!white,thick,densely dashed,smooth},
+{black!50!white,densely dashed,smooth},
+{black!50!white,densely dashed,smooth},
+{black,double,smooth},
+{black,double,thin,smooth}
+{black,double,thin,smooth}
 }
@@
 			 ymajorgrids,
+			 xmajorgrids,
 			 xtick={0,1,2,...,9},
@@
 			 cycle list name=fourlist,
-			 ytick={1,2.5,10,25,100,250,1000,2500,10000},
+			 ytick={1,2.5,10,25,100,250,1000,2500,10000,25000,100000,250000},
 			 ymode=log, log ticks with fixed point,
 			 ymajorgrids,
+			 xmajorgrids,
 			 xtick={0,1,2,...,9},
@@
 			 boxplot/draw direction=y,
-			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
+     		 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + (1/5)*mod(\plotnumofactualtype,4)},
+%			 boxplot/draw position={1/5*(\plotnumofactualtype+1) + 1/5*floor(\plotnumofactualtype/4)},
 			 boxplot/box extend=0.18,
 			 cycle list name=fourlist,
-			 ytick={0,25,50,75,100,125,150,175,200,225,250,275,300},
+			 ytick={0,50,100,150,200,250,300,350,400,450,500,550,600,650,700,750,800,850,900,950,1000},
 			 ymajorgrids,
-			 xtick={0,1,2,...,9},
+			 xmajorgrids,
+			 xtick={0,1,2,3,4,5,6,7,8,9},
 			 x tick label as interval,
@@
 			 cycle list name=fourlist,
-			 ytick={0,1000,2000,3000,4000,5000,6000,7000,8000},
+			 ytick={0,5000,10000,15000,20000,25000,30000,35000,40000,45000,50000,55000,  60000,65000},
 			 ymajorgrids,
+			 xmajorgrids,
 			 xtick={0,1,2,...,9},
@@
 
+\begin{figure*}[t]
+\centering
+\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
+\centering
+%======================
+% Crossing Resolution
+%======================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
+mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
+
+\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
+
+\addplot[name path=rm_u,black,smooth] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,black,smooth] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[black!50,fill opacity=0.5] fill between[of=rm_u and rm_d];
+
+\addplot[name path=rm-1_u,black!50!white,smooth] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,black!50!white,smooth] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[black!25,fill opacity=0.5] fill between[of=rm-1_u and rm-1_d];
+
+\addplot[name path=fa_u,black,smooth,densely dashed] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,black,smooth,densely dashed] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[black!50,fill opacity=0.5] fill between[of=fa_u and fa_d];
+
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}
+\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
+\centering
+%===================================
+% Crossing Resolution Aspect Ratio
+%===================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}
+
+\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
+\centering
+%======================================
+% Crossing Resolution Crossing Number
+%======================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}
+\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
+\centering
+%==================================
+% Crossing Resolution Iterations
+%==================================
+\begin{tikzpicture}
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
+\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
+\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
+\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
+\end{axis}
+\end{tikzpicture}}
+\caption{Experimental results on the crossing resolution for the Rome graphs.}
+\label{fig:cr-res}
+\end{figure*}
+
 \end{document}

### 2019-07-29 paper/data/boxPlotsWithTikz.tex
finished first set of figures

--- 2019-07-29 paper/data/plotsWithTikz.tex
@@
 \pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
+{blue,thick,smooth},
+{red,thick,smooth},
+{green!70!black,thick,smooth},
+{blue!50!red,thick,smooth},
 {black,double,smooth}
@@
 
-\addplot[name path=rm_u,black,smooth] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
-\addplot[name path=rm_d,black,smooth] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
-\addplot[black!50,fill opacity=0.5] fill between[of=rm_u and rm_d];
+\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];
 
-\addplot[name path=rm-1_u,black!50!white,smooth] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot[name path=rm-1_d,black!50!white,smooth] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot[black!25,fill opacity=0.5] fill between[of=rm-1_u and rm-1_d];
+\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];
 
-\addplot[name path=fa_u,black,smooth,densely dashed] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
-\addplot[name path=fa_d,black,smooth,densely dashed] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
-\addplot[black!50,fill opacity=0.5] fill between[of=fa_u and fa_d];
+\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];
+
+\addplot[blue,smooth,thick] table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[red,smooth,thick] table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[green!70!black,smooth,thick] table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!50!red,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
 

### 2019-07-29 paper/data/plotsWithTikz.tex
finished first set of figures

@@
 \pgfplotscreateplotcyclelist{mylist}{%
-{black,thick,smooth},
-{black!50!white,thick,smooth},
-{black,thick,densely dashed,smooth},
-{black!50!white,thick,densely dashed,smooth},
+{blue,thick,smooth},
+{red,thick,smooth},
+{green!70!black,thick,smooth},
+{red!40!blue,thick,smooth},
 {black,double,smooth}
@@
 \pgfkeys{/pgf/number format/.cd,1000 sep={}}
-\begin{figure}
-\centering
-\subfloat[\label{fig:crossingResolution}{}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================
-% Crossing Resolution
-%======================
-\begin{tikzpicture}
-\begin{axis}[
-			 width=\textwidth,
-			 boxplot/draw direction=y,
-			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
-			 boxplot/box extend=0.18,
-			 cycle list name=fourlist,
-			 ytick={0,10,20,30,40,50,60,70,80,90},
-			 ymajorgrids,
-			 xmajorgrids,
-			 xtick={0,1,2,...,9},
-			 x tick label as interval,
-     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
-     	     61--70,71--80,81--90,91--100},
-     	     ylabel={Crossing Resolution},
-			 xlabel={Number of Vertices},
-			 ylabel style={yshift=-0.2cm}]
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr11-20.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr21-30.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr21-30.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr31-40.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr31-40.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr41-50.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr41-50.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr51-60.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr51-60.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr61-70.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr61-70.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr71-80.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr71-80.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr81-90.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr81-90.csv};
-
-\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr91-100.csv};
-\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr91-100.csv};
-
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolution-aspectRatio}{}]{
-\centering
-\scalebox{.95}{
-\centering
-%===================================
-% Crossing Resolution Aspect Ratio
-%===================================
-\begin{tikzpicture}
-\begin{axis}[
-			 width=\textwidth,
-			 boxplot/draw direction=y,
-			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
-			 boxplot/box extend=0.18,
-			 cycle list name=fourlist,
-			 ytick={1,2.5,10,25,100,250,1000,2500,10000,25000,100000,250000},
-			 ymode=log, log ticks with fixed point,
-			 ymajorgrids,
-			 xmajorgrids,
-			 xtick={0,1,2,...,9},
-			 x tick label as interval,
-     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
-     	     61--70,71--80,81--90,91--100},
-     	     ylabel style={yshift=-0.2cm},
-			 ylabel={Aspect Ratio},
-			 xlabel={Number of Vertices}]
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr11-20.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr21-30.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr21-30.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr31-40.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr31-40.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr41-50.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr41-50.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr51-60.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr51-60.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr61-70.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr61-70.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr71-80.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr71-80.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr81-90.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr81-90.csv};
-
-\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr91-100.csv};
-\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr91-100.csv};
-
-\end{axis}
-\end{tikzpicture}}}
-
-\end{figure}
-
-\begin{figure}
-\centering
-
-\subfloat[\label{fig:crossingResolution-crossings}{}]{
-\centering
-\scalebox{.95}{
-\centering
-%======================================
-% Crossing Resolution Crossing Number
-%======================================
-\begin{tikzpicture}
-\begin{axis}[
-			 width=\textwidth,
-			 boxplot/draw direction=y,
-     		 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + (1/5)*mod(\plotnumofactualtype,4)},
-%			 boxplot/draw position={1/5*(\plotnumofactualtype+1) + 1/5*floor(\plotnumofactualtype/4)},
-			 boxplot/box extend=0.18,
-			 cycle list name=fourlist,
-			 ytick={0,50,100,150,200,250,300,350,400,450,500,550,600,650,700,750,800,850,900,950,1000},
-			 ymajorgrids,
-			 xmajorgrids,
-			 xtick={0,1,2,3,4,5,6,7,8,9},
-			 x tick label as interval,
-     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
-     	     61--70,71--80,81--90,91--100},
-     	     ylabel={Number of Crossings},
-			 xlabel={Number of Vertices},
-			 ylabel style={yshift=-0.2cm},
-			]
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr11-20.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr21-30.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr21-30.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr31-40.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr31-40.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr41-50.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr41-50.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr51-60.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr51-60.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr61-70.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr61-70.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr71-80.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr71-80.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr81-90.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr81-90.csv};
-
-\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr91-100.csv};
-\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr91-100.csv};
-\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr91-100.csv};
-
-\end{axis}
-\end{tikzpicture}}}
-
-\subfloat[\label{fig:crossingResolution-iterations}{}]{
-\centering
-\scalebox{.95}{
-\centering
-%==================================
-% Crossing Resolution Iterations
-%==================================
-\begin{tikzpicture}
-\begin{axis}[
-			 width=\textwidth,
-			 boxplot/draw direction=y,
-			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
-			 boxplot/box extend=0.18,
-			 cycle list name=fourlist,
-			 ytick={0,5000,10000,15000,20000,25000,30000,35000,40000,45000,50000,55000,  60000,65000},
-			 ymajorgrids,
-			 xmajorgrids,
-			 xtick={0,1,2,...,9},
-			 x tick label as interval,
-     	    xticklabels={11--20,21--30,31--40,41--50,51--60,
-     	     61--70,71--80,81--90,91--100},
-     	     ylabel={Iterations},
-     	     ylabel style={yshift=-0.2cm},
-			 xlabel={Number of Vertices}]
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr11-20.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr11-20.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr21-30.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr21-30.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr21-30.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr31-40.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr31-40.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr31-40.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr41-50.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr41-50.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr41-50.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr51-60.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr51-60.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr51-60.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr61-70.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr61-70.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr61-70.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr71-80.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr71-80.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr71-80.csv};
-
-\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr81-90.csv};
-\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr81-90.csv};
-\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr81-90.csv};
-
-\addplot+[mark = x, boxplot] table [y=iterations-rm, col sep=semicolon]  {cr91-100.csv};
-\addplot+[mark = x, boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr91-100.csv};
-\addplot+[mark = x, boxplot] table [y=iterations-fa, col sep=semicolon] {cr91-100.csv};
-\addplot+[mark = x, boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr91-100.csv};
-
-\end{axis}
-\end{tikzpicture}}}
-
-\end{figure}
 
@@
 
-\addplot[name path=rm_u,black,smooth] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
-\addplot[name path=rm_d,black,smooth] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
-\addplot[black!50,fill opacity=0.5] fill between[of=rm_u and rm_d];
+\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];
 
-\addplot[name path=rm-1_u,black!50!white,smooth] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot[name path=rm-1_d,black!50!white,smooth] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
-\addplot[black!25,fill opacity=0.5] fill between[of=rm-1_u and rm-1_d];
+\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];
 
-\addplot[name path=fa_u,black,smooth,densely dashed] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
-\addplot[name path=fa_d,black,smooth,densely dashed] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
-\addplot[black!50,fill opacity=0.5] fill between[of=fa_u and fa_d];
+\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];
+
+\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-crossingResolution.csv};
+\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-crossingResolution.csv};
+\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];
+
+\addplot[blue,smooth,thick] table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[red,smooth,thick] table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[green!70!black,smooth,thick] table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};
 
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+
 \addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
@@
 \addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
+
+\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=asp_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=asp_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];
+
+\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=asp_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=asp_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];
+
+\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=asp_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=asp_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];
+
+\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-aspectRatio.csv};
+\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-aspectRatio.csv};
+\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];
+
+\addplot[blue,smooth,thick] table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[red,smooth,thick] table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[green!70!black,smooth,thick] table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
+
 \legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,50,...,350},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+
 \addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
@@
 \addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
+
+\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=cn_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=cn_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];
+
+\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=cn_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=cn_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];
+
+\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=cn_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=cn_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];
+
+\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-crossings.csv};
+\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-crossings.csv};
+\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];
+
+\addplot[blue,smooth,thick] table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[red,smooth,thick] table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[green!70!black,smooth,thick] table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
+
 \legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
-xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+xmax=100, ytick={0,2000,4000,...,14000},tick pos=left, ymajorgrids]
+
 \addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
@@
 \addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
+
+\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=it_up rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=it_down rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];
+
+\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=it_up rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=it_down rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];
+
+\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=it_up fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=it_down fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];
+
+\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-iterations.csv};
+\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-iterations.csv};
+\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];
+
+\addplot[blue,smooth,thick] table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
+\addplot[red,smooth,thick] table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
+\addplot[green!70!black,smooth,thick] table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
+\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
+
 \legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}

### 2019-07-29 paper/data/plotsWithTikz.tex
agreed drawing style.

@@
 
-\pgfplotscreateplotcyclelist{mylist}{%
+\pgfplotscreateplotcyclelist{mycoloredlist}{%
 {blue,thick,smooth},
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mycoloredlist,
 mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
 
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,50,...,350},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,50,...,350},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
 
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
 xmax=100, ytick={0,2000,4000,...,14000},tick pos=left, ymajorgrids]

### 2019-08-01 paper/data/plotsWithTikz.tex
finished rome and north figures, only grid remaining

@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
+\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,3,10,30,100,300,1000,3000,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
 
@@
 \begin{tikzpicture}
-\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
+\begin{axis}[scaled y ticks = false,axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mycoloredlist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
 xmax=100, ytick={0,2000,4000,...,14000},tick pos=left, ymajorgrids]

### 2019-08-02 paper/comments/reply-letter.tex
Most of the reviewer's comments have been addressed.
Revisions done all together.

@@
 Details are provided below, where we list all reviews together with a detailed response to each suggestion by the reviewers.
+For convenience, we also attach a diff PDF highlighting the differences between the submitted version and the revision.
 
@@
 
-\tcomment{General comment: It should be also interesting to compare the crossing resolution between the "non-grid" and the "grid" versions of your algorithm}
+\rcomment{General comment: It should be also interesting to compare the crossing resolution between the "non-grid" and the "grid" versions of your algorithm}
 
-\response{\textcolor{red}{Which version of the algorithm do we use in the experiments?}}
+\response{In the experiments, since we are comparing our approach with spring embedding algorithms, which do not produce grid drawings, we indeed report results only for the non-grid variant of our algorithm.
+Only if there exist strong constraints on the size of the underlying grid, one can observe differences between the grid and non-grid variants  in terms of the crossings resolution of the produced drawings, which are again not big (less than 2 degrees).
+Hence, plotting these differences results in two almost overlapping curves.}
 \end{itemize}
@@
 The reason is that an experimental evaluation can only give some indications.
-Instead, we now explicitly mention which are the parameters that are interested~in.}
+Instead, we now explicitly mention which are the parameters that are interested~in (indicated by P.1-P.5 in the revision).}
 
@@
 \response{As mentioned in the previous reply, our experimental evaluation aims in giving some indications.
-The mean is a good metric in this direction, even though it is sensitive to outliers.
-\textcolor{red}{For completeness, we have added a diagram illustrating the standard deviation as suggested by the reviewer.}}
+The mean is a good metric in this direction, even though it is sensitive to outliers as the reviewer points out.
+For completeness, we added in our diagrams the standard deviation as suggested by the reviewer.
+Note that we had to use colors in our figures, as otherwise the amount of information to be visualized in grayscale made the figures hard to read.}
 
@@
 In the revised version, we are more precise in our description by explicitly mentioning this.
-We have revised accordingly.}
+}
 

### 2019-08-02 paper/journal.tex
agreed drawing style.
Revisions done all together.

@@
 If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position.
-Note at that in general the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.
+Note that that in general the vertex movement paradigm is an established technique and has successfully applied to graph-theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.
 
@@
 	\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=1]{figures/rome}}
+	\includegraphics[scale=0.99,page=1]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:cr-res-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=2]{figures/rome}}
+	\includegraphics[scale=0.99,page=2]{figures/rome_colored}}
 
 	\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=3]{figures/rome}}
+	\includegraphics[scale=0.99,page=3]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=4]{figures/rome}}
+	\includegraphics[scale=0.99,page=4]{figures/rome_colored}}
 	\caption{Experimental results on the crossing resolution for the Rome graphs.}
@@
 	\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=5]{figures/rome}}
+	\includegraphics[scale=0.99,page=9]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:to-res-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=6]{figures/rome}}
+	\includegraphics[scale=0.99,page=10]{figures/rome_colored}}
 
 	\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=7]{figures/rome}}
+	\includegraphics[scale=0.99,page=11]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=8]{figures/rome}}
+	\includegraphics[scale=0.99,page=12]{figures/rome_colored}}
 	\caption{Experimental results on the total resolution for the Rome graphs.}
@@
 	\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=9]{figures/rome}}
+	\includegraphics[scale=0.99,page=5]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:an-res-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=10]{figures/rome}}
+	\includegraphics[scale=0.99,page=6]{figures/rome_colored}}
 
 	\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=11]{figures/rome}}
+	\includegraphics[scale=0.99,page=7]{figures/rome_colored}}
 	\hfil
 	\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=12]{figures/rome}}
+	\includegraphics[scale=0.99,page=8]{figures/rome_colored}}
 	\caption{Experimental results on the angular resolution for the Rome graphs.}
@@
 Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution).
-A notable observation is that, for graphs with less than $30$ vertices the best results are achieved by the algorithm by Argyriou et al., while for the remaining graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}.
-For large graphs, the two algorithms tend to have the same performance.
+Our unrestricted algorithm yields drawings with better angular resolution than the corresponding ones produced by our restricted variant, and by the algorithms by Argyriou et al.\ and by Huang et al.; see Fig.~\ref{fig:an-res-1}.
+For graphs with 90 or more vertices, however, our unrestricted algorithms and the algorithm by Argyriou et al.\ tend to produce drawings with the same angular resolution on average.
 The restricted variant of our algorithm yields drawings with slightly worse angular resolution.

### 2019-08-05 paper/journal.tex
Last changes.

@@
 \end{enumerate}
+%
+Each bold-drawn curve corresponds to the mean values. The two same-colored curves above and below represent the upper and the lower standard deviation, respectively.
 
@@
 Note that we did not impose any grid constraint on our algorithms.
-Our results are summarized in Figures~\ref{fig:cr-res}, \ref{fig:an-res} and~\ref{fig:to-res}.
+Our results are summarized in Figures~\ref{fig:cr-res}, \ref{fig:to-res} and~\ref{fig:an-res}.
 
@@
 	\subfloat[\label{fig:northCrossing-1}{Crossing resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=1]{figures/north}}
+	\includegraphics[scale=0.99,page=1]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northCrossing-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=2]{figures/north}}
+	\includegraphics[scale=0.99,page=2]{figures/north_colored}}
 
 	\subfloat[\label{fig:northCrossing-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=3]{figures/north}}
+	\includegraphics[scale=0.99,page=3]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northCrossing-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=4]{figures/north}}
+	\includegraphics[scale=0.99,page=4]{figures/north_colored}}
 	\caption{Experimental results for the crossing resolution on the AT\&T graphs.}
@@
 	\subfloat[\label{fig:northTotal-1}{Total resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=5]{figures/north}}
+	\includegraphics[scale=0.99,page=9]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northTotal-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=6]{figures/north}}
+	\includegraphics[scale=0.99,page=10]{figures/north_colored}}
 
 	\subfloat[\label{fig:northTotal-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=7]{figures/north}}
+	\includegraphics[scale=0.99,page=11]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northTotal-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=8]{figures/north}}
+	\includegraphics[scale=0.99,page=12]{figures/north_colored}}
 	\caption{Experimental results for the total resolution on the AT\&T graphs.}
@@
 	\subfloat[\label{fig:northAngular-1}{Angular resolution vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=9]{figures/north}}
+	\includegraphics[scale=0.99,page=5]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northAngular-2}{Edge-length ratio vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=10]{figures/north}}
+	\includegraphics[scale=0.99,page=6]{figures/north_colored}}
 
 	\subfloat[\label{fig:northAngular-3}{No.~of crossings vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=11]{figures/north}}
+	\includegraphics[scale=0.99,page=7]{figures/north_colored}}
 	\hfil
 	\subfloat[\label{fig:northAngular-4}{No.~of iterations vs no.~of vertices}]{
-	\includegraphics[scale=0.99,page=12]{figures/north}}
+	\includegraphics[scale=0.99,page=8]{figures/north_colored}}
 	\caption{Experimental results for the angular resolution on the AT\&T graphs.}
@@
 % ==================================================================
-\input{grid-rome.tex}
+%\input{grid-rome.tex}
 % ==================================================================
+\begin{figure*}[t!]
+	\centering
+	\subfloat[\label{fig:crossingResolutionGrid}{Crossing resolution vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=1]{figures/grid_colored}}
+	\hfil
+	\subfloat[\label{fig:crossingResolutionAspectRatioGrid}{Edge-length ratio vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=2]{figures/grid_colored}}
+
+	\subfloat[\label{fig:crossingResolutionCrossingNumberGrid}{No.~of crossings vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=3]{figures/grid_colored}}
+	\hfil
+	\subfloat[\label{fig:crossingResolutionTimeGrid}{No.~of iterations vs no.~of vertices}]{
+	\includegraphics[scale=0.99,page=4]{figures/grid_colored}}
+	\caption{Experimental results for the angular resolution on the AT\&T graphs.}
+	\label{fig:experimentsGrid}
+\end{figure*}
 
